### 2014-10-17 TableSummarization/Paper/TableSummarization.tex
Added paper

%\documentclass{vldb}
\documentclass{sig-alternate}
%\documentclass[12pt,letterpaper]{article}
%\documentclass[12pt]{extreport}
%\documentstyle[amsmath,amsthm,amssymb,twocolumn]{article}
%\newcommand{\mytitle}{Evaluating the Crowd with Confidence}

%\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\pdfpagewidth=8.5in
\pdfpageheight=11in

\renewcommand{\baselinestretch}{1.0}

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\newcommand{\agp}[1]{ \textcolor{red}{\bf ||Ins: #1||}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

\newcommand{\squishlist}{
   \begin{list}{$\bullet$}
    { \setlength{\itemsep}{0pt}
      \setlength{\parsep}{2pt}
      \setlength{\topsep}{2pt}
      \setlength{\partopsep}{0pt}
    }
}
\newcommand{\squishend}{\end{list}}
\newcommand{\eat}[1]{}
\newcommand{\papertext}[1]{#1}
\newcommand{\techreporttext}[1]{}

\title{Interactive Summarization of Relational Tables}
% Can say multi-dimensional instead of relational
% Can say Dynamic Real-Time instead of Interactive?

\author{
Manas Joglekar\\Stanford University\\manasrj@stanford.edu
\and
Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
\and
Aditya Parameswaran\\Stanford University\\adityagp@cs.stanford.edu
}
\begin{document}
\maketitle

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints. The summary is in form of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of the subset. We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table.
% TODO: Do and mention that we do Experiments to show performance/usefulness.
\end{abstract}

% Aditya comments:

\begin{comment}
Meeting points:

\end{comment}

\section{Introduction}
% Something about data being ubiquitous. Data is only as valuable as the insights we can get from it. Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it. Big amounts of data hard for human to analyze, making it necessary to summarize the data. The summary needs to be fast and interactive to facilitate exploration. Help users make sense of data. 
% Mention our problem. Then features.

% TODO : Write introduction containing points below. 
% say that paper is on summarization. 
% Salient features 
% Interactive : = Dynamic + Real-Time (see below) 
% Flexible/Customizable : Because user can choose size/ weighting function (which columns/rule sizes to emphasize).
% Dynamic : Because it modifies according to user clicks.
% Real time : Because time taken is very small compared to explananion tables (and cubes?). 
% Coverage : We emphasize coverage over surprise. (Need motivation for this). Also motivate coverage based scoing over only giving top scoring rules i.e. MCount over Count.
% Interpretable : Meaning of rules is easy to understand. (But their significance? Not sure about this).
% Compare to existing work. OLAP/Explanation tables take more time, don't focus on coverage. OLAP requires more user clicks to get same info. Explanation tables is not interactive (verify this). Weighting functions also make our system highly flexible, allowing users to prefer rules of certain sizes, or with certain columns.

In this paper, we study the problem of table summarization.

Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of \textit{rules}, with the \textit{count} denoting the number of tuples that satisfy the rule, and the \textit{size} denoting the number of non-$\star$ values of the rule) for each rule.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
target & bicycles & $\star$ & $200$ & $1$ \\ \hline
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

The first rule tells the user that there are $6000$ tuples in all. 
The next rule says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles. The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores. The last rule tells the user that walmart is doing well in general over multiple products and states.

This gives the user an overall summary of the results. But after seeing this summary, the user may want to dig deeper to get more details on certain rules. For instance, the user may want to know which states walmart has more sales in, or which products they sell the most. In this case, the user can click on the rule with walmart, and will see an expanded summary as given in table~\ref{table:introexample2}.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}

This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.

This raises some questions regarding how to score the rules, and what operations to allow the user to perform to dig deeper into the rules.
We describe our user interface in Section \ref{sec:interface}. Regarding scoring, we generally want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible. Thus scoring rules involves a tradeoff between specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$). We further discuss our scoring scheme in Section~\ref{sec:weighting}.

% TODO: Overview of paper and list of contributions

\section{Preliminaries} 
\label{sec:notation}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table. Our objective (formally defined later) is to summarise the table with a list of {\em rules}. A {\em rule} is a tuple with one field per column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each field of the rule can either be one of the values in the corresponding column of the table, or $\star$, which is a wildcard character representing all values in the column. For a table column with numerical values, we can allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all fields. A rule-list is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

When a user clicks on a rule to know more about the part of the table covered by that rule, we display a list of sub-rules of the rule. This list is called a {\em rule-list}.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule. {\em Count}($r$) is defined as the total number of tuples $t \in T$ that are covered by $r$. Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list. Let {\em Size}($r$) refer to the number of non-starred values in rule $r$. We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} \text{MCount}(r)W(r)$$.

We now formally define our problem:

Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r)\text{MCount}(r)$$
subject to the following constraints:
\begin{enumerate}
\item If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{enumerate}

% Describe architecture here? (With a diagram). Show SampleHandler, QueryEngine, etc.

\section{Weighting Rules}
\label{sec:weighting}
We want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is. Given this weighting, our total score for our rule-list is given by : $$\sum_{r\in R} \text{MCount}(r)W(r)$$. Let us consider some simple kinds of $W(r)$s and their meanings:
\begin{enumerate}
\item $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :

Here we set weight equal to the size of the rule. Consider the examples in Table \ref{table:sizescoringexample}. Weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus when we are going to marginalize on one extra attribute ($B$ in this case) it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an interpretation for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with MCount $100$, we are going to get a $100$ of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those. Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-set. Since having more of the table pre-filled is better, maximising score gives us a desirable set of rules.

\item $W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the 'complexity' of the column. The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes log($|c|$) bits of information. A non-starred valued in a rule $r$ thus pre-fills $\text{MCount}(r) \text{log}(|c|)$ bits of the table. Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.

Note that this scoring function is closely related to the Minimum Description Length (MDL) of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score. 
\end{enumerate}

Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
\begin{enumerate}
\item Non-negativity : For all rules $r$, $W(r) \geq 0$ 
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\end{enumerate}

\begin{table}
\centering
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\end{table}

\section{Our interface}
\label{sec:interface}
We now explain how the user can interact with our system to explore the contents of a table.

When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}. At any point, the user can click on either a rule, or a star, to 'expand' the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of sub-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule would display table~\ref{table:introexample}. Clicking further on the second rule would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows sub-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Table with only the trivial rule \label{table:trivial}}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r$, we want to find the highest scoring list of rules to expand rule $r$ into. We now formally state the problem:

Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes the total score given by :
$$\sum_{r \in R}W(r)\text{MCount}(r)$$

$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily understandable summary of the table. We now describe a greedy algorithm for finding the approximately highest scoring rule-list.

The process of finding the rule-list can be broken down into two steps: Finding the set of rules to put in the list, and then ordering them. We first note that once the set of rules is given, then the optimal ordering is to order them in descending order by weight, since each tuple covered by one of the rules can contribute $1$ to the MCount of exactly one of the rules, and it can contribute the most to total score by assigning it the the rule with highest weight.

Thus the hard part of the problem is finding a set of rules that maximises total score when ordered in decreasing order by weight. We solve this part of the problem next.

\subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
Define the score of a set of rules $R$ to be equal to the score for the list obtained by sorting the set elements in decreasing order by weight. The key thing to note here is that the rule weight $W(r)$ depends on the rule alone and not on other rules or on distribution of column values in the database table. The total score is given by $\sum_{r \in R} \text{Count}(r)W(r)$ and thus rule $r$ 'contributes' MCount($r$)$W(r)$ to the total score. Thus the only part of the contribution that depends on the other rules in $R$ is MCount($r$) (the dependence is because other rules in $R$ may cover some rules that $r$ covers, causing MCount($r$) to go down). We assume that rules in the rule set are ordered in decreasing value of $W(r)$, and this a tuple is always counted for the first rule that covers it. In that case, note that if $r \in R_1 \subset R_2$, then Count($r$) in $R_1$ is going to be $\geq \text{Count}(r)$ in $R_2$. Thus the marginal value that $r$ adds to the total score is higher for $R_1$ than for $R_2$, and hence the total score for set $R$ is submodular.

\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}

Intuitively, this means that the marginal value of adding an element to a set cannot increase if we first add other elements to a set. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way. Specifically, consider the greedy algorithm:

\begin{framed}
\begin{enumerate}
\item Set $R = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($R \cup \left\lbrace r \right\rbrace$) is the highest.
\item $R = R \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\end{framed}

Because Score is a submodular function of the set $R$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm for Rule-Based Scoring Functions}. The algorithm takes four parameters as input: the table $T$, the number $n$ of rules required in the final solution list, a parameter $m_w$ which stands for \textit{Max Weight}, and the weight function $W$. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Smaller values of $m_w$ will cause the algorithm to run faster, but make it more likely to miss out on high-weight rules in the final solution.

The algorithm initializes the solution set $S$ to be empty, and then iterates for $n$ steps, adding the best marginal rule at each step. To find the best marginal rule, it calls the function in the box labelled \textbf{Find best marginal rule}.

Each call of \textbf{Find best marginal rule} needs to find the marginal counts of several rules to choose the best one. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory. In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}. We count the rule over several passes, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts for rules of size $j$. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted so far. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous pass. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

For the second pass onwards, we are more selective about which rules to consider for marginal value evaluation. We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$. Then for each rule $r$ from $C_n$, we consider the known marginal values of it's sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2. Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule. Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5. If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

\begin{framed}
\textbf{Find best marginal rule}

\textbf{Input :} $S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)

\textbf{Output :} $R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)
\begin{enumerate}
\item $H = 0$ // Threshold for deciding if to find count for a rule.
\item $C = C_o = C_n = \phi$ // Set of all, old and new candidate rules respectively.
\item For $j$ from $1$ to number of columns in $T$:
\begin{enumerate}[label*=\arabic*.]
\item if $j = 1$
\begin{enumerate}[label*=\arabic*.]
\item $C_n = \text{ all rules of size } 1$
\end{enumerate}
\item else 
\begin{enumerate}[label*=\arabic*.]
\item $C_n = \text{all size-}i$ super-rules of rules from $C_o$
\end{enumerate}
\item For each $R \in C_n$
\begin{enumerate}[label*=\arabic*.]
\item $M =\infty$ // Upper bound on marginal value count of $R$
\item For each $R$-sub-rule $R^{\prime} \in C$  
\begin{enumerate}[label*=\arabic*.]
\item $M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
\end{enumerate} 
\item if $(M < H)$ 
\begin{enumerate}[label*=\arabic*.]
\item $C_n = C_n \setminus \left\lbrace R \right\rbrace$ // Delete $R$ if its max count is too small for $R$ to possibly be in the solution.
\end{enumerate} 
\end{enumerate}
\item if $C_n = \phi$
\begin{enumerate}[label*=\arabic*.]
\item break;
\end{enumerate}
\item For each $R \in C_n$
\begin{enumerate}[label*=\arabic*.]
\item Count$(R) = 0$ // Initialize to zero
\item MarginalValue$(R) = 0$ // Initialize to zero
\end{enumerate}
\item For each tuple $t \in T$
\begin{enumerate}[label*=\arabic*.]
\item Let $R_S$ be the highest weight rule in $S$ that covers $t$
\item For each rule $R \in C_n$ that covers $t$
\begin{enumerate}[label*=\arabic*.]
\item Count($R$) $++$
\item MarginalValue($R$) $+= W(R) - \text{min}(W(R), W(R_S))$
\end{enumerate}
\end{enumerate}
\item $C_o = C_n$
\item $C_n = \phi$
\item $C = C \cup C_o$
\item $H = \text{max}_{R \in C}(\text{MarginalValue}(R))$
\end{enumerate}
\item return $\text{argmax}_{r \in C} \text{marginalValue}(r)$
\end{enumerate}
\end{framed}

\begin{framed}
\textbf{Greedy Algorithm for Rule-Based Scoring Functions}

\textbf{Input :} $n$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)

\textbf{Output :} $S$ (Solution set of rules)
\begin{enumerate}
\item $S = \phi$ // Initialize solution set to null set.
\item For $i$ from $1$ to $n$:
\begin{enumerate}[label*=\arabic*.]
\item $R_m = \text{Find best marginal rule}(S, T, m_w, W)$
\item $S = S \cup \left\lbrace R_m \right\rbrace$
\end{enumerate}
\item Return $S$
\end{enumerate}

\end{framed}

\subsection{Supporting user operations}
Our greedy algorithm finds the best set of rules to cover a table, as per some monotonic weight function $W$. We now describe how we use our algorithm to support the operations allowed in our user interface:
\begin{enumerate}
\item \textbf{Expand Rule} : When the user tries to expand rule $r$, we want to display the highest scoring set of sub-rules of $r$. In order to do this, we make a pass through table $T$ to select tuples covered by $r$, and load them into a new table $T^{\prime}$. Then we run the greedy algorithm on $T^{\prime}$ to get the set of rules to display to the user.
\item \textbf{Expand Star} : When the user tries to expand the $\star$ value in a column $c$ of rule $r$, we again create table $T^{\prime}$ consisting of tuples of $T$ that are covered by $r$. In addition, we use weight function $W^{\prime}$ such that $W^{\prime}(x) = 0$ if rule $x$ has a $\star$ in column $c$, and $W^{\prime}(x) = W(x)$ otherwise. Then we run the greedy algorithm on table $T^{\prime}$ with weight function $W^{\prime}$.
\end{enumerate}

\subsection{Sampling for big tables}
%TODO: Read and improve flow of this section.
Our greedy algorithm describes how to get a near optimal set of rules for a given table. But in addition to getting the best rules for the table, we want to find the best sub-rules of a given rule $r$ (when the user clicks on $r$). Iterating over the entire table to find counts when we only need the small part of the table that is covered by $r$ is wasteful. We can instead compute and materialize the set of tuples covered by $r$, and run our algorithm on that set.

Our algorithm also needs to make multiple passes through the table to find counts and marginal values of rules. This can be slow if the table size is high. In that case, if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.

We handle both the above problems, of materializing sets of tuples covered by a rule, and of tables that don't fit in main memory, by using sampling. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
The $minSS$ parameter is the minimum umber of sample tuples we are required to find counts on while running our greedy algorithm, and it determines how accurate our count estimates will be. We provide a way to find values of $minSS$ later.

At all points, the SampleHandler maintains a set of Table Samples in memory. Each table sample $s$ has the following attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate counts of a rule on the sample, into estimated counts on the entire table. The sum of $|T_s|$ is not allowed to exceed capacity $M$ at any point.

Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ with $f_s = r$ and $|T_s| \geq minSS$. Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get the estimated count on the entire table.

When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule. There are multiple ways it could do this:
\begin{enumerate}
\item \textbf{Find:} If it finds an existing sample with $r$ as the filter rule, and at least $minSS$ tuples, it simply returns that sample.
\item \textbf{Combine:} If \textbf{Find} doesn't work i.e. if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$. We can show that tuples from $T_{s^{\prime}}$ that are covered by $r$ can be taken to be uniformly sample tuples covered by $r$. If the union of such tuples over all existing samples exceeds $minSS$ in size, then the SampleHandler can create a sample using existing samples, instead of reading the entire table to create a new sample
\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the the SampleHandler needs to create a new sample by making a pass through the table. Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used. We can use reservoir sampling to get a uniformly random sample of given size in a single pass through the table. In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available). Making a larger sample is advantageous not only because due to higher accuracy, but because when the user later expands a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
\item \textbf{Delete/Shrink:} Finally, it may happen that there is not enough memory left to create a new sample. In that case, the SampleHandler needs to delete or shrink an existing sample. It can use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink. Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
%For shrink, prefer shrinking samples of the largest size (> minSS) first?
\end{enumerate}

When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user. If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table. But after expanding $r$, there is a high chance that the user goes further and expands one of the sub-rules $r^{\prime}$ of $r$. We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples. So while the user is reading the current rule-list from expanding $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background. That way, when the user expands $r^{\prime}$, the new loaded tuples for $r$ will increase that chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time. 
In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.

% Need to find and write about the exact algorithm to decide how much to load, what to shrink, and how much to shrink.

\subsubsection{Setting $minSS$}
% TODO: Write bound on reasonable value of minSS

% Additional optimizations. In sample tuples, store only those columns that are not in filter rule. If tuples are repeated across samples, then store pointer to the tuple instead.

%Independent. Can we bunch together rare values into a single value called 'rare value'?

\section{Experiments}
\subsection{Datasets}
The first dataset we use is a marketing dataset (MD), which contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

The columns (in order) are as follows:
(Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, How long have you lived in the San Fran./Oakland/San Jose Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home).

\subsection{User interface testing}
We restrict the table to the first $10$ columns in order to make the result tables fit into the page. Projecting onto $10$ columns reduces the number of distinct tuples to $2851$. We set the $n$ (number of rules) parameter to $4$, and $m_w$ to $5$. We now present the rule-based summaries displayed as a result of a few different user actions.

To begin with, the user sees an empty rule with the total number of tuples as the count. After this, suppose the user expands the rule. Then the user will see Table~\ref{table:uiexample1}. The first two new rules simply tell us that the table has $1501$ female and $1350$ male tuples. The next two rules also slightly more detailed, saying that there are $577$ females who have been in the Bay Area for $> 10$ years, and $158$ males aged $18-24$ who have never been married. Note that the latter two rules give very specific information which would require $3$ user clicks to find in an OLAP system, where our system displays that information to the user with a single click.

Now suppose the user decides to further explore the table, by looking at education related information of females in the dataset. So the user clicks on the $\star$ in the 'Education' column of the second rule. This opens up table~\ref{table:uiexamplestar}. It shows the number of females with different levels of education.

Instead of expanding the 'Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.

\subsection{Weighting functions}
Our system can display optimal rule summaries using any monotone weighting function. By default, we assign a rule weight equal to it's size. In this section, we try a couple of other weighting functions.

We first try the weighting function given by:
$$W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$$ where $|c|$ refers to the number of distinct possible values in column $c$. This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values. The rule summary for this weighting is in table~\ref{table:weigtingbitwise}. The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column. Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead.

The other weighting function we try is given by:
$$W(r) = \text{Min}(0, \text{Size}(r) - 1)$$
This gives us table~\ref{table:weightingsizeminusone}. This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values. As a result, we can see that our system displays rules having $2$ or $3$ non-$\star$ values.

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
\hline 
\end{tabular} 
\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
\end{table*}

\begin{table*}
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $378$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $271$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $416$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & Grad Study & $\star$ & $\star$ & $171$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
\hline 
\end{tabular} 
\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
\end{table*}

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $467$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $408$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Divorced/separated & $\star$ & $\star$ & $\star$ & $\star$ & $187$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $485$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
\hline 
\end{tabular} 
\caption{A Rule expansion \label{table:uiexamplerule}} 
\end{table*}

\begin{table*}
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $858$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & Married & $\star$ & $\star$ & $\star$ & $\star$ & $917$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $1062$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & $\star$ & $\star$ & $\star$ & Professional/Managerial & $\star$ & $637$ & $1$ \\
\hline 
\end{tabular} 
\caption{Bits scoring\label{table:weigtingbitwise}} 
\end{table*}

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 25-34 & $\star$ & $\star$ & $\star$ & $145$ & $3$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $137$ & $3$ \\
\hline 
\end{tabular} 
\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
\end{table*}

\begin{comment}
Notes based on Explanation Tables paper: Prove that our problem is NP-Hard to solve exactly. Something like the Set Cover Problem. Yes, that is easy.

The way we deal with overlap, is that the key differentiating factor?

We don't have the binary thing. Is it useful? Maybe we can summarize better without it. We are solving a different problem, optimising for coverage rather than explaining the binary attribute. They are optimising for KL divergence of max entropy distribution. They mention many other papers that optimize for cardinality or similar. We do weight times cardinality, which is a compromise between descriptiveness and coverage, and allow user to change the weight function. But should check those papers out anyway.

\end{comment}

\section{Extensions}
\subsection{Dealing with Numerical Attributes}
Our algorithm assumes that all attributes are categorical in nature. Attributes that have a large domain d to have a smaller tuple count per value, and hence don't appear in rule summaries. Thus our algorithm does not summarise information about numerical attributes.

However, we can modify the algorithm to deal with numerical attributes. Suppose we have a numerical attribute $A$. A simple approach is to create buckets for values of $A$. We choose a number of buckets $k$, and divide the range of values of $A$ into $k$ intervals, each corresponding to a bucket. We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket. Then we can use our algorithm, treating the bucket number as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).

The distribution of values for the $A$ may not be similar in different parts of the table. For instance, we may create buckets have approximately equal numbers of tuples in the table. But then suppose a user tries to expand a rule $r$. The part of the table covered by $r$ might have a very different distribution of age values. For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment. To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.

\section{Related Work}
%TODO: Repeat what we said in intro about interactive, etc.
There has been work on finding interesting rules in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}. The work mainly focuses on finding values that occur more often or less often that expected from a max entropy distribution. The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.

There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}. This work again focuses on displaying rules that will cause the max entropy distribution to best approximate the actual distribution of values.

Our algorithm uses ideas from the a priori algorithm~\cite{apriori}. Several extensions for the algorithm have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311}. We can potentially use these ideas to improve handing of numerical attributes in our work.

{\small 
\bibliographystyle{plain}
\bibliography{TableSummarization}
}

\end{document}

### 2014-10-20 TableSummarization/Paper/TableSummarization.tex
Updated writeup with intro and more on sampling

@@
 \newtheorem{definition}{Definition}
+\newtheorem{example}[definition]{Example}
 
@@
 }
+\newcommand{\stitle}[1]{\vspace{0.5em}\noindent\textbf{#1}}
 \newcommand{\squishend}{\end{list}}
@@
 We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table.
-% TODO: Do and mention that we do Experiments to show performance/usefulness.
+Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
+% TODO: Improve abstract.
 \end{abstract}
 
-% Aditya comments:
+\section{Introduction}
 
-\begin{comment}
-Meeting points:
+Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it.
+Large amounts of data hard for human to make sense of.
+This necessitates summarizing the data in an informative, interpretable way.
+The summary also needs to be fast and interactive to facilitate deeper exploration.
 
-\end{comment}
+In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints.
+A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}.
+Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a rule-scoring black box as user input.
+In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.
 
-\section{Introduction}
-% Something about data being ubiquitous.
-Data is only as valuable as the insights we can get from it.
-Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it.
-Big amounts of data hard for human to analyze, making it necessary to summarize the data.
-The summary needs to be fast and interactive to facilitate exploration.
-Help users make sense of data.
-% Mention our problem. Then features.
-
-% TODO : Write introduction containing points below. 
-% say that paper is on summarization. 
-% Salient features 
-% Interactive : = Dynamic + Real-Time (see below) 
-% Flexible/Customizable : Because user can choose size/ weighting function (which columns/rule sizes to emphasize).
-% Dynamic : Because it modifies according to user clicks.
-% Real time : Because time taken is very small compared to explananion tables (and cubes?). 
-% Coverage : We emphasize coverage over surprise. (Need motivation for this). Also motivate coverage based scoing over only giving top scoring rules i.e. MCount over Count.
-% Interpretable : Meaning of rules is easy to understand. (But their significance? Not sure about this).
-% Compare to existing work.
-OLAP/Explanation tables take more time, don't focus on coverage.
-OLAP requires more user clicks to get same info.
-Explanation tables is not interactive (verify this).
-Weighting functions also make our system highly flexible, allowing users to prefer rules of certain sizes, or with certain columns.
-
-In this paper, we study the problem of table summarization.
-
+\begin{example}\label{ex:introexample}
 Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'.
@@
 This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.
+\end{example}
 
-This raises some questions regarding how to score the rules, and what operations to allow the user to perform to dig deeper into the rules.
+The above example raises some questions regarding how to score the rules, and what operations to allow the user to perform to dig deeper into the rules.
 We describe our user interface in Section \ref{sec:interface}.
@@
 
-% TODO: Overview of paper and list of contributions
+There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}.
+This work mostly focuses on giving he most 'surprising' information to the user i.e.
+information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution.
+Our algorithm, on the other hand, focuses on coverage, describing as much of the table as possible.
+
+%??Something about how users can't imagine max-entropy distribution??.
+
+OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'.
+We focus on only obtaining counts for rules which cover a significant number of tuples, to get a fast response time.
+We compute more detailed rules on the fly, when the user explores the table further.
+In addition, we often display rules containing multiple non-$\star$ values in our summary (such as the (target, bicycles) rule in our example) is their score is high enough, whereas a user would need to multiple clicks in an OLAP system to get the same information.
+Finally, our algorithm has the flexibility of allowing the user to specify a rule-weighting function, in order to preferentially choose rules of a particular size, or containing information about a particular column that the user is interested in.
+
+% TODO : Need motivation for coverage over surprise. Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
+\stitle{Overview of paper:} 
+\squishlist
+
+\item In Section 2, we introduce the relevant notation and formally define our problem.
+
+\item In Section 3, we consider different schemes for weighting rules, along with their practical significance.
+
+\item In Section 4, we describe our user interface.
+
+\item In Section 5, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.
+
+\item In Section 6, we experimentally evaluate the performance of our algorithms a real dataset.
+\squishend
 
 \section{Preliminaries} 
-\label{sec:notation}
+\label{sec:preliminaries}
 Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table.
@@
 Our greedy algorithm finds the best set of rules to cover a table, as per some monotonic weight function $W$.
-We now describe how we use our algorithm to support the operations allowed in our user interface:
+We now describe how we can use our algorithm to support the operations allowed in our user interface:
 \begin{enumerate}
@@
 \subsection{Sampling for big tables}
-%TODO: Read and improve flow of this section.
-Our greedy algorithm describes how to get a near optimal set of rules for a given table.
-But in addition to getting the best rules for the table, we want to find the best sub-rules of a given rule $r$ (when the user clicks on $r$).
-Iterating over the entire table to find counts when we only need the small part of the table that is covered by $r$ is wasteful.
-We can instead compute and materialize the set of tuples covered by $r$, and run our algorithm on that set.
-
-Our algorithm also needs to make multiple passes through the table to find counts and marginal values of rules.
-This can be slow if the table size is high.
-In that case, if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
+Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
+These passes can be very expensive if the table is large, especially if it does not fit in main memory.
+If we want exact counts for rules, we have not choice but to read the entire table.
+But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
 If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.
 
-We handle both the above problems, of materializing sets of tuples covered by a rule, and of tables that don't fit in main memory, by using sampling.
+Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
 Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
 The $minSS$ parameter is the minimum umber of sample tuples we are required to find counts on while running our greedy algorithm, and it determines how accurate our count estimates will be.
-We provide a way to find values of $minSS$ later.
+We provide a way to find reasonable values of $minSS$ later.
 
@@
 \begin{enumerate}
-\item \textbf{Find:} If it finds an existing sample with $r$ as the filter rule, and at least $minSS$ tuples, it simply returns that sample.
+\item \textbf{Find:} If the SampleHandler finds an existing sample with $r$ as the filter rule, and at least $minSS$ tuples, it simply returns that sample. 
 \item \textbf{Combine:} If \textbf{Find} doesn't work i.e.
@@
 We can show that tuples from $T_{s^{\prime}}$ that are covered by $r$ can be taken to be uniformly sample tuples covered by $r$.
-If the union of such tuples over all existing samples exceeds $minSS$ in size, then the SampleHandler can create a sample using existing samples, instead of reading the entire table to create a new sample
-\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the the SampleHandler needs to create a new sample by making a pass through the table.
+That is, each tuple covered by $r$ is equally likely to appear in a uniform sample of $f_{s^{\prime}}$.
+If the union of such tuples over all such available $s^{\prime}$s exceeds $minSS$ in size, then the SampleHandler can create a sample using existing the samples, instead of reading the entire table to create a new sample.
+\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample by making a pass through the table.
 Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
@@
 In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available).
-Making a larger sample is advantageous not only because due to higher accuracy, but because when the user later expands a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
+Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later expands a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
+For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
 \item \textbf{Delete/Shrink:} Finally, it may happen that there is not enough memory left to create a new sample.
@@
 So while the user is reading the current rule-list from expanding $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
-That way, when the user expands $r^{\prime}$, the new loaded tuples for $r$ will increase that chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
+That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
 In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
 
-% Need to find and write about the exact algorithm to decide how much to load, what to shrink, and how much to shrink.
+% TODO: Need to find and write about the exact algorithm to decide how much to load, what to shrink, and how much to shrink. Different heuristic algos are a good experimental subject as well.
+
+\subsubsection{Additional optimizations}
+There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
+Suppose we have a sample $s$, and say it's filter rule $f_s$ has value $v$ in column $c$.
+Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$.
+So we do not need to explicitly store the column $c$ value of any tuple in $T_s$.
+We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
+In addition, we may have a tuple occur in multiple samples, especially when we use \textbf{Combine} to create a sample.
+Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
 
 \subsubsection{Setting $minSS$}
-% TODO: Write bound on reasonable value of minSS
+Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e.
+$x|T|$ tuples.
+Say we have a uniform random sample $s$ of $T$.
+The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$.
+Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$.
+In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$.
+That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.
+So the value of minSS must be at least $c\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary, and $c$ is a constant chosen by us based on how accurate we want the count estimate to be.
 
-% Additional optimizations. In sample tuples, store only those columns that are not in filter rule. If tuples are repeated across samples, then store pointer to the tuple instead.
-
-%Independent. Can we bunch together rare values into a single value called 'rare value'?
+Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
+This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule).
+Let $c$ be the column with the fewest distinct values.
+Say it has $|c|$ values.
+Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$.
+The highest scoring rule can have score at most $|C|$ (the total number of columns), the the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
+Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
 
@@
 
+% Aditya's idea of flexible ranges. Score depends on range length.
+
 \section{Related Work}

### 2014-10-25 TableSummarization/Paper/TableSummarization.tex
Added Sampling scheme

@@
 In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints.
-The summary is in form of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of the subset.
+The summary consists of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of that subset.
 We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table.
@@
 Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it.
-Large amounts of data hard for human to make sense of.
+Large amounts of data are hard for a human to make sense of.
 This necessitates summarizing the data in an informative, interpretable way.
@@
 \begin{example}\label{ex:introexample}
+% Need justification for how investor got data. Or change example.
 Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'.
@@
 For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}.
-The summary consists of a list of \textit{rules}, with the \textit{count} denoting the number of tuples that satisfy the rule, and the \textit{size} denoting the number of non-$\star$ values of the rule) for each rule.
+The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `size' denoting the number of non-$\star$ values of the rule for each rule.
 
+%TODO: Change indentation format. All indentation at start, number of indents equal to depth of rule
 \begin{table}
@@
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-target & bicycles & $\star$ & $200$ & $1$ \\ \hline
+target & bicycles & $\star$ & $200$ & $2$ \\ \hline
 $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
@@
 
+% Things to work into example : Number of rules. Ordering.
+
 The first rule tells the user that there are $6000$ tuples in all.
-The next rule says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles.
+We call it an `empty' rule, because it has only $\star$ values.
+The three additional rules give more information about the table.
+The number of additional rules ($3$ in this case) is a parameter that can be set by the user.
+The rules are shown in decreasing order by size.
+
+The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles.
 The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores.
@@
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
-target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
+target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
 $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
@@
 \end{table}
+% TODO: Should we display weight rather than size in table? After all, size is obvious. In intro example, we say that for this table, weight equals number of non-$\star$ columns.
 
+The three new rules in this table give the user additional information about the $1000$ walmart tuples.
 This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.
+
+In our summary, we want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible.
+Thus choosing rules involves a tradeoff between highly specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$).
+Our decision on what rules to choose is influenced by a rule-weighting scheme, which is further discussed in Section~\ref{sec:weighting}.
+Briefly, we allow the user to specify a weighting function $W$ as a black box, and display the optimal set of rules as per that weighting function.
+
+This example also raises questions on what operations to allow the user to perform to dig deeper into the rules. We describe our user interface in Section \ref{sec:interface}. 
 \end{example}
 
-The above example raises some questions regarding how to score the rules, and what operations to allow the user to perform to dig deeper into the rules.
-We describe our user interface in Section \ref{sec:interface}.
-Regarding scoring, we generally want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible.
-Thus scoring rules involves a tradeoff between specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$).
-We further discuss our scoring scheme in Section~\ref{sec:weighting}.
+To summarize, we attempt to summarize relational data in a way that is:
+\squishlist
+\item \textbf{Flexible : } We allow users to specify a weighting function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
+\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
+\item \textbf{Online : } Does not require any heavy pre-computation. 
+\squishend
 
 There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}.
-This work mostly focuses on giving he most 'surprising' information to the user i.e.
+These papers mostly focuses on giving the most 'surprising' information to the user i.e.
 information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution.
-Our algorithm, on the other hand, focuses on coverage, describing as much of the table as possible.
-
-%??Something about how users can't imagine max-entropy distribution??.
+For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user.
+In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
+% TODO: Say it's complementary.
+Also give an example where your thing is good.
+Mention MDL here.
+Or that max-entropy is hard to interpret for human, even though formally very clear.
+Our thing also works well when user doesn't know the set of column values.
+The set, at least number of column values, are needed for imagining the max-entropy distribution.
 
 OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'.
-We focus on only obtaining counts for rules which cover a significant number of tuples, to get a fast response time.
-We compute more detailed rules on the fly, when the user explores the table further.
-In addition, we often display rules containing multiple non-$\star$ values in our summary (such as the (target, bicycles) rule in our example) is their score is high enough, whereas a user would need to multiple clicks in an OLAP system to get the same information.
-Finally, our algorithm has the flexibility of allowing the user to specify a rule-weighting function, in order to preferentially choose rules of a particular size, or containing information about a particular column that the user is interested in.
+They have an operation called drill-down, which allows a user to get counts for rules having values in a given column.
+For instance in our example, a user can first drill down along the `Department Store' column, which will show the total number of tuples for each Department Store.
+After that, the user may drill down along the product column for `Walmart', which will show the user the number of tuples for each product in Walmart.
+Thus the user is required to decide which columns to drill down along in order to get interesting or frequent rules, and requires $k$ drill downs to get to rules of size $k$.
+In contrast, our summary may immediately display a size $k$ rule to the user if it has sufficiently high weight and count.
+There is no notion of ranking or ordering of rules in OLAP, since all possible rules are displayed for any drill down.
+In contrast, we only display the top few rules in terms of their score, and display them in decreasing order by weight.
+This makes the summary shorter and more readable by focusing only on the frequently occurring values in columns.
 
-% TODO : Need motivation for coverage over surprise. Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
+% TODO : Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
 \stitle{Overview of paper:} 
@@
 Our objective (formally defined later) is to summarise the table with a list of {\em rules}.
+
 A {\em rule} is a tuple with one field per column of the queried table.
@@
 A {\em trivial rule} is one that has a $\star$ value in all fields.
-A rule-list is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user.
 A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple.
 We abuse notation to write this as $t \in r$.
+
+A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user.
 
@@
 
-When a user clicks on a rule to know more about the part of the table covered by that rule, we display a list of sub-rules of the rule. This list is called a {\em rule-list}.
+When a user clicks on a rule to know more about the part of the table covered by that rule, we display a rule-list of sub-rules of the rule.
 For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule.
@@
 Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
+A high value of MCount indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules.
 We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list.
@@
 \label{sec:algorithms}
-When the user expands a rule $r$, we want to find the highest scoring list of rules to expand rule $r$ into. We now formally state the problem:
+When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. We now formally state the problem:
 
-Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes the total score given by :
-$$\sum_{r \in R}W(r)\text{MCount}(r)$$
+Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ sub-rules of $r^{\prime}$ that maximizes the total score given by :
+$$\text{Score}(R) = \sum_{r \in R}W(r)\text{MCount}(r)$$
 
 $W$ is required to be non-negative and monotonic.
-$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily understandable summary of the table.
-We now describe a greedy algorithm for finding the approximately highest scoring rule-list.
+$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.
 
-The process of finding the rule-list can be broken down into two steps: Finding the set of rules to put in the list, and then ordering them.
-We first note that once the set of rules is given, then the optimal ordering is to order them in descending order by weight, since each tuple covered by one of the rules can contribute $1$ to the MCount of exactly one of the rules, and it can contribute the most to total score by assigning it the the rule with highest weight.
+We will describe a greedy algorithm for finding the approximately highest scoring rule-list.
+First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
 
-Thus the hard part of the problem is finding a set of rules that maximises total score when ordered in decreasing order by weight. We solve this part of the problem next.
+\begin{lemma}\label{lemma:rule-ordering}
+Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
+$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
+\end{lemma}
+\begin{proof}
+We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$.
+For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to.
+Then Score can be rewritten as:
+$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
+Thus each tuple $t$ `contributes' $W(t)$ to the total score.
+Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$.
+Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
+$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
+as required.
+\end{proof}
+
+Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
+Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order.
+We want to find the $S$ that maximises $\text{Score}(S)$.
+We solve this part of the problem next.
 
 \subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
-Define the score of a set of rules $R$ to be equal to the score for the list obtained by sorting the set elements in decreasing order by weight.
-The key thing to note here is that the rule weight $W(r)$ depends on the rule alone and not on other rules or on distribution of column values in the database table.
-The total score is given by $\sum_{r \in R} \text{Count}(r)W(r)$ and thus rule $r$ 'contributes' MCount($r$)$W(r)$ to the total score.
-Thus the only part of the contribution that depends on the other rules in $R$ is MCount($r$) (the dependence is because other rules in $R$ may cover some rules that $r$ covers, causing MCount($r$) to go down).
-We assume that rules in the rule set are ordered in decreasing value of $W(r)$, and this a tuple is always counted for the first rule that covers it.
-In that case, note that if $r \in R_1 \subset R_2$, then Count($r$) in $R_1$ is going to be $\geq \text{Count}(r)$ in $R_2$.
-Thus the marginal value that $r$ adds to the total score is higher for $R_1$ than for $R_2$, and hence the total score for set $R$ is submodular.
-
+We will now show that the Score function over sets of rules has a property called submodularity, which gives us a greedy approximation algorithm for optimizing it. 
 \begin{definition}
@@
 \end{definition}
+Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead.
+For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.
 
-Intuitively, this means that the marginal value of adding an element to a set cannot increase if we first add other elements to a set.
-For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.
-Specifically, consider the greedy algorithm:
+\begin{lemma}\label{lemma:submodular}
+For a given table $T$, the Score function over sets $S$ of rules, defined by :
+$$\text{Score}(S) = \sum_{r \in S} \text{MCount}(r)W(r)$$
+is submodular.
+\end{lemma}
+\begin{proof}
+For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
+$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
+Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
+$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
+For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
+$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$. Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
+$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.
+
+Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
+$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
+$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
+which means that Score is submodular.
+\end{proof}
+
+Now consider the greedy algorithm:
 
@@
 \begin{enumerate}
-\item Set $R = \phi$
+\item Set $S = \phi$
 \item For $i$ from $1$ to $k$
 \begin{enumerate}
-\item Find the rule $r$ for which Score($R \cup \left\lbrace r \right\rbrace$) is the highest.
-\item $R = R \cup \left\lbrace r \right\rbrace$
+\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
+\item $S = S \cup \left\lbrace r \right\rbrace$
 \end{enumerate}
@@
 
-Because Score is a submodular function of the set $R$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
+Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
 
-Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm for Rule-Based Scoring Functions}.
-The algorithm takes four parameters as input: the table $T$, the number $n$ of rules required in the final solution list, a parameter $m_w$ which stands for \textit{Max Weight}, and the weight function $W$.
+Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm}.
+The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
+
+$m_w$ stands for \textit{Max Weight}.
 The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
-Smaller values of $m_w$ will cause the algorithm to run faster, but make it more likely to miss out on high-weight rules in the final solution.
+Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, our algorithm is guaranteed to return $S_o$.
+On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$.
+Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w \geq \textrm{max}_{r \in S_o}W(r)$.
+In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
+This is because as the size (and weight) of a rule increases, it's Count falls rapidly.
+The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$).
+Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
+Our experiments also show that the weights of rules in the optimal set are small.
 
-The algorithm initializes the solution set $S$ to be empty, and then iterates for $n$ steps, adding the best marginal rule at each step.
-To find the best marginal rule, it calls the function in the box labelled \textbf{Find best marginal rule}.
+% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary.
+Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.
 
-Each call of \textbf{Find best marginal rule} needs to find the marginal counts of several rules to choose the best one.
+The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty in Step $1$, and then iterates for $k$ steps in Step $2$, adding the best marginal rule at each step.
+To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.
+
+In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
+A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data.
 But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.
+
 In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}.
-We count the rule over several passes, with the maximum number of passes equal to the maximum size of a rule.
+We compute the best marginal rule over several passes (Step $3$), with the maximum number of passes equal to the maximum size of a rule.
 In the $j^{th}$ pass, we compute counts for rules of size $j$.
+To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$.
+In the first pass, we would compute the marginal values of all rules of size $1$.
+In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass.
+For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$.
+Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$.
+This implies that the maximum marginal value of any super-rule of $(\star, \star, c)$ having weight $\leq m_w$ is at most $100m_w$.
+If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$.
+Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.
+
+We now describe the function to find the best marginal rule.
+The pseudo-code for the function is in the box titled \textbf{Find best marginal rule}.
+The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
+The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration.
 We maintain three sets of rules.
@@
 \begin{framed}
-\textbf{Greedy Algorithm for Rule-Based Scoring Functions}
+\textbf{Greedy Algorithm}
 
-\textbf{Input :} $n$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)
+\textbf{Input :} $k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)
 
@@
 \item $S = \phi$ // Initialize solution set to null set.
-\item For $i$ from $1$ to $n$:
+\item For $i$ from $1$ to $k$:
 \begin{enumerate}[label*=\arabic*.]
@@
 These passes can be very expensive if the table is large, especially if it does not fit in main memory.
-If we want exact counts for rules, we have not choice but to read the entire table.
+If we want exact counts for rules, we have no choice but to read the entire table.
+
 But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
@@
 Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
-The $minSS$ parameter is the minimum umber of sample tuples we are required to find counts on while running our greedy algorithm, and it determines how accurate our count estimates will be.
+The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
 We provide a way to find reasonable values of $minSS$ later.
 
-At all points, the SampleHandler maintains a set of Table Samples in memory.
-Each table sample $s$ has the following attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
+At all points, the SampleHandler maintains a set of Samples in memory.
+Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
 The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
@@
 
-Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ with $f_s = r$ and $|T_s| \geq minSS$.
+Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
+The SampleHandler also computes $N_s$ when a sample is created.
 Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
@@
 
-When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule. There are multiple ways it could do this:
+When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule.
+At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample.
+Creating a new sample by making a pass through the table is called \textbf{Create} (further described below).
+At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
 \begin{enumerate}
-\item \textbf{Find:} If the SampleHandler finds an existing sample with $r$ as the filter rule, and at least $minSS$ tuples, it simply returns that sample. 
+\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as it's filter rule (i.e.
+$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq \text{minSS}$, then it simply returns sample $s$.
+Our greedy algorithm can then be run on $s$.
+
 \item \textbf{Combine:} If \textbf{Find} doesn't work i.e.
 if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-We can show that tuples from $T_{s^{\prime}}$ that are covered by $r$ can be taken to be uniformly sample tuples covered by $r$.
-That is, each tuple covered by $r$ is equally likely to appear in a uniform sample of $f_{s^{\prime}}$.
-If the union of such tuples over all such available $s^{\prime}$s exceeds $minSS$ in size, then the SampleHandler can create a sample using existing the samples, instead of reading the entire table to create a new sample.
+If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds minSS in size, then we can simply treat that set as our sample for rule $r$.
+We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
+That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
+
+Note that the \textbf{Combine} procedure doesn't really require additional memory.
+Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by the Greedy Algorithm.
+
 \item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample by making a pass through the table.
 Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-We can use reservoir sampling to get a uniformly random sample of given size in a single pass through the table.
+In addition, creating a new sample requires memory, and since capacity is limited, it may necessitate shrinking or deleting some existing samples.
+We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
+
 In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available).
@@
 For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
-\item \textbf{Delete/Shrink:} Finally, it may happen that there is not enough memory left to create a new sample.
+
+\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may happen that there is not enough memory left to create a new sample.
 In that case, the SampleHandler needs to delete or shrink an existing sample.
 It can use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
-Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
+Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
 %For shrink, prefer shrinking samples of the largest size (> minSS) first?
@@
 
-When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user.
+We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in section~\ref{sec:sampling_algorithms}.
+
+\textbf{Pre-fetching:} When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user.
 If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
@@
 
-% TODO: Need to find and write about the exact algorithm to decide how much to load, what to shrink, and how much to shrink. Different heuristic algos are a good experimental subject as well.
+\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
+We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
+Suppose at some stage, we have a tree $U$ of rules displayed to the user.
+Internal nodes of $U$ are rules that have been expanded, while leaves are rules that have not been expanded.
+Let $L$ be the set of leaves.
+Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for leaves.
+
+We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
+In the absence of additional information, we can assume a uniform distribution.
+That is, every leaf is equally likely to be expanded next.
+However, we can also use Machine Learning on past user behaviour, along with rule features such as `rule depth in tree' and `rule distance from last expanded rule' to get a better probability estimate of each rule being expanded.
+
+When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
+Since reading the table from Hard Disk is usually a bottleneck, it also do things like creating other samples for rules other than $r$, and augmenting existing samples.
+Hence, we assume that in a \textbf{Create} phase, the SampleHandler does a complete re-organization of existing samples.
+Based on the current leaves and their expansion probabilities, it tries to find an optimal set of samples to create for the next step.
+More specifically, for every displayed rule $r$, it determines an integer $n_r$, and then creates a sample $s_r$ with $f_{s_r} = r$ and $|T_{s_r}| = n_r$ while making it's pass through the table.
+
+Finally, for each rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
+Call this quantity $S(r_1, r_2)$.
+Define $S(r_1, r_2)$ to be zero if $r_1$ is not a sub-rule of $r_2$.
+Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$.
+For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$.
+If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
+We wish to maximise the probability that we can respond to the next user expansion without making another pass.
+
+Thus formally our problem is :
+Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximise}$ :
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
+where the $I$'s are indicator variables, subject to :
+$$\sum_{r \in U} n_r < \text{capacity}$$
+This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be okay to use an exponential algorithm to solve the problem.
+
+Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
+The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and it's immediate parent.
+That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
+$\text{ess}(r)$ is redefined according to the new value of $S$ as well.
+So $$\text{ess}(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
+
+Now consider a rule in $r_0 \in U \setminus L$ along with all it's children.
+Let $M_{r_0}$ denote the set containing $r_0$ and all it's leaf children.
+Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
+This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$.
+Thus for each non-leaf rule $r_0$ and all it's children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
+Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{\text{ess}(r) \geq \text{minSS}}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
+Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
+We describe both these steps in detail below:
+
+Let $r_0 \in U \setminus L$.
+Let $d$ be the number of leaf children of $r_0$.
+Let the children be $r_1, r_2, ...
+r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$.
+Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
+$$1 \leq i_m \leq i_M \leq d$$
+$$n_{r_0} = \frac{\text{minSS}}{S(r_0, r_{i_m})}$$
+$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
+$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
+$$i_M < j \Rightarrow n_{r_j} = 0$$
+And thus
+$$\text{ess}(r_j) \geq \text{minSS} \Leftrightarrow j \leq i_M$$
+
+Call this locally optimal solution $e$. Then we define its `probability value' to be
+$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
+and its `Sampling Cost' to be
+$$S(e) = \frac{\text{minSS}}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
+Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions.
+
+Then next step is to combine the solutions using dynamic programming.
+Let the $M$ sets be called $M_0, M_1, ...
+M_D$.
+Let our possible sample sizes range from $0$ to $\mathcal{S}$.
+Then we create a $D \times \mathcal{S}$ array $A$.
+The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ...
+M_i$ with total sample size at most $j$.
+We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$.
+Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$.
+Then we have,
+$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
+This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.
 
@@
 We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
-In addition, we may have a tuple occur in multiple samples, especially when we use \textbf{Combine} to create a sample.
+In addition, we may have a tuple occur in multiple samples.
 Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
@@
 That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.
-So the value of minSS must be at least $c\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary, and $c$ is a constant chosen by us based on how accurate we want the count estimate to be.
+
+We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = \text{minSS}$.
+If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be.
+Moreover, since we want good Count estimates for all rules displayed in the summary, we want $\text{minSS} >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.
 
@@
 Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$.
-The highest scoring rule can have score at most $|C|$ (the total number of columns), the the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
+For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times.
+So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.
+
+The highest scoring rule can have weight at most $|C|$ (the total number of columns).
+Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
 Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
+For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $\text{minSS} >> 5 \times 10$.
 
@@
 Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.
+
+%TODO: Add experiments on m_w parameter. Show small values are alright. And graph with time.
 

### 2014-10-29 TableSummarization/Paper/TableSummarization.tex
Added an experiment, and generated graph

@@
 \newtheorem{example}[definition]{Example}
+\newcounter{prob}
+\newtheorem{problem}[prob]{Problem}
 
@@
 \title{Interactive Summarization of Relational Tables}
-% Can say multi-dimensional instead of relational
-% Can say Dynamic Real-Time instead of Interactive?
 
@@
 \begin{abstract}
-In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints.
-The summary consists of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of that subset.
-We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table.
+In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints.
+The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that are covered by the rule.
+For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column.
+We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table.
+We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and a sampling based scheme for dealing with large tables.
 Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
-% TODO: Improve abstract.
 \end{abstract}
@@
 A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}.
-Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a rule-scoring black box as user input.
+Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a black box `weight function' as user input.
 In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.
@@
 For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}.
-The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `size' denoting the number of non-$\star$ values of the rule for each rule.
+The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `weight' in this case denoting the number of non-$\star$ values of the rule for each rule.
+The weight function can in general be customized by the user.
 
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Size \\
+\hline Store & Product & State & Count & Weight \\
 \hline
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-target & bicycles & $\star$ & $200$ & $2$ \\ \hline
-$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
-walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
+$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \hline
+$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
+$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
 \end{tabular}
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Size \\
+\hline Store & Product & State & Count & Weight \\
 \hline
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
-target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
-$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
-walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
-$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
-$\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
-$\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
+$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
+$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
+$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
 \end{tabular}
@@
 \end{table}
-% TODO: Should we display weight rather than size in table? After all, size is obvious. In intro example, we say that for this table, weight equals number of non-$\star$ columns.
 
@@
 \squishlist
-\item \textbf{Flexible : } We allow users to specify a weighting function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
+\item \textbf{Flexible : } We allow users to specify a weight function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
 \item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
@@
 In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
-% TODO: Say it's complementary.
-Also give an example where your thing is good.
-Mention MDL here.
-Or that max-entropy is hard to interpret for human, even though formally very clear.
+Our approaches are thus complementary and serve different purposes.
+% TODO: max-entropy is hard to interpret for human, even though formally very clear.
 Our thing also works well when user doesn't know the set of column values.
@@
 Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table.
-Our objective (formally defined later) is to summarise the table with a list of {\em rules}.
 
-A {\em rule} is a tuple with one field per column of the queried table.
+Our objective (formally defined later) is to summarize the table with a list of {\em rules}.
+A {\em rule} is a tuple with a value for each column of the queried table.
 In addition, a rule has other attributes, such as count and weight (which we define later) associated with it.
-The value in each field of the rule can either be one of the values in the corresponding column of the table, or $\star$, which is a wildcard character representing all values in the column.
-For a table column with numerical values, we can allow the corresponding rule-value to be a range instead of a single value.
-A {\em trivial rule} is one that has a $\star$ value in all fields.
-A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple.
+The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column.
+For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value.
+A {\em trivial rule} is one that has a $\star$ value in all columns.
+
+A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple.
 We abuse notation to write this as $t \in r$.
@@
 A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. 
+%TODO: Give name to out system (and paper)
 
@@
 Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
-We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$.
+We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e.
+there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not.
 If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.
@@
 Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.
-{\em Count}($r$) is defined as the total number of tuples $t \in T$ that are covered by $r$.
-Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
-A high value of MCount indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules.
-We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list.
-Let {\em Size}($r$) refer to the number of non-starred values in rule $r$.
-We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules.
+
+We now define some additional properties of rules. Let $r$ be a rule. Then:
+\squishlist
+\item {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
+\item {\em MCount}($r$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
+A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules.
+We want to pick rules with a high value of $MCount$ to display to the user, to increase the coverage of the rule-list.
+\item {\em Size}($r$) refers to the number of non-starred values in rule $r$. 
+\item We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules.
 A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$.
 We further describe our weighting functions in Section~\ref{sec:weighting}.
-The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} \text{MCount}(r)W(r)$$.
+\squishend
+
+The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$.
+Our problem is to choose the rule-list that maximises total score.
+We use $MCount$ rather than Count to increase coverage of the table.
+If we had defined total score as $$\sum_{r \in R} \text{Count}(r)W(r)$$, then our optimal rule-list would contain
+%TODO: Finish writing this part.
 
@@
 
+\begin{problem}\label{prob:optimal-subrule-list}
 Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
-$$\sum_{r \in R} W(r)\text{MCount}(r)$$
-subject to the following constraints:
+$$\sum_{r \in R} W(r)MCount(r)$$
+for one of the following interactions:
 \begin{enumerate}
-\item If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
-\item If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
+\item $[$Rule Interaction$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
+\item $[$Star Interaction$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
 \end{enumerate}
+\end{problem}
 
@@
 \label{sec:weighting}
-We want our rules to be as descriptive of the table as possible, i.e.
+We now describe how we score rules.
+At a high level, we want our rules to be as descriptive of the table as possible, i.e.
 given the rules, it should be as easy as possible to reproduce the table.
 We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is.
-Given this weighting, our total score for our rule-list is given by : $$\sum_{r\in R} \text{MCount}(r)W(r)$$.
-Let us consider some simple kinds of $W(r)$s and their meanings:
+Given this weight per rule, our total score for our rule-list is given by : $$\sum_{r\in R} MCount(r)W(r)$$.
+Let us consider some forms for the function $W(r)$ and their interpretation:
 \begin{enumerate}
@@
 Consider the examples in Table \ref{table:sizescoringexample}.
-Weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$.
-Thus total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$.
-If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$.
+The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$.
+Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$.
+If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$.
+Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$.
 If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus when we are going to marginalize on one extra attribute ($B$ in this case) it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
+Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
 
-To get an interpretation for this scoring function, imagine we are trying to reconstruct the table from the rules.
-Since we have rule ($a$, $b_1$) with MCount $100$, we are going to get a $100$ of the table's tuples from this rule.
+To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
+Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule.
 For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns).
@@
 Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total.
-Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-set.
-Since having more of the table pre-filled is better, maximising score gives us a desirable set of rules.
+Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list.
+Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.
 
@@
 Like the previous scoring function, this one adds some weight for every non-starred value in a rule.
-But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the 'complexity' of the column.
+But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column.
 The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
@@
 For a column $c$, specifying a value in the column takes log($|c|$) bits of information.
-A non-starred valued in a rule $r$ thus pre-fills $\text{MCount}(r) \text{log}(|c|)$ bits of the table.
+A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table.
 Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.
 
-Note that this scoring function is closely related to the Minimum Description Length (MDL) of a table.
+Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
 If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.
@@
 When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}.
-At any point, the user can click on either a rule, or a star, to 'expand' the rule.
-Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of sub-rules of $r$.
+At any point, the user can click on either a rule, or a star within a rule, to `expand' the rule.
+Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
 By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-As an example of this operation, clicking on the trivial rule would display table~\ref{table:introexample}.
+As an example of this operation, clicking on the trivial rule of table~\ref{table:trivial} would display table~\ref{table:introexample}.
 Clicking further on the second rule would display table~\ref{table:introexample2}.
@@
 This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows sub-rules of the walmart rule all specific to some product.
+For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
 This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Size \\
+\hline Store & Product & State & Count & Weight \\
 \hline
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Size \\
+\hline Store & Product & State & Count & Weight \\
 \hline
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
-target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
-$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
-walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
-$\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
-$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
-$\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
+$\triangleright$ target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
+$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
+$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
 \end{tabular}
@@
 \label{sec:algorithms}
-When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. We now formally state the problem:
+When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into.
+If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$.
+We can reduce Problem~\ref{prob:optimal-rule-list} to a simpler problem given below:
 
-Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ sub-rules of $r^{\prime}$ that maximizes the total score given by :
-$$\text{Score}(R) = \sum_{r \in R}W(r)\text{MCount}(r)$$
+\begin{problem}\label{prob:optimal-rule-list}
+Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
+$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
+\end{problem}
 
@@
 $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.
+
+Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
+\squishlist
+\item $[$Rule Interaction$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
+Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
+\item $[$Star Interaction$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$.
+In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.
+Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
+\squishend
 
@@
 Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order.
-We want to find the $S$ that maximises $\text{Score}(S)$.
-We solve this part of the problem next.
+This gives us
+\begin{problem}\label{prob:optimal-rule-set}
+Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
+$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
+\end{problem}
 
-\subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
-We will now show that the Score function over sets of rules has a property called submodularity, which gives us a greedy approximation algorithm for optimizing it. 
+The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.
+
+\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
+We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
 \begin{definition}
@@
 For a given table $T$, the Score function over sets $S$ of rules, defined by :
-$$\text{Score}(S) = \sum_{r \in S} \text{MCount}(r)W(r)$$
+$$\text{Score}(S) = \sum_{r \in S} MCount(r)W(r)$$
 is submodular.
@@
 
-Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm}.
+Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
 The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
 
-$m_w$ stands for \textit{Max Weight}.
+The parameter $m_w$ stands for \textit{Max Weight}.
 The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
@@
 
-The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty in Step $1$, and then iterates for $k$ steps in Step $2$, adding the best marginal rule at each step.
+The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
 To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.
@@
 In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}.
-We compute the best marginal rule over several passes (Step $3$), with the maximum number of passes equal to the maximum size of a rule.
+We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule.
 In the $j^{th}$ pass, we compute counts for rules of size $j$.
@@
 We now describe the function to find the best marginal rule.
-The pseudo-code for the function is in the box titled \textbf{Find best marginal rule}.
+The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.
 The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
@@
 
-\begin{framed}
-\textbf{Find best marginal rule}
+\begin{algorithm}\label{algo:best-rule-set}
+\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$S$ (Solution set of rules)}
+$S = \phi$
 
-\textbf{Input :} $S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)
+\For {$i$ from $1$ to $k$}{
+$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}
 
-\textbf{Output :} $R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)
-\begin{enumerate}
-\item $H = 0$ // Threshold for deciding if to find count for a rule.
-\item $C = C_o = C_n = \phi$ // Set of all, old and new candidate rules respectively.
-\item For $j$ from $1$ to number of columns in $T$:
-\begin{enumerate}[label*=\arabic*.]
-\item if $j = 1$
-\begin{enumerate}[label*=\arabic*.]
-\item $C_n = \text{ all rules of size } 1$
-\end{enumerate}
-\item else 
-\begin{enumerate}[label*=\arabic*.]
-\item $C_n = \text{all size-}i$ super-rules of rules from $C_o$
-\end{enumerate}
-\item For each $R \in C_n$
-\begin{enumerate}[label*=\arabic*.]
-\item $M =\infty$ // Upper bound on marginal value count of $R$
-\item For each $R$-sub-rule $R^{\prime} \in C$  
-\begin{enumerate}[label*=\arabic*.]
-\item $M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
-\end{enumerate} 
-\item if $(M < H)$ 
-\begin{enumerate}[label*=\arabic*.]
-\item $C_n = C_n \setminus \left\lbrace R \right\rbrace$ // Delete $R$ if its max count is too small for $R$ to possibly be in the solution.
-\end{enumerate} 
-\end{enumerate}
-\item if $C_n = \phi$
-\begin{enumerate}[label*=\arabic*.]
-\item break;
-\end{enumerate}
-\item For each $R \in C_n$
-\begin{enumerate}[label*=\arabic*.]
-\item Count$(R) = 0$ // Initialize to zero
-\item MarginalValue$(R) = 0$ // Initialize to zero
-\end{enumerate}
-\item For each tuple $t \in T$
-\begin{enumerate}[label*=\arabic*.]
-\item Let $R_S$ be the highest weight rule in $S$ that covers $t$
-\item For each rule $R \in C_n$ that covers $t$
-\begin{enumerate}[label*=\arabic*.]
-\item Count($R$) $++$
-\item MarginalValue($R$) $+= W(R) - \text{min}(W(R), W(R_S))$
-\end{enumerate}
-\end{enumerate}
-\item $C_o = C_n$
-\item $C_n = \phi$
-\item $C = C \cup C_o$
-\item $H = \text{max}_{R \in C}(\text{MarginalValue}(R))$
-\end{enumerate}
-\item return $\text{argmax}_{r \in C} \text{marginalValue}(r)$
-\end{enumerate}
-\end{framed}
+$S = S \cup \left\lbrace R_m \right\rbrace$
+}
+\Return $S$
+\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
+\end{algorithm}
 
-\begin{framed}
-\textbf{Greedy Algorithm}
+\begin{algorithm}\label{algo:best-marginal-rule}
+\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
+$H = 0$ \tcc*{Threshold for deciding if to find count for a rule.}
+$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
+\For{$j$ from $1$ to number of columns in $T$}{
+\If {$j = 1$}{
+$C_n = \text{ all rules of size } 1$
+}
+\Else{ 
+$C_n = \text{all size-}i$ super-rules of rules from $C_o$
+}
+\ForEach {$R \in C_n$}{
+$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
+\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
+$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
+}
+\If {$(M < H)$}{ 
+$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to possibly be in the solution}
+}
+}
+\If {$C_n = \phi$}{
+break;
+}
+\ForEach {$R \in C_n$}{
+Count$(R) = 0$ \tcc*{Initialize}
+MarginalValue$(R) = 0$ \tcc*{Initialize}
+}
+\ForEach {$t \in T$}{
+Let $R_S$ be the highest weight rule in $S$ that covers $t$
+\ForEach {$R \in C_n$ that covers $t$}{
+Count($R$) $++$
 
-\textbf{Input :} $k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)
+MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
+}
+}
+$C_o = C_n$
 
-\textbf{Output :} $S$ (Solution set of rules)
-\begin{enumerate}
-\item $S = \phi$ // Initialize solution set to null set.
-\item For $i$ from $1$ to $k$:
-\begin{enumerate}[label*=\arabic*.]
-\item $R_m = \text{Find best marginal rule}(S, T, m_w, W)$
-\item $S = S \cup \left\lbrace R_m \right\rbrace$
-\end{enumerate}
-\item Return $S$
-\end{enumerate}
+$C_n = \phi$
 
-\end{framed}
+$C = C \cup C_o$
 
-\subsection{Supporting user operations}
-Our greedy algorithm finds the best set of rules to cover a table, as per some monotonic weight function $W$.
-We now describe how we can use our algorithm to support the operations allowed in our user interface:
-\begin{enumerate}
-\item \textbf{Expand Rule} : When the user tries to expand rule $r$, we want to display the highest scoring set of sub-rules of $r$.
-In order to do this, we make a pass through table $T$ to select tuples covered by $r$, and load them into a new table $T^{\prime}$.
-Then we run the greedy algorithm on $T^{\prime}$ to get the set of rules to display to the user.
-\item \textbf{Expand Star} : When the user tries to expand the $\star$ value in a column $c$ of rule $r$, we again create table $T^{\prime}$ consisting of tuples of $T$ that are covered by $r$.
-In addition, we use weight function $W^{\prime}$ such that $W^{\prime}(x) = 0$ if rule $x$ has a $\star$ in column $c$, and $W^{\prime}(x) = W(x)$ otherwise.
-Then we run the greedy algorithm on table $T^{\prime}$ with weight function $W^{\prime}$.
-\end{enumerate}
+$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
+}
+\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
+\caption{Find best marginal rule}
+\end{algorithm}
 
@@
 
-%TODO: Add experiments on m_w parameter. Show small values are alright. And graph with time.
-
 The columns (in order) are as follows:
@@
 \subsection{User interface testing}
-We restrict the table to the first $10$ columns in order to make the result tables fit into the page.
-Projecting onto $10$ columns reduces the number of distinct tuples to $2851$.
-We set the $n$ (number of rules) parameter to $4$, and $m_w$ to $5$.
+We restrict the table to the first $7$ columns in order to make the result tables fit into the page.
+We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$.
 We now present the rule-based summaries displayed as a result of a few different user actions.
@@
 Then the user will see Table~\ref{table:uiexample1}.
-The first two new rules simply tell us that the table has $1501$ female and $1350$ male tuples.
-The next two rules also slightly more detailed, saying that there are $577$ females who have been in the Bay Area for $> 10$ years, and $158$ males aged $18-24$ who have never been married.
-Note that the latter two rules give very specific information which would require $3$ user clicks to find in an OLAP system, where our system displays that information to the user with a single click.
+The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples.
+The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Are for $> 10$ years.
+Note that the latter two rules give very specific information which would require upto $3$ user clicks to find in an OLAP system, whereas our system displays that information to the user with a single click.
 
@@
 
+\begin{table*} 
+\centering 
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\hline 
+\end{tabular} 
+\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
+\end{table*}
+
+\begin{table*} 
+\centering 
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\hline 
+\end{tabular} 
+\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
+\end{table*}
+
+\begin{table*} 
+\centering 
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\hline 
+\end{tabular} 
+\caption{A Rule expansion \label{table:uiexamplerule}} 
+\end{table*}
+
 \subsection{Weighting functions}
@@
 We first try the weighting function given by:
-$$W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$$ where $|c|$ refers to the number of distinct possible values in column $c$.
+$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct possible values in column $c$.
 This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values.
@@
 \begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
-\hline 
-\end{tabular} 
-\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
-\end{table*}
-
-\begin{table*}
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $378$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $271$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $416$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & Grad Study & $\star$ & $\star$ & $171$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
-\hline 
-\end{tabular} 
-\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1501$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $1350$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $467$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $408$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Divorced/separated & $\star$ & $\star$ & $\star$ & $\star$ & $187$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $485$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
-\hline 
-\end{tabular} 
-\caption{A Rule expansion \label{table:uiexamplerule}} 
-\end{table*}
-
-\begin{table*}
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $858$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & Married & $\star$ & $\star$ & $\star$ & $\star$ & $917$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $1062$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\star$ & $\star$ & $\star$ & $\star$ & Professional/Managerial & $\star$ & $637$ & $1$ \\
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
 \hline 
@@
 \begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Size \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $2851$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 25-34 & $\star$ & $\star$ & $\star$ & $145$ & $3$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $577$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Never married & 18-24 & $\star$ & $\star$ & $\star$ & $158$ & $3$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $137$ & $3$ \\
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
 \hline 
@@
 
+\subsection{Effects of $m_w$}
+Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
+The algorithm is guaranteed to find the best marginal rule as long as it's weight is $leq m_w$, but runs faster for smaller values of $m_w$.
+In this section, we study the effect of varying $m_w$ on the speed of our algorithm.
+
+We fix the weighting function $W$, and a value of $m_w$.
+Then we find the time taken for expanding the empty rule with the weighting function $W$ and $m_w$ parameter, and average this time across $10$ iterations.
+This time is plotted against $m_w$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:mw_speed}.
+The figure shows that running time seems to be approximately linear in $m_w$.
+
+The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule.
+Thus for size scoring, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
+For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
+At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively.
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
+\vspace{-13pt}
+\end{figure}
+
+% TODO: Experiments on m_w. Experiments on Sampling schemes.
+
 \begin{comment}
-Notes based on Explanation Tables paper: Prove that our problem is NP-Hard to solve exactly. Something like the Set Cover Problem. Yes, that is easy.
+Notes based on Explanation Tables paper: Prove that our problem is NP-Hard to solve exactly. Something like the Set Cover Problem.
 
@@
 
-% Aditya's idea of flexible ranges. Score depends on range length.
-
 \section{Related Work}
@@
 
+% Related Work summary :
+% Explanation tables takes lots of papers and talks about them in one go in related work.
+saying they optimize for different opbjective functions.
+We can say that as well.
+We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity.
+% Efficient and Effective... Tableaux
+
 {\small 

### 2014-10-29 TableSummarization/src/dataextraction/Marketing.java

@@ -43,7 +43,7 @@ public class Marketing {
 		maritalStatus.put("column", "Marital Status");
 		maritalStatus.put("1", "Married");
 		maritalStatus.put("2", "Living together");
-		maritalStatus.put("3", "Divorced/separated");
+		maritalStatus.put("3", "Divorced / separated");
 		maritalStatus.put("4", "Widowed");
 		maritalStatus.put("5", "Never married");
 		maritalStatus.put("NA", "NULL");
@@ -74,10 +74,10 @@ public class Marketing {
 		
 		Map<String, String> occupation = new HashMap<String, String>();
 		occupation.put("column", "Occupation");
-		occupation.put("1", "Professional/Managerial");
+		occupation.put("1", "Professional / Managerial");
 		occupation.put("2", "Sales Worker");
-		occupation.put("3", "Factory Worker/Laborer/Driver");
-		occupation.put("4", "Clerical/Service Worker");
+		occupation.put("3", "Factory Worker / Laborer / Driver");
+		occupation.put("4", "Clerical / Service Worker");
 		occupation.put("5", "Homemaker");
 		occupation.put("6", "Student");
 		occupation.put("7", "Military");

### 2014-10-29 TableSummarization/src/solvers/Rule.java

@@ -357,7 +357,7 @@ public class Rule implements Comparable<Rule> {
 		if (marginalCounts.get(this) != null) {
 			marginal = marginalCounts.get(this).toString();
 		}
-		ruleString = ruleString + ",\"" + marginal + "\",\"" + count + "\",\"" + size + "\"";
+		ruleString = ruleString + ",\"" + marginal + "\",\"" + count + "\",\"" + score + "\"";
 		return ruleString;
 	}
 
@@ -377,7 +377,7 @@ public class Rule implements Comparable<Rule> {
 				ruleString = ruleString + "\"" + table.names.get(i).get(table.dictionary.get(i).get(val)) + "\"";
 			}
 		}
-		ruleString = ruleString + ",\"" + count + "\",\"" + size + "\"";
+		ruleString = ruleString + ",\"" + count + "\",\"" + score + "\"";
 		return ruleString;
 	}
 	
@@ -402,7 +402,7 @@ public class Rule implements Comparable<Rule> {
 		if (marginalCounts.get(this) != null) {
 			marginal = marginalCounts.get(this).toString();
 		}
-		ruleString = ruleString + " & $" + marginal + "$ & $" + count + "$ & $" + size + "$ \\\\";
+		ruleString = ruleString + " & $" + marginal + "$ & $" + count + "$ & $" + score + "$ \\\\";
 		return ruleString;
 	}
 
@@ -423,7 +423,7 @@ public class Rule implements Comparable<Rule> {
 				ruleString = ruleString + (table.names.get(index).get(table.dictionary.get(index).get(val))).replaceAll("\\$", "\\\\\\$");
 			}
 		}
-		ruleString = ruleString + " & $" + count + "$ & $" + size + "$ \\\\";
+		ruleString = ruleString + " & $" + count + "$ & $" + score + "$ \\\\";
 		return ruleString;
 	}
 
@@ -448,7 +448,7 @@ public class Rule implements Comparable<Rule> {
 		if (marginalCounts.get(this) != null) {
 			marginal = marginalCounts.get(this).toString();
 		}
-		ruleString = ruleString + " & $" + marginal + "$ & $" + count + "$ & $" + size + "$ \\\\";
+		ruleString = ruleString + " & $" + marginal + "$ & $" + count + "$ & $" + score + "$ \\\\";
 		return ruleString;
 	}
 
@@ -469,7 +469,7 @@ public class Rule implements Comparable<Rule> {
[…]

### 2014-10-29 TableSummarization/src/solvers/RuleTree.java

@@ -51,6 +51,7 @@ public class RuleTree {
 		Map<Integer, Integer> valueMap = new HashMap<Integer, Integer>();
 		final Integer length = table.dictionary.size();
 		Rule emptyRule = new Rule(valueMap, length);
+		emptyRule.score = 0;
 		emptyRule.count = table.contents.size();
 		emptyRule.counted = true;
 		nodeMap = new HashMap<Rule, RuleNode>();

### 2014-10-29 TableSummarization/src/solvers/RuleTreeDisplay.java

@@ -13,8 +13,11 @@ import java.util.Stack;
 
 import dataextraction.Marketing;
 import dataextraction.TableInfo;
+import solvers.Rule.sizeBitsScorer;
+import solvers.Rule.sizeMinusKScorer;
 import solvers.Rule.sizeScorer;
 import solvers.RuleTree.RuleNode;
+import utils.Experiments;
 
 public class RuleTreeDisplay {
 	RuleTree ruleTree;
@@ -98,7 +101,7 @@ public class RuleTreeDisplay {
 
 	private String tableHeaderTex () {
 		String answer = "";
-		answer = answer + "\\begin{table} \n\\centering \n\\begin{tabular}{|";
+		answer = answer + "\\begin{table*} \n\\centering \n\\begin{tabular}{|";
 		for (int i = 0; i < columnOrder.size(); i++) {
 			answer = answer + " p{1.5cm} |";
 		}
@@ -111,12 +114,12 @@ public class RuleTreeDisplay {
 			answer = answer + ruleTree.table.names.get(columnOrder.get(i)).get("column");
 			first = false;
 		}
-		answer = answer + " & Count & Size \\\\ \\hline \n";
+		answer = answer + " & Count & Weight \\\\ \\hline \n";
 		return answer;
 	}
 	
 	private String tableFooterTex () {
-		String answer = "\\hline \n\\end{tabular} \n\\caption{ADD CAPTION} \n\\end{table} \n";
+		String answer = "\\hline \n\\end{tabular} \n\\caption{ADD CAPTION} \n\\end{table*} \n";
 		return answer;
 	}
 	
@@ -141,7 +144,10 @@ public class RuleTreeDisplay {
 				cline = cline + "\\cline{" + (columnOrder.size() + 1) + "-" + (columnOrder.size() + 2) + "} ";
 			}
 			answer = answer + cline;
-			answer = answer + currentNode.rule.fullRuleStringTex(ruleTree.table, columnOrder, parentRule) + "\n";
+			for (int i = 0; i < currentNode.depth; i++) {
+				answer = answer + "$\\triangleright$ ";
+			}
+			answer = answer + currentNode.rule.fullRuleStringTex(ruleTree.table, columnOrder) + "\n";
 			List<RuleNode> toAdd = new ArrayList<RuleNode>();
 			for (RuleNode childNode : currentNode.children) {
 				if (!hiddenRows.contains(childNode)) {
[…]

### 2014-10-29 TableSummarization/src/utils/Experiments.java

@@ -1,0 +1,31 @@
+package utils;
+
+import java.io.FileWriter;
+import java.io.IOException;
+import java.io.PrintWriter;
+import java.util.HashMap;
+
+import solvers.Rule;
+import solvers.RuleTree;
+import solvers.RuleTreeDisplay;
+import solvers.Scorer;
+import dataextraction.TableInfo;
+
+public class Experiments {
+	public static void mwSpeedTest (TableInfo table, int mwMin, int mwMax, int ruleNums, Scorer scorer, String outFile) throws IOException {
+		PrintWriter pw = new PrintWriter(new FileWriter(outFile));
+		Rule emptyRule = new Rule(new HashMap<Integer, Integer>(), table.dictionary.size());
+		int numIters = 10;
+		for (int mw = mwMin; mw < mwMax; mw++) {
+			long totalTime = 0;
+			for (int iter = 0; iter < numIters; iter++ ) {
+				RuleTree ruleTree = new RuleTree(table);
+				long timer = System.currentTimeMillis();
+				ruleTree.expandRow(emptyRule, ruleNums, mw, scorer);
+				totalTime += System.currentTimeMillis() - timer;	
+			}
+			pw.println(mw + "\t" + totalTime/numIters);
+		}
+		pw.close();	
+	}
+}

### 2014-10-29 TableSummarization/Paper/TableSummarization.tex
Changes to writeup

@@
 We use $MCount$ rather than Count to increase coverage of the table.
-If we had defined total score as $$\sum_{r \in R} \text{Count}(r)W(r)$$, then our optimal rule-list would contain
-%TODO: Finish writing this part.
+If we had defined total score as $$\sum_{r \in R} \text{Count}(r)W(r)$$, then our optimal rule-list would contain rules that repeatedly refer to the most `summarizable' part of the table.
+For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then the summary could consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.
+%TODO: Write an actual example? To also verify what actually happens.
 

### 2014-10-30 TableSummarization/Paper/TableSummarization.tex
Added related work

@@
 In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints.
-The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that are covered by the rule.
+The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that satisfy that pattern.
 For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column.
 We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table.
-We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and a sampling based scheme for dealing with large tables.
+We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and a sampling based scheme for efficiently summarizing large tables.
 Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
@@
 
-%TODO: Change indentation format. All indentation at start, number of indents equal to depth of rule
 \begin{table}
@@
 
-% TODO : Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
+We further discuss related work in Section~\ref{sec:related}.
+
 \stitle{Overview of paper:} 
@@
 
-\item In Section 6, we experimentally evaluate the performance of our algorithms a real dataset.
+\item In Section 6, we experimentally evaluate the performance of our algorithms on a real dataset.
 \squishend
@@
 For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then the summary could consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.
-%TODO: Write an actual example? To also verify what actually happens.
 
@@
 
-% TODO: Experiments on m_w. Experiments on Sampling schemes.
+% TODO: Experiments on Sampling schemes.
 
@@
 
-\section{Related Work}
+\section{Related Work}\label{sec:related}
 %TODO: Repeat what we said in intro about interactive, etc.
 There has been work on finding interesting rules in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
-The work mainly focuses on finding values that occur more often or less often that expected from a max entropy distribution.
+This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max entropy distribution.
 The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
@@
 
+Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
+There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
+All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
+
 Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
@@
 
-% Related Work summary :
+% Related Work summary : Is ours the only only that is flexible (due to weight function black box)?
 % Explanation tables takes lots of papers and talks about them in one go in related work.
@@
 We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity.
-% Efficient and Effective... Tableaux
+% Succintly summarizing with itemsets-- Mampaey:2011:TMI:2020408.2020499 -- Also seemed to optimize according to max entropy criterion
+% Efficient and Effective... Tableaux -- Golab_efficientand -- Similar to conditional functional dep. paper.
+% Near Optimal Tableaux Conditional functional dep.
+-- Golab:2008:GNT:1453856.1453900 -- Finding a smallest sized tableau that provides some confidence and support for a conditional functional dependency.
+% MDL Summarization with holes -- Bu:2005:MSH:1083592.1083644 -- Min sized summary with a rectangle and exceptions (holes). 
+% Generalized MDL Approach -- Lakshmanan:2002:GMA:1287369.1287435 -- Similar to MDL Approach.
+% Overlapped Hyperrectangle -- Xiang_succinctsummarization -- Efficiently find the set of hyperrectangles that covers the data (sets of itemsets)
+% Tiling Databases Geerts04tilingdatabases Simlar to hyper-rectangles. Probabilistic interpretation of frequent itemsets.
 

### 2014-11-06 TableSummarization/Paper/TableSummarization.tex
Added more experiments. And changes in paper

@@
 We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table.
-We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and a sampling based scheme for efficiently summarizing large tables.
+We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and an optimized sampling scheme for efficiently summarizing large tables.
 Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
@@
 \begin{example}\label{ex:introexample}
-% Need justification for how investor got data. Or change example.
 Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'.
@@
 The number of additional rules ($3$ in this case) is a parameter that can be set by the user.
-The rules are shown in decreasing order by size.
+The rules are shown in decreasing order by weight.
 
-The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles.
+The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that Target is selling a lot of bicycles.
 The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores.
@@
 
-\item In Section 2, we introduce the relevant notation and formally define our problem.
+\item In Section~\ref{sec:preliminaries}, we introduce the relevant notation and formally define our problem.
 
-\item In Section 3, we consider different schemes for weighting rules, along with their practical significance.
+\item In Section~\ref{sec:weighting}, we consider different schemes for weighting rules, along with their practical significance.
 
-\item In Section 4, we describe our user interface.
+\item In Section~\ref{sec:interface}, we describe our user interface.
 
-\item In Section 5, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.
+\item In Section~\ref{sec:algorithms}, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.
 
-\item In Section 6, we experimentally evaluate the performance of our algorithms on a real dataset.
+\item In Section~\ref{sec:experiments}, we experimentally evaluate the performance of our algorithms on a real dataset.
+
+\item In Section~\ref{sec:related}, we describe related work.
 \squishend
@@
 
-The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$.
-Our problem is to choose the rule-list that maximises total score.
+The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$ Our problem is to choose the rule-list that maximises total score.
 We use $MCount$ rather than Count to increase coverage of the table.
-If we had defined total score as $$\sum_{r \in R} \text{Count}(r)W(r)$$, then our optimal rule-list would contain rules that repeatedly refer to the most `summarizable' part of the table.
+If we had defined total score as $\sum_{r \in R} \text{Count}(r)W(r)$, then our optimal rule-list would contain rules that repeatedly refer to the most `summarizable' part of the table.
 For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then the summary could consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.
@@
 We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is.
-Given this weight per rule, our total score for our rule-list is given by : $$\sum_{r\in R} MCount(r)W(r)$$.
+Given this weight per rule, our total score for our rule-list is given by : $\sum_{r\in R} MCount(r)W(r)$.
 Let us consider some forms for the function $W(r)$ and their interpretation:
@@
 If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$.
-We can reduce Problem~\ref{prob:optimal-rule-list} to a simpler problem given below:
+We can reduce Problem~\ref{prob:optimal-subrule-list} to the following simpler problem by removing the user-interaction based constraints:
 
@@
 For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
-$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$. Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
+$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$ 
+Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
 $$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.
@@
 On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$.
-Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w \geq \textrm{max}_{r \in S_o}W(r)$.
+Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
 In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
@@
 Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
-Our experiments also show that the weights of rules in the optimal set are small.
+Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set are small.
 
@@
 Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$.
-This implies that the maximum marginal value of any super-rule of $(\star, \star, c)$ having weight $\leq m_w$ is at most $100m_w$.
+This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$.
 If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$.
@@
 We maintain three sets of rules.
-$C$ is the set of all rules whose marginal values have been counted so far.
+$C$ is the set of all rules whose marginal values have been counted in all previous iterations.
 $C_n$ is the set of rules whose marginal values are going to be counted in the current pass.
-And $C_o$ is the set of rules whose marginal values were counted in the previous pass.
+And $C_o$ is the set of rules whose marginal values were counted in the previous iteration.
 For the first pass, we set $C_n$ to be all rules of size $1$.
@@
 We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$.
-Then for each rule $r$ from $C_n$, we consider the known marginal values of it's sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2.
+Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2.
 Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule.
@@
 When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
-Since reading the table from Hard Disk is usually a bottleneck, it also do things like creating other samples for rules other than $r$, and augmenting existing samples.
-Hence, we assume that in a \textbf{Create} phase, the SampleHandler does a complete re-organization of existing samples.
-Based on the current leaves and their expansion probabilities, it tries to find an optimal set of samples to create for the next step.
-More specifically, for every displayed rule $r$, it determines an integer $n_r$, and then creates a sample $s_r$ with $f_{s_r} = r$ and $|T_{s_r}| = n_r$ while making it's pass through the table.
+Since making a pass over the the table from the Hard Disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
+Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.
+For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table.
+Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule.
+More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making it's pass through the table.
 
-Finally, for each rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
+We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user click can be satisfied using samples available in memory.
+We define some terms required to formalise this problem next.
+
+For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
 Call this quantity $S(r_1, r_2)$.
-Define $S(r_1, r_2)$ to be zero if $r_1$ is not a sub-rule of $r_2$.
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$.
-For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$.
+We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
+Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$.
 If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
@@
 Thus formally our problem is :
-Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximise}$ :
+Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
 $$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
@@
 $$\sum_{r \in U} n_r < \text{capacity}$$
-This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be okay to use an exponential algorithm to solve the problem.
+This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
 
@@
 
-\section{Experiments}
-\subsection{Datasets}
-The first dataset we use is a marketing dataset (MD), which contains demographic information about potential customers.
+\section{Experiments}\label{sec:experiments}
+\subsection{Dataset}
+The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers.
 The source of the dataset is Impact Resources, Inc., Columbus, OH (1987).
@@
 The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Are for $> 10$ years.
-Note that the latter two rules give very specific information which would require upto $3$ user clicks to find in an OLAP system, whereas our system displays that information to the user with a single click.
+Note that the latter two rules give very specific information which would require upto $3$ user clicks to find using Data Cube Drill-Downs, whereas our system displays that information to the user with a single click.
 
@@
 This opens up table~\ref{table:uiexamplestar}.
-It shows the number of females with different levels of education.
+It shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
 
@@
 This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values.
-The rule summary for this weighting is in table~\ref{table:weigtingbitwise}.
+The rule summary for this weighting is in table~\ref{table:weigtingbitwise} (contrast with Table~\ref{table:uiexample1}).
 The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column.
-Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead.
+Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.
 
@@
 This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values.
-As a result, we can see that our system displays rules having $2$ or $3$ non-$\star$ values.
+As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Table~\ref{table:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
+
+%TODO: Give application examples for different weighting functions
+%TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.
 
@@
 Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
-The algorithm is guaranteed to find the best marginal rule as long as it's weight is $leq m_w$, but runs faster for smaller values of $m_w$.
+The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
 In this section, we study the effect of varying $m_w$ on the speed of our algorithm.
+Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.
 
-We fix the weighting function $W$, and a value of $m_w$.
-Then we find the time taken for expanding the empty rule with the weighting function $W$ and $m_w$ parameter, and average this time across $10$ iterations.
+We consider a weighting function $W$, and a value of $m_w$.
+For that value of the $W$ and $m_w$ parameters, we find the time taken for expanding the empty rule.
+We repeat this procedure $10$ times and take the average value of the running times across the $10$ iterations.
 This time is plotted against $m_w$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:mw_speed}.
@@
 At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively.
+Of course, the minimum value of $m_w$ we can use is not known to us beforehand.
+But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.
 
@@
 
-\section{Extensions}
+\section{Extensions}\label{sec:extensions}
 \subsection{Dealing with Numerical Attributes}

### 2014-11-06 TableSummarization/src/dataextraction/SampleHandler.java

@@ -12,9 +12,9 @@ import solvers.Scorer;
  * when to discard a sample to free memory, and which sample(s) to use for evaluating a query. 
  */
 public class SampleHandler {
-	TableInfo table;
-	int minSampleSize; // Minimum sample size we need (depends on how accurate we want our numbers to be, and how low selectivity rules we want to detect.
-	int capacity; // Maximum number of rows we can store across all samples. Maybe change this to memory instead of rows.
+	final public TableInfo table;
+	final public int minSampleSize; // Minimum sample size we need (depends on how accurate we want our numbers to be, and how low selectivity rules we want to detect.
+	final public int capacity; // Maximum number of rows we can store across all samples. Maybe change this to memory instead of rows.
 	Set<TableSample> samples; // Set of current samples.
 	
 	

### 2014-11-06 TableSummarization/src/solvers/NonStarCountSolvers.java

@@ -17,6 +17,7 @@ import java.util.Set;
 
 import solvers.Rule.sizeScorer;
 import dataextraction.Marketing;
+import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
 import dataextraction.TableSample;
 
@@ -1148,8 +1149,8 @@ public class NonStarCountSolvers {
 	 * modifies results to add in the components of baseRule.
 	 */
 	public static Set<Rule> getSolution (TableInfo table, Rule baseRule, Integer ruleNums, Integer maxRuleScore, final Scorer scorer,
-			Integer requiredColumn) throws IOException {
-		Integer sampleSize = Integer.MAX_VALUE;
+			Integer requiredColumn, SampleHandler sampleHandler) throws IOException {
+		Integer sampleSize = sampleHandler.minSampleSize;
 		final TableSample sample = TableSample.createSample(table, baseRule, sampleSize);
 		final TableInfo origTable = table;
 		Scorer sampleScorer = new Scorer () {
@@ -1234,7 +1235,10 @@ public class NonStarCountSolvers {
 		Integer ruleNums = 5;
 		//getSolution (table, ruleNums, maxRuleScore);
 		Integer requiredColumn = -1;
-		Set<Rule> solutionSet = getSolution (table, bigRule, ruleNums, maxRuleScore, scorer, requiredColumn);
+		int minSampleSize = Integer.MAX_VALUE;
+		int capacity = Integer.MAX_VALUE;
+		SampleHandler sampleHandler = new SampleHandler(table, capacity, minSampleSize);
+		Set<Rule> solutionSet = getSolution (table, bigRule, ruleNums, maxRuleScore, scorer, requiredColumn, sampleHandler);
 		out.println(solutionSet.toString());
 		if(1!=2)return;
 		List<Rule> rules = new ArrayList<Rule>();

### 2014-11-06 TableSummarization/src/solvers/RuleTree.java

@@ -17,14 +17,15 @@ import java.util.Set;
 import java.util.Stack;
 
 import solvers.RuleTree.RuleNode;
+import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
 
 public class RuleTree {
 	public static class RuleNode {
-		final Rule rule;
-		List<RuleNode> children;
-		RuleNode parent;
-		Integer depth;
+		final public Rule rule;
+		public List<RuleNode> children;
+		public RuleNode parent;
+		public Integer depth;
 		
 		public RuleNode (Rule rule) {
 			this.rule = rule;
@@ -41,7 +42,7 @@ public class RuleTree {
 	
 	TableInfo table;
 	Map<Rule, RuleNode> nodeMap;
-	RuleNode root;
+	final public RuleNode root;
 	
 	/**
 	 * Creates a new ruletree with only a single node, which is the empty rule. 
@@ -153,20 +154,16 @@ public class RuleTree {
 		deleteChildren(rule);
 	}
 	
-	public void expandStar (Rule rule, Integer ruleNums, Integer maxRuleScore, final Scorer scorer, final Integer col) throws IOException {
-		Set<Rule> solutionSet = NonStarCountSolvers.getSolution (table, rule, ruleNums, maxRuleScore, scorer, col);
+	public void expandStar (Rule rule, Integer ruleNums, Integer maxRuleScore, final Scorer scorer, final Integer col, SampleHandler sampleHandler) throws IOException {
+		Set<Rule> solutionSet = NonStarCountSolvers.getSolution (table, rule, ruleNums, maxRuleScore, scorer, col, sampleHandler);
 		for (Rule solutionRule : solutionSet) {
 			addChild (rule, solutionRule);
 		}
 	}
 
-	public void expandColumn (Integer col) {
-		// To implement.  Should simply make call to QueryEngine and add results.
-	}
-	
-	public void expandRow (Rule rule, Integer ruleNums, Integer maxRuleScore, Scorer scorer) throws IOException {
+	public void expandRow (Rule rule, Integer ruleNums, Integer maxRuleScore, Scorer scorer, SampleHandler sampleHandler) throws IOException {
 		// Modify this later to intelligently use exiting samples, etc.
-		Set<Rule> solutionSet = NonStarCountSolvers.getSolution (table, rule, ruleNums, maxRuleScore, scorer, -1);
[…]

### 2014-11-06 TableSummarization/src/solvers/RuleTreeDisplay.java

@@ -12,6 +12,7 @@ import java.util.Set;
 import java.util.Stack;
 
 import dataextraction.Marketing;
+import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
 import solvers.Rule.sizeBitsScorer;
 import solvers.Rule.sizeMinusKScorer;
@@ -188,7 +189,12 @@ public class RuleTreeDisplay {
 	}
 	
 	public static void main (String[] args) throws IOException {
-		String outFile = "Data_Graphs/mw_speed_bits";
+		final String mw_bits_outFile = "Data_Graphs/mw_speed_bits";
+		final String mw_size_outFile = "Data_Graphs/mw_speed_size";
+		final String minSS_size_outFile = "Data_Graphs/minSS_speed_size";
+		final String minSS_bits_outFile = "Data_Graphs/minSS_speed_bits";
+		final String minSS_size_error_outFile = "Data_Graphs/minSS_error_size";
+		final String minSS_bits_error_outFile = "Data_Graphs/minSS_error_bits";
 		List<Integer> columns = new ArrayList<Integer>();
 		final Integer firstNumColumns = 7;
 		for (int i = 1; i < firstNumColumns; i++) {
@@ -196,14 +202,25 @@ public class RuleTreeDisplay {
 		}
 		TableInfo fullTable = Marketing.parseData();
 		TableInfo table = fullTable.getSubTable(columns);
+		Integer ruleNums = 4;
+		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_outFile);
+		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeScorer(), mw_size_outFile);
+		
+		Experiments.minSSSpeedTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_outFile);
+		Experiments.minSSSpeedTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_outFile);
+		
+		//Experiments.minSSErrorTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_error_outFile);
+		//Experiments.minSSErrorTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_error_outFile);
+				
+		if(1!=2)return;
[…]

### 2014-11-06 TableSummarization/src/utils/Experiments.java

@@ -4,15 +4,21 @@ import java.io.FileWriter;
 import java.io.IOException;
 import java.io.PrintWriter;
 import java.util.HashMap;
+import java.util.Map;
 
 import solvers.Rule;
 import solvers.RuleTree;
+import solvers.RuleTree.RuleNode;
 import solvers.RuleTreeDisplay;
 import solvers.Scorer;
+import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
 
 public class Experiments {
 	public static void mwSpeedTest (TableInfo table, int mwMin, int mwMax, int ruleNums, Scorer scorer, String outFile) throws IOException {
+		int minSampleSize = Integer.MAX_VALUE;
+		int capacity = Integer.MAX_VALUE;
+		SampleHandler sampleHandler = new SampleHandler(table, capacity, minSampleSize);
 		PrintWriter pw = new PrintWriter(new FileWriter(outFile));
 		Rule emptyRule = new Rule(new HashMap<Integer, Integer>(), table.dictionary.size());
 		int numIters = 10;
@@ -21,10 +27,64 @@ public class Experiments {
 			for (int iter = 0; iter < numIters; iter++ ) {
 				RuleTree ruleTree = new RuleTree(table);
 				long timer = System.currentTimeMillis();
-				ruleTree.expandRow(emptyRule, ruleNums, mw, scorer);
+				ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
 				totalTime += System.currentTimeMillis() - timer;	
 			}
 			pw.println(mw + "\t" + totalTime/numIters);
+		}
+		pw.close();	
+	}
+	
+	public static void minSSSpeedTest (TableInfo table, int mw, int minSSMin, int minSSMax, int minSSStep, int ruleNums, Scorer scorer, String outFile) throws IOException {
+		int capacity = Integer.MAX_VALUE;
+		PrintWriter pw = new PrintWriter(new FileWriter(outFile));
+		Rule emptyRule = new Rule(new HashMap<Integer, Integer>(), table.dictionary.size());
+		int numIters = 50;
+		for (int minSS = minSSMin; minSS < minSSMax; minSS += minSSStep) {
+			SampleHandler sampleHandler = new SampleHandler(table, capacity, minSS);
+			long totalTime = 0;
+			for (int iter = 0; iter < numIters; iter++ ) {
[…]

### 2014-11-07 TableSummarization/Paper/TableSummarization.tex
Added experiments to paper

@@
 
+\subsection{Effects of $minSS$}
+We now study the effects of the sampling parameter $minSS$.
+This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.
+Higher values of $minSS$ give us more accurate counts for rules, but also increase computation time.
+
+We consider a value of $minSS$, and a weight function $W$.
+For those parameters, we expand the empty rule and find the time taken to expand it, as well the the error in counts of the displayed rules.
+The error is the absolute value of the difference between the actual count of the rule on the entire table, and it's estimated count from the sample.
+We repeat this $50$ times for each $minSS$ and $W$, and find the average time and error per $minSS$ and $W$.
+The average time over these $50$ iterations is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}.
+The average error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error}.
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
+\vspace{-13pt}
+\end{figure}
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/minSS_error_count.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error}}
+\vspace{-13pt}
+\end{figure}
+
 % TODO: Experiments on Sampling schemes.

### 2014-11-14 TableSummarization/Paper/TableSummarization.tex
Changes in paper

--- 2014-11-06 TableSummarization/Paper/TableSummarization.tex
@@
 \begin{abstract}
-In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints.
-The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that satisfy that pattern.
-For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column.
-We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table.
-We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and an optimized sampling scheme for efficiently summarizing large tables.
-Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
+In this paper we present {\em smart drill down},
+a system for interactively exploring a relational table
+to discover ``interesting'' groups of tuples.
+Each group of tuples is described by a {\em rule}.
+For instance, the rule $(a, b, \star, 1000)$ tells us that
+there are a $1000$ tuples with value $a$ in the first column and $b$
+in the second column (and any value in the third column).
+Smart drill down presents an analyst with a list of rules that
+together describe interesting aspects of the table.
+The analyst can tailor the definition of interesting,
+and can interactively drill down on an existing rule to
+explore that part of the table.
+We describe an
+algorithm for finding the approximately optimal list of rules, and an optimized
+sampling scheme for efficiently summarizing large tables. Finally, we
+demonstrate the usefulness of smart drill down by performing experiments on
+a real dataset.
 \end{abstract}
@@
 
-Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it.
-Large amounts of data are hard for a human to make sense of.
-This necessitates summarizing the data in an informative, interpretable way.
-The summary also needs to be fast and interactive to facilitate deeper exploration.
-
-In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints.
-A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}.
-Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a black box `weight function' as user input.
-In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.
+Drill down (and roll up) [citation] are very useful tools for exploring
+relational tables.
+In this paper we study an extension of drill down
+that lets users zoom into the more ``interesting'' parts of a table,
+with fewer operations and without having to examine as much data
+as traditional drill down.
+We call our approach {\em smart drill down};
+it does not replace traditional drill down but we will argue is a very
+useful complement.
+The best way to explain smart drill down is through a simple example.
 
 \begin{example}\label{ex:introexample}
-Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'.
-Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money.
-The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it.
-We propose to summarize the result of the query with a set of `rules'.
-For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}.
-The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `weight' in this case denoting the number of non-$\star$ values of the rule for each rule.
-The weight function can in general be customized by the user.
+Consider a table with columns `Department Store', `Product', `Region'
+and `Sales'. Suppose an analyst queries for tuples
+where Sales were higher than some threshold, in order
+to find the best selling products.
+If the resulting table has many tuples,
+the analyst can use traditional drill down to explore it.
+For instance, the system may initially tell the analyst there are
+6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
+as shown in Table~\ref{table:introexample0}.
+The $\star$ character is a wildcard that matches any value in the database.
+The Count attribute can be replaced by another aggregate,
+e.g., the total sales.
+The right-most Weight attribute is the number of non-$\star$ attributes; 
+its significance will be discussed shortly.
+If the analyst drills down on the Store attribute (first $\star$),
+then the system displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
+where $X$ is a Store in the answer table, and $C$
+is the number of tuples for $X$ (or the aggregate sales for $X$).
+
+Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
+he obtains Table~\ref{table:introexample}.
+The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
+that display noteworthy or interesting drill downs.
+The number $3$ is a user specified parameter, which we call $k$.
+
+For example, the tuple (Target, bicycles, $\star$, $200$, $2$)
+says that there are $200$ tuples (out of the 6000) with
+Target as the first column value and bicycle as the second.
+This fact tells the analyst that Target is selling a lot of bicycles.
+The next tuple tells the analyst that comforters are selling well in
+the MA-3 region, across multiple stores. The last tuple
+states that Walmart is doing well in general over multiple products and regions.
+We call each tuple in Table~\ref{table:introexample} a {\em rule}
+to distinguish it from the tuples in the original table that is being explored.
+Each rule summarizes the set of tuples that are described by it.
+Again, instead of Count, the system can display some other aggregate, such as
+the total Sales.
 
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
+\hline Store & Product & Region & Count & Weight \\
 \hline
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \hline
-$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
-$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
 \end{tabular}
-\caption{Example of a rule based table summary \label{table:introexample}}
+\caption{Initial summary}\label{table:introexample0}
 \end{table}
-
-% Things to work into example : Number of rules. Ordering.
-
-The first rule tells the user that there are $6000$ tuples in all.
-We call it an `empty' rule, because it has only $\star$ values.
-The three additional rules give more information about the table.
-The number of additional rules ($3$ in this case) is a parameter that can be set by the user.
-The rules are shown in decreasing order by weight.
-
-The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that Target is selling a lot of bicycles.
-The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores.
-The last rule tells the user that walmart is doing well in general over multiple products and states.
-
-This gives the user an overall summary of the results.
-But after seeing this summary, the user may want to dig deeper to get more details on certain rules.
-For instance, the user may want to know which states walmart has more sales in, or which products they sell the most.
-In this case, the user can click on the rule with walmart, and will see an expanded summary as given in table~\ref{table:introexample2}.
 
@@
 \begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
+\hline Store & Product & Region & Count & Weight \\
+\hline
+$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
+$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \hline
+$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \hline
+$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
+\end{tabular}
+\caption{Result after first smart drill down}\label{table:introexample}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{| l | l | l | l | l |}
+\hline Store & Product & Region & Count & Weight \\
 \hline
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
-$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
-$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
-$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
+$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
+$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \cline{1-5}
+$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ Walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ Walmart & $\star$ & CA-1 & $150$ & $2$ \\ \cline{2-5}
+$\triangleright$ $\triangleright$ Walmart & $\star$ & WA-5 & $130$ & $2$ \\ \hline
 \end{tabular}
-\caption{Example of a rule based table summary \label{table:introexample2}}
+\caption{Result after second smart drill down} \label{table:introexample2}
 \end{table}
 
-The three new rules in this table give the user additional information about the $1000$ walmart tuples.
-This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.
-
-In our summary, we want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible.
-Thus choosing rules involves a tradeoff between highly specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$).
-Our decision on what rules to choose is influenced by a rule-weighting scheme, which is further discussed in Section~\ref{sec:weighting}.
-Briefly, we allow the user to specify a weighting function $W$ as a black box, and display the optimal set of rules as per that weighting function.
-
-This example also raises questions on what operations to allow the user to perform to dig deeper into the rules. We describe our user interface in Section \ref{sec:interface}. 
+Say that after seeing the results of Table~\ref{table:introexample},
+the analyst wishes to dig deeper into the Walmart tuples
+represented by the last rule.
+For instance, the analyst may want to know
+which states Walmart has more sales in, or which products they sell
+the most. In this case, the analyst clicks on the Walmart rule,
+obtaining the expanded summary in Table~\ref{table:introexample2}.
+The three new rules in this table provide additional information
+about the $1000$ Walmart tuples.
+In particular, one of the new rules shows that
+Walmart sells a lot of cookies and the others show it sells a lot of products in
+the regions CA-1 and WA-5.
+
+When the analyst clicks on a rule $r$, smart drill down
+expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
+(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
+There are three factors that make a rule set interesting.
+One is if contains rules with high Count (or total sales) fields,
+since the larger the count, the more tuples are summarized.
+A second factor is if the rules have high weight (number of non-$\star$ attributes).
+For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
+seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
+since the former rule tells us the high sales are concentrated in a single region.
+A third desirability factor is diversity:
+For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
+in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
+than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
+describes tuples that are not described by the first rule.
+
+In this paper we describe how to combine or blend these three factors
+in order to obtain a single desirability score for a set of rules.
+Our score function can actually be tuned by the analyst
+(by specifying how weights are computed),
+providing significant flexibility in what is considered a good set of rules.
+We also present an efficient optimization procedure to maximize score, invoked
+by smart drill down to select the set of $k$ rules to display.
+
 \end{example}
 
-To summarize, we attempt to summarize relational data in a way that is:
-\squishlist
-\item \textbf{Flexible : } We allow users to specify a weight function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
-\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
-\item \textbf{Online : } Does not require any heavy pre-computation. 
-\squishend
-
-There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}.
-These papers mostly focuses on giving the most 'surprising' information to the user i.e.
-information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution.
-For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user.
-In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
-Our approaches are thus complementary and serve different purposes.
-% TODO: max-entropy is hard to interpret for human, even though formally very clear.
-Our thing also works well when user doesn't know the set of column values.
-The set, at least number of column values, are needed for imagining the max-entropy distribution.
-
-OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'.
-They have an operation called drill-down, which allows a user to get counts for rules having values in a given column.
-For instance in our example, a user can first drill down along the `Department Store' column, which will show the total number of tuples for each Department Store.
-After that, the user may drill down along the product column for `Walmart', which will show the user the number of tuples for each product in Walmart.
-Thus the user is required to decide which columns to drill down along in order to get interesting or frequent rules, and requires $k$ drill downs to get to rules of size $k$.
-In contrast, our summary may immediately display a size $k$ rule to the user if it has sufficiently high weight and count.
-There is no notion of ranking or ordering of rules in OLAP, since all possible rules are displayed for any drill down.
-In contrast, we only display the top few rules in terms of their score, and display them in decreasing order by weight.
-This makes the summary shorter and more readable by focusing only on the frequently occurring values in columns.
+Compared to traditional drill down, our smart drill down has two important advantages:
+\begin{itemize}
+\item
+Smart drill down limits the information displayed
+to the most interesting $k$ facts (rules).
+With traditional drill down, a column is expanded and {\em all}
+attribute values are displayed in arbitrary order.
+In our example, if we drill down on say the store attribute,
+we would see all stores listed, which may be a very large number.
+\item
+Smart drill down explores several attributes to open up together,
+and automatically selects combinations that are interesting.
+For example, in Table~\ref{table:introexample},
+the rule (Target, bicycles, $\star$, $200$, $2$)
+is obtained after a single drill down;
+with a traditional approach, the analyst would first have to drill down on
+Store, examine the results, drill down on Product,
+look through all the displayed rules and then find the interesting rule
+(Target, bicycles, $\star$, $200$, $2$).
+\end{itemize}
+
+Incidentally, note that in the example we only described one type of smart drill down,
+where the analyst selects a {\em rule} to drill down on
+(e.g., the Walmart rule going from Table~\ref{table:introexample}
+Table~\ref{table:introexample2}.
+In Section~\ref{sec:interface} we describe another option
+where the analyst clicks an a {\em *} to obtain
+rules that have values in that location.
+
+Our work on smart drill down is related
+to table summarization and anomaly
+detection~\cite{Sarawagi:2001:UMA:767141.767148,
+Sarawagi00user-adaptiveexploration,
+Sarawagi98discovery-drivenexploration,
+DBLP:journals/pvldb/GebalyAGKS14}.
+These papers mostly focus on
+giving the most ``surprising'' information to the user, i.e., information
+that would minimize the Kullback-Liebler(KL) divergence of the
+resulting maximum entropy distribution. For instance, if a certain set of
+values occur together in an unexpectedly small number of tuples, that
+set of values may be displayed to the user. In contrast, our algorithm
+focuses on rules with high counts, covering as
+much of the table as possible.
+Furthermore, our summarization is couched in
+an interactive environment, where the analyst
+directs the drill down and can tailor the optimization criteria.
+Nevertheless, one can envision extending traditional
+and smart drill down to provide an additional option of anomaly detection.
 
@@
 
-\item In Section~\ref{sec:preliminaries}, we introduce the relevant notation and formally define our problem.
-
-\item In Section~\ref{sec:weighting}, we consider different schemes for weighting rules, along with their practical significance.
-
-\item In Section~\ref{sec:interface}, we describe our user interface.
-
-\item In Section~\ref{sec:algorithms}, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.
-
-\item In Section~\ref{sec:experiments}, we experimentally evaluate the performance of our algorithms on a real dataset.
+\item In Section~\ref{sec:preliminaries}, we introduce the relevant
+notation and formally define smart drill down.
+
+\item In Section~\ref{sec:weighting}, we consider different schemes
+for weighting rules, along with their practical significance.
+
+\item In Section~\ref{sec:interface}, we describe our
+interactive user interface.
+
+\item In Section~\ref{sec:algorithms}, we present our algorithms for
+finding optimal sets of rules, as well as our sampling schemes
+for dealing with large tables.
+
+\item Based on our implemented smart drill down,
+in Section~\ref{sec:experiments} we experimentally evaluate
+performance on a real dataset,
+and show additional examples of smart drill down in action.
+
+\item In Section~\ref{sec:extensions}, we describe how to extend our results to summarize numerical attributes, and to use the $Sum$ aggregate instead of $Count$.
 
@@
 
-A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. 
-%TODO: Give name to out system (and paper)
+A {\em rule-list} is an ordered list of rules returned by our system as a result of a smart drill-down by the user.
 
@@
 
-When a user clicks on a rule to know more about the part of the table covered by that rule, we display a rule-list of sub-rules of the rule.
-For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule.
+When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
+For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule.
 Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.
@@
 
+Throughout this paper, we use the {\em Count} aggregate of a rule when determining what rules to display to the user.
+We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead.
+We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in Section~\ref{sec:extensions-sum}.
+
 The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$ Our problem is to choose the rule-list that maximises total score.
@@
 $$\sum_{r \in R} W(r)MCount(r)$$
-for one of the following interactions:
+for one of the following smart drill-downs:
 \begin{enumerate}
-\item $[$Rule Interaction$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
-\item $[$Star Interaction$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
+\item $[$Rule Drill-Down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
+\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
 \end{enumerate}
@@
 \label{sec:interface}
-We now explain how the user can interact with our system to explore the contents of a table.
-
-When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}.
-At any point, the user can click on either a rule, or a star within a rule, to `expand' the rule.
+We now explain how the user can use smart drill-downs in our system to explore the contents of a table.
+
+When the user starts using our system, he sees a table with a single trivial rule as shown in table~\ref{table:introexample0}.
+At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill-down' on the rule.
 Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
 By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-As an example of this operation, clicking on the trivial rule of table~\ref{table:trivial} would display table~\ref{table:introexample}.
-Clicking further on the second rule would display table~\ref{table:introexample2}.
+As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}.
+Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}.
 The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
@@
 For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
-\hline
-$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-\end{tabular}
-\caption{Table with only the trivial rule \label{table:trivial}}
-\end{table}
 
@@
 $\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
-$\triangleright$ target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
+$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
 $\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
@@
 \label{sec:algorithms}
-When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into.
+When the user drills down on a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into.
 If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$.
@@
 \squishlist
-\item $[$Rule Interaction$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
+\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
 Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
-\item $[$Star Interaction$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$.
+\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$.
 In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.
@@
 The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
-Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, our algorithm is guaranteed to return $S_o$.
-On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$.
-Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
-In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
+Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, Algorithm~\ref{algo:best-rule-set} is guaranteed to return $S_o$.
+On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by Algorithm~\ref{algo:best-rule-set} does not contain $r$.
+Algorithm~\ref{algo:best-rule-set} runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
+In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
 This is because as the size (and weight) of a rule increases, it's Count falls rapidly.
@@
 Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
-Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set are small.
+Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
 
@@
 
-The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
+Algorithm~\ref{algo:best-rule-set} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
 To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.
@@
 The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
-The scaling factor $N_s$ is used to translate counts of a rule on the sample, into estimated counts on the entire table.
-The sum of $|T_s|$ is not allowed to exceed capacity $M$ at any point.
-
-Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
+The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
+The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
+
+Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
 The SampleHandler also computes $N_s$ when a sample is created.
-Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
-The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get the estimated count on the entire table.
+Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
+The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
 
@@
 $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq \text{minSS}$, then it simply returns sample $s$.
-Our greedy algorithm can then be run on $s$.
+Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
 
@@
 if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds minSS in size, then we can simply treat that set as our sample for rule $r$.
+If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
 We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
@@
 Note that the \textbf{Combine} procedure doesn't really require additional memory.
-Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by the Greedy Algorithm.
+Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
 
@@
 Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-In addition, creating a new sample requires memory, and since capacity is limited, it may necessitate shrinking or deleting some existing samples.
+In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples.
 We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
@@
 In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available).
-Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later expands a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
+Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
 For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
@@
 
-We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in section~\ref{sec:sampling_algorithms}.
+We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in Section~\ref{sec:sampling_algorithms}.
 
@@
 If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
-But after expanding $r$, there is a high chance that the user goes further and expands one of the sub-rules $r^{\prime}$ of $r$.
+But after expanding $r$, there is a high chance that the user goes further and drills down on of the sub-rules $r^{\prime}$ of $r$.
 We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
-So while the user is reading the current rule-list from expanding $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
+So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
 That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
@@
 Let $L$ be the set of leaves.
-Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for leaves.
+Each leaf is something that the user can potentially expand (drill down on) in the next step, and thus we would like to have pre-fetched samples for leaves.
 
 We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
-In the absence of additional information, we can assume a uniform distribution.
-That is, every leaf is equally likely to be expanded next.
+In the absence of additional information, we can assume a uniform probability distribution.
+That is, we can assume that every leaf is equally likely to be expanded next.
 However, we can also use Machine Learning on past user behaviour, along with rule features such as `rule depth in tree' and `rule distance from last expanded rule' to get a better probability estimate of each rule being expanded.
@@
 
-We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user click can be satisfied using samples available in memory.
+We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user drill down can be satisfied using samples available in memory.
 We define some terms required to formalise this problem next.
@@
 
+Unless otherwise specified, in all our experiments, we restrict the table to the first $7$ columns in order to make the result tables fit into the page.
+We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
+
 \subsection{User interface testing}
-We restrict the table to the first $7$ columns in order to make the result tables fit into the page.
-We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$.
 We now present the rule-based summaries displayed as a result of a few different user actions.
@@
 
+\subsection{Effects of $minSS$}
+We now study the effects of the sampling parameter $minSS$.
+This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.
+Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.
+
+We consider one value of $minSS$ and one weight function $W$ at a time.
+For those values of $minSS$ and $W$, we drill down on the empty rule and measure the time taken for this operation.
+We also measure the percent error in the estimated counts of the displayed rules.
+That is, for each displayed rule $r$, if the displayed (estimated) count if $c_1$ and the actual count (computed separately on the entire table) is $c_2$, then the percent error for rule $r$ is $\frac{100 \times |c_1-c_2|}{c_2}$.
+We consider the average of percent errors over all displayed rules.
+For each value of $minSS$ and $W$, we drill down on the empty rule and find the computation time and percent error $50$ times, and take the average value for time and error over those $50$ iterations.
+This average time is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}.
+The average percent error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error_percent}.
+
+The computation time increases approximately linearly with $minSS$.
+This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, which takes time linear in the size of the sample.
+The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$.
+
+In addition, we measure the number of incorrect rules per iteration.
+If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule.
+We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}.
+This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function.
+Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
+\vspace{-13pt}
+\end{figure}
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
+\vspace{-13pt}
+\end{figure}
+
+\begin{figure}
+\hspace{-20pt}
+  \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+\vspace{-10pt}
+\caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
+\vspace{-13pt}
+\end{figure}
+
 % TODO: Experiments on Sampling schemes.
@@
 \section{Extensions}\label{sec:extensions}
-\subsection{Dealing with Numerical Attributes}
+\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
 Our algorithm assumes that all attributes are categorical in nature.
@@
 To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.
+
+\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
+Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
+However, if we have a numerical column (i.e.
+a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead.
+Suppose we are given a measure column $c_m$.
+Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule.
+{\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$.
+The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r)W(r)$.
+Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.
 

### 2014-11-14 TableSummarization/src/solvers/RuleTreeDisplay.java

@@ -206,11 +206,11 @@ public class RuleTreeDisplay {
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_outFile);
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeScorer(), mw_size_outFile);
 		
-		Experiments.minSSSpeedTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_outFile);
-		Experiments.minSSSpeedTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_outFile);
+		//Experiments.minSSSpeedTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_outFile);
+		//Experiments.minSSSpeedTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_outFile);
 		
-		//Experiments.minSSErrorTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_error_outFile);
-		//Experiments.minSSErrorTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_error_outFile);
+		Experiments.minSSErrorTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_error_outFile);
+		Experiments.minSSErrorTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_error_outFile);
 				
 		if(1!=2)return;
 		int minSampleSize = Integer.MAX_VALUE;

### 2014-11-14 TableSummarization/src/utils/Experiments.java

@@ -4,7 +4,9 @@ import java.io.FileWriter;
 import java.io.IOException;
 import java.io.PrintWriter;
 import java.util.HashMap;
+import java.util.HashSet;
 import java.util.Map;
+import java.util.Set;
 
 import solvers.Rule;
 import solvers.RuleTree;
@@ -51,7 +53,7 @@ public class Experiments {
 			}
 			pw.println(minSS + "\t" + totalTime/numIters);
 		}
-		pw.close();	
+		pw.close();
 	}
 	
 	public static void minSSErrorTest (TableInfo table, int mw, int minSSMin, int minSSMax, int minSSStep, int ruleNums, Scorer scorer, String outFile) throws IOException {
@@ -60,31 +62,42 @@ public class Experiments {
 		Rule emptyRule = new Rule(new HashMap<Integer, Integer>(), table.dictionary.size());
 		int numIters = 50;
 		Map<Rule, Integer> ruleCounts = new HashMap<Rule, Integer>();
+		Set<Rule> selectedRules = new HashSet<Rule>();
 		{
 			final int minSS = Integer.MAX_VALUE;
 			SampleHandler sampleHandler = new SampleHandler(table, capacity, minSS);
 			RuleTree ruleTree = new RuleTree(table);
-			ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
+			ruleTree.expandRow(emptyRule, 2 * ruleNums, mw, scorer, sampleHandler);
 			for (RuleNode child : ruleTree.root.children) {
 				ruleCounts.put(child.rule, child.rule.count);
+			}
+			ruleTree.contractRow(emptyRule);
+			ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
+			for (RuleNode child : ruleTree.root.children) {
+				selectedRules.add(child.rule);
 			}
 		}
 		for (int minSS = minSSMin; minSS < minSSMax; minSS += minSSStep) {
 			SampleHandler sampleHandler = new SampleHandler(table, capacity, minSS);
 			long countError = 0;
-			int ruleError = 0;
+			Double percentCountError = 0.0;
+			Double ruleError = 0.0;
 			for (int iter = 0; iter < numIters; iter++ ) {
 				RuleTree ruleTree = new RuleTree(table);
 				ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
 				for (RuleNode node : ruleTree.root.children) {
[…]

### 2014-11-14 TableSummarization/Paper/TableSummarization.tex
split, first pass on formal

--- 2014-11-06 TableSummarization/Paper/TableSummarization.tex
@@
 \usepackage{framed}
-
+\usepackage[normalem]{ulem}
 \newtheorem{theorem}{Theorem}
@@
 
-\newcommand{\agp}[1]{ \textcolor{red}{\bf ||Ins: #1||}}
+\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
 \newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}
@@
 
+\newcommand{\calD}{\mathcal{D}\xspace}
+
 \title{Interactive Summarization of Relational Tables}
@@
 \begin{abstract}
-In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints.
-The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that satisfy that pattern.
-For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column.
-We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table.
-We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and an optimized sampling scheme for efficiently summarizing large tables.
-Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
+In this paper we present {\em smart drill down},
+a system for interactively exploring a relational table
+to discover ``interesting'' groups of tuples.
+Each group of tuples is described by a {\em rule}.
+For instance, the rule $(a, b, \star, 1000)$ tells us that
+there are a $1000$ tuples with value $a$ in the first column and $b$
+in the second column (and any value in the third column).
+Smart drill down presents an analyst with a list of rules that
+together describe interesting aspects of the table.
+The analyst can tailor the definition of interesting,
+and can interactively drill down on an existing rule to
+explore that part of the table.
+We describe an
+algorithm for finding the approximately optimal list of rules, and an optimized
+sampling scheme for efficiently summarizing large tables. Finally, we
+demonstrate the usefulness of smart drill down by performing experiments on
+a real dataset.
 \end{abstract}
 
-\section{Introduction}
+\input{intro}
 
-Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it.
-Large amounts of data are hard for a human to make sense of.
-This necessitates summarizing the data in an informative, interpretable way.
-The summary also needs to be fast and interactive to facilitate deeper exploration.
+\input{formal}
 
-In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints.
-A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}.
-Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a black box `weight function' as user input.
-In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.
+\input{algorithms}
 
-\begin{example}\label{ex:introexample}
-Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'.
-Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money.
-The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it.
-We propose to summarize the result of the query with a set of `rules'.
-For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}.
-The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `weight' in this case denoting the number of non-$\star$ values of the rule for each rule.
-The weight function can in general be customized by the user.
+\input{exp}
 
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
-\hline
-$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \hline
-$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
-$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
-\end{tabular}
-\caption{Example of a rule based table summary \label{table:introexample}}
-\end{table}
+\input{extensions}
 
-% Things to work into example : Number of rules. Ordering.
-
-The first rule tells the user that there are $6000$ tuples in all.
-We call it an `empty' rule, because it has only $\star$ values.
-The three additional rules give more information about the table.
-The number of additional rules ($3$ in this case) is a parameter that can be set by the user.
-The rules are shown in decreasing order by weight.
-
-The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that Target is selling a lot of bicycles.
-The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores.
-The last rule tells the user that walmart is doing well in general over multiple products and states.
-
-This gives the user an overall summary of the results.
-But after seeing this summary, the user may want to dig deeper to get more details on certain rules.
-For instance, the user may want to know which states walmart has more sales in, or which products they sell the most.
-In this case, the user can click on the rule with walmart, and will see an expanded summary as given in table~\ref{table:introexample2}.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
-\hline
-$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
-$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
-$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
-$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
-\end{tabular}
-\caption{Example of a rule based table summary \label{table:introexample2}}
-\end{table}
-
-The three new rules in this table give the user additional information about the $1000$ walmart tuples.
-This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.
-
-In our summary, we want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible.
-Thus choosing rules involves a tradeoff between highly specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$).
-Our decision on what rules to choose is influenced by a rule-weighting scheme, which is further discussed in Section~\ref{sec:weighting}.
-Briefly, we allow the user to specify a weighting function $W$ as a black box, and display the optimal set of rules as per that weighting function.
-
-This example also raises questions on what operations to allow the user to perform to dig deeper into the rules. We describe our user interface in Section \ref{sec:interface}. 
-\end{example}
-
-To summarize, we attempt to summarize relational data in a way that is:
-\squishlist
-\item \textbf{Flexible : } We allow users to specify a weight function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
-\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
-\item \textbf{Online : } Does not require any heavy pre-computation. 
-\squishend
-
-There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}.
-These papers mostly focuses on giving the most 'surprising' information to the user i.e.
-information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution.
-For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user.
-In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
-Our approaches are thus complementary and serve different purposes.
-% TODO: max-entropy is hard to interpret for human, even though formally very clear.
-Our thing also works well when user doesn't know the set of column values.
-The set, at least number of column values, are needed for imagining the max-entropy distribution.
-
-OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'.
-They have an operation called drill-down, which allows a user to get counts for rules having values in a given column.
-For instance in our example, a user can first drill down along the `Department Store' column, which will show the total number of tuples for each Department Store.
-After that, the user may drill down along the product column for `Walmart', which will show the user the number of tuples for each product in Walmart.
-Thus the user is required to decide which columns to drill down along in order to get interesting or frequent rules, and requires $k$ drill downs to get to rules of size $k$.
-In contrast, our summary may immediately display a size $k$ rule to the user if it has sufficiently high weight and count.
-There is no notion of ranking or ordering of rules in OLAP, since all possible rules are displayed for any drill down.
-In contrast, we only display the top few rules in terms of their score, and display them in decreasing order by weight.
-This makes the summary shorter and more readable by focusing only on the frequently occurring values in columns.
-
-We further discuss related work in Section~\ref{sec:related}.
-
-\stitle{Overview of paper:} 
-\squishlist
-
-\item In Section~\ref{sec:preliminaries}, we introduce the relevant notation and formally define our problem.
-
-\item In Section~\ref{sec:weighting}, we consider different schemes for weighting rules, along with their practical significance.
-
-\item In Section~\ref{sec:interface}, we describe our user interface.
-
-\item In Section~\ref{sec:algorithms}, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.
-
-\item In Section~\ref{sec:experiments}, we experimentally evaluate the performance of our algorithms on a real dataset.
-
-\item In Section~\ref{sec:related}, we describe related work.
-\squishend
-
-\section{Preliminaries} 
-\label{sec:preliminaries}
-Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table.
-
-Our objective (formally defined later) is to summarize the table with a list of {\em rules}.
-A {\em rule} is a tuple with a value for each column of the queried table.
-In addition, a rule has other attributes, such as count and weight (which we define later) associated with it.
-The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column.
-For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value.
-A {\em trivial rule} is one that has a $\star$ value in all columns.
-
-A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple.
-We abuse notation to write this as $t \in r$.
-
-A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. 
-%TODO: Give name to out system (and paper)
-
-We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values.
-For example, rule ($a$, $\star$) subsumes rule ($a$, $b$).
-Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
-We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e.
-there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not.
-If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.
-
-When a user clicks on a rule to know more about the part of the table covered by that rule, we display a rule-list of sub-rules of the rule.
-For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule.
-Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.
-
-Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
-Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.
-
-We now define some additional properties of rules. Let $r$ be a rule. Then:
-\squishlist
-\item {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
-\item {\em MCount}($r$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
-A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules.
-We want to pick rules with a high value of $MCount$ to display to the user, to increase the coverage of the rule-list.
-\item {\em Size}($r$) refers to the number of non-starred values in rule $r$. 
-\item We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules.
-A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$.
-We further describe our weighting functions in Section~\ref{sec:weighting}.
-\squishend
-
-The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$ Our problem is to choose the rule-list that maximises total score.
-We use $MCount$ rather than Count to increase coverage of the table.
-If we had defined total score as $\sum_{r \in R} \text{Count}(r)W(r)$, then our optimal rule-list would contain rules that repeatedly refer to the most `summarizable' part of the table.
-For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then the summary could consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.
-
-We now formally define our problem:
-
-\begin{problem}\label{prob:optimal-subrule-list}
-Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
-$$\sum_{r \in R} W(r)MCount(r)$$
-for one of the following interactions:
-\begin{enumerate}
-\item $[$Rule Interaction$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
-\item $[$Star Interaction$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
-\end{enumerate}
-\end{problem}
-
-% Describe architecture here? (With a diagram). Show SampleHandler, QueryEngine, etc.
-
-\section{Weighting Rules}
-\label{sec:weighting}
-We now describe how we score rules.
-At a high level, we want our rules to be as descriptive of the table as possible, i.e.
-given the rules, it should be as easy as possible to reproduce the table.
-We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is.
-Given this weight per rule, our total score for our rule-list is given by : $\sum_{r\in R} MCount(r)W(r)$.
-Let us consider some forms for the function $W(r)$ and their interpretation:
-\begin{enumerate}
-\item $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :
-
-Here we set weight equal to the size of the rule.
-Consider the examples in Table \ref{table:sizescoringexample}.
-The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$.
-Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$.
-If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$.
-Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$.
-If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
-
-To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
-Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule.
-For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns).
-Thus this rule contributes $200$ to the score.
-For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those.
-Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total.
-Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list.
-Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.
-
-\item $W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$.
-Like the previous scoring function, this one adds some weight for every non-starred value in a rule.
-But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column.
-The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
-Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$.
-Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.
-
-The interpretation for this function is similar to the one for the last function.
-Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules.
-But this time, we count the number of `bits' that are pre-filled.
-For a column $c$, specifying a value in the column takes log($|c|$) bits of information.
-A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table.
-Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.
-
-Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
-If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.
-\end{enumerate}
-
-Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
-\begin{enumerate}
-\item Non-negativity : For all rules $r$, $W(r) \geq 0$ 
-\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
-\end{enumerate}
-
-\begin{table}
-\centering
-\begin{tabular}{ | l | c | }
- \hline Rule-MCount list & Score \\ \hline
-  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
-  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
-  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
-\end{tabular}
-\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
-\end{table}
-
-\section{Our interface}
-\label{sec:interface}
-We now explain how the user can interact with our system to explore the contents of a table.
-
-When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}.
-At any point, the user can click on either a rule, or a star within a rule, to `expand' the rule.
-Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
-By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-As an example of this operation, clicking on the trivial rule of table~\ref{table:trivial} would display table~\ref{table:introexample}.
-Clicking further on the second rule would display table~\ref{table:introexample2}.
-The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
-
-Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$.
-This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
-This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
-
-Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation i.e.
-collapses it.
-For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
-\hline
-$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
-\end{tabular}
-\caption{Table with only the trivial rule \label{table:trivial}}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l | l | l |}
-\hline Store & Product & State & Count & Weight \\
-\hline
-$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
-$\triangleright$ target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
-$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
-$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
-$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
-\end{tabular}
-\caption{Result of clicking on a $\star$ \label{table:introexample3}}
-\end{table}
-
-\section{Algorithms}
-\label{sec:algorithms}
-When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into.
-If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$.
-We can reduce Problem~\ref{prob:optimal-subrule-list} to the following simpler problem by removing the user-interaction based constraints:
-
-\begin{problem}\label{prob:optimal-rule-list}
-Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
-$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
-\end{problem}
-
-$W$ is required to be non-negative and monotonic.
-$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.
-
-Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
-\squishlist
-\item $[$Rule Interaction$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
-Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
-\item $[$Star Interaction$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$.
-In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.
-Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
-\squishend
-
-We will describe a greedy algorithm for finding the approximately highest scoring rule-list.
-First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
-
-\begin{lemma}\label{lemma:rule-ordering}
-Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
-$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
-\end{lemma}
-\begin{proof}
-We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$.
-For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to.
-Then Score can be rewritten as:
-$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
-Thus each tuple $t$ `contributes' $W(t)$ to the total score.
-Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$.
-Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
-$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
-as required.
-\end{proof}
-
-Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
-Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order.
-This gives us
-\begin{problem}\label{prob:optimal-rule-set}
-Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
-$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
-\end{problem}
-
-The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.
-
-\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
-We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
-\begin{definition}
-A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
-$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
-\end{definition}
-Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead.
-For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.
-
-\begin{lemma}\label{lemma:submodular}
-For a given table $T$, the Score function over sets $S$ of rules, defined by :
-$$\text{Score}(S) = \sum_{r \in S} MCount(r)W(r)$$
-is submodular.
-\end{lemma}
-\begin{proof}
-For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
-$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
-Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
-$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
-For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
-$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$ 
-Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
-$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.
-
-Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
-$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
-$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
-which means that Score is submodular.
-\end{proof}
-
-Now consider the greedy algorithm:
-
-\begin{framed}
-\begin{enumerate}
-\item Set $S = \phi$
-\item For $i$ from $1$ to $k$
-\begin{enumerate}
-\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
-\item $S = S \cup \left\lbrace r \right\rbrace$
-\end{enumerate}
-\end{enumerate}
-\end{framed}
-
-Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
-
-Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
-The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
-
-The parameter $m_w$ stands for \textit{Max Weight}.
-The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
-Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, our algorithm is guaranteed to return $S_o$.
-On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$.
-Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
-In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
-This is because as the size (and weight) of a rule increases, it's Count falls rapidly.
-The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$).
-Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
-Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set are small.
-
-% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary.
-Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.
-
-The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
-To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.
-
-In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
-A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data.
-But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.
-
-In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}.
-We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule.
-In the $j^{th}$ pass, we compute counts for rules of size $j$.
-To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$.
-In the first pass, we would compute the marginal values of all rules of size $1$.
-In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass.
-For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$.
-Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$.
-This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$.
-If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$.
-Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.
-
-We now describe the function to find the best marginal rule.
-The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.
-The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
-The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration.
-We maintain three sets of rules.
-$C$ is the set of all rules whose marginal values have been counted in all previous iterations.
-$C_n$ is the set of rules whose marginal values are going to be counted in the current pass.
-And $C_o$ is the set of rules whose marginal values were counted in the previous iteration.
-For the first pass, we set $C_n$ to be all rules of size $1$.
-Then we compute marginal values for those rules, and set $C = C_o = C_n$.
-
-For the second pass onwards, we are more selective about which rules to consider for marginal value evaluation.
-We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$.
-Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2.
-Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule.
-Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5.
-If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.
-
-\begin{algorithm}\label{algo:best-rule-set}
-\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$S$ (Solution set of rules)}
-$S = \phi$
-
-\For {$i$ from $1$ to $k$}{
-$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}
-
-$S = S \cup \left\lbrace R_m \right\rbrace$
-}
-\Return $S$
-\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
-\end{algorithm}
-
-\begin{algorithm}\label{algo:best-marginal-rule}
-\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
-$H = 0$ \tcc*{Threshold for deciding if to find count for a rule.}
-$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
-\For{$j$ from $1$ to number of columns in $T$}{
-\If {$j = 1$}{
-$C_n = \text{ all rules of size } 1$
-}
-\Else{ 
-$C_n = \text{all size-}i$ super-rules of rules from $C_o$
-}
-\ForEach {$R \in C_n$}{
-$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
-\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
-$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
-}
-\If {$(M < H)$}{ 
-$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to possibly be in the solution}
-}
-}
-\If {$C_n = \phi$}{
-break;
-}
-\ForEach {$R \in C_n$}{
-Count$(R) = 0$ \tcc*{Initialize}
-MarginalValue$(R) = 0$ \tcc*{Initialize}
-}
-\ForEach {$t \in T$}{
-Let $R_S$ be the highest weight rule in $S$ that covers $t$
-\ForEach {$R \in C_n$ that covers $t$}{
-Count($R$) $++$
-
-MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
-}
-}
-$C_o = C_n$
-
-$C_n = \phi$
-
-$C = C \cup C_o$
-
-$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
-}
-\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
-\caption{Find best marginal rule}
-\end{algorithm}
-
-\subsection{Sampling for big tables}
-Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
-These passes can be very expensive if the table is large, especially if it does not fit in main memory.
-If we want exact counts for rules, we have no choice but to read the entire table.
-
-But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.
-
-Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
-Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
-The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
-We provide a way to find reasonable values of $minSS$ later.
-
-At all points, the SampleHandler maintains a set of Samples in memory.
-Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
-The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
-The scaling factor $N_s$ is used to translate counts of a rule on the sample, into estimated counts on the entire table.
-The sum of $|T_s|$ is not allowed to exceed capacity $M$ at any point.
-
-Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
-The SampleHandler also computes $N_s$ when a sample is created.
-Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
-The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get the estimated count on the entire table.
-
-When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule.
-At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample.
-Creating a new sample by making a pass through the table is called \textbf{Create} (further described below).
-At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
-\begin{enumerate}
-\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as it's filter rule (i.e.
-$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq \text{minSS}$, then it simply returns sample $s$.
-Our greedy algorithm can then be run on $s$.
-
-\item \textbf{Combine:} If \textbf{Find} doesn't work i.e.
-if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds minSS in size, then we can simply treat that set as our sample for rule $r$.
-We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
-That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
-
-Note that the \textbf{Combine} procedure doesn't really require additional memory.
-Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by the Greedy Algorithm.
-
-\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample by making a pass through the table.
-Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-In addition, creating a new sample requires memory, and since capacity is limited, it may necessitate shrinking or deleting some existing samples.
-We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
-
-In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available).
-Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later expands a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
-For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
-
-\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may happen that there is not enough memory left to create a new sample.
-In that case, the SampleHandler needs to delete or shrink an existing sample.
-It can use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
-Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
-%For shrink, prefer shrinking samples of the largest size (> minSS) first?
-\end{enumerate}
-
-We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in section~\ref{sec:sampling_algorithms}.
-
-\textbf{Pre-fetching:} When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user.
-If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
-But after expanding $r$, there is a high chance that the user goes further and expands one of the sub-rules $r^{\prime}$ of $r$.
-We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
-So while the user is reading the current rule-list from expanding $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
-That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
-In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
-
-\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
-We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
-Suppose at some stage, we have a tree $U$ of rules displayed to the user.
-Internal nodes of $U$ are rules that have been expanded, while leaves are rules that have not been expanded.
-Let $L$ be the set of leaves.
-Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for leaves.
-
-We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
-In the absence of additional information, we can assume a uniform distribution.
-That is, every leaf is equally likely to be expanded next.
-However, we can also use Machine Learning on past user behaviour, along with rule features such as `rule depth in tree' and `rule distance from last expanded rule' to get a better probability estimate of each rule being expanded.
-
-When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
-Since making a pass over the the table from the Hard Disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
-Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.
-For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table.
-Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule.
-More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making it's pass through the table.
-
-We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user click can be satisfied using samples available in memory.
-We define some terms required to formalise this problem next.
-
-For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
-Call this quantity $S(r_1, r_2)$.
-We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$.
-If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
-We wish to maximise the probability that we can respond to the next user expansion without making another pass.
-
-Thus formally our problem is :
-Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
-where the $I$'s are indicator variables, subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
-This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
-
-Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
-The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and it's immediate parent.
-That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
-$\text{ess}(r)$ is redefined according to the new value of $S$ as well.
-So $$\text{ess}(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
-
-Now consider a rule in $r_0 \in U \setminus L$ along with all it's children.
-Let $M_{r_0}$ denote the set containing $r_0$ and all it's leaf children.
-Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
-This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$.
-Thus for each non-leaf rule $r_0$ and all it's children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
-Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{\text{ess}(r) \geq \text{minSS}}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
-Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
-We describe both these steps in detail below:
-
-Let $r_0 \in U \setminus L$.
-Let $d$ be the number of leaf children of $r_0$.
-Let the children be $r_1, r_2, ...
-r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$.
-Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
-$$1 \leq i_m \leq i_M \leq d$$
-$$n_{r_0} = \frac{\text{minSS}}{S(r_0, r_{i_m})}$$
-$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
-$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
-$$i_M < j \Rightarrow n_{r_j} = 0$$
-And thus
-$$\text{ess}(r_j) \geq \text{minSS} \Leftrightarrow j \leq i_M$$
-
-Call this locally optimal solution $e$. Then we define its `probability value' to be
-$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
-and its `Sampling Cost' to be
-$$S(e) = \frac{\text{minSS}}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
-Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions.
-
-Then next step is to combine the solutions using dynamic programming.
-Let the $M$ sets be called $M_0, M_1, ...
-M_D$.
-Let our possible sample sizes range from $0$ to $\mathcal{S}$.
-Then we create a $D \times \mathcal{S}$ array $A$.
-The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ...
-M_i$ with total sample size at most $j$.
-We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$.
-Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$.
-Then we have,
-$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
-This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.
-
-\subsubsection{Additional optimizations}
-There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
-Suppose we have a sample $s$, and say it's filter rule $f_s$ has value $v$ in column $c$.
-Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$.
-So we do not need to explicitly store the column $c$ value of any tuple in $T_s$.
-We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
-In addition, we may have a tuple occur in multiple samples.
-Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
-
-\subsubsection{Setting $minSS$}
-Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e.
-$x|T|$ tuples.
-Say we have a uniform random sample $s$ of $T$.
-The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$.
-Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$.
-In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$.
-That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.
-
-We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = \text{minSS}$.
-If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be.
-Moreover, since we want good Count estimates for all rules displayed in the summary, we want $\text{minSS} >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.
-
-Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
-This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule).
-Let $c$ be the column with the fewest distinct values.
-Say it has $|c|$ values.
-Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$.
-For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times.
-So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.
-
-The highest scoring rule can have weight at most $|C|$ (the total number of columns).
-Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
-Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
-For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $\text{minSS} >> 5 \times 10$.
-
-\section{Experiments}\label{sec:experiments}
-\subsection{Dataset}
-The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers.
-The source of the dataset is Impact Resources, Inc., Columbus, OH (1987).
-A total of $N=9409$ questionnaires containing $502$ questions were
-filled out by shopping mall customers in the San Francisco Bay area.
-This dataset is an extract from this survey.
-Each tuple in the data table describes a single person.
-There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on.
-Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.
-
-The columns (in order) are as follows:
-(Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, How long have you lived in the San Fran./Oakland/San Jose Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home).
-
-\subsection{User interface testing}
-We restrict the table to the first $7$ columns in order to make the result tables fit into the page.
-We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$.
-We now present the rule-based summaries displayed as a result of a few different user actions.
-
-To begin with, the user sees an empty rule with the total number of tuples as the count.
-After this, suppose the user expands the rule.
-Then the user will see Table~\ref{table:uiexample1}.
-The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples.
-The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Are for $> 10$ years.
-Note that the latter two rules give very specific information which would require upto $3$ user clicks to find using Data Cube Drill-Downs, whereas our system displays that information to the user with a single click.
-
-Now suppose the user decides to further explore the table, by looking at education related information of females in the dataset.
-So the user clicks on the $\star$ in the 'Education' column of the second rule.
-This opens up table~\ref{table:uiexamplestar}.
-It shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
-
-Instead of expanding the 'Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\hline 
-\end{tabular} 
-\caption{A Rule expansion \label{table:uiexamplerule}} 
-\end{table*}
-
-\subsection{Weighting functions}
-Our system can display optimal rule summaries using any monotone weighting function.
-By default, we assign a rule weight equal to it's size.
-In this section, we try a couple of other weighting functions.
-
-We first try the weighting function given by:
-$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct possible values in column $c$.
-This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values.
-The rule summary for this weighting is in table~\ref{table:weigtingbitwise} (contrast with Table~\ref{table:uiexample1}).
-The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column.
-Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.
-
-The other weighting function we try is given by:
-$$W(r) = \text{Min}(0, \text{Size}(r) - 1)$$
-This gives us table~\ref{table:weightingsizeminusone}.
-This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values.
-As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Table~\ref{table:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
-
-%TODO: Give application examples for different weighting functions
-%TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
-\hline 
-\end{tabular} 
-\caption{Bits scoring\label{table:weigtingbitwise}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
-\hline 
-\end{tabular} 
-\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
-\end{table*}
-
-\subsection{Effects of $m_w$}
-Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
-The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
-In this section, we study the effect of varying $m_w$ on the speed of our algorithm.
-Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.
-
-We consider a weighting function $W$, and a value of $m_w$.
-For that value of the $W$ and $m_w$ parameters, we find the time taken for expanding the empty rule.
-We repeat this procedure $10$ times and take the average value of the running times across the $10$ iterations.
-This time is plotted against $m_w$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:mw_speed}.
-The figure shows that running time seems to be approximately linear in $m_w$.
-
-The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule.
-Thus for size scoring, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
-For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
-At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively.
-Of course, the minimum value of $m_w$ we can use is not known to us beforehand.
-But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.
-
-\begin{figure}
-\hspace{-20pt}
-  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-10pt}
-  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
-\vspace{-13pt}
-\end{figure}
+\input{related}
 
@@
 
-\section{Extensions}\label{sec:extensions}
-\subsection{Dealing with Numerical Attributes}
-Our algorithm assumes that all attributes are categorical in nature.
-Attributes that have a large domain d to have a smaller tuple count per value, and hence don't appear in rule summaries.
-Thus our algorithm does not summarise information about numerical attributes.
-
-However, we can modify the algorithm to deal with numerical attributes.
-Suppose we have a numerical attribute $A$.
-A simple approach is to create buckets for values of $A$.
-We choose a number of buckets $k$, and divide the range of values of $A$ into $k$ intervals, each corresponding to a bucket.
-We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket.
-Then we can use our algorithm, treating the bucket number as a categorical attribute.
-This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).
-
-The distribution of values for the $A$ may not be similar in different parts of the table.
-For instance, we may create buckets have approximately equal numbers of tuples in the table.
-But then suppose a user tries to expand a rule $r$.
-The part of the table covered by $r$ might have a very different distribution of age values.
-For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment.
-To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.
-
-\section{Related Work}\label{sec:related}
-%TODO: Repeat what we said in intro about interactive, etc.
-There has been work on finding interesting rules in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
-This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max entropy distribution.
-The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
-
-There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}.
-This work again focuses on displaying rules that will cause the max entropy distribution to best approximate the actual distribution of values.
-
-Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
-There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
-All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
-
-Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
-Several extensions for the algorithm have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311}.
-We can potentially use these ideas to improve handing of numerical attributes in our work.
-
-% Related Work summary : Is ours the only only that is flexible (due to weight function black box)?
-% Explanation tables takes lots of papers and talks about them in one go in related work.
-saying they optimize for different opbjective functions.
-We can say that as well.
-We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity.
-% Succintly summarizing with itemsets-- Mampaey:2011:TMI:2020408.2020499 -- Also seemed to optimize according to max entropy criterion
-% Efficient and Effective... Tableaux -- Golab_efficientand -- Similar to conditional functional dep. paper.
-% Near Optimal Tableaux Conditional functional dep.
--- Golab:2008:GNT:1453856.1453900 -- Finding a smallest sized tableau that provides some confidence and support for a conditional functional dependency.
-% MDL Summarization with holes -- Bu:2005:MSH:1083592.1083644 -- Min sized summary with a rectangle and exceptions (holes). 
-% Generalized MDL Approach -- Lakshmanan:2002:GMA:1287369.1287435 -- Similar to MDL Approach.
-% Overlapped Hyperrectangle -- Xiang_succinctsummarization -- Efficiently find the set of hyperrectangles that covers the data (sets of itemsets)
-% Tiling Databases Geerts04tilingdatabases Simlar to hyper-rectangles. Probabilistic interpretation of frequent itemsets.
-
 {\small 

### 2014-11-14 TableSummarization/Paper/algorithms.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Algorithms}
\label{sec:algorithms}
When the user drills down on a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$. We can reduce Problem~\ref{prob:optimal-subrule-list} to the following simpler problem by removing the user-interaction based constraints:

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
\end{problem}

$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.

Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\squishlist
\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\squishend

We will describe a greedy algorithm for finding the approximately highest scoring rule-list. First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
\end{lemma}
\begin{proof}
We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$. For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to. Then Score can be rewritten as:
$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
Thus each tuple $t$ `contributes' $W(t)$ to the total score. Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$. Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
as required.
\end{proof}

Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. This gives us 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
\end{problem}

The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.

\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

\begin{lemma}\label{lemma:submodular}
For a given table $T$, the Score function over sets $S$ of rules, defined by :
$$\text{Score}(S) = \sum_{r \in S} MCount(r)W(r)$$
is submodular.
\end{lemma}
\begin{proof}
For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$ 
Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.

Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
which means that Score is submodular.
\end{proof}

Now consider the greedy algorithm:

\begin{framed}
\begin{enumerate}
\item Set $S = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
\item $S = S \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\end{framed}

Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}. The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, Algorithm~\ref{algo:best-rule-set} is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by Algorithm~\ref{algo:best-rule-set} does not contain $r$. Algorithm~\ref{algo:best-rule-set} runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, it's Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.

% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary. Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.

Algorithm~\ref{algo:best-rule-set} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step. To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.

In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}. We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

We now describe the function to find the best marginal rule. The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}. The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far. The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted in all previous iterations. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous iteration. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

For the second pass onwards, we are more selective about which rules to consider for marginal value evaluation. We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$. Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2. Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule. Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5. If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

\begin{algorithm}\label{algo:best-rule-set}
\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
\KwOut{$S$ (Solution set of rules)}
$S = \phi$

\For {$i$ from $1$ to $k$}{
$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}

$S = S \cup \left\lbrace R_m \right\rbrace$
}
\Return $S$
\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
\end{algorithm}

\begin{algorithm}\label{algo:best-marginal-rule}
\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
$H = 0$ \tcc*{Threshold for deciding if to find count for a rule.}
$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
\For{$j$ from $1$ to number of columns in $T$}{
\If {$j = 1$}{
$C_n = \text{ all rules of size } 1$
}
\Else{ 
$C_n = \text{all size-}i$ super-rules of rules from $C_o$
}
\ForEach {$R \in C_n$}{
$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
}
\If {$(M < H)$}{ 
$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to possibly be in the solution}
}
}
\If {$C_n = \phi$}{
break;
}
\ForEach {$R \in C_n$}{
Count$(R) = 0$ \tcc*{Initialize}
MarginalValue$(R) = 0$ \tcc*{Initialize}
}
\ForEach {$t \in T$}{
Let $R_S$ be the highest weight rule in $S$ that covers $t$
\ForEach {$R \in C_n$ that covers $t$}{
Count($R$) $++$

MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
}
}
$C_o = C_n$

$C_n = \phi$

$C = C \cup C_o$

$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
}
\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
\caption{Find best marginal rule}
\end{algorithm}

\subsection{Sampling for big tables}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if it does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be. We provide a way to find reasonable values of $minSS$ later.

At all points, the SampleHandler maintains a set of Samples in memory. Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. The SampleHandler also computes $N_s$ when a sample is created. Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.

When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule. At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample. Creating a new sample by making a pass through the table is called \textbf{Create} (further described below). At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
\begin{enumerate}
\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as it's filter rule (i.e. $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq \text{minSS}$, then it simply returns sample $s$. Algorithm~\ref{algo:best-rule-set} can then be run on $s$.

\item \textbf{Combine:} If \textbf{Find} doesn't work i.e. if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$. If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution. That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.

Note that the \textbf{Combine} procedure doesn't really require additional memory. Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.

\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample by making a pass through the table. Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used. In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples. 
We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.

In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available). Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table. For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.

\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may happen that there is not enough memory left to create a new sample. In that case, the SampleHandler needs to delete or shrink an existing sample. It can use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink. Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
%For shrink, prefer shrinking samples of the largest size (> minSS) first?
\end{enumerate}

We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in Section~\ref{sec:sampling_algorithms}.

\textbf{Pre-fetching:} When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user. If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table. But after expanding $r$, there is a high chance that the user goes further and drills down on of the sub-rules $r^{\prime}$ of $r$. We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples. So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background. That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time. 
In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.

\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory. Suppose at some stage, we have a tree $U$ of rules displayed to the user. Internal nodes of $U$ are rules that have been expanded, while leaves are rules that have not been expanded. Let $L$ be the set of leaves. Each leaf is something that the user can potentially expand (drill down on) in the next step, and thus we would like to have pre-fetched samples for leaves.

We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded. In the absence of additional information, we can assume a uniform probability distribution. That is, we can assume that every leaf is equally likely to be expanded next. However, we can also use Machine Learning on past user behaviour, along with rule features such as `rule depth in tree' and `rule distance from last expanded rule' to get a better probability estimate of each rule being expanded.

When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory. Since making a pass over the the table from the Hard Disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass. Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples. For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table. Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule. More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making it's pass through the table.

We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user drill down can be satisfied using samples available in memory. We define some terms required to formalise this problem next.

For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples. Call this quantity $S(r_1, r_2)$. We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$. If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table. We wish to maximise the probability that we can respond to the next user expansion without making another pass.

Thus formally our problem is :
Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
where the $I$'s are indicator variables, subject to :
$$\sum_{r \in U} n_r < \text{capacity}$$
This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.

Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution. The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and it's immediate parent. That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$. $\text{ess}(r)$ is redefined according to the new value of $S$ as well. So $$\text{ess}(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.

Now consider a rule in $r_0 \in U \setminus L$ along with all it's children. Let $M_{r_0}$ denote the set containing $r_0$ and all it's leaf children. Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$. This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$. Thus for each non-leaf rule $r_0$ and all it's children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$. Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{\text{ess}(r) \geq \text{minSS}}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$. Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s. We describe both these steps in detail below:

Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$. Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
$$1 \leq i_m \leq i_M \leq d$$
$$n_{r_0} = \frac{\text{minSS}}{S(r_0, r_{i_m})}$$
$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
$$i_M < j \Rightarrow n_{r_j} = 0$$
And thus
$$\text{ess}(r_j) \geq \text{minSS} \Leftrightarrow j \leq i_M$$

Call this locally optimal solution $e$. Then we define its `probability value' to be
$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
and its `Sampling Cost' to be
$$S(e) = \frac{\text{minSS}}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions.

Then next step is to combine the solutions using dynamic programming. Let the $M$ sets be called $M_0, M_1, ... M_D$. Let our possible sample sizes range from $0$ to $\mathcal{S}$. Then we create a $D \times \mathcal{S}$ array $A$. The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ... M_i$ with total sample size at most $j$. We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$. Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$. Then we have,
$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.

\subsubsection{Additional optimizations}
There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
Suppose we have a sample $s$, and say it's filter rule $f_s$ has value $v$ in column $c$. Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$. So we do not need to explicitly store the column $c$ value of any tuple in $T_s$. We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
In addition, we may have a tuple occur in multiple samples. Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.

\subsubsection{Setting $minSS$}
Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e. $x|T|$ tuples. Say we have a uniform random sample $s$ of $T$. The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$. Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$. In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$. That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.

We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = \text{minSS}$. If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be. Moreover, since we want good Count estimates for all rules displayed in the summary, we want $\text{minSS} >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.

Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$. This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule). Let $c$ be the column with the fewest distinct values. Say it has $|c|$ values. Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$. For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times. So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.

The highest scoring rule can have weight at most $|C|$ (the total number of columns). Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$. Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS. For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $\text{minSS} >> 5 \times 10$.

### 2014-11-14 TableSummarization/Paper/exp.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Experiments}\label{sec:experiments}
\subsection{Dataset}
The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

The columns (in order) are as follows:
(Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, How long have you lived in the San Fran./Oakland/San Jose Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home).

Unless otherwise specified, in all our experiments, we restrict the table to the first $7$ columns in order to make the result tables fit into the page. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.

\subsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

To begin with, the user sees an empty rule with the total number of tuples as the count. After this, suppose the user expands the rule. Then the user will see Table~\ref{table:uiexample1}. The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples. The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Are for $> 10$ years. Note that the latter two rules give very specific information which would require upto $3$ user clicks to find using Data Cube Drill-Downs, whereas our system displays that information to the user with a single click.

Now suppose the user decides to further explore the table, by looking at education related information of females in the dataset. So the user clicks on the $\star$ in the 'Education' column of the second rule. This opens up table~\ref{table:uiexamplestar}. It shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.

Instead of expanding the 'Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
\hline 
\end{tabular} 
\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
\end{table*}

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
\hline 
\end{tabular} 
\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
\end{table*}

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
\hline 
\end{tabular} 
\caption{A Rule expansion \label{table:uiexamplerule}} 
\end{table*}

\subsection{Weighting functions}
Our system can display optimal rule summaries using any monotone weighting function. By default, we assign a rule weight equal to it's size. In this section, we try a couple of other weighting functions.

We first try the weighting function given by:
$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct possible values in column $c$. This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values. The rule summary for this weighting is in table~\ref{table:weigtingbitwise} (contrast with Table~\ref{table:uiexample1}). The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column. Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.

The other weighting function we try is given by:
$$W(r) = \text{Min}(0, \text{Size}(r) - 1)$$
This gives us table~\ref{table:weightingsizeminusone}. This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values. As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Table~\ref{table:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.

%TODO: Give application examples for different weighting functions
%TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
\hline 
\end{tabular} 
\caption{Bits scoring\label{table:weigtingbitwise}} 
\end{table*}

\begin{table*} 
\centering 
\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
\hline 
\end{tabular} 
\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
\end{table*}

\subsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$. In this section, we study the effect of varying $m_w$ on the speed of our algorithm. Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

We consider a weighting function $W$, and a value of $m_w$. For that value of the $W$ and $m_w$ parameters, we find the time taken for expanding the empty rule. We repeat this procedure $10$ times and take the average value of the running times across the $10$ iterations. This time is plotted against $m_w$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:mw_speed}. The figure shows that running time seems to be approximately linear in $m_w$.

The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule. Thus for size scoring, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$. For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$. At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively. Of course, the minimum value of $m_w$ we can use is not known to us beforehand. But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
\vspace{-13pt}
\end{figure}

\subsection{Effects of $minSS$}
We now study the effects of the sampling parameter $minSS$. This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.   Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.

We consider one value of $minSS$ and one weight function $W$ at a time. For those values of $minSS$ and $W$, we drill down on the empty rule and measure the time taken for this operation. We also measure the percent error in the estimated counts of the displayed rules. That is, for each displayed rule $r$, if the displayed (estimated) count if $c_1$ and the actual count (computed separately on the entire table) is $c_2$, then the percent error for rule $r$ is $\frac{100 \times |c_1-c_2|}{c_2}$. We consider the average of percent errors over all displayed rules. For each value of $minSS$ and $W$, we drill down on the empty rule and find the computation time and percent error $50$ times, and take the average value for time and error over those $50$ iterations. This average time is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}. The average percent error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error_percent}.

The computation time increases approximately linearly with $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, which takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$.

In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}. This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
\vspace{-13pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
\vspace{-13pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
\caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
\vspace{-13pt}
\end{figure}

### 2014-11-14 TableSummarization/Paper/extensions.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Extensions}\label{sec:extensions}
\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
Our algorithm assumes that all attributes are categorical in nature. Attributes that have a large domain d to have a smaller tuple count per value, and hence don't appear in rule summaries. Thus our algorithm does not summarise information about numerical attributes.

However, we can modify the algorithm to deal with numerical attributes. Suppose we have a numerical attribute $A$. A simple approach is to create buckets for values of $A$. We choose a number of buckets $k$, and divide the range of values of $A$ into $k$ intervals, each corresponding to a bucket. We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket. Then we can use our algorithm, treating the bucket number as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).

The distribution of values for the $A$ may not be similar in different parts of the table. For instance, we may create buckets have approximately equal numbers of tuples in the table. But then suppose a user tries to expand a rule $r$. The part of the table covered by $r$ might have a very different distribution of age values. For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment. To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.

\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.

### 2014-11-14 TableSummarization/Paper/formal.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Formal Description}

\subsection{Preliminaries and Definitions}
\label{sec:preliminaries}

\stitle{Tables and Rules:} As in a traditional OLAP setting, we assume we are given a 
star or snowflake schema;
for simplicity, we represent this schema using a single denormalized relational table,
which we call $\calD$. 
For the purpose of the rest of the discussion, we will operate on this
table $\calD$.
We let $T$ denote the set of tuples in $\calD$, and $C$ denote 
the set of columns in $\calD$.

Our objective (formally defined later) is to 
enable smart drill-downs on this table or on portions of it:
the result of our drill-downs are lists of {\em rules}. 
A {\em rule} is a tuple with a value for each column of the table. 
In addition, a rule has other attributes, such as count and weight 
(which we define later) associated with it. 
The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. The {\em trivial rule} is one that has a $\star$ value in all columns.

\stitle{Coverage:} A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$. At a high level, we are interested
in identifying rules that cover many tuples. We next define the concept of subsumption that allow us to 
relate the coverage of different rules to each other.

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e. there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system as a result of a smart drill-down by the user. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

\agp{seems useless to describe here.}  
\sout{Let $R$ denote a rule-list. 
For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.}

\stitle{Scoring:} We now define some additional properties of rules; these properties
help us ``score'' individual rules as part of a rule-list.

There are two portions that constitute our scores for a rule as part of a rule list. 
The first portion dictates how much the rule $r$ ``covers'' the tuples in $\calD$;
the second portion dictates how ``good'' the rule $r$ is (independent of how many
tuples it covers). 
The reason why we separate the scoring into these two portions is
that they allow us to separate the inherent goodness of a rule from
how much it captures of the data in $\calD$.

We now describe the first portion:
we define {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user
as part of the smart drill-down result, to increase the coverage of the rule-list.

\agp{not sure you are using size at this point. not essential to describe now:} 
\sout{{\em Size}($r$) refers to the number of non-starred values in rule $r$.}

Now, onto the second portion: we let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could, as we will see later, 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$.
We further describe our weighting functions in Section~\ref{sec:weighting}.

Thus, the total score for our list of rules is given by 
$$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
The goal, therefore, is to choose the rule-list that maximizes the 
total score.

We use $MCount$ rather than Count in the above equation to ensure that we don't redundantly cover
the same tuples multiple times using multiple rules, and thereby increase coverage of the table. 
If we had defined total score as $\sum_{r \in R} \text{Count}(r)W(r)$, then our optimal rule-list could 
contain rules that repeatedly refer to the most `summarizable' part of the table. 
For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then 
for some weighting functions $W$, 
the summary may potentially consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.

\stitle{Formal Problem:} We now formally define our problem:

\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r)$$
for one of the following smart drill-down operations:
\begin{enumerate}
\item $[$Rule Drill-Down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{enumerate}
\end{problem}
Throughout this paper, we use the {\em Count} aggregate of a rule when determining what rules to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in Section~\ref{sec:extensions-sum}.

% Describe architecture here? (With a diagram). Show SampleHandler, QueryEngine, etc.

\subsection{Weighting Rules}
\label{sec:weighting}
We now describe our weighting function $W$ that is used to score individual rules. 
At a high level, we want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is. Let us consider some forms for the function $W(r)$ and their interpretation:

\stitle{Weighting Function 1:} 
$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those. Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.

\stitle{Weighting Function 2:}
$W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column. The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes log($|c|$) bits of information. A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table. Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.

\smallskip

\stitle{Other Weighting Functions:}
Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
\begin{enumerate}
\item Non-negativity : For all rules $r$, $W(r) \geq 0$ 
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\end{enumerate}

\begin{table}
\centering
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\end{table}

\subsection{Our Smart Drill-Down Interface}
\label{sec:interface}
We now explain how the user can use smart drill-downs in our system to explore the contents of a table.

When the user starts using our system, he sees a table with a single trivial rule as shown in table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill-down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}

### 2014-11-14 TableSummarization/Paper/intro.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Introduction}

Drill down (and roll up) [citation] are very useful tools for exploring
relational tables.
In this paper we study an extension of drill down
that lets users zoom into the more ``interesting'' parts of a table,
with fewer operations and without having to examine as much data
as traditional drill down.
We call our approach {\em smart drill down};
it does not replace traditional drill down but we will argue is a very
useful complement.
The best way to explain smart drill down is through a simple example.

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by another aggregate,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the system displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
he obtains Table~\ref{table:introexample}.
The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
that display noteworthy or interesting drill downs.
The number $3$ is a user specified parameter, which we call $k$.

For example, the tuple (Target, bicycles, $\star$, $200$, $2$)
says that there are $200$ tuples (out of the 6000) with
Target as the first column value and bicycle as the second.
This fact tells the analyst that Target is selling a lot of bicycles.
The next tuple tells the analyst that comforters are selling well in
the MA-3 region, across multiple stores. The last tuple
states that Walmart is doing well in general over multiple products and regions.
We call each tuple in Table~\ref{table:introexample} a {\em rule}
to distinguish it from the tuples in the original table that is being explored.
Each rule summarizes the set of tuples that are described by it.
Again, instead of Count, the system can display some other aggregate, such as
the total Sales.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Initial summary}\label{table:introexample0}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \hline
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Result after first smart drill down}\label{table:introexample}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & CA-1 & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & WA-5 & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Result after second smart drill down} \label{table:introexample2}
\end{table}

Say that after seeing the results of Table~\ref{table:introexample},
the analyst wishes to dig deeper into the Walmart tuples
represented by the last rule.
For instance, the analyst may want to know
which states Walmart has more sales in, or which products they sell
the most. In this case, the analyst clicks on the Walmart rule,
obtaining the expanded summary in Table~\ref{table:introexample2}.
The three new rules in this table provide additional information
about the $1000$ Walmart tuples.
In particular, one of the new rules shows that
Walmart sells a lot of cookies and the others show it sells a lot of products in
the regions CA-1 and WA-5.

When the analyst clicks on a rule $r$, smart drill down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former rule tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.

In this paper we describe how to combine or blend these three factors
in order to obtain a single desirability score for a set of rules.
Our score function can actually be tuned by the analyst
(by specifying how weights are computed),
providing significant flexibility in what is considered a good set of rules.
We also present an efficient optimization procedure to maximize score, invoked
by smart drill down to select the set of $k$ rules to display.

\end{example}

Compared to traditional drill down, our smart drill down has two important advantages:
\begin{itemize}
\item
Smart drill down limits the information displayed
to the most interesting $k$ facts (rules).
With traditional drill down, a column is expanded and {\em all}
attribute values are displayed in arbitrary order.
In our example, if we drill down on say the store attribute,
we would see all stores listed, which may be a very large number.
\item
Smart drill down explores several attributes to open up together,
and automatically selects combinations that are interesting.
For example, in Table~\ref{table:introexample},
the rule (Target, bicycles, $\star$, $200$, $2$)
is obtained after a single drill down;
with a traditional approach, the analyst would first have to drill down on
Store, examine the results, drill down on Product,
look through all the displayed rules and then find the interesting rule
(Target, bicycles, $\star$, $200$, $2$).
\end{itemize}

Incidentally, note that in the example we only described one type of smart drill down,
where the analyst selects a {\em rule} to drill down on
(e.g., the Walmart rule going from Table~\ref{table:introexample}
Table~\ref{table:introexample2}.
In Section~\ref{sec:interface} we describe another option
where the analyst clicks an a {\em *} to obtain
rules that have values in that location.

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence of the
resulting maximum entropy distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill down to provide an additional option of anomaly detection.

We further discuss related work in Section~\ref{sec:related}.

\stitle{Overview of paper:} 
\squishlist

\item In Section~\ref{sec:preliminaries}, we introduce the relevant
notation and formally define smart drill down.

\item In Section~\ref{sec:weighting}, we consider different schemes
for weighting rules, along with their practical significance.

\item In Section~\ref{sec:interface}, we describe our
interactive user interface.

\item In Section~\ref{sec:algorithms}, we present our algorithms for
finding optimal sets of rules, as well as our sampling schemes
for dealing with large tables.

\item Based on our implemented smart drill down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on a real dataset,
and show additional examples of smart drill down in action.

\item In Section~\ref{sec:extensions}, we describe how to extend our results to summarize numerical attributes, and to use the $Sum$ aggregate instead of $Count$.

\item In Section~\ref{sec:related}, we describe related work.
\squishend

### 2014-11-14 TableSummarization/Paper/related.tex
split, first pass on formal

%!TEX root = TableSummarization.tex

\section{Related Work}\label{sec:related}
%TODO: Repeat what we said in intro about interactive, etc.
There has been work on finding interesting rules in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}. This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max entropy distribution. The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.

There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}. This work again focuses on displaying rules that will cause the max entropy distribution to best approximate the actual distribution of values.

Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies. There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables. All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.

Our algorithm uses ideas from the a priori algorithm~\cite{apriori}. Several extensions for the algorithm have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311}. We can potentially use these ideas to improve handing of numerical attributes in our work.

% Related Work summary : Is ours the only only that is flexible (due to weight function black box)?
% Explanation tables takes lots of papers and talks about them in one go in related work. saying they optimize for different opbjective functions. We can say that as well. We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity. 
% Succintly summarizing with itemsets-- Mampaey:2011:TMI:2020408.2020499 -- Also seemed to optimize according to max entropy criterion
% Efficient and Effective... Tableaux -- Golab_efficientand -- Similar to conditional functional dep. paper.
% Near Optimal Tableaux Conditional functional dep. -- Golab:2008:GNT:1453856.1453900 -- Finding a smallest sized tableau that provides some confidence and support for a conditional functional dependency.
% MDL Summarization with holes -- Bu:2005:MSH:1083592.1083644 -- Min sized summary with a rectangle and exceptions (holes). 
% Generalized MDL Approach -- Lakshmanan:2002:GMA:1287369.1287435 -- Similar to MDL Approach.
% Overlapped Hyperrectangle -- Xiang_succinctsummarization -- Efficiently find the set of hyperrectangles that covers the data (sets of itemsets)
% Tiling Databases Geerts04tilingdatabases Simlar to hyper-rectangles. Probabilistic interpretation of frequent itemsets.

### 2014-11-15 TableSummarization/Paper/algorithms.tex
Many small changes

@@
 Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
-$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
+$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r,R)$$
 \end{problem}
@@
 Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
-$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
+$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r,R)$$
 \end{problem}
@@
 For a given table $T$, the Score function over sets $S$ of rules, defined by :
-$$\text{Score}(S) = \sum_{r \in S} MCount(r)W(r)$$
+$$\text{Score}(S) = \sum_{r \in S} MCount(r,S)W(r)$$
 is submodular.

### 2014-11-15 TableSummarization/Paper/formal.tex
Many small changes

@@
 The {\em trivial rule} is one that has a $\star$ value in all columns.
+The {\em Size} of a rule is defined as the number of non-starred values in that rule.
 
@@
 
-\agp{seems useless to describe here.}  
-\sout{Let $R$ denote a rule-list. 
-For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
-Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.}
-
 \stitle{Scoring:} We now define some additional properties of rules; these properties
@@
 as part of the smart drill-down result, to increase the coverage of the rule-list.
-
-\agp{not sure you are using size at this point. not essential to describe now:} 
-\sout{{\em Size}($r$) refers to the number of non-starred values in rule $r$.}
 
@@
 Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
-$$\sum_{r \in R} W(r) \times MCount(r)$$
+$$\sum_{r \in R} W(r) \times MCount(r,R)$$
 for one of the following smart drill-down operations:
@@
 \stitle{Weighting Function 2:}
-$W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$.
+$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
 Like the previous scoring function, this one adds some weight for every non-starred value in a rule.
@@
 But this time, we count the number of `bits' that are pre-filled.
-For a column $c$, specifying a value in the column takes log($|c|$) bits of information.
-A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table.
+For a column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information.
+A non-starred valued in a rule $r$ thus pre-fills $MCount(r,R) \lceil \text{log}(|c|) \rceil$ bits of the table.
 Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.
@@
 Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
-If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.
+If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.
 
@@
 \end{enumerate}
+
+A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).
 

### 2014-11-15 TableSummarization/Paper/intro.tex
Many small changes

@@
 
-Drill down (and roll up) [citation] are very useful tools for exploring
+Drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
 relational tables.
@@
 where the analyst selects a {\em rule} to drill down on
-(e.g., the Walmart rule going from Table~\ref{table:introexample}
+(e.g., the Walmart rule going from Table~\ref{table:introexample} to
 Table~\ref{table:introexample2}.
 In Section~\ref{sec:interface} we describe another option
-where the analyst clicks an a {\em *} to obtain
+where the analyst clicks an a $\star$ to obtain
 rules that have values in that location.
@@
 giving the most ``surprising'' information to the user, i.e., information
-that would minimize the Kullback-Liebler(KL) divergence of the
-resulting maximum entropy distribution. For instance, if a certain set of
+that would minimize the Kullback-Liebler(KL) divergence between the
+resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
 values occur together in an unexpectedly small number of tuples, that

### 2014-11-15 TableSummarization/Paper/related.tex
Many small changes

@@
 There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
+Several existing works also deal with the problem of frequent itemset mining~\cite{apriori, 1411744, Han:2000:MFP:342009.335372}.
 All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
@@
 Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
-Several extensions for the algorithm have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311}.
+Several extensions have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311, Miller:1997:ARO:253260.253361}.
 We can potentially use these ideas to improve handing of numerical attributes in our work.

### 2014-11-15 TableSummarization/Paper/intro.tex
small change to intro

@@
 \section{Introduction}
+Analysts often use OLAP (Online Analytical Processing) operations
+such as drill-down (and roll-up)~\cite{export:69578} to explore
+relational databases. 
+These operations are very useful for analytics and data exploration and have stood the test of time;
+all commercial database systems in existence support these operations.
 
-Drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
-relational tables.
-In this paper we study an extension of drill down
-that lets users zoom into the more ``interesting'' parts of a table,
-with fewer operations and without having to examine as much data
-as traditional drill down.
-We call our approach {\em smart drill down};
-it does not replace traditional drill down but we will argue is a very
-useful complement.
+However, there are cases where drill-down is ineffective; 
+for example, when the number of distinct values
+in a column is large, vanilla drill-down 
+could easily overwhelm analysts by presenting them with too many 
+results (i.e., aggregates). 
+Further, drill-down only allows us to instantiate values
+in one column at a time, instead of allowing simultaneous drill-downs
+on multiple columns---simultaneous drill-down on multiple columns 
+could once again suffer from the problem
+of having too many results, stemming from distinct combinations of column values.
+
+In this paper, we present a new interaction operator 
+that is an extension to a traditional 
+drill-down operator, aimed at providing {\em complementary}
+functionality to drill-down in cases where drill-down is
+ineffective. We call our operator {\em smart drill-down}.
+At a high level, smart drill-down lets analysts zoom into
+the more ``interesting'' parts of a table or a database,
+with fewer operations, and without having to examine as much
+data as traditional drill-down.
+Note that our goal is to {\em not} replace traditional 
+drill-down functionality, which we believe is essential;
+instead, our goal is to provide auxiliary functionality 
+which the analyst is free to use whenever they find 
+traditional drill-downs ineffective.
+
+In addition to presenting this new operator called smart drill-down, in this paper, 
+we also present techniques to compute the results for this
+operator {\em in an interactive fashion} on increasingly larger databases. 
+Unlike the traditional OLAP setting, these computations 
+require no pre-materialization, and can be implemented 
+within any relational database system with no additional storage.
+
+% Drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
+% relational tables.
+% In this paper we study an extension of drill down
+% that lets users zoom into the more ``interesting'' parts of a table,
+% with fewer operations and without having to examine as much data
+% as traditional drill down.
+% We call our approach {\em smart drill down};
+% it does not replace traditional drill down but we will argue is a very
+% useful complement.
 The best way to explain smart drill down is through a simple example.

### 2014-11-15 TableSummarization/Paper/intro.tex
changes

@@
 could once again suffer from the problem
-of having too many results, stemming from distinct combinations of column values.
+of having too many results, stemming from many distinct combinations of column values.
 
@@
 data as traditional drill-down.
-Note that our goal is to {\em not} replace traditional 
-drill-down functionality, which we believe is essential;
+Note that our goal is {\em not} to replace traditional 
+drill-down functionality, which we believe is fundamental;
 instead, our goal is to provide auxiliary functionality 

### 2014-11-15 TableSummarization/Paper/intro.tex
changes to intro to beef up

@@
 These operations are very useful for analytics and data exploration and have stood the test of time;
-all commercial database systems in existence support these operations.
+all commercial OLAP systems 
+in existence support these operations. (Recent reports estimate the size of the OLAP market to be \$10+ Billion~\cite{gartner}.)
 
@@
 Further, drill-down only allows us to instantiate values
-in one column at a time, instead of allowing simultaneous drill-downs
-on multiple columns---simultaneous drill-down on multiple columns 
+ one column at a time, instead of allowing simultaneous drill-downs
+on multiple columns---this simultaneous drill-down on multiple columns 
 could once again suffer from the problem

### 2014-11-15 TableSummarization/Paper/formal.tex
small changes

@@
 \section{Formal Description} 
-
+We describe our formal problem
+in Section~\ref{sec:preliminaries},
+describe different scoring functions
+in Section~\ref{sec:weighting},
+and describe our operator interfaces
+in Section~\ref{sec:interface}. 
 \subsection{Preliminaries and Definitions}
@@
 as well as the number of distinct values in each column of $\calD$.
-A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$.
+A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
+on monotonic weighting functions because we prefer 
+rules that are more ``specific''
+rather than those that are more ``general'' 
+(thereby conveying less information).
 We further describe our weighting functions in Section~\ref{sec:weighting}.
@@
 $$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
-The goal, therefore, is to choose the rule-list that maximizes the 
+Overall, our goal is to choose the rule-list that maximizes the 
 total score.
 
-We use $MCount$ rather than Count in the above equation to ensure that we don't redundantly cover
+We use $MCount$ rather than Count in the above equation to ensure that we do not redundantly cover
 the same tuples multiple times using multiple rules, and thereby increase coverage of the table. 
@@
 \stitle{Formal Problem:} We now formally define our problem:
-
 \begin{problem}\label{prob:optimal-subrule-list}
@@
 given the rules, it should be as easy as possible to reproduce the table.
-We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is.
-Let us consider some forms for the function $W(r)$ and their interpretation:
+We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is (i.e., how much information
+it conveys). Let us consider some canonical forms for the function $W(r)$ and their interpretation; later, we describe the full
+space of weighting functions our techniques can handle:
 
@@
 But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column.
-The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
+The reason behind this variation is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
 Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$.
@@
 \begin{enumerate}
-\item Non-negativity : For all rules $r$, $W(r) \geq 0$ 
+\item Non-negativity: For all rules $r$, $W(r) \geq 0$.
 \item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
 \end{enumerate}
-
 A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).
@@
 
-\subsection{Our Smart Drill-Down Interface}
+\subsection{Example: Smart Drill-Down Operations}
 \label{sec:interface}
-We now explain how the user can use smart drill-downs in our system to explore the contents of a table.
+We now explain how a user can use smart drill-downs to explore the contents of a table.
 
-When the user starts using our system, he sees a table with a single trivial rule as shown in table~\ref{table:introexample0}.
+When the user starts using a system equipped with
+the smart drill-down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}.
 At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill-down' on the rule.
@@
 This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
+For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
 This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
 
-Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation i.e.
+Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e.
 collapses it.
 For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.
+This operation is equivalent to a traditional
+roll-up, but for smart drill-downs instead of traditional drill-downs.
 

### 2014-11-15 TableSummarization/Paper/algorithms.tex
small changes
small suggestions

@@
 
-\section{Algorithms}
+\section{Smart Drill-Down Algorithms}
 \label{sec:algorithms}
+We now describe online algorithms for implementing
+the smart drill-down operator.
+
+\subsection{Problem Reduction and Important Property} \label{sec:reduction}
 When the user drills down on a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into.
@@
 
-$W$ is required to be non-negative and monotonic.
-$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.
+\agp{not essential to mention here.}
+\sout{$W$ is required to be non-negative and monotonic.
+$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.}
 
-Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
+\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
 \squishlist
-\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
+\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
 Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
-\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$.
+\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$.
 In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.
@@
 
-We will describe a greedy algorithm for finding the approximately highest scoring rule-list.
+\agp{not essential to mention here}\sout{We will describe a greedy algorithm for finding the approximately highest scoring rule-list.}
+
 First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
+\agp{mention that this is independent of MCount. Also give intuition--this is because
+we can transfer tuples from one rule to a better rule adn only increase score.}
 
@@
 
-Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
+\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
 Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order.
@@
 
-The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.
+\agp{Confusing since you're conflating lists and sets. Explicitly define the score of a set of rules
+is the score of the best rule list (and in the equation above, have max over all R' where R' is an ordered
+version of S)}
 
-\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
+\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.
+
+\subsection{Algorithm Overview}\label{sec:alg-overview}
+\agp{Give a high-level perspective of the algorithm; 
+Basics: greedy approximation + a-priori; 
+Advanced: Sampling for larger datasets: comprising two parts
+how to best use samples, and how to select samples}
+
+\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
+\agp{can we show hardness in data or schema complexity?}
 We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
@@
 \begin{framed}
+\vspace{-15pt}
 \begin{enumerate}
@@
 \end{enumerate}
+\vspace{-15pt}
 \end{framed}
@@
 \subsection{Sampling for big tables}
+
+\agp{Here also, there is a need for some organization. There are clearly
+some good ideas here, but you need to give a high-level picture first before
+diving in.}
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
@@
 
+\agp{not clear what problem you're trying to solve here. consider raising it up as a subsection
+and giving it a name. }
 Thus formally our problem is :

### 2014-11-15 TableSummarization/Paper/exp.tex
small suggestions

@@
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
+
+\agp{I propose using two sections: one is Qualitative Study, which will contain 4.2, 4.3, and 
+Quantitiative Study, whcih will contain the rest of the sections.
+For the quantitative study, mention that the goal of the experimental evaluation is to study
+the impact of the two parameters on time and accuracy before you get into details.
+Maybe also good to fix mw, and show the performance with or without sampling,
+and with or without a-priori (one rule per phase, for instance).
+}
 

### 2014-11-15 TableSummarization/Paper/exp.tex
small

--- 2014-11-14 TableSummarization/Paper/exp.tex
@@
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
+
+\agp{I propose using two sections: one is Qualitative Study, which will contain 4.2, 4.3, and 
+Quantitiative Study, whcih will contain the rest of the sections.
+For the quantitative study, mention that the goal of the experimental evaluation is to study
+the impact of the two parameters on time and accuracy before you get into details.}
+
+\agp{Additional experiments:
+Maybe also good to fix mw, and show the performance with or without sampling,
+and with or without a-priori (one rule per phase, for instance).
+Also, would be good to see performance on increasing dataset size.
+}
 

### 2014-11-18 TableSummarization/Paper/algorithms.tex
Addressed Comments. Added NP-Hardness proof

@@
 \begin{problem}\label{prob:optimal-rule-list}
-Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
+Given a table $T$, a monotonic weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
 $$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r,R)$$
@@
 
-\agp{not essential to mention here.}
-\sout{$W$ is required to be non-negative and monotonic.
-$k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.}
-
 \noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
-\squishlist
+\begin{enumerate}
 \item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
@@
 Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
-\squishend
-
-\agp{not essential to mention here}\sout{We will describe a greedy algorithm for finding the approximately highest scoring rule-list.}
+\end{enumerate}
 
 First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
-\agp{mention that this is independent of MCount. Also give intuition--this is because
-we can transfer tuples from one rule to a better rule adn only increase score.}
+Note that the weight of a rule is independent of its $MCount$.
+The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score.
+Thus if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.
 
@@
 
-\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
-Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order.
-This gives us
+\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a set of rules as follows:
+
+\begin{definition}\label{def:set-score}
+Let $R$ be a set of rules. Then the Score of $R$ is defined by
+$$\text{Score}(R) = \text{Score}(R^{\prime})$$
+where $R^{\prime}$ is the list of rules obtained by ordering the rules in $R$ in decreasing order by weight.
+\end{definition}
+
+This gives us a reduced version of Problem~\ref{prob:optimal-rule-list}: 
 \begin{problem}\label{prob:optimal-rule-set}
-Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
-$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r,R)$$
+Given a table $T$, a monotonic weight function $W$, and a number $k$, find the set $R$ of $k$ rules which maximizes Score($R$) as defined in Definition~\ref{def:set-score}.
 \end{problem}
 
-\agp{Confusing since you're conflating lists and sets. Explicitly define the score of a set of rules
-is the score of the best rule list (and in the equation above, have max over all R' where R' is an ordered
-version of S)}
+\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear.
+We now first show that Problem~\ref{prob:optimal-rule-set}, and consequently Problem~\ref{prob:optimal-subrule-list} and Problem~\ref{prob:optimal-rule-list} are NP-Hard, and then present an approximation algorithm for solving Problem~\ref{prob:optimal-rule-set}.
 
-\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.
+\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
+We reduce the well known NP-Hard problem called the {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set}, which will prove NP-Hardness for Problem~\ref{prob:optimal-rule-set}.
+The Maximum Coverage Problem is given below:
+\begin{problem}\label{prob:maximum-coverage}
+Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ...
+S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximises $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
+\end{problem}
+Thus the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
+We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.
+
+Then we construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns.
+Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$.
+For each $u \in U$, the tuple $t_u$ corresponding to $u$ defined as follows: For each $S_i \in S$, $$t_u(i) = 1 \textrm{ if } u \in S_i \textrm{ else } t_u(i) = 0$$
+
+Define the weight function $W$ as follows:
+$$W(r) = \textrm{min}(1, |\left\lbrace i \in \left\lbrace 1, 2, ... m \right\rbrace : r(i) = 1 \right\rbrace|)$$
+Thus $W$ gives weight $0$ to rules that do not have value $1$ in any column, and weight $1$ to all other rules.
+An optimal rule set for weight function $W$ on table $T$ need only contain rules having a single $1$ value and no $0$ values, because adding more values can only reduce $MCount$, without increasing weight.
+
+Then a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance.
+This completes our reduction.
 
 \subsection{Algorithm Overview}\label{sec:alg-overview}
-\agp{Give a high-level perspective of the algorithm; 
-Basics: greedy approximation + a-priori; 
-Advanced: Sampling for larger datasets: comprising two parts
-how to best use samples, and how to select samples}
+We now present our algorithms for approximating the solution to Problem~\ref{prob:optimal-rule-set}.
+The problem consists of finding a set of given size $k$, that maximises Score.
+We first show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm.
+We start with an empty rule set $R$, and then run the greedy algorithm for $k$ steps, where in each step, we add the rule $r$ to $R$ that maximises Score($R \cup \left\lbrace r \right\rbrace$).
+In order to find the rule $r$ to add in each step, we need to measure $Score(R \cup \left\lbrace r \right\rbrace) - Score(R)$ for each $r$.
+This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent itemset mining.
+
+In addition, the greedy algorithm can be run over a sample of the table instead of over the entire table, to reduce computation time for large tables.
+We later describe a scheme for maintaining multiple samples in memory and using them to improve response time for different drill down operations performed by the user.
 
 \subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
-\agp{can we show hardness in data or schema complexity?}
 We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
@@
 \subsection{Sampling for big tables}
+\subsubsection{Sampling Overview}
+We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.
+First in Section~\ref{sec:sample-using}, we motivate the need for sampling.
+Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
+The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next.
+Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk.
+Finally, we mention some additional optimizations we can make, and heuristics for setting the minimum sample size required from the SampleHandler.
 
-\agp{Here also, there is a need for some organization. There are clearly
-some good ideas here, but you need to give a high-level picture first before
-diving in.}
+\subsubsection{Using samples to respond to user operations}\label{sec:sample-using}
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
@@
 Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
-Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
+Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
 The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
-We provide a way to find reasonable values of $minSS$ later.
+We later provide a way to find reasonable values of $minSS$.
 
@@
 We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$.
+Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as below:
+\begin{definition}\label{def:ess}
+$$\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
+\end{definition}
 If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
 We wish to maximise the probability that we can respond to the next user expansion without making another pass.
-
-\agp{not clear what problem you're trying to solve here. consider raising it up as a subsection
-and giving it a name. }
-Thus formally our problem is :
-Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
+We now formally define our problem below:
+\begin{problem}\label{prob:sample-sizes}
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
 $$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
@@
 $$\sum_{r \in U} n_r < \text{capacity}$$
-This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
+\end{problem}
+Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables.
+If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
 

### 2014-11-18 TableSummarization/Paper/exp.tex
Addressed Comments. Added NP-Hardness proof

--- 2014-11-14 TableSummarization/Paper/exp.tex
@@
 \section{Experiments}\label{sec:experiments}
+
+\begin{comment} ask about below:
+Additional experiments:
+Maybe also good to fix mw, and show the performance with or without sampling,
+and with or without a-priori (one rule per phase, for instance).
+Also, would be good to see performance on increasing dataset size.
+\end{comment}
+
 \subsection{Dataset}
@@
 The columns (in order) are as follows:
-(Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, How long have you lived in the San Fran./Oakland/San Jose Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home).
+Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, Time lived in Bay Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home.
 
@@
 
-\subsection{User interface testing}
+\subsection{Qualitative Study}
+We first perform a qualitative study of our Smart Drill Down system.
+We observe the effects of various user interface operations on our dataset, and then try out a few different weight functions to study their effects.
+\subsubsection{User interface testing}
 We now present the rule-based summaries displayed as a result of a few different user actions.
@@
 
-\subsection{Weighting functions}
-Our system can display optimal rule summaries using any monotone weighting function.
+\subsubsection{Weighting functions}
+Our system can display optimal rule summaries using any monotonic weighting function.
 By default, we assign a rule weight equal to it's size.
-In this section, we try a couple of other weighting functions.
+In this section, we try out a couple of other weighting functions.
 
@@
 
-\subsection{Effects of $m_w$}
+\subsection{Quantitative Study}
+The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size).
+In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm.
+
+\subsubsection{Effects of $m_w$}
 Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
@@
 
-\subsection{Effects of $minSS$}
+\subsubsection{Effects of $minSS$}
 We now study the effects of the sampling parameter $minSS$.
@@
 
-The computation time increases approximately linearly with $minSS$.
-This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, which takes time linear in the size of the sample.
-The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$.
+The computation time increases approximately linearly with parameter $minSS$.
+This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample.
+The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
 

### 2014-11-18 TableSummarization/Paper/extensions.tex
Many small changes
small suggestions
Addressed Comments. Added NP-Hardness proof

@@
 Our algorithm assumes that all attributes are categorical in nature.
-Attributes that have a large domain d to have a smaller tuple count per value, and hence don't appear in rule summaries.
+Attributes that have a large domain tend to have a smaller tuple count per value, and hence don't appear in rule summaries.
 Thus our algorithm does not summarise information about numerical attributes.
@@
 
+\begin{comment}
 The distribution of values for the $A$ may not be similar in different parts of the table.
@@
 To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.
+
+ask : we should mention this right up front when we're describing the problem.
+\end{comment}
 
@@
 {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$.
-The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r)W(r)$.
+The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$.
 Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.

### 2014-11-18 TableSummarization/Paper/formal.tex
Addressed Comments. Added NP-Hardness proof

@@
 
-\section{Formal Description} 
+\section{Formal Description}\label{sec:formal}
 We describe our formal problem
@@
 
-Now, onto the second portion: we let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
+Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
 As we will see, the weighting function does not depend on the specific
@@
 Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
-\begin{enumerate}
+\squishlist
 \item Non-negativity: For all rules $r$, $W(r) \geq 0$.
 \item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
-\end{enumerate}
+\squishend
 A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).

### 2014-11-18 TableSummarization/Paper/intro.tex
small changes
Addressed Comments. Added NP-Hardness proof

@@
 Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection.
+and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}..
 
-We further discuss related work in Section~\ref{sec:related}.
+Briefly, the contributions of our paper can be summarized are follows: We develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation.
+In addition, we develop a scheme for maintaining and using multiple samples of the table in memory to improve response time on large tables.
 
@@
 
-\item In Section~\ref{sec:preliminaries}, we introduce the relevant
-notation and formally define smart drill down.
-
-\item In Section~\ref{sec:weighting}, we consider different schemes
-for weighting rules, along with their practical significance.
-
-\item In Section~\ref{sec:interface}, we describe our
-interactive user interface.
+\item In Section~\ref{sec:formal}, we introduce the relevant notation and formally define smart drill down.
+After that, we discuss different schemes for weighting rules, and describe our interactive user interface.
 

### 2014-11-21 TableSummarization/Paper/exp.tex
New graphs. Yet to be added to writeup

@@
 \end{figure}
+
+\begin{comment}
+more experiment stuff
+vanilla a priori size scoring, first 7 columns (maybe we should include 13, to show suffering on english rule
+Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+$\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $1$ \\
+Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
+$\star$ & Married & $\star$ & $\star$ & $\star$ & > 10 years & $2053$ & $2$ \\
+Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+Female & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1966$ & $2$ \\
+$\star$ & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $1964$ & $2$ \\
+
+\end{comment}

### 2014-11-21 TableSummarization/src/dataextraction/USCensus1990.java

@@ -1,0 +1,127 @@
+package dataextraction;
+
+import static java.lang.System.out;
+
+import java.io.BufferedReader;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+
+import solvers.Rule;
+
+public class USCensus1990 {
+	final static String DATAFILELOCATION = "TestDatasets/USCensus1990/USCensus1990.data.txt";
+
+	public static void addNames (TableInfo table) {
+		List<Map<String, String>> names = new ArrayList<Map<String, String>>();
+		
+		for (List<String> col : table.dictionary) {
+			Map<String, String> colNames = new HashMap<String, String>();
+			colNames.put("column", "C");
+			for (String name : col) {
+				colNames.put(name, name);
+			}
+			names.add(colNames);
+		}
+		table.names = names;
+	}
+	
+	public static TableInfo parseData(Double sampleProb) throws IOException {
+		BufferedReader br = new BufferedReader(new FileReader(DATAFILELOCATION));
+		List<List<String>> dictionary = new ArrayList<List<String>>();
+		List<Map<String, Integer>> reverseDictionary = new ArrayList<Map<String, Integer>>();
+		List<List<Integer>> contents = new ArrayList<List<Integer>>();
+		
+		String line = br.readLine();
+		{
+			String[] vals = line.split(",");
+			for (int i = 0; i < vals.length; i++) {
+				dictionary.add(new ArrayList<String>());
+				reverseDictionary.add(new HashMap<String, Integer>());
+			}
+		}
+		
+		while ((line = br.readLine()) != null) {
+			if (Math.random() > sampleProb) {
+				continue;
+			}
+			String[] vals = line.split(",");
+			List<Integer> tuple = new ArrayList<Integer>(vals.length);
+			for (int i = 0; i < vals.length; i++) {
+				final String value = vals[i];
+				/*if (value.equals("03")) {
+					out.println(line);
+					System.exit(1);
+				}*/
+				Map<String, Integer> columnDictionary = reverseDictionary.get(i);
+				if (columnDictionary.containsKey(value)) {
+					tuple.add(columnDictionary.get(value));
+				} else {
+					columnDictionary.put(value, columnDictionary.keySet().size());
[…]

### 2014-11-21 TableSummarization/src/solvers/NonStarCountSolvers.java

@@ -900,6 +900,97 @@ public class NonStarCountSolvers {
 	
 	/**
 	 * Finds rule that adds the most marginal value, given the chosen solution rules. 
+	 * This function starts afresh, not accessing the old values of ruleMap or superRules.
+	 */
+	public static List<Rule> getTopKRules (TableInfo table, Integer maxRuleScore, Scorer scorer, int numRules) throws IOException {
+		List<Rule> solution = new ArrayList<Rule>();
+		final Integer length = table.dictionary.size();
+		List<List<Rule>> singleRules = getSingleRulesWithMarginalValues (table, new HashSet<Rule>(), scorer);
+		Map<Rule, Set<Rule>> superRules = new HashMap<Rule, Set<Rule>>();
+		Map<Rule, Rule> ruleMap = new HashMap<Rule, Rule>();
+		Map<List<Integer>, Rule> ruleListMap = new HashMap<List<Integer>, Rule>();
+		Integer valueThreshold = -1;
+		List<Rule> candidateRules = new ArrayList<Rule>();
+		
+		for (int col = 0; col < singleRules.size(); col++) {
+			final List<Rule> colRules = singleRules.get(col);
+			for (int val = 0; val < colRules.size(); val++) {
+				Rule rule = colRules.get(val);
+				ruleMap.put(rule, rule);
+				ruleListMap.put(rule.values, rule);
+				superRules.put(rule, new HashSet<Rule>());
+				candidateRules.add(rule);
+			}
+		}			
+		Collections.sort(candidateRules);
+		Collections.reverse(candidateRules);
+		valueThreshold = candidateRules.get(numRules).maxMarginalValue;
+				
+		Set<Rule> latestSuperRules = ruleMap.keySet();
+		
+		for (int currentSize = 2; currentSize < 7; currentSize++) {
+			//out.println(bestMarginalRuleValue);
+			Set<Rule> nextSuperRules = new HashSet<Rule>();
+			for (Rule r : latestSuperRules) {
+				Rule rule = ruleMap.get(r);
+				if (rule.maxMarginalValue + rule.count * (maxRuleScore - rule.score) >= valueThreshold) {
+					nextSuperRules.addAll(rule.findSuperRules(table, currentSize));					
+				}
+			}
+			//out.println("Generated " + nextSuperRules.size() + " candidate rules of size " + currentSize);
+			Iterator<Rule> ruleIter = nextSuperRules.iterator();
+			while (ruleIter.hasNext()) {
[…]

### 2014-11-21 TableSummarization/src/solvers/RuleTreeDisplay.java

@@ -14,6 +14,7 @@ import java.util.Stack;
 import dataextraction.Marketing;
 import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
+import dataextraction.USCensus1990;
 import solvers.Rule.sizeBitsScorer;
 import solvers.Rule.sizeMinusKScorer;
 import solvers.Rule.sizeScorer;
@@ -189,28 +190,41 @@ public class RuleTreeDisplay {
 	}
 	
 	public static void main (String[] args) throws IOException {
-		final String mw_bits_outFile = "Data_Graphs/mw_speed_bits";
-		final String mw_size_outFile = "Data_Graphs/mw_speed_size";
-		final String minSS_size_outFile = "Data_Graphs/minSS_speed_size";
-		final String minSS_bits_outFile = "Data_Graphs/minSS_speed_bits";
-		final String minSS_size_error_outFile = "Data_Graphs/minSS_error_size";
-		final String minSS_bits_error_outFile = "Data_Graphs/minSS_error_bits";
 		List<Integer> columns = new ArrayList<Integer>();
 		final Integer firstNumColumns = 7;
 		for (int i = 1; i < firstNumColumns; i++) {
 			columns.add(i);
 		}
-		TableInfo fullTable = Marketing.parseData();
+		//TableInfo fullTable = Marketing.parseData();
+		TableInfo fullTable = USCensus1990.parseData(0.01);
 		TableInfo table = fullTable.getSubTable(columns);
 		Integer ruleNums = 4;
+		final String mw_bits_outFile = "Data_Graphs/mw_speed_bits";
+		final String mw_size_outFile = "Data_Graphs/mw_speed_size";
+		final String mw_bits_USCensus_outFile = "Data_Graphs/mw_speed_USCensus_bits";
+		final String mw_size_USCensus_outFile = "Data_Graphs/mw_speed_USCensus_size";
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_outFile);
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeScorer(), mw_size_outFile);
+		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_USCensus_outFile);
[…]

### 2014-11-21 TableSummarization/src/utils/Experiments.java

@@ -5,6 +5,7 @@ import java.io.IOException;
 import java.io.PrintWriter;
 import java.util.HashMap;
 import java.util.HashSet;
+import java.util.List;
 import java.util.Map;
 import java.util.Set;
 
@@ -15,6 +16,7 @@ import solvers.RuleTreeDisplay;
 import solvers.Scorer;
 import dataextraction.SampleHandler;
 import dataextraction.TableInfo;
+import dataextraction.USCensus1990;
 
 public class Experiments {
 	public static void mwSpeedTest (TableInfo table, int mwMin, int mwMax, int ruleNums, Scorer scorer, String outFile) throws IOException {
@@ -70,6 +72,60 @@ public class Experiments {
 			ruleTree.expandRow(emptyRule, 2 * ruleNums, mw, scorer, sampleHandler);
 			for (RuleNode child : ruleTree.root.children) {
 				ruleCounts.put(child.rule, child.rule.count);
+			}
+			ruleTree.contractRow(emptyRule);
+			ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
+			for (RuleNode child : ruleTree.root.children) {
+				selectedRules.add(child.rule);
+			}
+		}
+		for (int minSS = minSSMin; minSS < minSSMax; minSS += minSSStep) {
+			SampleHandler sampleHandler = new SampleHandler(table, capacity, minSS);
+			long countError = 0;
+			Double percentCountError = 0.0;
+			Double ruleError = 0.0;
+			for (int iter = 0; iter < numIters; iter++ ) {
+				RuleTree ruleTree = new RuleTree(table);
+				ruleTree.expandRow(emptyRule, ruleNums, mw, scorer, sampleHandler);
+				for (RuleNode node : ruleTree.root.children) {
+					if (ruleCounts.containsKey(node.rule)) {
+						countError += Math.abs(node.rule.count - ruleCounts.get(node.rule));
+						percentCountError += Math.abs((node.rule.count - ruleCounts.get(node.rule)) * 100.0 /ruleCounts.get(node.rule));
+					}
+					if (!selectedRules.contains(node.rule)) {
+						ruleError++;
+					}
+				}
+				percentCountError /= ruleTree.root.children.size();
+			}
+			pw.println(minSS + "\t" + countError/numIters + "\t" + percentCountError/numIters + "\t" + (0.0 + ruleError)/numIters);
[…]

### 2014-11-21 TableSummarization/Paper/algorithms.tex
minor changes
changes to greedy sections

@@
 the smart drill-down operator.
+We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
+We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}.
 
@@
 \subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
-We reduce the well known NP-Hard problem called the {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set}, which will prove NP-Hardness for Problem~\ref{prob:optimal-rule-set}.
-The Maximum Coverage Problem is given below:
+We reduce the well known NP-Hard {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
+thus demonstrating the NP-Hardness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
 \begin{problem}\label{prob:maximum-coverage}
@@
 \end{problem}
-Thus the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
+Thus, the goal of the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
 We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.
 
-Then we construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns.
+\begin{proof}
+We construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns.
 Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$.
@@
 
-Then a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance.
+Then, a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance.
 This completes our reduction.
+\end{proof}
 
 \subsection{Algorithm Overview}\label{sec:alg-overview}
-We now present our algorithms for approximating the solution to Problem~\ref{prob:optimal-rule-set}.
-The problem consists of finding a set of given size $k$, that maximises Score.
-We first show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm.
-We start with an empty rule set $R$, and then run the greedy algorithm for $k$ steps, where in each step, we add the rule $r$ to $R$ that maximises Score($R \cup \left\lbrace r \right\rbrace$).
-In order to find the rule $r$ to add in each step, we need to measure $Score(R \cup \left\lbrace r \right\rbrace) - Score(R)$ for each $r$.
-This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent itemset mining.
+Given that the problem is NP-Hard, we now present our algorithms 
+for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
+The problem consists of finding a set of given size $k$, that maximizes 
+Score.
 
-In addition, the greedy algorithm can be run over a sample of the table instead of over the entire table, to reduce computation time for large tables.
-We later describe a scheme for maintaining multiple samples in memory and using them to improve response time for different drill down operations performed by the user.
+The next few sections fully develop the
+details of our solution:
+
+\squishlist
+\item We show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm.
+At a high level, this greedy algorithm is simple to state.
+The algorithm runs for $k$ steps;
+we start with an empty rule set $R$, and then at each step, we add the next best rule that maximizes Score
+
+\item In order to find the rule $r$ to add in each step, we need to measure the impact on Score for each $r$.
+This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent item-set mining.
+\squishend
+
+In some cases, the dataset may still be too large for us to return a good rule set in
+a reasonable time; in such cases, we may want to run our algorithm on a sample of the table
+rather than the entire table. In Section~\ref{sec:sampling}, we describe a scheme 
+for maintaining multiple samples in memory and using them to improve response 
+time for different drill down operations performed by the user.
 
 \subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
-We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
+\stitle{Submodularity:} We will now show that the Score function over sets of rules has a property called {\em submodularity}, giving us a greedy approximation algorithm for optimizing it. 
 \begin{definition}
@@
 
-Now consider the greedy algorithm:
-
+\stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
+as listed below, has desirable approximation guarantees:
 \begin{framed}
@@
 \end{framed}
+Since Score is a submodular function of the set $S$, 
+this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
 
-Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
+The expensive step in the above procedure is the step where the Score is computed for every
+single rule. Given the number of rules can be very large, this can be an especially time-consuming process.
 
-Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
+Instead of using the procedure described above directly, we instead develop a ``parameterized'' version 
+that will admit further approximation (depending on the parameter) in order to reduce computation further. We describe
+this algorithm next.
+
+\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
 The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
@@
 
-In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
+\stitle{Finding the Best Marginal Rule:} In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
 A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data.
@@
 
-\subsection{Sampling for big tables}
+\subsection{Sampling for big tables}\label{sec:sampling}
 \subsubsection{Sampling Overview}

### 2014-11-21 TableSummarization/Paper/TableSummarization.tex
small changes
Addressed Comments. Added NP-Hardness proof
added new graphs to writeup

@@
 
-%\usepackage{amsmath}
+\usepackage{amsmath}
 \makeatletter
@@
 \pdfpageheight=11in
-
-\renewcommand{\baselinestretch}{1.0}
 
@@
 \usepackage[normalem]{ulem}
+
+\renewcommand{\baselinestretch}{1.0}
+
+\renewcommand*\ttdefault{cmvtt}
+\usepackage[T1]{fontenc}
+
 \newtheorem{theorem}{Theorem}
@@
 \newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
+\newcommand{\mrj}[1]{\textcolor{green}{#1}}
 \newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}
@@
 
-\title{Interactive Summarization of Relational Tables}
+\title{Smart Drill-Down: Interactive Summarization of Relational Tables}
 
@@
 \and
-Aditya Parameswaran\\Stanford University\\adityagp@cs.stanford.edu
+Aditya Parameswaran\\University of Illinois\\adityagp@illinois.edu
 }
@@
 demonstrate the usefulness of smart drill down by performing experiments on
-a real dataset.
+real datasets.
 \end{abstract}

### 2014-11-21 TableSummarization/Paper/exp.tex
added new graphs to writeup

@@
 
-\subsection{Dataset}
-The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers.
+\subsection{Datasets}
+The first dataset, denoted `Marketing', contains demographic information about potential customers.
 The source of the dataset is Impact Resources, Inc., Columbus, OH (1987).
@@
 
-Unless otherwise specified, in all our experiments, we restrict the table to the first $7$ columns in order to make the result tables fit into the page.
+The second dataset, denoted `Census', is a US 1990 Census dataset from the UCI Machine Learning repository~\cite{uciml}.
+It consists of about $2.5$ million tuples, with each tuple corresponding to a person.
+It has $68$ columns, including ancestry, age, and citizenship.
+Numerical columns, such as age, have been bucketized beforehand in the dataset.
+We use this dataset in the experiments in Section~\ref{sec:quantitative_study} in order to study the accuracy and performance of sampling on a large dataset.
+
+Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page.
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
@@
 We first perform a qualitative study of our Smart Drill Down system.
-We observe the effects of various user interface operations on our dataset, and then try out a few different weight functions to study their effects.
+We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
 \subsubsection{User interface testing}
@@
 
-\subsection{Quantitative Study}
+% TODO: Add the experiment where you compare to vanilla a priori.
+
+\subsection{Quantitative Study}\label{sec:quantitative_study}
 The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size).
 In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm.
+We use both the Marketing and Census datasets.
+The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples.
+We observe that the accuracy and running time of our algorithm depends largely on the $m_w$ and $minSS$ parameters, rather than on the underlying database size.
 
@@
 The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule.
-Thus for size scoring, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
+Thus for size scoring on the Marketing dataset, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
 For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
@@
 The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
+% TODO redo the census bits weighting experiments with proper m_w (forgot to set it correctly earlier)
 
+% TODO, find for census dataset.
 In addition, we measure the number of incorrect rules per iteration.

### 2014-11-21 TableSummarization/Paper/intro.tex
New graphs. Yet to be added to writeup
added new graphs to writeup

@@
 The $\star$ character is a wildcard that matches any value in the database.
-The Count attribute can be replaced by another aggregate,
+The Count attribute can be replaced by a Sum aggregate over some measure column,
 e.g., the total sales.
@@
 in Section~\ref{sec:experiments} we experimentally evaluate
-performance on a real dataset,
+performance on real datasets,
 and show additional examples of smart drill down in action.

### 2014-11-24 TableSummarization/Paper/TableSummarization.tex
Addressed Aditya's comments. Changes in experiments.

@@
 \usepackage{subfigure}
-\usepackage{amsmath,amssymb}
+\let\proof\relax
+\let\endproof\relax
+\usepackage{amsmath,amssymb,amsthm}
 \usepackage{graphicx,color}
@@
 
+\usepackage{floatrow}
+\floatsetup[table]{font=scriptsize}
+
+\newtheorem*{theorem*}{Theorem}
 \newtheorem{theorem}{Theorem}
+\newtheorem*{lemma*}{Lemma}
 \newtheorem{lemma}{Lemma}
@@
 \newcommand{\papertext}[1]{#1}
-\newcommand{\techreporttext}[1]{}
+\newcommand{\techreporttext}[1]{#1}
 
@@
 }
+
 \begin{document}
+
 \maketitle
@@
 
-% TODO: Experiments on Sampling schemes.
-
-\begin{comment}
-Notes based on Explanation Tables paper: Prove that our problem is NP-Hard to solve exactly. Something like the Set Cover Problem.
-
-The way we deal with overlap, is that the key differentiating factor?
-
-We don't have the binary thing.
-Is it useful?
-Maybe we can summarize better without it.
-We are solving a different problem, optimising for coverage rather than explaining the binary attribute.
-They are optimising for KL divergence of max entropy distribution.
-They mention many other papers that optimize for cardinality or similar.
-We do weight times cardinality, which is a compromise between descriptiveness and coverage, and allow user to change the weight function.
-But should check those papers out anyway.
-
-\end{comment}
-
 {\small 
@@
 }
+\techreporttext{\input{appendix}}
 

### 2014-11-24 TableSummarization/Paper/algorithms.tex
Addressed Aditya's comments. Changes in experiments.

@@
 
-\section{Smart Drill-Down Algorithms}
-\label{sec:algorithms}
+\section{Smart Drill-Down Algorithms} \label{sec:algorithms}
 We now describe online algorithms for implementing
@@
 \end{lemma}
-\begin{proof}
-We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$.
-For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to.
-Then Score can be rewritten as:
-$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
-Thus each tuple $t$ `contributes' $W(t)$ to the total score.
-Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$.
-Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
-$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
-as required.
-\end{proof}
+Proofs of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}.
 
@@
 Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ...
-S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximises $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
+S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximizes $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
 \end{problem}
 Thus, the goal of the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
-We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.
-
-\begin{proof}
-We construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns.
-Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$.
-For each $u \in U$, the tuple $t_u$ corresponding to $u$ defined as follows: For each $S_i \in S$, $$t_u(i) = 1 \textrm{ if } u \in S_i \textrm{ else } t_u(i) = 0$$
-
-Define the weight function $W$ as follows:
-$$W(r) = \textrm{min}(1, |\left\lbrace i \in \left\lbrace 1, 2, ... m \right\rbrace : r(i) = 1 \right\rbrace|)$$
-Thus $W$ gives weight $0$ to rules that do not have value $1$ in any column, and weight $1$ to all other rules.
-An optimal rule set for weight function $W$ on table $T$ need only contain rules having a single $1$ value and no $0$ values, because adding more values can only reduce $MCount$, without increasing weight.
-
-Then, a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance.
-This completes our reduction.
-\end{proof}
+We can reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
+\begin{lemma}
+Problem~\ref{prob:optimal-rule-set} is NP-Hard
+\end{lemma}
 
@@
 \end{lemma}
-\begin{proof}
-For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
-$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
-Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
-$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
-For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
-$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$ 
-Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
-$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.
 
-Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
-$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
-$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
-which means that Score is submodular.
-\end{proof}
-
+%TODO: Give  better name to `Greedy Algorithm', and use that name to reference in later as well, in sampling/experimental/extension sections. 
 \stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
@@
 In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
-This is because as the size (and weight) of a rule increases, it's Count falls rapidly.
+This is because as the size (and weight) of a rule increases, its Count falls rapidly.
 The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$).
@@
 
-\subsection{Sampling for big tables}\label{sec:sampling}
+\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
 \subsubsection{Sampling Overview}
@@
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
-These passes can be very expensive if the table is large, especially if it does not fit in main memory.
+These passes can be very expensive if the table is large, especially if the table does not fit in main memory.
 If we want exact counts for rules, we have no choice but to read the entire table.
@@
 But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.
+Thankfully since our goal is to find a representative coverage of the table,
+if we miss out on a few rare tuples, it does not hurt us.
+If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
 
@@
 Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
-The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
+The $minSS$ parameter is the minimum number of sample tuples that are used to determine counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
 We later provide a way to find reasonable values of $minSS$.
@@
 At all points, the SampleHandler maintains a set of Samples in memory.
+For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user.
 Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
@@
 The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
+% TODO: Give example of what samples we'd keep.
 
 Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
+Thus the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
 The SampleHandler also computes $N_s$ when a sample is created.
@@
 The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
+In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.
 
@@
 \begin{enumerate}
-\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as it's filter rule (i.e.
-$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq \text{minSS}$, then it simply returns sample $s$.
+\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e.
+$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
 Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
@@
 
-Note that the \textbf{Combine} procedure doesn't really require additional memory.
-Since all the tuples in the `new' sample are already present in existing samples, it can simply store a set of pointers to the tuples, and the pointers can be deleted as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
+Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}.
+Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
+In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
+%TODO: Not sure how clear/necessary this is.
 
-\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample by making a pass through the table.
+\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
 Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
@@
 
-In addition, the SampleHandler can choose to create a sample of $minSS$ size, or larger (if enough memory is available).
+In addition, the SampleHandler need to create a sample of size exactly equal to $minSS$.
+It can create a sample of $< minSS$ size if there are some tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
 Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
-For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of it's sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
+For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
+We describe our algorithms for deciding what sized samples to create, and what samples to shrink, in Section~\ref{sec:sampling_algorithms}.
 
-\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may happen that there is not enough memory left to create a new sample.
+\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample.
 In that case, the SampleHandler needs to delete or shrink an existing sample.
@@
 
-\textbf{Pre-fetching:} When the user clicks on rule $r$ (the rule itself or a $\star$ in the rule), we need to get a sample, run the greedy algorithm, and display a rule-list to the user.
+\textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user.
 If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
-But after expanding $r$, there is a high chance that the user goes further and drills down on of the sub-rules $r^{\prime}$ of $r$.
+But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$.
 We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
@@
 We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
+Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the Hard Disk.
+
 Suppose at some stage, we have a tree $U$ of rules displayed to the user.
-Internal nodes of $U$ are rules that have been expanded, while leaves are rules that have not been expanded.
+The root of the tree is the trivial rule, and the parent of each rule $r$ in the tree is the rule that was drilled down on that caused $r$ to be displayed.
+Internal nodes of $U$ are rules that have been expanded (drilled down on), while leaves are rules that have not been expanded.
 Let $L$ be the set of leaves.
-Each leaf is something that the user can potentially expand (drill down on) in the next step, and thus we would like to have pre-fetched samples for leaves.
+Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for leaves.
 
@@
 
+%TODO: Comment 15 of Aditya. I don't quite understand it. 
 When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
@@
 Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule.
-More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making it's pass through the table.
+More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making its pass through the table.
 
-We choose the $n_{r^{\prime}}$ values so as to maximise the probability that the next user drill down can be satisfied using samples available in memory.
+We choose the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory.
 We define some terms required to formalise this problem next.
 
+Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule.
 For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
@@
 \begin{definition}\label{def:ess}
-$$\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
+$$ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
 \end{definition}
-If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
-We wish to maximise the probability that we can respond to the next user expansion without making another pass.
+If $ess(r^{\prime}) \geq minSS$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
+We wish to maximize the probability that we can respond to the next user expansion without making another pass.
 We now formally define our problem below:
@@
 choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{ess(r^{\prime}) \geq minSS}$$
 where the $I$'s are indicator variables, subject to :
@@
 Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
-The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and it's immediate parent.
+The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
-$\text{ess}(r)$ is redefined according to the new value of $S$ as well.
-So $$\text{ess}(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
+$ess(r)$ is redefined according to the new value of $S$ as well.
+So $$ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
 
-Now consider a rule in $r_0 \in U \setminus L$ along with all it's children.
-Let $M_{r_0}$ denote the set containing $r_0$ and all it's leaf children.
+Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
+Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
 Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
 This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$.
-Thus for each non-leaf rule $r_0$ and all it's children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
-Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{\text{ess}(r) \geq \text{minSS}}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
+Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
+Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
 Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
@@
 Let the children be $r_1, r_2, ...
+r_d$.
+For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
+r_d$.
+Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
+\squishlist
+\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
+\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
+This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$.
+Both these decreases would decrease sampling cost without affecting probability score.
+\squishend
+Thus there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
+There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution.
+Consider any such locally optimal solution $e$.
+For $e$ let children $r_{i_1}, r_{i_2}, ...
+r_{i_m}$ be in the first category, $r_{i_{m+1}}, ..
+r_{i_M}$ be in the second category, and $r_{i_{M+1}}, ..
+r_{i_d}$ in the third.
+Then the `probability value' of solution $e$ is given by :
+$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
+and its `Sampling Cost' is
+$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
+Thus there are at most $3^d$ locally optimal solutions.
+$d$ is usually small, even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
+
+\begin{comment}
+Let $r_0 \in U \setminus L$.
+Let $d$ be the number of leaf children of $r_0$.
+Let the children be $r_1, r_2, ...
 r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$.
@@
 $$1 \leq i_m \leq i_M \leq d$$
-$$n_{r_0} = \frac{\text{minSS}}{S(r_0, r_{i_m})}$$
+$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
 $$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
-$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
+$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
 $$i_M < j \Rightarrow n_{r_j} = 0$$
 And thus
-$$\text{ess}(r_j) \geq \text{minSS} \Leftrightarrow j \leq i_M$$
+$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$
 
@@
 and its `Sampling Cost' to be
-$$S(e) = \frac{\text{minSS}}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} \text{minSS} - \frac{\text{minSS}}{S(r_0, r_{i_j})}$$
+$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
 Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions. 
+\end{comment}
 
@@
 Let our possible sample sizes range from $0$ to $\mathcal{S}$.
+The number of sample sizes can be pretty large ($\mathcal{S}$), but we can make it smaller by discretizing the sample sizes, say to have granularity $100$.
 Then we create a $D \times \mathcal{S}$ array $A$.
@@
 There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
-Suppose we have a sample $s$, and say it's filter rule $f_s$ has value $v$ in column $c$.
+Suppose we have a sample $s$, and say its filter rule $f_s$ has value $v$ in column $c$.
 Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$.
@@
 
-We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = \text{minSS}$.
+We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = minSS$.
 If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be.
-Moreover, since we want good Count estimates for all rules displayed in the summary, we want $\text{minSS} >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.
+Moreover, since we want good Count estimates for all rules displayed in the summary, we want $minSS >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.
 
@@
 Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
-For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $\text{minSS} >> 5 \times 10$.
+For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $minSS >> 5 \times 10$.

### 2014-11-24 TableSummarization/Paper/exp.tex
Addressed Aditya's comments. Changes in experiments.

@@
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
+Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.
 
@@
 Our system can display optimal rule summaries using any monotonic weighting function.
-By default, we assign a rule weight equal to it's size.
+By default, we assign a rule weight equal to its size.
 In this section, we try out a couple of other weighting functions.
@@
 Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
-The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
+The algorithm is guaranteed to find the best marginal rule as long as its weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
 In this section, we study the effect of varying $m_w$ on the speed of our algorithm.
@@
 The figure shows that running time seems to be approximately linear in $m_w$.
+
+For the Census dataset, the running time is dominated by time spent in making a pass through the $2.5$ million tuples to create the first sample.
+The response time for the next user click should be quite small, as the sample created for the first expansion can usually be re-used for the next rule expansion.
 
@@
   \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-10pt}
+\vspace{-5pt}
   \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
-\vspace{-13pt}
+\vspace{-20pt}
 \end{figure}
@@
 This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample.
+The running time at very small values of $minSS$ is simply the time taken to read the table and create the sample.
 The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
-% TODO redo the census bits weighting experiments with proper m_w (forgot to set it correctly earlier)
 
-% TODO, find for census dataset.
 In addition, we measure the number of incorrect rules per iteration.
@@
 We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}.
-This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function.
+This number for the Marketing dataset is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function.
+For the Census dataset, it is around $1$ for $minSS \leq 1000$ and falls to about $0.3$ for larger values of $minSS$.
 Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.
@@
   \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-10pt}
+\vspace{-5pt}
   \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
-\vspace{-13pt}
+\vspace{-20pt}
 \end{figure}
@@
   \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-10pt}
+\vspace{-5pt}
   \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
-\vspace{-13pt}
+\vspace{-20pt}
 \end{figure}
@@
   \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-10pt}
+\vspace{-5pt}
 \caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
-\vspace{-13pt}
+\vspace{-20pt}
 \end{figure}

### 2014-11-24 TableSummarization/Paper/formal.tex
Addressed Aditya's comments. Changes in experiments.

@@
 Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
-If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.
+If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.
 

### 2014-11-24 TableSummarization/Paper/related.tex
Addressed Aditya's comments. Changes in experiments.

@@
 All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
+
+We use sampling to find approximate estimates of rule counts.
+Various other database systems~\cite{Acharya:1999:AAQ:304182.304581, Agarwal:2013:BQB:2465351.2465355} use samples to find approximate results to SQL aggregation queries.
+These systems create samples in advance and only update them when the database changes.
+In contrast, we keep updating our samples on the fly, as the user interacts with our system.
 

### 2014-11-24 TableSummarization/src/solvers/RuleTreeDisplay.java

@@ -191,12 +191,15 @@ public class RuleTreeDisplay {
 	
 	public static void main (String[] args) throws IOException {
 		List<Integer> columns = new ArrayList<Integer>();
+		Long timer;
 		final Integer firstNumColumns = 7;
 		for (int i = 1; i < firstNumColumns; i++) {
 			columns.add(i);
 		}
 		//TableInfo fullTable = Marketing.parseData();
-		TableInfo fullTable = USCensus1990.parseData(0.01);
+		timer = System.currentTimeMillis();
+		TableInfo fullTable = USCensus1990.parseData(0.04);
+		//out.println(System.currentTimeMillis() - timer);
 		TableInfo table = fullTable.getSubTable(columns);
 		Integer ruleNums = 4;
 		final String mw_bits_outFile = "Data_Graphs/mw_speed_bits";
@@ -205,6 +208,7 @@ public class RuleTreeDisplay {
 		final String mw_size_USCensus_outFile = "Data_Graphs/mw_speed_USCensus_size";
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_outFile);
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeScorer(), mw_size_outFile);
+		// Note: for USCensus experiments. Manually add the time for reading the table from file to all outputs.
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeBitsScorer(), mw_bits_USCensus_outFile);
 		//Experiments.mwSpeedTest(table, 1, 20, ruleNums, new Rule.sizeScorer(), mw_size_USCensus_outFile);
 		
@@ -214,6 +218,7 @@ public class RuleTreeDisplay {
 		final String minSS_bits_USCensus_outFile = "Data_Graphs/minSS_speed_USCensus_bits";
 		//Experiments.minSSSpeedTest(table, 8, 500, 8000, 500, ruleNums, new Rule.sizeScorer(), minSS_size_outFile);
 		//Experiments.minSSSpeedTest(table, 20, 500, 8000, 500, ruleNums, new Rule.sizeBitsScorer(), minSS_bits_outFile);
+		// Note: for USCensus experiments. Manually add the time for reading the table from file to all outputs.
[…]

### 2014-11-27 TableSummarization/Paper/algorithms.tex
Switched to VLDB format

@@
 We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
-We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}.
+We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions} in the technical report~\cite{tr}.
 
@@
 We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as below:
+Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{\prime}$, define the function {\em ess} (for `effective sample size') as below:
 \begin{definition}\label{def:ess}

### 2014-11-27 TableSummarization/Paper/extensions.tex
Switched to VLDB format

@@
 A simple approach is to create buckets for values of $A$.
-We choose a number of buckets $k$, and divide the range of values of $A$ into $k$ intervals, each corresponding to a bucket.
+We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, each corresponding to a bucket.
 We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket.
@@
 To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.
-
-ask : we should mention this right up front when we're describing the problem.
 \end{comment}

### 2014-11-27 TableSummarization/Paper/formal.tex
Switched to VLDB format

@@
 We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead.
-We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in Section~\ref{sec:extensions-sum}.
+We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in Section~\ref{sec:extensions-sum} in the technical report~\cite{tr}.
 
@@
 
-\stitle{Weighting Function 1:} 
+\stitle{Size Weighting Function:} 
 $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
@@
 
-\stitle{Weighting Function 2:}
+\stitle{Bits Weighting Function:}
 $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.

### 2014-11-27 TableSummarization/Paper/intro.tex
Switched to VLDB format

@@
 In addition to presenting this new operator called smart drill-down, in this paper, 
-we also present techniques to compute the results for this
+we also present novel sampling techniques to compute the results for this
 operator {\em in an interactive fashion} on increasingly larger databases. 
@@
 Briefly, the contributions of our paper can be summarized are follows: We develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation.
-In addition, we develop a scheme for maintaining and using multiple samples of the table in memory to improve response time on large tables.
+In addition, we develop a scheme for dynamically maintaining and using multiple samples of the table in memory to improve response time on large tables.
 
@@
 \item In Section~\ref{sec:algorithms}, we present our algorithms for
-finding optimal sets of rules, as well as our sampling schemes
+finding optimal sets of rules, as well as our dynamic sampling schemes
 for dealing with large tables.
@@
 
-\item In Section~\ref{sec:extensions}, we describe how to extend our results to summarize numerical attributes, and to use the $Sum$ aggregate instead of $Count$.
-
 \item In Section~\ref{sec:related}, we describe related work.

### 2014-11-28 TableSummarization/Paper/TableSummarization.tex
Switched to VLDB format
Several changes. Also added conclusion

@@
-%\documentclass{vldb}
-\documentclass{sig-alternate}
-%\documentclass[12pt,letterpaper]{article}
-%\documentclass[12pt]{extreport}
-%\documentstyle[amsmath,amsthm,amssymb,twocolumn]{article}
-%\newcommand{\mytitle}{Evaluating the Crowd with Confidence}
-
-\usepackage{amsmath}
-\makeatletter
-\let\@copyrightspace\relax
-\makeatother
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-
+\documentclass{vldb}
+\usepackage{graphicx}
+\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
 \usepackage{enumitem}
@@
 \newcommand{\papertext}[1]{#1}
-\newcommand{\techreporttext}[1]{#1}
+\newcommand{\techreporttext}[1]{}
 
@@
 
-\title{Smart Drill-Down: Interactive Summarization of Relational Tables}
-
-\author{
-Manas Joglekar\\Stanford University\\manasrj@stanford.edu
-\and
-Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
-\and
-Aditya Parameswaran\\University of Illinois\\adityagp@illinois.edu
-}
-
 \begin{document}
 
+\title{Smart Drill-Down}
+\numberofauthors{3} 
+\author{
+\alignauthor
+Manas Joglekar\\
+       \affaddr{Stanford University}\\
+       \affaddr{353 Serra Mall}\\
+	   \affaddr{Stanford, California 94305}\\
+       \email{manasrj@stanford.edu}
+\alignauthor
+Hector Garcia-Molina\\
+       \affaddr{Stanford University}\\
+       \affaddr{353 Serra Mall}\\
+	   \affaddr{Stanford, California 94305}\\
+       \email{hector@cs.stanford.edu}
+\alignauthor 
+Aditya Parameswaran\\
+       \affaddr{University of Illinois at Urbana-Champaign}\\
+       \affaddr{Champaign, Illinois 61820}\\
+       \email{adityagp@illinois.edu}
+}
 \maketitle
@@
 In this paper we present {\em smart drill down},
-a system for interactively exploring a relational table
+an operator for interactively exploring a relational table
 to discover ``interesting'' groups of tuples.
@@
 The analyst can tailor the definition of interesting,
-and can interactively drill down on an existing rule to
+and can interactively apply smart drill down on an existing rule to
 explore that part of the table.
 We describe an
-algorithm for finding the approximately optimal list of rules, and an optimized
-sampling scheme for efficiently summarizing large tables. Finally, we
-demonstrate the usefulness of smart drill down by performing experiments on
-real datasets.
+algorithm for finding the approximately optimal list of rules to display when the user uses a smart drill down, and a dynamic sampling scheme for efficiently dealing with large tables. Finally, we
+perform experiments on real datasets to demonstrate the usefulness of smart drill down and to study the performance of our algorithms.
 \end{abstract}
@@
 
-\input{extensions}
+\techreporttext{\input{extensions}}
 
@@
 
-{\small 
-\bibliographystyle{plain}
+\input{conc}
+
+\bibliographystyle{abbrv}
 \bibliography{TableSummarization}
-}
+
+% Example of an appendix; typically would start on a new page
+%pagebreak
+
 \techreporttext{\input{appendix}}

### 2014-11-28 TableSummarization/Paper/algorithms.tex
Several changes. Also added conclusion

@@
 We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
-We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions} in the technical report~\cite{tr}.
+We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}.
 
@@
 \item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
-Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
+Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
 \item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$.
 In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.
-Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
+Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
 \end{enumerate}
 
-First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
+As a first step towards solving Problem~\ref{prob:optimal-rule-list}, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
 Note that the weight of a rule is independent of its $MCount$.
@@
 \end{lemma}
-Proofs of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}.
+The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}.
 
-\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a set of rules as follows:
+\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:
 
@@
 \begin{problem}\label{prob:optimal-rule-set}
-Given a table $T$, a monotonic weight function $W$, and a number $k$, find the set $R$ of $k$ rules which maximizes Score($R$) as defined in Definition~\ref{def:set-score}.
+Given a table $T$, a monotonic weight function $W$, and a number $k$, find the set (not list) $R$ of $k$ rules which maximizes Score($R$) as defined in Definition~\ref{def:set-score}.
 \end{problem}
@@
 \item In order to find the rule $r$ to add in each step, we need to measure the impact on Score for each $r$.
-This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent item-set mining.
+This is done in several passes over the table, using ideas from the a-priori algorithm~\cite{apriori} for frequent item-set mining.
 \squishend
@@
 time for different drill down operations performed by the user.
+Our sampling scheme dynamically adapts to the current interaction scenario that the user is in; drawing from ideas in approximation algorithms and optimization theory.
 
@@
 Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead.
-For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.
+For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.
 
@@
 The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$).
-Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
+Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
@@
 
-In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}.
+In order to avoid counting too many rules, we use an idea from the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}.
+Recall that the a-priori algorithm is used to find all frequent itemsets that have a support greater than a threshold.
+Unlike the frequent itemset mining algorithm, our goal is to find the single best marginal rule.
+Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset.
 We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule.
-In the $j^{th}$ pass, we compute counts for rules of size $j$.
+In the $j^{th}$ pass, we compute counts and marginal values for rules of size $j$.
 To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$.
-In the first pass, we would compute the marginal values of all rules of size $1$.
-In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass.
+In the first pass, we would compute the counts and marginal values of all rules of size $1$.
+In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts and marginal values of size $2$ rules, and be more selective about which rules to count in the second pass.
 For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$.
@@
 If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.
+
+Note that the reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently.
+This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.
 
@@
 }
+$C = C \cup C_n$
+
 $C_o = C_n$
@@
 $C_n = \phi$
-
-$C = C \cup C_o$
 
@@
 \subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
-\subsubsection{Sampling Overview}
 We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.
@@
 
-At all points, the SampleHandler maintains a set of Samples in memory.
+At all points, the SampleHandler maintains a set of samples in memory.
 For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user.
@@
 We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
-Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the Hard Disk.
+Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the hard disk.
 
@@
 When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
-Since making a pass over the the table from the Hard Disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
+Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
 Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.

### 2014-11-28 TableSummarization/Paper/conc.tex
Several changes. Also added conclusion

%!TEX root = TableSummarization.tex

\section{Conclusion}\label{sec:conclusion}
We have presented a new data exploration operator called smart drill down.
Like traditional drill down, it allows an analyst
to quickly discover interesting value patterns (rules) that occur frequently
(or that represent high values of some metric attribute)
across diverse parts of a table.

We presented an algorithm for optimally selecting rules
to display, as well as a scheme for performing such selections
based on data samples. Working with samples makes smart drill down
relatively insensitive to the size of the table.

Our experimental results show that smart drill down
is fast enough to be interactive under various
realistic scenarios. We also showed that the accuracy is high
when sampling is used, and when the maximum weight ($m_w$)
approximation is used.

### 2014-11-28 TableSummarization/Paper/exp.tex
Switched to VLDB format
Several changes. Also added conclusion

@@
 Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page.
+We use our current implementation of our algorithms, which directly outputs tables in latex format.
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
@@
 \end{tabular} 
+\vspace{-15pt}
 \caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
@@
 \end{tabular} 
+\vspace{-15pt}
 \caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
@@
 \end{tabular} 
+\vspace{-15pt}
 \caption{A Rule expansion \label{table:uiexamplerule}} 
@@
 \end{tabular} 
+\vspace{-15pt}
 \caption{Bits scoring\label{table:weigtingbitwise}} 
@@
 \end{tabular} 
+\vspace{-15pt}
 \caption{Size minus one weighting \label{table:weightingsizeminusone}} 
@@
 \begin{figure}
-\hspace{-20pt}
+\hspace{-15pt}
   \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
@@
   \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
-\vspace{-20pt}
+\vspace{-15pt}
 \end{figure}
@@
 For each value of $minSS$ and $W$, we drill down on the empty rule and find the computation time and percent error $50$ times, and take the average value for time and error over those $50$ iterations.
-This average time is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}.
-The average percent error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error_percent}.
+This average time is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}(a).
+The average percent error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error_percent}(b).
 
@@
 If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule.
-We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}.
+We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}(c).
 This number for the Marketing dataset is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function.
@@
 
-\begin{figure}
-\hspace{-20pt}
-  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-5pt}
-  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
-\vspace{-20pt}
-\end{figure}
-
-\begin{figure}
-\hspace{-20pt}
-  \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-5pt}
-  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
-\vspace{-20pt}
-\end{figure}
-
-\begin{figure}
-\hspace{-20pt}
-  \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-5pt}
-\caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
-\vspace{-20pt}
-\end{figure}
+\begin{figure*}[!ht]
+\centering
+\includegraphics[scale=0.46]{graphs/minSS_speed.pdf}
+\includegraphics[scale=0.46]{graphs/minSS_error_percent.pdf}
+\includegraphics[scale=0.46]{graphs/minSS_error_rule.pdf}
+\vspace{-15pt}
+\caption{(a) Running time for different values of parameter $minSS$ (b) Error in Count for different values of parameter $minSS$ (c) Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_speed}\label{fig:minSS_error_percent} \label{fig:minSS_error_rule}}
+\vspace{-10pt}
+\end{figure*}
 
@@
 $\star$ & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $1964$ & $2$ \\
-
 \end{comment}

### 2014-11-28 TableSummarization/Paper/formal.tex
Several changes. Also added conclusion

@@
 that they allow us to separate the inherent goodness of a rule from
-how much it captures of the data in $\calD$.
+how much it captures the data in $\calD$.
 
 We now describe the first portion:
-we define {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
-Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
+we define {\em Count}($r$) as the total number of tuples $t \in T$ that are covered by $r$. 
+Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') as the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$.
 A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules.
@@
 
+Our smart drill downs still display the Count of each rule rather than the $MCount$.
+This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret.
+In any case, it would be a simple extension to display MCount in another column.
+
 \stitle{Formal Problem:} We now formally define our problem:
@@
 \end{problem}
-Throughout this paper, we use the {\em Count} aggregate of a rule when determining what rules to display to the user.
+Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user.
 We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead.
-We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in Section~\ref{sec:extensions-sum} in the technical report~\cite{tr}.
-
-% Describe architecture here? (With a diagram). Show SampleHandler, QueryEngine, etc.
+We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in the `Extensions' section of the technical report~\cite{tr}.
 
@@
 given the rules, it should be as easy as possible to reproduce the table.
-We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is (i.e., how much information
+We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ depending on how expressive the rule is (i.e., how much information
 it conveys). Let us consider some canonical forms for the function $W(r)$ and their interpretation; later, we describe the full
@@
 Thus this rule contributes $200$ to the score.
-For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those.
+For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those tuples.
 Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total.
@@
 Like the previous scoring function, this one adds some weight for every non-starred value in a rule.
-But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column.
+But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the inherent complexity of the column.
 The reason behind this variation is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
@@
 A non-starred valued in a rule $r$ thus pre-fills $MCount(r,R) \lceil \text{log}(|c|) \rceil$ bits of the table.
-Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.
+Hence this scoring function gives us the the number of bits of the table that are pre-filled by the rule set.
 
@@
 \stitle{Other Weighting Functions:}
-Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
+Even though we have given two example weighting functions here, our algorithms allow the user to leverage any weighting function $W$, subject to only two conditions:
 \squishlist
@@
 \squishend
-A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).
+A weight function can be used in several ways, including expressing a higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or expressing indifference towards a column (by adding zero weight for having non-$\star$ value in that column).
 
@@
 \end{tabular}
+\vspace{-15pt}
 \caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
@@
 
-\subsection{Example: Smart Drill-Down Operations}
+\subsection{Smart Drill-Down Operations}
 \label{sec:interface}
-We now explain how a user can use smart drill-downs to explore the contents of a table.
-
 When the user starts using a system equipped with
@@
 \end{tabular}
+\vspace{-15pt}
 \caption{Result of clicking on a $\star$ \label{table:introexample3}}

### 2014-11-28 TableSummarization/Paper/intro.tex
Several changes. Also added conclusion

@@
 Each rule summarizes the set of tuples that are described by it.
-Again, instead of Count, the system can display some other aggregate, such as
+Again, instead of Count, the system can display a sum aggregate, such as
 the total Sales.
@@
 \end{tabular}
+\vspace{-15pt}
 \caption{Initial summary}\label{table:introexample0}
@@
 \end{tabular}
+\vspace{-15pt}
 \caption{Result after first smart drill down}\label{table:introexample}
@@
 \centering
+\vspace{-10pt}
 \begin{tabular}{| l | l | l | l | l |}
@@
 \end{tabular}
+\vspace{-15pt}
 \caption{Result after second smart drill down} \label{table:introexample2}
@@
 In Section~\ref{sec:interface} we describe another option
-where the analyst clicks an a $\star$ to obtain
-rules that have values in that location.
+where the analyst clicks on a $\star$ in a column to obtain
+rules that have non-$\star$ values in that column.
 
@@
 Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}..
+and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
 
@@
 \item In Section~\ref{sec:formal}, we introduce the relevant notation and formally define smart drill down.
-After that, we discuss different schemes for weighting rules, and describe our interactive user interface.
+After that, we describe different schemes for weighting rules, and our interactive user interface.
 
@@
 
-\item In Section~\ref{sec:related}, we describe related work.
+\item We describe related work in Section~\ref{sec:related}, and conclude in Section~\ref{sec:conclusion}.
 \squishend

### 2014-11-29 TableSummarization/Paper/algorithms.tex
Added hinge loss, and clarified tree structure.

@@
 
+% Example from situation in a previous table. And good and bad samples to keep there.
 At all points, the SampleHandler maintains a set of samples in memory.
@@
 
-In addition, the SampleHandler need to create a sample of size exactly equal to $minSS$.
-It can create a sample of $< minSS$ size if there are some tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
+In addition, the SampleHandler needs to return a sample of size equal to $minSS$.
+It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
 Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
@@
 In that case, the SampleHandler needs to delete or shrink an existing sample.
-It can use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
+It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
 Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
@@
 Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the hard disk.
+We leverage techniques from approximation algorithms and optimization theory.
 
-Suppose at some stage, we have a tree $U$ of rules displayed to the user.
-The root of the tree is the trivial rule, and the parent of each rule $r$ in the tree is the rule that was drilled down on that caused $r$ to be displayed.
-Internal nodes of $U$ are rules that have been expanded (drilled down on), while leaves are rules that have not been expanded.
+At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule.
+The tree is formed as follows: The root of the tree corresponds to the trivial rule.
+And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, ..
+r_k$ being displayed.
+Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, ..
+r_k$.
+Note that multiple nodes of the tree may correspond to the same rule.
+Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded.
 Let $L$ be the set of leaves.
-Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for leaves.
+Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
+In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.
 
@@
 That is, we can assume that every leaf is equally likely to be expanded next.
-However, we can also use Machine Learning on past user behaviour, along with rule features such as `rule depth in tree' and `rule distance from last expanded rule' to get a better probability estimate of each rule being expanded.
+We can improve on this by assigning zero probability to leaves whose rule is also displayed in an internal node of the tree.
+We can also use Machine Learning on past user behaviour, along with node features such as `node depth in tree' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.
 
-%TODO: Comment 15 of Aditya. I don't quite understand it. 
 When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
@@
 We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
+If the same rule occurs in multiple nodes $r_1$, $r_2$ of the tree, then $S(r_1, r_2)$ is naturally $1$.
 Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{\prime}$, define the function {\em ess} (for `effective sample size') as below:
@@
 If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
+
+\begin{comment}
+However, we can make the problem convex (and hence tractable) with two simplifications.
+The first simplification is, we modify our objective function to use hinge-loss instead of a step function.
+That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
+Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$.
+The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers.
+After obtaining our optimal sample sizes, we can round them up to get integer sample sizes.
+This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
+$|U|$ is usually negligible compared to the memory capacity, or $minSS$.
+
+In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function).
+Thus our new optimization problem becomes
+\begin{problem}\label{prob:sample-sizes-hinge-loss}
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
+subject to :
+$$\sum_{r \in U} n_r < \text{capacity}$$
+\end{problem}
+The constraint is linear in the $n_r$ variables, and hence convex.
+Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex.
+The constant function $-1$ is convex as well.
+Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum.
+Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
+
+The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus
+\end{comment}
 

### 2014-11-29 TableSummarization/Paper/intro.tex
Added hinge loss, and clarified tree structure.

@@
 If the analyst drills down on the Store attribute (first $\star$),
-then the system displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
+then the operator displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
 where $X$ is a Store in the answer table, and $C$
@@
 Each rule summarizes the set of tuples that are described by it.
-Again, instead of Count, the system can display a sum aggregate, such as
+Again, instead of Count, the operator can display a sum aggregate, such as
 the total Sales.

### 2014-11-29 TableSummarization/Paper/algorithms.tex
small changes

@@
 
-\begin{comment}
+-----Hinge-loss begin-----
+
 However, we can make the problem convex (and hence tractable) with two simplifications.
@@
 
-The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus
-\end{comment}
+The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.
+
+----Hinge-loss-end-----
 

### 2014-11-29 TableSummarization/Paper/algorithms.tex
small length reduction changes

@@
 \end{enumerate}
+
+\begin{algorithm}\label{algo:best-rule-set}
+\scriptsize
+\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$S$ (Solution set of rules)}
+$S = \phi$
+
+\For {$i$ from $1$ to $k$}{
+$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}
+
+$S = S \cup \left\lbrace R_m \right\rbrace$
+}
+\Return $S$
+\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
+\end{algorithm}
 
@@
 \squishend
+
+\begin{algorithm}\label{algo:best-marginal-rule}
+\scriptsize
+\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
+$H = 0$ \tcc*{Threshold for deciding if to count for a rule.}
+$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
+\For{$j$ from $1$ to number of columns in $T$}{
+\If {$j = 1$}{
+$C_n = \text{ all rules of size } 1$
+}
+\Else{ 
+$C_n = \text{all size-}i$ super-rules of rules from $C_o$
+}
+\ForEach {$R \in C_n$}{
+$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
+\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
+$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
+}
+\If {$(M < H)$}{ 
+$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to be in the solution}
+}
+}
+\If {$C_n = \phi$}{
+break;
+}
+\ForEach {$R \in C_n$}{
+Count$(R) = 0$ \tcc*{Initialize}
+MarginalValue$(R) = 0$ \tcc*{Initialize}
+}
+\ForEach {$t \in T$}{
+Let $R_S$ be the highest weight rule in $S$ that covers $t$
+
+\ForEach {$R \in C_n$ that covers $t$}{
+Count($R$) $++$
+
+MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
+}
+}
+$C = C \cup C_n$
+
+$C_o = C_n$
+
+$C_n = \phi$
+
+$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
+}
+\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
+\caption{Find best marginal rule}
+\end{algorithm}
 
@@
 This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.
-
-\begin{algorithm}\label{algo:best-rule-set}
-\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$S$ (Solution set of rules)}
-$S = \phi$
-
-\For {$i$ from $1$ to $k$}{
-$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}
-
-$S = S \cup \left\lbrace R_m \right\rbrace$
-}
-\Return $S$
-\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
-\end{algorithm}
-
-\begin{algorithm}\label{algo:best-marginal-rule}
-\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
-$H = 0$ \tcc*{Threshold for deciding if to find count for a rule.}
-$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
-\For{$j$ from $1$ to number of columns in $T$}{
-\If {$j = 1$}{
-$C_n = \text{ all rules of size } 1$
-}
-\Else{ 
-$C_n = \text{all size-}i$ super-rules of rules from $C_o$
-}
-\ForEach {$R \in C_n$}{
-$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
-\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
-$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
-}
-\If {$(M < H)$}{ 
-$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to possibly be in the solution}
-}
-}
-\If {$C_n = \phi$}{
-break;
-}
-\ForEach {$R \in C_n$}{
-Count$(R) = 0$ \tcc*{Initialize}
-MarginalValue$(R) = 0$ \tcc*{Initialize}
-}
-\ForEach {$t \in T$}{
-Let $R_S$ be the highest weight rule in $S$ that covers $t$
-\ForEach {$R \in C_n$ that covers $t$}{
-Count($R$) $++$
-
-MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
-}
-}
-$C = C \cup C_n$
-
-$C_o = C_n$
-
-$C_n = \phi$
-
-$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
-}
-\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
-\caption{Find best marginal rule}
-\end{algorithm}
 

### 2014-11-29 TableSummarization/Paper/algorithms.tex
small length reduction changes

@@
 \noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear.
-We now first show that Problem~\ref{prob:optimal-rule-set}, and consequently Problem~\ref{prob:optimal-subrule-list} and Problem~\ref{prob:optimal-rule-list} are NP-Hard, and then present an approximation algorithm for solving Problem~\ref{prob:optimal-rule-set}.
+We now first show that Problem~\ref{prob:optimal-rule-set}, and consequently Problem~\ref{prob:optimal-subrule-list} and Problem~\ref{prob:optimal-rule-list} are {\sc NP-Hard}, and then present an approximation algorithm for solving Problem~\ref{prob:optimal-rule-set}.
 
 \subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
-We reduce the well known NP-Hard {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
-thus demonstrating the NP-Hardness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
+We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
+thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
 \begin{problem}\label{prob:maximum-coverage}
@@
 \begin{lemma}
-Problem~\ref{prob:optimal-rule-set} is NP-Hard
+Problem~\ref{prob:optimal-rule-set} is {\sc NP-Hard}.
 \end{lemma}
@@
 \subsection{Algorithm Overview}\label{sec:alg-overview}
-Given that the problem is NP-Hard, we now present our algorithms 
+Given that the problem is {\sc NP-Hard}, we now present our algorithms 
 for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
@@
 Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk.
-Finally, we mention some additional optimizations we can make, and heuristics for setting the minimum sample size required from the SampleHandler.
+Finally, we
+\techreporttext{mention some additional optimizations we can make, and} 
+describe how we can set the minimum sample size required from the SampleHandler.
 
@@
 
+\techreporttext{
 \subsubsection{Additional optimizations}
@@
 Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
+}
 

### 2014-11-29 TableSummarization/Paper/intro.tex
small length reduction changes

@@
 require no pre-materialization, and can be implemented 
-within any relational database system with no additional storage.
+within any relational database system.
 
@@
 Compared to traditional drill down, our smart drill down has two important advantages:
-\begin{itemize}
+\squishlist
 \item
@@
 (Target, bicycles, $\star$, $200$, $2$).
-\end{itemize}
+\squishend
 
@@
 
-Briefly, the contributions of our paper can be summarized are follows: We develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation.
-In addition, we develop a scheme for dynamically maintaining and using multiple samples of the table in memory to improve response time on large tables.
+Our chief contribution in this paper is the {\em smart drill-down} interaction operator,
+an extension of traditional drill-down, aimed at allowing analysts to zoom into the
+more ``interesting'' parts of a dataset. In addition to this operator,
+we develop techniques to support this operator on increasingly larger datasets:
+\squishlist
+\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
+and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
+\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
+maintaining samples in memory to support smart drill-down. 
+We show that optimal identification of samples is once again {\sc NP-Hard}, \agp{??} and 
+we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
+% to improve response time on large tables.
+\squishend
 

### 2014-11-29 TableSummarization/Paper/algorithms.tex
stuff

@@
 
-\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
-We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.
-First in Section~\ref{sec:sample-using}, we motivate the need for sampling.
-Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
-The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next.
-Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk.
-Finally, we
-\techreporttext{mention some additional optimizations we can make, and} 
-describe how we can set the minimum sample size required from the SampleHandler.
-
-\subsubsection{Using samples to respond to user operations}\label{sec:sample-using}
-Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
-These passes can be very expensive if the table is large, especially if the table does not fit in main memory.
-If we want exact counts for rules, we have no choice but to read the entire table.
-
-But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-Thankfully since our goal is to find a representative coverage of the table,
-if we miss out on a few rare tuples, it does not hurt us.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
-
-Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
-Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
-The $minSS$ parameter is the minimum number of sample tuples that are used to determine counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
-We later provide a way to find reasonable values of $minSS$.
-
-% Example from situation in a previous table. And good and bad samples to keep there.
-At all points, the SampleHandler maintains a set of samples in memory.
-For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user.
-Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
-The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
-The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
-The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
-% TODO: Give example of what samples we'd keep.
-
-Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
-Thus the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
-The SampleHandler also computes $N_s$ when a sample is created.
-Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
-The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
-In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.
-
-When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule.
-At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample.
-Creating a new sample by making a pass through the table is called \textbf{Create} (further described below).
-At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
-\begin{enumerate}
-\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e.
-$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
-Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
-
-\item \textbf{Combine:} If \textbf{Find} doesn't work i.e.
-if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
-We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
-That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
-
-Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}.
-Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
-In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
-%TODO: Not sure how clear/necessary this is.
-
-\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
-Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples.
-We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
-
-In addition, the SampleHandler needs to return a sample of size equal to $minSS$.
-It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
-Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
-For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
-We describe our algorithms for deciding what sized samples to create, and what samples to shrink, in Section~\ref{sec:sampling_algorithms}.
-
-\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample.
-In that case, the SampleHandler needs to delete or shrink an existing sample.
-It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
-Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
-%For shrink, prefer shrinking samples of the largest size (> minSS) first?
-\end{enumerate}
-
-We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in Section~\ref{sec:sampling_algorithms}.
-
-\textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user.
-If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
-But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$.
-We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
-So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
-That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
-In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
-
-\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
-We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
-Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the hard disk.
-We leverage techniques from approximation algorithms and optimization theory.
-
-At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule.
-The tree is formed as follows: The root of the tree corresponds to the trivial rule.
-And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, ..
-r_k$ being displayed.
-Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, ..
-r_k$.
-Note that multiple nodes of the tree may correspond to the same rule.
-Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded.
-Let $L$ be the set of leaves.
-Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
-In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.
-
-We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
-In the absence of additional information, we can assume a uniform probability distribution.
-That is, we can assume that every leaf is equally likely to be expanded next.
-We can improve on this by assigning zero probability to leaves whose rule is also displayed in an internal node of the tree.
-We can also use Machine Learning on past user behaviour, along with node features such as `node depth in tree' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.
-
-When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
-Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
-Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.
-For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table.
-Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule.
-More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making its pass through the table.
-
-We choose the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory.
-We define some terms required to formalise this problem next.
-
-Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule.
-For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
-Call this quantity $S(r_1, r_2)$.
-We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$.
-If the same rule occurs in multiple nodes $r_1$, $r_2$ of the tree, then $S(r_1, r_2)$ is naturally $1$.
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{\prime}$, define the function {\em ess} (for `effective sample size') as below:
-\begin{definition}\label{def:ess}
-$$ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
-\end{definition}
-If $ess(r^{\prime}) \geq minSS$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
-We wish to maximize the probability that we can respond to the next user expansion without making another pass.
-We now formally define our problem below:
-\begin{problem}\label{prob:sample-sizes}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
-choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{ess(r^{\prime}) \geq minSS}$$
-where the $I$'s are indicator variables, subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
-\end{problem}
-Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables.
-If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
-
------Hinge-loss begin-----
-
-However, we can make the problem convex (and hence tractable) with two simplifications.
-The first simplification is, we modify our objective function to use hinge-loss instead of a step function.
-That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
-Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$.
-The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers.
-After obtaining our optimal sample sizes, we can round them up to get integer sample sizes.
-This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
-$|U|$ is usually negligible compared to the memory capacity, or $minSS$.
-
-In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function).
-Thus our new optimization problem becomes
-\begin{problem}\label{prob:sample-sizes-hinge-loss}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
-choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
-subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
-\end{problem}
-The constraint is linear in the $n_r$ variables, and hence convex.
-Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex.
-The constant function $-1$ is convex as well.
-Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum.
-Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
-
-The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.
-
-----Hinge-loss-end-----
-
-Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
-The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
-That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
-$ess(r)$ is redefined according to the new value of $S$ as well.
-So $$ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
-
-Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
-Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
-Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
-This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$.
-Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
-Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
-Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
-We describe both these steps in detail below:
-
-Let $r_0 \in U \setminus L$.
-Let $d$ be the number of leaf children of $r_0$.
-Let the children be $r_1, r_2, ...
-r_d$.
-For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
-r_d$.
-Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
-\squishlist
-\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
-\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
-This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$.
-Both these decreases would decrease sampling cost without affecting probability score.
-\squishend
-Thus there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
-There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution.
-Consider any such locally optimal solution $e$.
-For $e$ let children $r_{i_1}, r_{i_2}, ...
-r_{i_m}$ be in the first category, $r_{i_{m+1}}, ..
-r_{i_M}$ be in the second category, and $r_{i_{M+1}}, ..
-r_{i_d}$ in the third.
-Then the `probability value' of solution $e$ is given by :
-$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
-and its `Sampling Cost' is
-$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Thus there are at most $3^d$ locally optimal solutions.
-$d$ is usually small, even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
-
-\begin{comment}
-Let $r_0 \in U \setminus L$.
-Let $d$ be the number of leaf children of $r_0$.
-Let the children be $r_1, r_2, ...
-r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$.
-Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
-$$1 \leq i_m \leq i_M \leq d$$
-$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
-$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
-$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-$$i_M < j \Rightarrow n_{r_j} = 0$$
-And thus
-$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$
-
-Call this locally optimal solution $e$. Then we define its `probability value' to be
-$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
-and its `Sampling Cost' to be
-$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions. 
-\end{comment}
-
-Then next step is to combine the solutions using dynamic programming.
-Let the $M$ sets be called $M_0, M_1, ...
-M_D$.
-Let our possible sample sizes range from $0$ to $\mathcal{S}$.
-The number of sample sizes can be pretty large ($\mathcal{S}$), but we can make it smaller by discretizing the sample sizes, say to have granularity $100$.
-Then we create a $D \times \mathcal{S}$ array $A$.
-The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ...
-M_i$ with total sample size at most $j$.
-We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$.
-Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$.
-Then we have,
-$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
-This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.
-
-\techreporttext{
-\subsubsection{Additional optimizations}
-There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
-Suppose we have a sample $s$, and say its filter rule $f_s$ has value $v$ in column $c$.
-Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$.
-So we do not need to explicitly store the column $c$ value of any tuple in $T_s$.
-We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
-In addition, we may have a tuple occur in multiple samples.
-Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
-}
-
-\subsubsection{Setting $minSS$}
-Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e.
-$x|T|$ tuples.
-Say we have a uniform random sample $s$ of $T$.
-The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$.
-Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$.
-In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$.
-That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.
-
-We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = minSS$.
-If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be.
-Moreover, since we want good Count estimates for all rules displayed in the summary, we want $minSS >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.
-
-Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
-This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule).
-Let $c$ be the column with the fewest distinct values.
-Say it has $|c|$ values.
-Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$.
-For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times.
-So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.
-
-The highest scoring rule can have weight at most $|C|$ (the total number of columns).
-Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
-Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
-For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $minSS >> 5 \times 10$.
+\input{sampling}

### 2014-11-29 TableSummarization/Paper/sampling.tex
sampling

%!TEX root = TableSummarization.tex

\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. 
\papertext{In the technical report,}
\techreporttext{Finally,}
we mention some additional optimizations we can make, and
describe how we can set the minimum sample size required from the SampleHandler.

\subsubsection{Using samples to respond to user operations}\label{sec:sample-using}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if the table does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. Thankfully since our goal is to find a representative coverage of the table,
if we miss out on a few rare tuples, it does not hurt us. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
The $minSS$ parameter is the minimum number of sample tuples that are used to determine counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be. We later provide a way to find reasonable values of $minSS$.

% Example from situation in a previous table. And good and bad samples to keep there.
At all points, the SampleHandler maintains a set of samples in memory. For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user. Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point. 
% TODO: Give example of what samples we'd keep.

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. Thus the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$. The SampleHandler also computes $N_s$ when a sample is created. Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table. In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.

When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule. At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample. Creating a new sample by making a pass through the table is called \textbf{Create} (further described below). At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
\begin{enumerate}
\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e. $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$. Algorithm~\ref{algo:best-rule-set} can then be run on $s$.

\item \textbf{Combine:} If \textbf{Find} doesn't work i.e. if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$. If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution. That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.

Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}. Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}. In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
%TODO: Not sure how clear/necessary this is.

\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table. Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used. In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples. 
We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.

In addition, the SampleHandler needs to return a sample of size equal to $minSS$. It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available). Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table. For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table. We describe our algorithms for deciding what sized samples to create, and what samples to shrink, in Section~\ref{sec:sampling_algorithms}.

\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample. In that case, the SampleHandler needs to delete or shrink an existing sample. It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink. Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
%For shrink, prefer shrinking samples of the largest size (> minSS) first?
\end{enumerate}

We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in Section~\ref{sec:sampling_algorithms}.

\textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user. If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table. But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$. We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples. So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background. That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.

\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory. Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the hard disk. We leverage techniques from approximation algorithms and optimization theory.

At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule. The tree is formed as follows: The root of the tree corresponds to the trivial rule. And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$. Note that multiple nodes of the tree may correspond to the same rule. Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded. Let $L$ be the set of leaves. Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.

We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded. In the absence of additional information, we can assume a uniform probability distribution. That is, we can assume that every leaf is equally likely to be expanded next. We can improve on this by assigning zero probability to leaves whose rule is also displayed in an internal node of the tree. We can also use Machine Learning on past user behaviour, along with node features such as `node depth in tree' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.

When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory. Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass. Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples. For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table. Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule. More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making its pass through the table.

We choose the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory. We define some terms required to formalise this problem next.

Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule. For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples. Call this quantity $S(r_1, r_2)$. We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. If the same rule occurs in multiple nodes $r_1$, $r_2$ of the tree, then $S(r_1, r_2)$ is naturally $1$. Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{\prime}$, define the function {\em ess} (for `effective sample size') as below: 
\begin{definition}\label{def:ess}
$$ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
\end{definition}
If $ess(r^{\prime}) \geq minSS$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table. We wish to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{ess(r^{\prime}) \geq minSS}$$
where the $I$'s are indicator variables, subject to :
$$\sum_{r \in U} n_r < \text{capacity}$$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.

-----Hinge-loss begin-----

However, we can make the problem convex (and hence tractable) with two simplifications. The first simplification is, we modify our objective function to use hinge-loss instead of a step function. That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$. The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers. After obtaining our optimal sample sizes, we can round them up to get integer sample sizes. This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree. $|U|$ is usually negligible compared to the memory capacity, or $minSS$.

In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function). Thus our new optimization problem becomes
\begin{problem}\label{prob:sample-sizes-hinge-loss}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
subject to :
$$\sum_{r \in U} n_r < \text{capacity}$$
\end{problem}
The constraint is linear in the $n_r$ variables, and hence convex. Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex. The constant function $-1$ is convex as well. Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum. Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.

The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.

----Hinge-loss-end-----

Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution. The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent. That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$. $ess(r)$ is redefined according to the new value of $S$ as well. So $$ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.

Now consider a rule in $r_0 \in U \setminus L$ along with all its children. Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children. Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$. This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$. Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$. Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$. Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s. We describe both these steps in detail below:

Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$. For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ... r_d$. Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
\squishlist
\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$. Both these decreases would decrease sampling cost without affecting probability score. 
\squishend
Thus there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution. Consider any such locally optimal solution $e$. For $e$ let children $r_{i_1}, r_{i_2}, ... r_{i_m}$ be in the first category, $r_{i_{m+1}}, .. r_{i_M}$ be in the second category, and $r_{i_{M+1}}, .. r_{i_d}$ in the third.  Then the `probability value' of solution $e$ is given by : 
$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
and its `Sampling Cost' is
$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
Thus there are at most $3^d$ locally optimal solutions. $d$ is usually small, even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.

\begin{comment}
Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$. Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
$$1 \leq i_m \leq i_M \leq d$$
$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
$$i_M < j \Rightarrow n_{r_j} = 0$$
And thus
$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$

Call this locally optimal solution $e$. Then we define its `probability value' to be
$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
and its `Sampling Cost' to be
$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions. 
\end{comment}

Then next step is to combine the solutions using dynamic programming. Let the $M$ sets be called $M_0, M_1, ... M_D$. Let our possible sample sizes range from $0$ to $\mathcal{S}$. The number of sample sizes can be pretty large ($\mathcal{S}$), but we can make it smaller by discretizing the sample sizes, say to have granularity $100$. Then we create a $D \times \mathcal{S}$ array $A$. The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ... M_i$ with total sample size at most $j$. We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$. Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$. Then we have,
$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.

\techreporttext{
\subsubsection{Additional optimizations}
There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
Suppose we have a sample $s$, and say its filter rule $f_s$ has value $v$ in column $c$. Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$. So we do not need to explicitly store the column $c$ value of any tuple in $T_s$. We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
In addition, we may have a tuple occur in multiple samples. Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.

\subsubsection{Setting $minSS$}
Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e. $x|T|$ tuples. Say we have a uniform random sample $s$ of $T$. The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$. Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$. In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$. That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.

We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = minSS$. If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be. Moreover, since we want good Count estimates for all rules displayed in the summary, we want $minSS >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.

Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$. This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule). Let $c$ be the column with the fewest distinct values. Say it has $|c|$ values. Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$. For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times. So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.

The highest scoring rule can have weight at most $|C|$ (the total number of columns). Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$. Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS. For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $minSS >> 5 \times 10$.
}

### 2014-11-29 TableSummarization/Paper/TableSummarization.tex
small length reduction changes
small changes for length

@@
 
-\usepackage{floatrow}
-\floatsetup[table]{font=scriptsize}
+% \usepackage{floatrow}
+% \floatsetup[table]{font=scriptsize}
+% \renewcommand\FBbskip{-10pt}
 
@@
 
+\makeatletter
+\def\@copyrightspace{\relax}
+\makeatother
+
 \begin{document}
@@
        \affaddr{Stanford University}\\
-       \affaddr{353 Serra Mall}\\
-	   \affaddr{Stanford, California 94305}\\
+%       \affaddr{353 Serra Mall}\\
+%	   \affaddr{Stanford, California 94305}\\
        \email{manasrj@stanford.edu}
@@
        \affaddr{Stanford University}\\
-       \affaddr{353 Serra Mall}\\
-	   \affaddr{Stanford, California 94305}\\
+%       \affaddr{353 Serra Mall}\\
+%	   \affaddr{Stanford, California 94305}\\
        \email{hector@cs.stanford.edu}
@@
 Aditya Parameswaran\\
-       \affaddr{University of Illinois at Urbana-Champaign}\\
-       \affaddr{Champaign, Illinois 61820}\\
+       \affaddr{University of Illinois (UIUC)}\\
+%       \affaddr{Champaign, Illinois 61820}\\
        \email{adityagp@illinois.edu}
@@
 \begin{abstract}
-In this paper we present {\em smart drill down},
+We present {\em smart drill-down},
 an operator for interactively exploring a relational table
-to discover ``interesting'' groups of tuples.
+to discover and summarize ``interesting'' groups of tuples.
 Each group of tuples is described by a {\em rule}.
 For instance, the rule $(a, b, \star, 1000)$ tells us that
-there are a $1000$ tuples with value $a$ in the first column and $b$
+there are a thousand tuples with value $a$ in the first column and $b$
 in the second column (and any value in the third column).
-Smart drill down presents an analyst with a list of rules that
+Smart drill-down presents an analyst with a list of rules that
 together describe interesting aspects of the table.
 The analyst can tailor the definition of interesting,
-and can interactively apply smart drill down on an existing rule to
+and can interactively apply smart drill-down on an existing rule to
 explore that part of the table.
-We describe an
-algorithm for finding the approximately optimal list of rules to display when the user uses a smart drill down, and a dynamic sampling scheme for efficiently dealing with large tables. Finally, we
-perform experiments on real datasets to demonstrate the usefulness of smart drill down and to study the performance of our algorithms.
+We demonstrate that the underlying optimization problems 
+are {\sc NP-Hard}, and describe an
+algorithm for finding the approximately optimal list of rules to display when the user uses a smart drill-down, and a dynamic sampling scheme for efficiently interacting with large tables. Finally, we
+perform experiments on real datasets to demonstrate the usefulness of smart drill-down and study the performance of our algorithms.
 \end{abstract}
@@
 
+{\small
 \bibliographystyle{abbrv}
 \bibliography{TableSummarization}  
+}
 

### 2014-11-29 TableSummarization/Paper/exp.tex
Added hinge loss, and clarified tree structure.
small changes for length

@@
 \section{Experiments}\label{sec:experiments}
-
-\begin{comment} ask about below:
-Additional experiments:
-Maybe also good to fix mw, and show the performance with or without sampling,
-and with or without a-priori (one rule per phase, for instance).
-Also, would be good to see performance on increasing dataset size.
-\end{comment}
 
@@
 Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.
+\begin{table*} 
+\centering 
+\scriptsize
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\hline 
+\end{tabular} 
+\vspace{-10pt}
+\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
+\vspace{-10pt}
+\end{table*}
+
+\begin{table*} 
+\centering 
+\scriptsize
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\hline 
+\end{tabular} 
+\vspace{-10pt}
+\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
+\vspace{-10pt}
+\end{table*}
+
+\begin{table*} 
+\centering 
+\scriptsize
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
+\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
+\hline 
+\end{tabular} 
+\vspace{-10pt}
+\caption{A Rule expansion \label{table:uiexamplerule}} 
+\vspace{-10pt}
+\end{table*}
 
@@
 
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\vspace{-15pt}
-\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\vspace{-15pt}
-\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\hline 
-\end{tabular} 
-\vspace{-15pt}
-\caption{A Rule expansion \label{table:uiexamplerule}} 
-\end{table*}
-
 \subsubsection{Weighting functions}
@@
 %TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
-\hline 
-\end{tabular} 
-\vspace{-15pt}
-\caption{Bits scoring\label{table:weigtingbitwise}} 
-\end{table*}
-
-\begin{table*} 
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
-\hline 
-\end{tabular} 
-\vspace{-15pt}
-\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
-\end{table*}
 
@@
 \begin{figure}
-\hspace{-15pt}
-  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+%\hspace{-15pt}
+\centering
+  \includegraphics[scale=0.6]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
 \vspace{-5pt}
@@
 The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
+
+\begin{table*} 
+\scriptsize
+\centering 
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
+\hline 
+\end{tabular} 
+\vspace{-10pt}
+\caption{Bits scoring\label{table:weigtingbitwise}} 
+\vspace{-10pt}
+\end{table*}
+
+\begin{table*} 
+\scriptsize
+\centering 
+\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
+\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
+\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
+\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
+\hline 
+\end{tabular}
+\vspace{-10pt}
+\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
+\vspace{-10pt}
+\end{table*}
 

### 2014-11-29 TableSummarization/Paper/intro.tex
small changes for length

@@
 require no pre-materialization, and can be implemented 
-within any relational database system.
+within or on top of any relational database system.
 
-% Drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
+% drill-down (and roll up)~\cite{export:69578} are very useful tools for exploring
 % relational tables.
-% In this paper we study an extension of drill down
+% In this paper we study an extension of drill-down
 % that lets users zoom into the more ``interesting'' parts of a table,
 % with fewer operations and without having to examine as much data
-% as traditional drill down.
-% We call our approach {\em smart drill down};
-% it does not replace traditional drill down but we will argue is a very
+% as traditional drill-down.
+% We call our approach {\em smart drill-down};
+% it does not replace traditional drill-down but we will argue is a very
 % useful complement.
-The best way to explain smart drill down is through a simple example.
+The best way to explain smart drill-down is through a simple example.
 
@@
 If the resulting table has many tuples,
-the analyst can use traditional drill down to explore it.
+the analyst can use traditional drill-down to explore it.
 For instance, the system may initially tell the analyst there are
@@
 
-Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
+Instead, when the analyst uses smart drill-down on Table~\ref{table:introexample0},
 he obtains Table~\ref{table:introexample}.
 The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
-that display noteworthy or interesting drill downs.
+that display noteworthy or interesting drill-downs.
 The number $3$ is a user specified parameter, which we call $k$.
@@
 Each rule summarizes the set of tuples that are described by it.
-Again, instead of Count, the operator can display a sum aggregate, such as
+Again, instead of Count, the operator can display a Sum aggregate, such as
 the total Sales.
@@
 \begin{table}
+\scriptsize
 \centering
@@
 \end{tabular}
-\vspace{-15pt}
+\vspace{-10pt}
 \caption{Initial summary}\label{table:introexample0}
@@
 \end{tabular}
-\vspace{-15pt}
-\caption{Result after first smart drill down}\label{table:introexample}
+\vspace{-10pt}
+\caption{Result after first smart drill-down}\label{table:introexample}
+\vspace{-10pt}
 \end{table}
@@
 \begin{table}
+\scriptsize
 \centering
-\vspace{-10pt}
+\vspace{-5pt}
 \begin{tabular}{| l | l | l | l | l |}
@@
 \end{tabular}
-\vspace{-15pt}
-\caption{Result after second smart drill down} \label{table:introexample2}
+\vspace{-10pt}
+\caption{Result after second smart drill-down} \label{table:introexample2}
+\vspace{-10pt}
 \end{table}
@@
 
-When the analyst clicks on a rule $r$, smart drill down
+When the analyst clicks on a rule $r$, smart drill-down
 expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
-(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
+(We discuss other smart drill-down operations in Section~\ref{sec:interface}.)
 There are three factors that make a rule set interesting.
@@
 We also present an efficient optimization procedure to maximize score, invoked
-by smart drill down to select the set of $k$ rules to display.
+by smart drill-down to select the set of $k$ rules to display.
 
@@
 
-Compared to traditional drill down, our smart drill down has two important advantages:
+Compared to traditional drill-down, our smart drill-down has two important advantages:
 \squishlist
 \item
-Smart drill down limits the information displayed
+Smart drill-down limits the information displayed
 to the most interesting $k$ facts (rules).
-With traditional drill down, a column is expanded and {\em all}
+With traditional drill-down, a column is expanded and {\em all}
 attribute values are displayed in arbitrary order.
-In our example, if we drill down on say the store attribute,
+In our example, if we drill-down on say the store attribute,
 we would see all stores listed, which may be a very large number.
 \item
-Smart drill down explores several attributes to open up together,
+Smart drill-down explores several attributes to open up together,
 and automatically selects combinations that are interesting.
@@
 the rule (Target, bicycles, $\star$, $200$, $2$)
-is obtained after a single drill down;
-with a traditional approach, the analyst would first have to drill down on
-Store, examine the results, drill down on Product,
+is obtained after a single drill-down;
+with a traditional approach, the analyst would first have to drill-down on
+Store, examine the results, drill-down on Product,
 look through all the displayed rules and then find the interesting rule
@@
 
-Incidentally, note that in the example we only described one type of smart drill down,
-where the analyst selects a {\em rule} to drill down on
+Incidentally, note that in the example we only described one type of smart drill-down,
+where the analyst selects a {\em rule} to drill-down on
 (e.g., the Walmart rule going from Table~\ref{table:introexample} to
@@
 
-Our work on smart drill down is related
+Our work on smart drill-down is related
 to table summarization and anomaly
@@
 an interactive environment, where the analyst
-directs the drill down and can tailor the optimization criteria.
+directs the drill-down and can tailor the optimization criteria.
 Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
+and smart drill-down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
 
-Our chief contribution in this paper is the {\em smart drill-down} interaction operator,
+To reiterate, our chief contribution in this paper is the {\em smart drill-down} interaction operator,
 an extension of traditional drill-down, aimed at allowing analysts to zoom into the
@@
 \item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
-and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
+and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill-down operation. b
 \item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
@@
 
-\item In Section~\ref{sec:formal}, we introduce the relevant notation and formally define smart drill down.
-After that, we describe different schemes for weighting rules, and our interactive user interface.
+\item In Section~\ref{sec:formal}, we formally define smart drill-down. After that, we describe different schemes for weighting rules, and our interactive user interface.
 
@@
 
-\item Based on our implemented smart drill down,
+\item Based on our implemented smart drill-down,
 in Section~\ref{sec:experiments} we experimentally evaluate
 performance on real datasets,
-and show additional examples of smart drill down in action.
+and show additional examples of smart drill-down in action.
 

### 2014-11-29 TableSummarization/Paper/formal.tex
Added hinge loss, and clarified tree structure.
small changes for length
length reduction and small changes

@@
 For example, rule ($a$, $\star$) subsumes rule ($a$, $b$).
-Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
+Thus, if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
 We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e.
@@
 \stitle{Rule Lists:}
-A {\em rule-list} is an ordered list of rules returned by our system as a result of a smart drill-down by the user. 
+A {\em rule-list} is an ordered list of rules returned by our system when a smart drill-down operation concludes. 
 When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
@@
 The reason why we separate the scoring into these two portions is
-that they allow us to separate the inherent goodness of a rule from
+that they allow us to separate the inherent ``goodness'' of a rule from
 how much it captures the data in $\calD$.
@@
 $$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
-Overall, our goal is to choose the rule-list that maximizes the 
+Overall, our goal is to choose the rule-list maximizing 
 total score.
@@
 
-Our smart drill downs still display the Count of each rule rather than the $MCount$.
+Our smart drill-downs still display the Count of each rule rather than the $MCount$.
 This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret.
@@
 for one of the following smart drill-down operations:
-\begin{enumerate}
+\begin{denselist}
 \item $[$Rule Drill-Down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
 \item $[$Star Drill-Down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
-\end{enumerate}
+\end{denselist}
 \end{problem}
-Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user.
+\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user.
 We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead.
-We discuss how to modify our algorithms to use $Sum$ instead of $Count$ in the `Extensions' section of the technical report~\cite{tr}.
+We discuss how to modify our algorithms to use $Sum$ instead of $Count$
+\techreporttext{in the `Extensions' section}
+\papertext{in the technical report~\cite{tr}}.
 
@@
 it conveys). Let us consider some canonical forms for the function $W(r)$ and their interpretation; later, we describe the full
-space of weighting functions our techniques can handle:
+family of weighting functions our techniques can handle:
 
@@
 The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$.
-Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$.
+Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$.
 If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$.
-Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$.
+Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$.
 If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
+Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
 
 To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
-Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule.
+Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule.
 For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns).
-Thus this rule contributes $200$ to the score.
+Thus, this rule contributes $200$ to the score.
 For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those tuples.
-Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total.
-Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list.
+Thus, $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total.
+Thus, this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list.
 Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.
@@
 \stitle{Other Weighting Functions:}
-Even though we have given two example weighting functions here, our algorithms allow the user to leverage any weighting function $W$, subject to only two conditions:
+Even though we have given two example weighting functions here, our algorithms allow the user to leverage any weighting function $W$, subject to two conditions:
 \squishlist
@@
 \centering
+\scriptsize
 \begin{tabular}{ | l | c | }
@@
 \end{tabular}
+\vspace{-10pt}
+\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
 \vspace{-15pt}
-\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
 \end{table}
@@
 \end{tabular}
-\vspace{-15pt}
+\vspace{-10pt}
 \caption{Result of clicking on a $\star$ \label{table:introexample3}}
+\vspace{-10pt}
 \end{table}

### 2014-11-29 TableSummarization/Paper/exp.tex
exp release

@@
 \section{Experiments}\label{sec:experiments}
+We have implemented a fully-functional interactive system instrumented with the smart drill-down operator.
+We now describe our experiments on this system with real datasets.
 
-\subsection{Datasets}
-The first dataset, denoted `Marketing', contains demographic information about potential customers.
-The source of the dataset is Impact Resources, Inc., Columbus, OH (1987).
-A total of $N=9409$ questionnaires containing $502$ questions were
-filled out by shopping mall customers in the San Francisco Bay area.
-This dataset is an extract from this survey.
-Each tuple in the data table describes a single person.
-There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on.
-Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.
+\stitle{Datasets.} The first dataset, denoted `Marketing', contains demographic information about potential customers~\cite{dataset1}.
+A total of $N=9409$ questionnaires containing $502$ questions were filled out by shopping mall customers in the San Francisco Bay area.
+This dataset is the summarized result of this survey.
+Each tuple in the table describes a single person.
+There are $14$ columns, each of which is a demographic attribute, such as annual income, gender, marital status, age, education, and so on.
+Continuous values, such as income, have been bucketized in the dataset, and each column has up to $10$ distinct values.
 \begin{table*} 
@@
 \end{table*} 
-
 \begin{table*} 
@@
 \vspace{-10pt}
-\caption{Star expansion on 'Education' Column \label{table:uiexamplestar}} 
+\caption{Star expansion on `Education' Column \label{table:uiexamplestar}} 
 \vspace{-10pt}
 \end{table*} 
-
 \begin{table*} 
@@
 \vspace{-10pt}
-\caption{A Rule expansion \label{table:uiexamplerule}} 
+\caption{A rule expansion \label{table:uiexamplerule}} 
 \vspace{-10pt}
 \end{table*} 
+The columns (in order) are as follows:
+annual household income, gender, marital status, age, education, occupation, time lived in the Bay Area, dual incomes?, persons in household, persons in household under $18$, householder status, type of home, ethnic classification, language most spoken in home.
 
-The columns (in order) are as follows:
-Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, Time lived in Bay Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home.
-
-The second dataset, denoted `Census', is a US 1990 Census dataset from the UCI Machine Learning repository~\cite{uciml}.
-It consists of about $2.5$ million tuples, with each tuple corresponding to a person.
+The second dataset, denoted `Census', is a US 1990 Census dataset from the UCI Machine Learning repository~\cite{uciml},
+consisting of about $2.5$ million tuples, with each tuple corresponding to a person.
 It has $68$ columns, including ancestry, age, and citizenship.
 Numerical columns, such as age, have been bucketized beforehand in the dataset.
-We use this dataset in the experiments in Section~\ref{sec:quantitative_study} in order to study the accuracy and performance of sampling on a large dataset.
+We use this dataset in Section~\ref{sec:quantitative_study} in order to study the accuracy and performance of sampling on a large dataset.
 
-Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page.
-We use our current implementation of our algorithms, which directly outputs tables in latex format.
-We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.
+Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page.
+We use our current implementation of our the smart drill-down operator, which can directly output tables in \LaTeX \ format.
+We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}).
 Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.
@@
 \subsection{Qualitative Study}
-We first perform a qualitative study of our Smart Drill Down system.
+We first perform a qualitative study of smart drill-down.
 We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
@@
 To begin with, the user sees an empty rule with the total number of tuples as the count.
-After this, suppose the user expands the rule.
+Suppose the user expands the rule.
 Then the user will see Table~\ref{table:uiexample1}.
 The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples.
-The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Are for $> 10$ years.
-Note that the latter two rules give very specific information which would require upto $3$ user clicks to find using Data Cube Drill-Downs, whereas our system displays that information to the user with a single click.
+The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Area for $> 10$ years.
+Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill-down, whereas smart drill-down displays that information to the user with a single click.
 
 Now suppose the user decides to further explore the table, by looking at education related information of females in the dataset.
-So the user clicks on the $\star$ in the 'Education' column of the second rule.
-This opens up table~\ref{table:uiexamplestar}.
-It shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
+Say the user clicks on the $\star$ in the `Education' column of the second rule.
+This opens up table~\ref{table:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
 
-Instead of expanding the 'Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.
+Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.
 
 \subsubsection{Weighting functions}
-Our system can display optimal rule summaries using any monotonic weighting function.
-By default, we assign a rule weight equal to its size.
-In this section, we try out a couple of other weighting functions.
+Our system can display optimal rule lists for any monotonic weighting function. By default, we assign a rule weight equal to its size. In this section, we consider other weighting functions.
 
@@
 The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column.
-Thus this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.
+Thus, this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.
 
@@
 The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples.
-We observe that the accuracy and running time of our algorithm depends largely on the $m_w$ and $minSS$ parameters, rather than on the underlying database size.
+We observe that the accuracy and running time of our algorithm depends largely on $m_w$ and $minSS$, rather than the underlying database size.
 
@@
 The algorithm is guaranteed to find the best marginal rule as long as its weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
-In this section, we study the effect of varying $m_w$ on the speed of our algorithm.
-Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.
+We now study the effect of varying $m_w$ on the speed of our algorithm running on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.
 
-We consider a weighting function $W$, and a value of $m_w$.
+We fix a weighting function $W$, and a value of $m_w$.
 For that value of the $W$ and $m_w$ parameters, we find the time taken for expanding the empty rule.
@@
 The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule.
-Thus for size scoring on the Marketing dataset, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
+Thus, for size scoring on the Marketing dataset, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
 For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
@@
 \begin{figure}
-%\hspace{-15pt}
+\vspace{-5pt}
 \centering
   \includegraphics[scale=0.6]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
-\vspace{-5pt}
+\vspace{-10pt}
   \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}

### 2014-11-29 TableSummarization/Paper/algorithms.tex
changes to related, to intro

@@
 
-\begin{algorithm}\label{algo:best-rule-set}
+\begin{algorithm}
 \scriptsize
@@
 \Return $S$
-\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
+\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}\label{algo:best-rule-set}}
 \end{algorithm}
@@
 \begin{definition}\label{def:set-score}
-Let $R$ be a set of rules. Then the Score of $R$ is defined by
+Let $R$ be a set of rules. Then the Score of $R$ is
 $$\text{Score}(R) = \text{Score}(R^{\prime})$$
@@
 
-\begin{algorithm}\label{algo:best-marginal-rule}
+\begin{algorithm}
 \scriptsize
@@
 \Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
-\caption{Find best marginal rule}
+\caption{Find best marginal rule\label{algo:best-marginal-rule}}
 \end{algorithm}
@@
 \begin{lemma}\label{lemma:submodular}
-For a given table $T$, the Score function over sets $S$ of rules, defined by :
+For a given table $T$, the Score function over sets $S$ of rules, defined by the following is submodular:
 $$\text{Score}(S) = \sum_{r \in S} MCount(r,S)W(r)$$
-is submodular.
 \end{lemma}
@@
 as listed below, has desirable approximation guarantees:
+\vspace{-10pt}
 \begin{framed}
-\vspace{-15pt}
-\begin{enumerate}
+\vspace{-5pt}
+\begin{denselist}
 \item Set $S = \phi$
 \item For $i$ from $1$ to $k$
-\begin{enumerate}
+\begin{denselist}
 \item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
 \item $S = S \cup \left\lbrace r \right\rbrace$
-\end{enumerate}
-\end{enumerate}
-\vspace{-15pt}
+\end{denselist}
+\end{denselist}
+\vspace{-5pt}
 \end{framed}
+\vspace{-10pt}
 Since Score is a submodular function of the set $S$, 
@@
 The expensive step in the above procedure is the step where the Score is computed for every
-single rule. Given the number of rules can be very large, this can be an especially time-consuming process.
+single rule. Given the number of rules can be large, this can be an especially time-consuming process.
 

### 2014-11-29 TableSummarization/Paper/intro.tex
changes to related, to intro

@@
 \caption{Initial summary}\label{table:introexample0}
+\vspace{-5pt}
 \end{table}
@@
 In particular, one of the new rules shows that
-Walmart sells a lot of cookies and the others show it sells a lot of products in
+Walmart sells a lot of cookies; the others show it sells a lot of products in
 the regions CA-1 and WA-5.
@@
 There are three factors that make a rule set interesting.
-One is if contains rules with high Count (or total sales) fields,
+One is if it contains rules with high Count (or total sales) fields,
 since the larger the count, the more tuples are summarized.
@@
 seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
-since the former rule tells us the high sales are concentrated in a single region.
+since the former tells us the high sales are concentrated in a single region.
 A third desirability factor is diversity:

### 2014-11-29 TableSummarization/Paper/related.tex
changes to related, to intro

@@
 %TODO: Repeat what we said in intro about interactive, etc.
-There has been work on finding interesting rules in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
-This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max entropy distribution.
+There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
+This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution.
 The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
@@
 There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
-Several existing works also deal with the problem of frequent itemset mining~\cite{apriori, 1411744, Han:2000:MFP:342009.335372}.
-All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
+In both these cases, the emphasis is on completely covering or
+summarizing the table, suffering from the same problems as traditional drill-down in that the user may be presented with
+too many results.
+The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case only work for a restricted number of attributes (in their case, 4).
+\agp{check}
+
+Several existing papers also deal with the problem of frequent itemset mining~\cite{apriori, 1411744, Han:2000:MFP:342009.335372}.
+Vanilla frequent itemset mining is not directly applicable to our problem because the flexible user-specified objective function emphasizes coverage of the table rather than simply frequent itemsets.
+However, we do leverage ideas from the a-priori algorithm~\cite{apriori} as applicable.
+Several extensions have been proposed to the a-priori algorithm, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311, Miller:1997:ARO:253260.253361}.
+We can potentially use these ideas to improve handing of numerical attributes in our work.
+Unlike our paper, there has been no work on dynamically maintaining samples for interaction in the frequent itemset literature, since frequent itemset mining is a one-shot problem.
+
+% All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
 
@@
 
-Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
-Several extensions have been proposed, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311, Miller:1997:ARO:253260.253361}.
-We can potentially use these ideas to improve handing of numerical attributes in our work.
+% Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
 

### 2014-11-29 TableSummarization/Paper/intro.tex
Added proof sketch of NP-hardness

@@
 maintaining samples in memory to support smart drill-down. 
-We show that optimal identification of samples is once again {\sc NP-Hard}, \agp{??} and 
+We show that optimal identification of samples is once again {\sc NP-Hard}, and 
 we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.

### 2014-11-29 TableSummarization/Paper/sampling.tex
Added proof sketch of NP-hardness

@@
 Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
-%For shrink, prefer shrinking samples of the largest size (> minSS) first?
 \end{enumerate}
@@
 Note that multiple nodes of the tree may correspond to the same rule.
+Thus even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering.
 Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded.
@@
 We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
-In the absence of additional information, we can assume a uniform probability distribution.
+In the absence of additional data, we can assume a uniform probability distribution.
 That is, we can assume that every leaf is equally likely to be expanded next.
-We can improve on this by assigning zero probability to leaves whose rule is also displayed in an internal node of the tree.
-We can also use Machine Learning on past user behaviour, along with node features such as `node depth in tree' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.
+If we have data on past user behaviour available, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.
 
@@
 If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
+We can show that Problem~\ref{prob:sample-sizes} is NP-Hard by reducing the knapsack problem to it.
+\begin{lemma}
+Problem~\ref{prob:sample-sizes} is NP-Hard.
+\end{lemma}
+\papertext{
+\begin{proof}(Sketch)
+Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
+We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
+We will reduce this to an instance of Problem~\ref{prob:sample-sizes}.
 
------Hinge-loss begin-----
+We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem.
+For Problem~\ref{prob:sample-sizes}, we set $d$ (branching factor) equal to $2$.
+$capacity$ is set to $(m + W) \times minSS$.
+Tree $U$ has $m$ special nodes $r_1, r_2, ...
+r_m$, and each $r_i$ has two children $r_{i,1}, r_{i,2}$.
+These $2m$ children are all leaves, and all leaves other than these have expansion probability $0$.
+The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$.
+In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
+Thus each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
+In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$.
+Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
+Thus each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
+Thus in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
+The latter option consumes $w_i \times minSS$ and gives and extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
+Thus having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
+Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
+Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$s for which $ess(r_{i,2}) = minSS$ gives a solution to the original knapsack problem.
+\end{proof}}
+\techreporttext{
+\begin{proof}
 
-However, we can make the problem convex (and hence tractable) with two simplifications.
-The first simplification is, we modify our objective function to use hinge-loss instead of a step function.
-That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
-Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$.
-The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers.
-After obtaining our optimal sample sizes, we can round them up to get integer sample sizes.
-This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
-$|U|$ is usually negligible compared to the memory capacity, or $minSS$.
+\end{proof}
+}
 
-In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function).
-Thus our new optimization problem becomes
-\begin{problem}\label{prob:sample-sizes-hinge-loss}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
-choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
-subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
-\end{problem}
-The constraint is linear in the $n_r$ variables, and hence convex.
-Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex.
-The constant function $-1$ is convex as well.
-Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum.
-Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
-
-The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.
-
-----Hinge-loss-end-----
-
-Otherwise, with an additional simplification, we can reduce it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
+However, with an additional simplifying assumption, we can make the problem approximately solvable.
+it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
 The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
@@
 \techreporttext{
+\subsubsection{Alternative Convex-Optimization based solution}
+We noted earlier that Problem~\ref{prob:sample-sizes} is NP-Hard, but can be approximately solved with an additional simplifying assumption regarding the $S(r_1, r_2)$ values.
+Instead of making this simplification, we can make the problem convex (and hence tractable) with two different simplifications.
+The first simplification is, we modify our objective function to use hinge-loss instead of a step function.
+That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
+Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$.
+The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers.
+After obtaining our optimal sample sizes, we can round them up to get integer sample sizes.
+This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
+$|U|$ is usually negligible compared to the memory capacity, or $minSS$.
+
+In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function).
+Thus our new optimization problem becomes
+\begin{problem}\label{prob:sample-sizes-hinge-loss}
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
+subject to :
+$$\sum_{r \in U} n_r < \text{capacity}$$
+\end{problem}
+The constraint is linear in the $n_r$ variables, and hence convex.
+Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex.
+The constant function $-1$ is convex as well.
+Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum.
+Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
+
+The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.
+
 \subsubsection{Additional optimizations}

### 2014-11-30 TableSummarization/Paper/algorithms.tex
Replaced drill-down with drill down everywhere

@@
 
-\section{Smart Drill-Down Algorithms} \label{sec:algorithms}
+\section{Smart drill down Algorithms} \label{sec:algorithms}
 We now describe online algorithms for implementing
-the smart drill-down operator.
+the smart drill down operator.
 We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
@@
 \begin{enumerate}
-\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
+\item $[$Rule drill down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$.
 Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
-\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$.
+\item $[$Star drill down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$.
 In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise.

### 2014-11-30 TableSummarization/Paper/exp.tex
Replaced drill-down with drill down everywhere

@@
 \section{Experiments}\label{sec:experiments}
-We have implemented a fully-functional interactive system instrumented with the smart drill-down operator.
+We have implemented a fully-functional interactive system instrumented with the smart drill down operator.
 We now describe our experiments on this system with real datasets.
@@
 Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page.
-We use our current implementation of our the smart drill-down operator, which can directly output tables in \LaTeX \ format.
+We use our current implementation of our the smart drill down operator, which can directly output tables in \LaTeX \ format.
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}).
@@
 \subsection{Qualitative Study}
-We first perform a qualitative study of smart drill-down.
+We first perform a qualitative study of smart drill down.
 We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
@@
 The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Area for $> 10$ years.
-Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill-down, whereas smart drill-down displays that information to the user with a single click.
+Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill down, whereas smart drill down displays that information to the user with a single click.
 

### 2014-11-30 TableSummarization/Paper/formal.tex
Replaced drill-down with drill down everywhere

@@
 Our objective (formally defined later) is to 
-enable smart drill-downs on this table or on portions of it:
-the result of our drill-downs are lists of {\em rules}. 
+enable smart drill downs on this table or on portions of it:
+the result of our drill downs are lists of {\em rules}. 
 A {\em rule} is a tuple with a value for each column of the table. 
@@
 \stitle{Rule Lists:}
-A {\em rule-list} is an ordered list of rules returned by our system when a smart drill-down operation concludes. 
+A {\em rule-list} is an ordered list of rules returned by our system when a smart drill down operation concludes. 
 When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
@@
 We want to pick rules with a high value of $MCount$ to display to the user
-as part of the smart drill-down result, to increase the coverage of the rule-list.
+as part of the smart drill down result, to increase the coverage of the rule-list.
 
@@
 
-Our smart drill-downs still display the Count of each rule rather than the $MCount$.
+Our smart drill downs still display the Count of each rule rather than the $MCount$.
 This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret.
@@
 $$\sum_{r \in R} W(r) \times MCount(r,R)$$
-for one of the following smart drill-down operations:
+for one of the following smart drill down operations:
 \begin{denselist}
-\item $[$Rule Drill-Down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
-\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
+\item $[$Rule drill down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
+\item $[$Star drill down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
 \end{denselist}
@@
 
-\subsection{Smart Drill-Down Operations}
+\subsection{Smart drill down Operations}
 \label{sec:interface}
 When the user starts using a system equipped with
-the smart drill-down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}.
-At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill-down' on the rule.
+the smart drill down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}.
+At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule.
 Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
@@
 This operation is equivalent to a traditional
-roll-up, but for smart drill-downs instead of traditional drill-downs.
+roll-up, but for smart drill downs instead of traditional drill downs.
 

### 2014-11-30 TableSummarization/Paper/intro.tex
Replaced drill-down with drill down everywhere

@@
 Analysts often use OLAP (Online Analytical Processing) operations
-such as drill-down (and roll-up)~\cite{export:69578} to explore
+such as drill down (and roll-up)~\cite{export:69578} to explore
 relational databases. 
@@
 
-However, there are cases where drill-down is ineffective; 
+However, there are cases where drill down is ineffective; 
 for example, when the number of distinct values
-in a column is large, vanilla drill-down 
+in a column is large, vanilla drill down 
 could easily overwhelm analysts by presenting them with too many 
 results (i.e., aggregates). 
-Further, drill-down only allows us to instantiate values
- one column at a time, instead of allowing simultaneous drill-downs
-on multiple columns---this simultaneous drill-down on multiple columns 
+Further, drill down only allows us to instantiate values
+ one column at a time, instead of allowing simultaneous drill downs
+on multiple columns---this simultaneous drill down on multiple columns 
 could once again suffer from the problem
@@
 that is an extension to a traditional 
-drill-down operator, aimed at providing {\em complementary}
-functionality to drill-down in cases where drill-down is
-ineffective. We call our operator {\em smart drill-down}.
-At a high level, smart drill-down lets analysts zoom into
+drill down operator, aimed at providing {\em complementary}
+functionality to drill down in cases where drill down is
+ineffective. We call our operator {\em smart drill down}.
+At a high level, smart drill down lets analysts zoom into
 the more ``interesting'' parts of a table or a database,
 with fewer operations, and without having to examine as much
-data as traditional drill-down.
+data as traditional drill down.
 Note that our goal is {\em not} to replace traditional 
-drill-down functionality, which we believe is fundamental;
+drill down functionality, which we believe is fundamental;
 instead, our goal is to provide auxiliary functionality 
 which the analyst is free to use whenever they find 
-traditional drill-downs ineffective.
+traditional drill downs ineffective.
 
-In addition to presenting this new operator called smart drill-down, in this paper, 
+In addition to presenting this new operator called smart drill down, in this paper, 
 we also present novel sampling techniques to compute the results for this
@@
 
-% drill-down (and roll up)~\cite{export:69578} are very useful tools for exploring
+% drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
 % relational tables.
-% In this paper we study an extension of drill-down
+% In this paper we study an extension of drill down
 % that lets users zoom into the more ``interesting'' parts of a table,
 % with fewer operations and without having to examine as much data
-% as traditional drill-down.
-% We call our approach {\em smart drill-down};
-% it does not replace traditional drill-down but we will argue is a very
+% as traditional drill down.
+% We call our approach {\em smart drill down};
+% it does not replace traditional drill down but we will argue is a very
 % useful complement.
-The best way to explain smart drill-down is through a simple example.
+The best way to explain smart drill down is through a simple example.
 
@@
 If the resulting table has many tuples,
-the analyst can use traditional drill-down to explore it.
+the analyst can use traditional drill down to explore it.
 For instance, the system may initially tell the analyst there are
@@
 
-Instead, when the analyst uses smart drill-down on Table~\ref{table:introexample0},
+Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
 he obtains Table~\ref{table:introexample}.
 The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
-that display noteworthy or interesting drill-downs.
+that display noteworthy or interesting drill downs.
 The number $3$ is a user specified parameter, which we call $k$.
@@
 \vspace{-10pt}
-\caption{Result after first smart drill-down}\label{table:introexample}
+\caption{Result after first smart drill down}\label{table:introexample}
 \vspace{-10pt}
@@
 \vspace{-10pt}
-\caption{Result after second smart drill-down} \label{table:introexample2}
+\caption{Result after second smart drill down} \label{table:introexample2}
 \vspace{-10pt}
@@
 
-When the analyst clicks on a rule $r$, smart drill-down
+When the analyst clicks on a rule $r$, smart drill down
 expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
-(We discuss other smart drill-down operations in Section~\ref{sec:interface}.)
+(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
 There are three factors that make a rule set interesting.
@@
 We also present an efficient optimization procedure to maximize score, invoked
-by smart drill-down to select the set of $k$ rules to display.
+by smart drill down to select the set of $k$ rules to display.
 
@@
 
-Compared to traditional drill-down, our smart drill-down has two important advantages:
+Compared to traditional drill down, our smart drill down has two important advantages:
 \squishlist
 \item
-Smart drill-down limits the information displayed
+Smart drill down limits the information displayed
 to the most interesting $k$ facts (rules).
-With traditional drill-down, a column is expanded and {\em all}
+With traditional drill down, a column is expanded and {\em all}
 attribute values are displayed in arbitrary order.
-In our example, if we drill-down on say the store attribute,
+In our example, if we drill down on say the store attribute,
 we would see all stores listed, which may be a very large number.
 \item
-Smart drill-down explores several attributes to open up together,
+Smart drill down explores several attributes to open up together,
 and automatically selects combinations that are interesting.
@@
 the rule (Target, bicycles, $\star$, $200$, $2$)
-is obtained after a single drill-down;
-with a traditional approach, the analyst would first have to drill-down on
-Store, examine the results, drill-down on Product,
+is obtained after a single drill down;
+with a traditional approach, the analyst would first have to drill down on
+Store, examine the results, drill down on Product,
 look through all the displayed rules and then find the interesting rule
@@
 
-Incidentally, note that in the example we only described one type of smart drill-down,
-where the analyst selects a {\em rule} to drill-down on
+Incidentally, note that in the example we only described one type of smart drill down,
+where the analyst selects a {\em rule} to drill down on
 (e.g., the Walmart rule going from Table~\ref{table:introexample} to
@@
 
-Our work on smart drill-down is related
+Our work on smart drill down is related
 to table summarization and anomaly
@@
 an interactive environment, where the analyst
-directs the drill-down and can tailor the optimization criteria.
+directs the drill down and can tailor the optimization criteria.
 Nevertheless, one can envision extending traditional
-and smart drill-down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
+and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
 
-To reiterate, our chief contribution in this paper is the {\em smart drill-down} interaction operator,
-an extension of traditional drill-down, aimed at allowing analysts to zoom into the
+To reiterate, our chief contribution in this paper is the {\em smart drill down} interaction operator,
+an extension of traditional drill down, aimed at allowing analysts to zoom into the
 more ``interesting'' parts of a dataset. In addition to this operator,
@@
 \item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
-and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill-down operation. b
+and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. b
 \item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
-maintaining samples in memory to support smart drill-down. 
+maintaining samples in memory to support smart drill down. 
 We show that optimal identification of samples is once again {\sc NP-Hard}, and 
@@
 
-\item In Section~\ref{sec:formal}, we formally define smart drill-down. After that, we describe different schemes for weighting rules, and our interactive user interface.
+\item In Section~\ref{sec:formal}, we formally define smart drill down. After that, we describe different schemes for weighting rules, and our interactive user interface.
 
@@
 
-\item Based on our implemented smart drill-down,
+\item Based on our implemented smart drill down,
 in Section~\ref{sec:experiments} we experimentally evaluate
 performance on real datasets,
-and show additional examples of smart drill-down in action.
+and show additional examples of smart drill down in action.
 

### 2014-11-30 TableSummarization/Paper/related.tex
Replaced drill-down with drill down everywhere

@@
 In both these cases, the emphasis is on completely covering or
-summarizing the table, suffering from the same problems as traditional drill-down in that the user may be presented with
+summarizing the table, suffering from the same problems as traditional drill down in that the user may be presented with
 too many results.

### 2014-11-30 TableSummarization/Paper/formal.tex
small changes

@@
 Thus, if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
-We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e.
-there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not.
 If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.
@@
 This operation is equivalent to a traditional
-roll-up, but for smart drill downs instead of traditional drill downs.
+roll up, but for smart drill downs instead of traditional drill downs.
 

### 2014-11-30 TableSummarization/Paper/intro.tex
small changes

@@
 Analysts often use OLAP (Online Analytical Processing) operations
-such as drill down (and roll-up)~\cite{export:69578} to explore
+such as drill down (and roll up)~\cite{export:69578} to explore
 relational databases. 
@@
 instead, our goal is to provide auxiliary functionality 
-which the analyst is free to use whenever they find 
+which analysts are free to use whenever they find 
 traditional drill downs ineffective.
@@
 \item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
-and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. b
+and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
 \item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
@@
 we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
-% to improve response time on large tables.
 \squishend

### 2014-11-30 TableSummarization/Paper/sampling.tex
small changes

@@
 \begin{problem}\label{prob:sample-sizes}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
 choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
@@
 where the $I$'s are indicator variables, subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
+$$\sum_{r \in U} n_r < M$$
 \end{problem}
@@
 \end{lemma}
-\papertext{
 \begin{proof}(Sketch)
@@
 For Problem~\ref{prob:sample-sizes}, we set $d$ (branching factor) equal to $2$.
-$capacity$ is set to $(m + W) \times minSS$.
+$M$ is set to $(m + W) \times minSS$.
 Tree $U$ has $m$ special nodes $r_1, r_2, ...
@@
 Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$s for which $ess(r_{i,2}) = minSS$ gives a solution to the original knapsack problem.
-\end{proof}}
-\techreporttext{
-\begin{proof}
-
 \end{proof}
-}
 
@@
 This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
-$|U|$ is usually negligible compared to the memory capacity, or $minSS$.
+$|U|$ is usually negligible compared to the memory capacity $M$, or $minSS$.
 
@@
 \begin{problem}\label{prob:sample-sizes-hinge-loss}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
 choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
@@
 subject to :
-$$\sum_{r \in U} n_r < \text{capacity}$$
+$$\sum_{r \in U} n_r \leq M$$
 \end{problem}

### 2014-11-30 TableSummarization/Paper/sampling.tex
more changes

@@
 That is, we can assume that every leaf is equally likely to be expanded next.
-If we have data on past user behaviour available, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `node distance from last expanded node' to get a better probability estimate of each node being expanded next.
+If we have data on past user behaviour available, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `distance from last expanded node' to get a better probability estimate of each node being expanded next.
 
@@
 We can show that Problem~\ref{prob:sample-sizes} is NP-Hard by reducing the knapsack problem to it.
-\begin{lemma}
+\begin{lemma}\label{lemma:sampling-hardness}
 Problem~\ref{prob:sample-sizes} is NP-Hard.
@@
 We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem.
-For Problem~\ref{prob:sample-sizes}, we set $d$ (branching factor) equal to $2$.
-$M$ is set to $(m + W) \times minSS$.
+For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$.
 Tree $U$ has $m$ special nodes $r_1, r_2, ...
@@
 Thus each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
+Thus $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$.
 In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$.
@@
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
-The latter option consumes $w_i \times minSS$ and gives and extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
+The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
 Thus having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
 Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
-Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$s for which $ess(r_{i,2}) = minSS$ gives a solution to the original knapsack problem.
+Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
 \end{proof}
@@
 However, with an additional simplifying assumption, we can make the problem approximately solvable.
-it to the knapsack-like problem, and use a PTAS (Polynomial Time Approximation Scheme) to find the approximately optimal solution.
-The simplification is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
+The assumption is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
-$ess(r)$ is redefined according to the new value of $S$ as well.
-So $$ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$$ where $r$ is the parent of $r^{\prime}$.
+This is similar to what we had for tree $U$ in our proof of Lemma~\ref{lemma:sampling-hardness}.
+$ess(r)$ now becomes $ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$ where $r$ is the parent of $r^{\prime}$.
 
@@
 Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
-This allows us to effectively split the problem into multiple subproblems, one per $M_{r_0}$.
+This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
 Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
@@
 Thus there are at most $3^d$ locally optimal solutions.
-$d$ is usually small, even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
+$d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
 
@@
 $$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
-This can be solved using dynamic programming, in $O(D\mathcal{S}d^2)$ time.
+This can be solved using dynamic programming, in $O(D\mathcal{S}3^d)$ time.
 

### 2014-11-30 TableSummarization/Paper/algorithms.tex
small fixes

@@
 The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score.
-Thus if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.
+Thus, if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.
 
@@
 
-\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:
+\noindent Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
+Or equivalently, we can define Score for a \emph{set} of rules as follows:
 
@@
 $$\text{Score}(R) = \text{Score}(R^{\prime})$$
-where $R^{\prime}$ is the list of rules obtained by ordering the rules in $R$ in decreasing order by weight.
+where $R^{\prime}$ is the list of rules obtained by ordering the rules in the set $R$ in decreasing order by weight.
 \end{definition}
@@
 for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
-The problem consists of finding a set of given size $k$, that maximizes 
+The problem consists of finding a set of rules, given size $k$, that maximizes 
 Score.
@@
 \begin{definition}
-A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
-$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
+A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$: $f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$
 \end{definition}
 Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead.
-For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.
+For monotonic non-negative submodular functions, it is well known that the solution to the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.
 
@@
 
-In order to avoid counting too many rules, we use an idea from the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}.
+In order to avoid counting too many rules, we leverage a technique inspired by the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}.
 Recall that the a-priori algorithm is used to find all frequent itemsets that have a support greater than a threshold.
-Unlike the frequent itemset mining algorithm, our goal is to find the single best marginal rule.
+Unlike the a-priori algorithm, our goal is to find the single best marginal rule.
 Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset.
-We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule.
+
+We compute the best marginal rule over multiple passes on the dataset, with the maximum number of passes equal to the maximum size of a rule.
 In the $j^{th}$ pass, we compute counts and marginal values for rules of size $j$.

### 2014-11-30 TableSummarization/Paper/sampling.tex
small fixes

@@
 describe how we can set the minimum sample size required from the SampleHandler.
+\agp{revisit}
 
@@
 Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
-Thus the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
+Thus, the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
 The SampleHandler also computes $N_s$ when a sample is created.
@@
 Note that multiple nodes of the tree may correspond to the same rule.
-Thus even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering.
+Thus, even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering.
 Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded.
@@
 In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
-Thus each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
-Thus $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$.
+Thus, each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
+Thus, $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$.
 In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$.
 Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
-Thus each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
-Thus in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
+Thus, in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
 The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
-Thus having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
+Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
 Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
@@
 This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
-Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
+Thus, for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
 Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
@@
 \squishend
-Thus there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
+Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
 There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution.
@@
 $$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Thus there are at most $3^d$ locally optimal solutions.
+Thus, there are at most $3^d$ locally optimal solutions.
 $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
@@
 $$i_M < j \Rightarrow n_{r_j} = 0$$
-And thus
+And Thus,
 $$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$
@@
 In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function).
-Thus our new optimization problem becomes
+Thus, our new optimization problem becomes
 \begin{problem}\label{prob:sample-sizes-hinge-loss}
@@
 Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum.
-Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
+Thus, we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.
 
@@
 
-Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
+Thus, a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
 This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule).

### 2014-11-30 TableSummarization/Paper/sampling.tex
sampling reorg

@@
 \subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
+
+Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
+These passes can be rather expensive if the table is large, especially if the table does not fit in main memory.
+If we want exact counts for rules, we have no choice but to read the entire table.
+
+But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
+Thankfully since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us.
+If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
+
+Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
 We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.
+
+\agp{revisit}
 First in Section~\ref{sec:sample-using}, we motivate the need for sampling.
@@
 describe how we can set the minimum sample size required from the SampleHandler.
-\agp{revisit}
-
-\subsubsection{Using samples to respond to user operations}\label{sec:sample-using}
-Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
-These passes can be very expensive if the table is large, especially if the table does not fit in main memory.
-If we want exact counts for rules, we have no choice but to read the entire table.
-
-But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-Thankfully since our goal is to find a representative coverage of the table,
-if we miss out on a few rare tuples, it does not hurt us.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
-
-Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
-Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
-The $minSS$ parameter is the minimum number of sample tuples that are used to determine counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be.
-We later provide a way to find reasonable values of $minSS$.
-
-% Example from situation in a previous table. And good and bad samples to keep there.
-At all points, the SampleHandler maintains a set of samples in memory.
-For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user.
-Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table.
-The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
-The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
-The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
-% TODO: Give example of what samples we'd keep.
-
-Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
-Thus, the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
-The SampleHandler also computes $N_s$ when a sample is created.
-Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
-The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
-In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.
-
-When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule.
-At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample.
-Creating a new sample by making a pass through the table is called \textbf{Create} (further described below).
-At later stages, when there are potentially multiple samples available, there are multiple ways it could return a sample for rule $r$:
-\begin{enumerate}
-\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e.
-$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
-Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
-
-\item \textbf{Combine:} If \textbf{Find} doesn't work i.e.
-if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
-We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
-That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
-
-Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}.
-Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
-In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
-%TODO: Not sure how clear/necessary this is.
-
-\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
-Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples.
-We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
-
-In addition, the SampleHandler needs to return a sample of size equal to $minSS$.
-It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
-Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
-For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
-We describe our algorithms for deciding what sized samples to create, and what samples to shrink, in Section~\ref{sec:sampling_algorithms}.
-
-\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample.
-In that case, the SampleHandler needs to delete or shrink an existing sample.
-It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
-Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
-\end{enumerate}
-
-We describe our algorithm for deciding what samples to shrink, and what sized samples to create, in Section~\ref{sec:sampling_algorithms}.
-
-\textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user.
-If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
-But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$.
-We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
-So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
-That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
-In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
 
 \subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
-We now discuss algorithms for the SampleHandler to decide what sizes of Samples to maintain in memory.
-Briefly, we want to determine sample sizes that maximize the probability that the next user click can be answered using the existing samples, without reading from the hard disk.
-We leverage techniques from approximation algorithms and optimization theory.
 
-At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule.
+We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
+The parameter $minSS$ is the minimum number of sample tuples needed
+such that we can use the sample in memory 
+instead of having to resort to the entire table stored
+on hard disk. 
+This parameter determines how accurate our count estimates will be, and also
+how quickly we can return results to the user. 
+We describe how to set $minSS$ in the technical report.
+
+We now consider the following problem: say we currently have no
+samples in memory (we describe the scenario where there are already 
+some samples in Section~\ref{sec:sample-handler}), and say the user is currently viewing some rules; how do we
+materialize the ``best possible'' samples in memory that fit within the capacity $M$,
+such that we can respond to as many user interactions using the stored samples,
+without having to resort to retrieving the entire table.
+That is, we want to maximize the probability that the next user interaction
+can be answered using the existing samples, without reading from the hard disk. 
+We leverage techniques from approximation algorithms and optimization theory, formalized
+below.
+
+\stitle{Tree of Rules.}
+At any stage, we have a tree $U$ of rules displayed to the user, 
+with each node of the tree corresponding to a displayed rule. 
+In the rest of this section, we will refer to nodes in $U$ and rules interchangeably.
 The tree is formed as follows: The root of the tree corresponds to the trivial rule. 
-And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, ..
-r_k$ being displayed.
-Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, ..
-r_k$.
-Note that multiple nodes of the tree may correspond to the same rule.
-Thus, even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering.
-Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded.
+And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. 
+Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$, and so on.
+
+Even though the rules displayed to the user can be considered a tree, 
+multiple nodes of the tree may correspond to the same rule. 
+For example, $(a, b)$ may be a child of both $(\star, b)$ and $(a, \star)$.
+Thus, even though the structure displayed to the user is a tree, 
+the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering. 
+However, for the purposes of this section, we focus on the tree representation.
+
+Internal nodes of the tree $U$ are ones that have been expanded (drilled down on), 
+while the leaves are nodes that have not been expanded. 
 Let $L$ be the set of leaves. 
 Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
-In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.
 
+\stitle{Probability Distribution.} 
 We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
@@
 
-When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
+% When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
 Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
@@
 
-We choose the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory.
-We define some terms required to formalise this problem next.
+\stitle{Sampling Strategy.} The sampling strategy we adopt involves storing, 
+for every displayed rule $r'$, a sample of $r'$ of size $n_{r'}$, i.e., 
+containing $n_{r'}$ randomly chosen tuples $\in r'$.
+Note that we may choose to not store a sample for some rules, in which case $n_{r'}$
+will be zero for those rules. 
+We need to pick the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory. 
+As it turns out, picking a sample for some rules can help not just that rule, but
+also all of its sub-rules, all the while preserving uniform randomness of samples. We formalize this notion next.
 
+\stitle{Selectivity.} 
 Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule. 
 For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
-Call this quantity $S(r_1, r_2)$.
+We denote this quantity as $S(r_1, r_2)$.
 We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. 
 If the same rule occurs in multiple nodes $r_1$, $r_2$ of the tree, then $S(r_1, r_2)$ is naturally $1$. 
-Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{\prime}$, define the function {\em ess} (for `effective sample size') as below:
-\begin{definition}\label{def:ess}
-$$ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
-\end{definition}
-If $ess(r^{\prime}) \geq minSS$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table.
-We wish to maximize the probability that we can respond to the next user expansion without making another pass.
-We now formally define our problem below:
+Essentially, $S(r_1, r_2)$ denotes how much $r_1$'s sample helps $r_2$. 
+Naturally, if $r_1$ is a super-rule of $r_2$, then $S$ will be $0$ since using a sample of $r_1$
+for $r_2$ will lead to bias.
+
+Then, if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is denoted as $ess(r')$ (for `effective sample size'), as below: 
+\begin{equation} \label{def:ess}
+ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r
+\end{equation}
+Basically, $ess(r')$ captures how much of an unbiased sample of $r'$ can be retrieved using all the samples for the rules in $U$.
+If $ess(r^{\prime}) \geq minSS$, then if the user expands $r^{\prime}$, we do not need to make another 
+pass through the table.
+
+We wish to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
 \begin{problem}\label{prob:sample-sizes}
@@
 choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{ess(r^{\prime}) \geq minSS}$$
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
 where the $I$'s are indicator variables, subject to :
-$$\sum_{r \in U} n_r < M$$
+$\sum_{r \in U} n_r < M$
 \end{problem}
 Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
-If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.
-We can show that Problem~\ref{prob:sample-sizes} is NP-Hard by reducing the knapsack problem to it.
+%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
+We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
 \begin{lemma}\label{lemma:sampling-hardness}
-Problem~\ref{prob:sample-sizes} is NP-Hard.
+Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
 \end{lemma}
@@
 We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
-We will reduce this to an instance of Problem~\ref{prob:sample-sizes}.
+We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.
 
@@
 In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
-Thus, each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
+Therefore, each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
 Thus, $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$.
@@
 Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
-Thus, in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+As a result, in any optimal solution, $ess(r_{i,1})$ must be equal to $minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
@@
 Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
-Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
-Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
+Moreover, the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
+Hence, solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
 \end{proof}
 
-However, with an additional simplifying assumption, we can make the problem approximately solvable.
-The assumption is: For each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
+\stitle{Approximate DP Solution.} Even though the problem as stated is {\sc NP-Hard}, with an additional simplifying assumption, we can make the problem approximately solvable using a Dynamic Programming algorithm.
+The assumption is: for each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
@@
 Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
-Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
+Then, by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
 This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
@@
 Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
-Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
+Then, we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
 We describe both these steps in detail below:
@@
 r_d$.
-For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
+For any child $r_i$, $n_{r_i}$ only contributes to its own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
 r_d$.
@@
 Then the `probability value' of solution $e$ is given by :
-$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
+$P(e) = \sum_{j = 1}^{i_M}p_{j}$,
 and its `Sampling Cost' is
 $$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Thus, there are at most $3^d$ locally optimal solutions.
-$d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
+Thus, there are at most $3^d$ locally optimal solutions; $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
 
@@
 }
+
+\subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
+We now describe the design of the {\em SampleHandler}, which given
+a certain memory capacity $M$, and a minimum sample size $minSS$,
+creates, maintains, retrieves, and removes samples, all in
+response to user interactions on the table.
+It uses the algorithms from Section~\ref{sec:sampling_algorithms} 
+to decide which samples to create, as we will see below.
+
+% Example from situation in a previous table. And good and bad samples to keep there.
+At all points, the SampleHandler maintains a set of samples in memory.
+For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user.
+Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ and (c) a set $T_s$ of tuples from the table.
+The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
+The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
+The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
+% TODO: Give example of what samples we'd keep.
+
+Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
+Thus, the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
+The SampleHandler also computes $N_s$ when a sample is created.
+Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
+The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
+In addition, since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.
+
+When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule.
+At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample.
+Creating a new sample by making a pass through the table is called \textbf{Create} (further described below).
+At later stages, when there are potentially multiple samples available, there are multiple mechanisms it could use to return a sample for rule $r$:
+\squishlist
+\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e.
+$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
+Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
+
+\item \textbf{Combine:} If \textbf{Find} doesn't work i.e., if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
+If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
+We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
+That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
+
+Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}.
+Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
+In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
+%TODO: Not sure how clear/necessary this is.
+
+\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
+Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
+In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples.
+We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
+
+In addition, the SampleHandler needs to return a sample of size equal to $minSS$.
+It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
+Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
+For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
+
+When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
+Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
+Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.
+For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table, using
+the algorithm from Section~\ref{sec:sampling_algorithms}.
+
+\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample.
+In that case, the SampleHandler needs to delete or shrink an existing sample.
+It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
+Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
+\squishend
+
+\smallskip
+\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user.
+If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
+But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$.
+We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
+So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
+That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
+In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.

### 2014-11-30 TableSummarization/Paper/sampling.tex
sampling small changes

@@
 
-\agp{revisit}
-First in Section~\ref{sec:sample-using}, we motivate the need for sampling.
-Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
+We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms}.
+Then, in Section~\ref{sec:sample-handler}, we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
 The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next.
-Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk.
 \papertext{In the technical report,}

### 2014-11-30 TableSummarization/Paper/related.tex
Corrected section 3.5.2

@@
 There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}.
-This work again focuses on displaying rules that will cause the max entropy distribution to best approximate the actual distribution of values.
+This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values.
 
@@
 too many results.
-The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case only work for a restricted number of attributes (in their case, 4).
-\agp{check}
+The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case do not scale well to a large number of attributes (in their case, $\geq 4$).
 

### 2014-11-30 TableSummarization/Paper/sampling.tex
Corrected section 3.5.2

@@
 
-But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
+But if we are willing to accept approximate counts rather than exact counts, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
 Thankfully since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us.
@@
 materialize the ``best possible'' samples in memory that fit within the capacity $M$,
-such that we can respond to as many user interactions using the stored samples,
+such that we can respond to as many user interactions as possible using the stored samples,
 without having to resort to retrieving the entire table.
@@
 Essentially, $S(r_1, r_2)$ denotes how much $r_1$'s sample helps $r_2$. 
-Naturally, if $r_1$ is a super-rule of $r_2$, then $S$ will be $0$ since using a sample of $r_1$
+If $r_1$ is a strict super-rule of $r_2$, then $S$ will be $0$ since using a sample of $r_1$
 for $r_2$ will lead to bias.
 
-Then, if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is denoted as $ess(r')$ (for `effective sample size'), as below: 
+Then, if we have an $n_r$ sized uniformly random sample of tuples covered by $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is denoted as $ess(r')$ (for `effective sample size'), as below:
 \begin{equation} \label{def:ess}
@@
 where the $I$'s are indicator variables, subject to :
-$\sum_{r \in U} n_r < M$
+$\sum_{r \in U} n_r \leq M$
 \end{problem}
@@
 Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
-As a result, in any optimal solution, $ess(r_{i,1})$ must be equal to $minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+As a result, in any optimal solution, $ess(r_{i,1})$ must be equal to $minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
@@
 \stitle{Approximate DP Solution.} Even though the problem as stated is {\sc NP-Hard}, with an additional simplifying assumption, we can make the problem approximately solvable using a Dynamic Programming algorithm.
-The assumption is: for each $r \in L$, we assume that it will get tuples only from samples obtained for itself (filter $= r$) and its immediate parent.
+The assumption is: for each $r \in L$, we assume that it will get tuples only from samples obtained for itself and its immediate parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
@@
 Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
-Then, by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$.
+Then, by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the $ess$ value of rules in $M_{r_0}$.
 This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
@@
 \squishend
-Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
+Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess = minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
 There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution.
@@
 Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used.
-In addition, creating a new sample requires memory, and since memory capacity is limited (given by parameter $M$), it may necessitate shrinking or deleting some existing samples.
 We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.
 
-In addition, the SampleHandler needs to return a sample of size equal to $minSS$.
-It can create a new sample of $< minSS$ size if there are additional tuples covered by $r$ present in existing samples, or $> minSS$ (if enough memory is available).
-Making a larger sample is advantageous not only to get higher accuracy, but also because, when the user later drills down on a sub-rule $r^{\prime}$ of $r$, having a large $r$ sample increases the chance that the \textbf{Combine} strategy will work for $r^{\prime}$, which can let us avoid making another expensive pass through the table.
-For example, if $minSS = 500$, but we get a size $2000$ sample $s$ for the empty rule, then when the user clicks on one of its sub-rules, say $r$, there is a good chance the $2000$ tuples from $T_s$ contain at least $500$ tuples covered by $r$ and that allows us to display the rule-list expanding $r$ quickly instead of making another pass through the table.
-
-When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory.
-Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
-Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples.
-For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table, using
-the algorithm from Section~\ref{sec:sampling_algorithms}.
-
-\item \textbf{Delete/Shrink:} When the SampleHandler uses \textbf{Create}, it may not have enough memory left to create a new sample.
-In that case, the SampleHandler needs to delete or shrink an existing sample.
-It could use a heuristic, like a LRU (Least-Recently-Used) policy to decide which sample to delete or shrink.
-Note that shrinking a sample by randomly deleting some tuples (even if the sample size goes below $minSS$) may be better than deleting it entirely, because in future, we may be able to recreate a sample using \textbf{Combine} on the existing shrinked sample along with other samples.
+When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to access the hard disk to make a pass through the entire table.
+Since accessing the hard disk and making a pass through the entire table is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass.
+Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine the new optimal allocation of memory $n_r$ for each displayed rule $r$.
+Then in a single pass, it creates a sample of size $n_r$ for each displayed $r$.
 \squishend
@@
 We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples.
-So while the user is reading the current rule-list obtained from drilling down on $r$, we can start making a pass through the table to create a bigger sample of $r$ in the background.
-That way, when the user expands $r^{\prime}$, some of the newly loaded tuples for $r$ will also be covered by $r^{\prime}$, increasing the chance that we can use \textbf{Combine} on $r^{\prime}$, and reduce our response time.
+So while the user is busy reading the current rule-list obtained from drilling down on $r$, we can start running the algorithm from Section~\ref{sec:sampling_algorithms} in the background, and then making a pass through the table to create a new samples.
+That way, when the user expands the next rule $r^{\prime}$, there will be a high chance of a sample being pre-fetched for $r^{\prime}$, increasing the chance that we can use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ and reducing our response time.
 In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.

### 2015-02-10 TableSummarization/Paper/algorithms.tex
Changes to paper

@@
 We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
-We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}.
+\papertext{We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}}\techreporttext{We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}}.
 
@@
 
+\techreporttext{
 \begin{algorithm}
@@
 \end{algorithm}
+}
 
@@
 \end{lemma}
-The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}.
+\papertext{The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}}\techreporttext{The proof of this lemma, as well as other proofs, can be found in the appendix}.
 
@@
 
-%TODO: Give  better name to `Greedy Algorithm', and use that name to reference in later as well, in sampling/experimental/extension sections. 
 \stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
 as listed below, has desirable approximation guarantees:
-\vspace{-10pt}
+\vspace{-5pt}
 \begin{framed}
@@
 \end{framed}
-\vspace{-10pt}
+\vspace{-5pt}
 Since Score is a submodular function of the set $S$, 
@@
 
-\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
-The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
+\stitle{Parametrized Algorithm:} \techreporttext{Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.} \papertext{We describe our algorithm in detail below.
+The pseudo-code for our algorithm is in the technical report~\cite{tr}.} We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et).
+BRS takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
 
@@
 The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
-Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, Algorithm~\ref{algo:best-rule-set} is guaranteed to return $S_o$.
-On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by Algorithm~\ref{algo:best-rule-set} does not contain $r$.
-Algorithm~\ref{algo:best-rule-set} runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
+Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$.
+On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$.
+BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$.
 In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small.
@@
 
-Algorithm~\ref{algo:best-rule-set} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
+BRS initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step.
 To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.
@@
 We now describe the function to find the best marginal rule.
-The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.
-The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
-The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration.
+\techreporttext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.} \papertext{The pseudo-code for the function is in the technical report.} The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
+The function makes several iterations \techreporttext{(Step $3$)}, counting marginal values for size $j$ rules in the $j^{th}$ iteration.
 We maintain three sets of rules.
@@
 We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$.
-Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2.
+Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$\techreporttext{, as shown in Step 3.3.2}.
 Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule.
-Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5.
+Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$\techreporttext{, as shown in Step 3.5}.
 If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

### 2015-02-10 TableSummarization/Paper/exp.tex
small changes
Changes to paper

@@
 \section{Experiments}\label{sec:experiments}
-We have implemented a fully-functional interactive system instrumented with the smart drill down operator.
+We have implemented a fully-functional interactive system instrumented with the smart drill down operator, with a web interface.
 We now describe our experiments on this system with real datasets.
@@
 Continuous values, such as income, have been bucketized in the dataset, and each column has up to $10$ distinct values.
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
+\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
+\end{figure}
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
+\vspace{-10pt}
+\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
+\vspace{-15pt}
+\end{figure}
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
+\vspace{-10pt}
+\caption{A rule expansion \label{fig:uiexamplerule}}
+\vspace{-15pt}
+\end{figure}
+
+\begin{comment}
 \begin{table*}
@@
 \end{table*} 
+\end{comment}
+
 The columns (in order) are as follows:
@@
 Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page.
-We use our current implementation of our the smart drill down operator, which can directly output tables in \LaTeX \ format.
+We use the current implementation of our the smart drill down operator, and insert cropped screenshots of its output in this paper.
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}).
@@
 Suppose the user expands the rule.
-Then the user will see Table~\ref{table:uiexample1}.
+Then the user will see Figure~\ref{fig:uiexample1}.
 The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples.
@@
 Say the user clicks on the $\star$ in the `Education' column of the second rule.
-This opens up table~\ref{table:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
+This opens up Figure~\ref{fig:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
 
-Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed table~\ref{table:uiexamplerule}.
+Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed Figure~\ref{fig:uiexamplerule}.
 
+\vspace{-3pt}
 \subsubsection{Weighting functions}
@@
 This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values.
-The rule summary for this weighting is in table~\ref{table:weigtingbitwise} (contrast with Table~\ref{table:uiexample1}).
+The rule summary for this weighting is in Figure~\ref{fig:weigtingbitwise} (contrast with Figure~\ref{fig:uiexample1}).
 The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column.
-Thus, this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Table~\ref{table:uiexample1}.
+Thus, this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Figure~\ref{fig:uiexample1}.
 
@@
 $$W(r) = \text{Min}(0, \text{Size}(r) - 1)$$
-This gives us table~\ref{table:weightingsizeminusone}.
+This gives us Figure~\ref{fig:weightingsizeminusone}.
 This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values.
-As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Table~\ref{table:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
+As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Figure~\ref{fig:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
 
@@
 % TODO: Add the experiment where you compare to vanilla a priori.
-
+\vspace{-3pt}
 \subsection{Quantitative Study}\label{sec:quantitative_study}
@@
 \subsubsection{Effects of $m_w$}
-Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$.
+Our algorithm for finding the best marginal rule takes an input parameter called $m_w$.
 The algorithm is guaranteed to find the best marginal rule as long as its weight is $\leq m_w$, but runs faster for smaller values of $m_w$.
@@
 The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule.
-Thus, for size scoring on the Marketing dataset, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$.
-For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
+Thus, for size scoring on the Marketing dataset, according to Figure~\ref{fig:uiexample1}, we require $m_w \geq 3$.
+For the second weighting function, according to Figure~\ref{fig:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
 At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively.
@@
 \centering
-  \includegraphics[scale=0.6]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+  \includegraphics[scale=0.46]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
 \vspace{-10pt}
@@
 We now study the effects of the sampling parameter $minSS$.
-This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.
+This parameter tells the SampleHandler the minimum sample size on which we run BRS.
 Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.
@@
 The computation time increases approximately linearly with parameter $minSS$.
-This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample.
+This is expected, because BRS makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample.
 The running time at very small values of $minSS$ is simply the time taken to read the table and create the sample.
+This also increases linearly with table size, and is only about $2.5$ seconds for a table with $2.5$ million tuples.
+The running time when $minSS$ equals table size, is the time we would require without sampling.
+Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
 The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
 
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/bits_scoring.jpg}
+\caption{Bits scoring\label{fig:weigtingbitwise}}
+\end{figure}
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/min2_scoring.jpg}
+\vspace{-10pt}
+\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
+\vspace{-15pt}
+\end{figure}
+
+\begin{comment}
 \begin{table*} 
@@
 \end{table*}  
+\end{comment}
 
@@
 \end{figure*}
-
-\begin{comment}
-more experiment stuff
-vanilla a priori size scoring, first 7 columns (maybe we should include 13, to show suffering on english rule
-Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-$\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $1$ \\
-Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
-$\star$ & Married & $\star$ & $\star$ & $\star$ & > 10 years & $2053$ & $2$ \\
-Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-Female & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1966$ & $2$ \\
-$\star$ & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $1964$ & $2$ \\
-\end{comment}

### 2015-02-10 TableSummarization/Paper/formal.tex
Changes to paper

@@
 
-We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values.
-For example, rule ($a$, $\star$) subsumes rule ($a$, $b$).
-Thus, if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
-If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.
+We say that rule $r_1$ is a {\em sub-rule} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values.
+For example, rule ($a$, $\star$) is a sub-rule of ($a$, $b$).
+If $r_1$ is a sub-rule of $r_2$, then we also say that $r_2$ is a {\em super-rule} of $r_1$.
+If $r_1$ is a sub-rule of $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.
 
 \stitle{Rule Lists:}
-A {\em rule-list} is an ordered list of rules returned by our system when a smart drill down operation concludes. 
-When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
+A {\em rule-list} is an ordered list of rules returned by our system in response to a smart drill down operation. 
+When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a new rule-list below $r$.
 For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule.
@@
 \stitle{Size Weighting Function:} 
-$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
+\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
 Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule.
@@
 If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).
+Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).}
 
-To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
+\papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
+Here we set weight equal to the number of non-starred values in the rule $r$ i.e.
+the {\em size} of the rule.
+For example, in Table~\ref{table:introexample3}, the rule (Target, bicycles, $\star$) has weight $2$.
+To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules.
+We start with a blank value in every column of every table tuple.
+Then for each displayed rule $r$, we fill each of its values in the corresponding column of $MCount(r)$ tuples.
+We do this for each rule, filling a total of $\sum_{r} MCount(r)W(r)$ values.
+Thus maximising score leads to choosing the set of rules that will pre-fill as much of the table as possible.}
+
+\techreporttext{To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
 Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule.
@@
 Thus, this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list.
-Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.
+Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.}
 
@@
 $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
-Like the previous scoring function, this one adds some weight for every non-starred value in a rule.
-But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the inherent complexity of the column.
-The reason behind this variation is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
+This scoring function weighs each column depending on its inherent complexity, instead of equally like the Size function.
+The reason behind is: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values.
 Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$.
@@
 The interpretation for this function is similar to the one for the last function.
-Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules.
-But this time, we count the number of `bits' that are pre-filled.
-For a column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information.
-A non-starred valued in a rule $r$ thus pre-fills $MCount(r,R) \lceil \text{log}(|c|) \rceil$ bits of the table.
-Hence this scoring function gives us the the number of bits of the table that are pre-filled by the rule set.
-
-Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
+Except that this time, we count the number of `bits' of the table that are pre-filled by the rules.
+This is because for column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information.
+This scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
 If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.
@@
 
+\techreporttext{
 \begin{table}
@@
 \end{table}
+}
 
@@
 When the user starts using a system equipped with
-the smart drill down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}.
+the smart drill down operator, they first see a table with a single trivial rule as shown in Table~\ref{table:introexample0}.
 At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule.
@@
 By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}.
-Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}.
+As an example of this operation, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
+Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.
 The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
@@
 This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
-This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
-
+\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.} This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
 Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e.
 collapses it.
-For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.
+\techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}.
 This operation is equivalent to a traditional
-roll up, but for smart drill downs instead of traditional drill downs.
+roll up, but for smart drill downs instead of traditional drill downs.}
 
+\techreporttext{
 \begin{table}
@@
 \end{table}
+}

### 2015-02-10 TableSummarization/Paper/intro.tex
Changes to paper

@@
 
-In addition to presenting this new operator called smart drill down, in this paper, 
-we also present novel sampling techniques to compute the results for this
+In addition to presenting this new operator called smart drill down, we also present novel sampling techniques to compute the results for this
 operator {\em in an interactive fashion} on increasingly larger databases. 

### 2015-02-10 TableSummarization/Paper/related.tex
Changes to paper

@@
 \section{Related Work}\label{sec:related}
-%TODO: Repeat what we said in intro about interactive, etc.
+
 There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
@@
 The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
+Some other data exploration work~\cite{sellam:meet} focuses on finding attribute values that divide the database in equal sized parts, while we focus on values that occur as frequently as possible.
 
@@
 In contrast, we keep updating our samples on the fly, as the user interacts with our system.
-
-% Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.
-
-% Related Work summary : Is ours the only only that is flexible (due to weight function black box)?
-% Explanation tables takes lots of papers and talks about them in one go in related work.
-saying they optimize for different opbjective functions.
-We can say that as well.
-We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity.
-% Succintly summarizing with itemsets-- Mampaey:2011:TMI:2020408.2020499 -- Also seemed to optimize according to max entropy criterion
-% Efficient and Effective... Tableaux -- Golab_efficientand -- Similar to conditional functional dep. paper.
-% Near Optimal Tableaux Conditional functional dep.
--- Golab:2008:GNT:1453856.1453900 -- Finding a smallest sized tableau that provides some confidence and support for a conditional functional dependency.
-% MDL Summarization with holes -- Bu:2005:MSH:1083592.1083644 -- Min sized summary with a rectangle and exceptions (holes). 
-% Generalized MDL Approach -- Lakshmanan:2002:GMA:1287369.1287435 -- Similar to MDL Approach.
-% Overlapped Hyperrectangle -- Xiang_succinctsummarization -- Efficiently find the set of hyperrectangles that covers the data (sets of itemsets)
-% Tiling Databases Geerts04tilingdatabases Simlar to hyper-rectangles. Probabilistic interpretation of frequent itemsets.
+There is work on using weighted sampling~\cite{6691587} to create samples favouring data that is of interest to a user, based on the user's history.
+In contrast, we create samples at run time in response to the user's commands.

### 2015-02-10 TableSummarization/Paper/sampling.tex
Changes to paper

@@
 
+\papertext{
+BRS makes multiple passes over the table in order to determine the best set of rules to display.
+This can be really slow when the table is too large to fit in main memory.
+We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance.
+If we had obtained a sample $s$ by selecting each table tuple with probability $p$, and run BRS on $s$, then we multiply the count of each rule found by BRS, by $\frac{1}{p}$ to estimate its count over the full table.
+
+In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples.
+We show that the problem is NP-Hard, and mention an approximate solution using an additional assumption.
+In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process.
+Detailed descriptions of the sampling problem and the SampleHandler are in the technical report.
+The technical report also contains some additional optimizations, and ways to set the minimum sample size.
+
+\subsubsection{Deciding what to sample}\label{sec:sampling_algorithms}
+We are given a memory capacity $M$, and a minimum sample size $minSS$, both specified by the user.
+$minSS$ is the minimum number of tuples on which we are allowed to run BRS, without accessing the hard disk.
+A higher value of $minSS$ increases both accuracy and computation cost of our system.
+
+At any point, we have a tree $U$ of rules displayed to the user.
+Initially, the tree consists of a single node corresponding to the empty rule.
+When the user drills down on a rule $r$, the sub-rules of $r$ obtained by running BRS are added as children of node $r$.
+Our system maintains multiple samples in memory, with upto one sample per rule in $U$.
+Specifically, for each rule $r \in U$, we choose an integer $n_r$, and create a sample $s_r$ consisting of $n_r$ uniformly randomly chosen tuples covered by $r$ from the table.
+Because of the memory constraint, we must have $\sum_{r\in U} n_r \leq M$.
+
+When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$.
+If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down.
+If not, then we need to access the hard disk to obtain more tuples covered by $r$.
+
+Leaves of tree $U$ correspond to rules that may be drilled down on next.
+We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next.
+This can be a uniform distribution, or a machine learned distribution using past user data.
+We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing the hard disk.
+
+If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$.
+We denote the ratio of selectivities $x$ by $S(r',r)$.
+$S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$.
+If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$.
+We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$.
+We now formally define our problem:
+\begin{problem}\label{prob:sample-sizes}
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
+where the $I$'s are indicator variables, subject to :
+$\sum_{r \in U} n_r \leq M$
+\end{problem}
+Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
+%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
+We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
+\begin{lemma}\label{lemma:sampling-hardness}
+Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
+\end{lemma}
+\begin{proof}(Sketch)
+Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
+We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
+We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.
+
+We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$.
+For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$.
+Tree $U$ has $m$ special nodes $r_1, r_2, ...
+r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$.
+All other leaves have expansion probability $0$.
+The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$.
+In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes.
+Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$.
+In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$.
+Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
+So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$.
+Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
+Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem.
+Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
+\end{proof}
+A more detailed proof sketch is given in the technical report.
+}
+
+\techreporttext{
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
-These passes can be rather expensive if the table is large, especially if the table does not fit in main memory.
+These passes can be expensive if the table is large and the table does not fit in main memory.
 If we want exact counts for rules, we have no choice but to read the entire table.
@@
 The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next.
-\papertext{In the technical report,}
-\techreporttext{Finally,}
-we mention some additional optimizations we can make, and
+Finally, we mention some additional optimizations we can make, and
 describe how we can set the minimum sample size required from the SampleHandler.
@@
 how quickly we can return results to the user. 
-We describe how to set $minSS$ in the technical report.
+\papertext{We describe how to set $minSS$ in the technical report.}
 
@@
 
-We wish to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
+We wish to set the sample size $n_r$ of each rule so as to maximize the probability that we can respond to the next user expansion without making another pass.
+We now formally define our problem below:
 \begin{problem}\label{prob:sample-sizes}
@@
 \end{proof}
+}
 
 \stitle{Approximate DP Solution.} Even though the problem as stated is {\sc NP-Hard}, with an additional simplifying assumption, we can make the problem approximately solvable using a Dynamic Programming algorithm.
-The assumption is: for each $r \in L$, we assume that it will get tuples only from samples obtained for itself and its immediate parent.
+The assumption is: for each $r \in L$, we assume that its $ess$ can get sample tuples only from samples obtained for itself and its immediate parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
@@
 
-Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
+\papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
+\squishlist
+\item \textbf{Locally Optimal Solutions :} For each rule $r_0 \in U \setminus L$, let $M_{r_0}$ be the set consisting of $r_0$ and its leaf children.
+We consider the number of tuples assigned to members of $M_{r_0}$.
+Each such assignment of numbers has a `cost' equal to total number of tuples assigned, and `value' equal to sum of probabilities of $M_{r_0}$ members with $ess > minSS$.
+We consider all `locally optimal' assignments, where probability value cannot be increased without increasing the cost.
+\item \textbf{Combining Solutions :} Then, we use dynamic programming to pick one optimal solution per $r_0$ so as to get maximum probability value with the given cost constraint.
+We discretize costs, say to have granularity $100$.
+Let $M_{0}, M_{1}, ...
+M_{D}$ be the sets for which we found locally optimal solutions.
+We make $D$ iterations in all, where in the $i^{th}$ iteration, we try different locally optimal solutions for $M_i$ and use it to update the maximum attainable value for each cost.
+\squishend
+}
+
+\techreporttext{Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
 Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
@@
 
-\begin{comment}
-Let $r_0 \in U \setminus L$.
-Let $d$ be the number of leaf children of $r_0$.
-Let the children be $r_1, r_2, ...
-r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$.
-Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
-$$1 \leq i_m \leq i_M \leq d$$
-$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
-$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
-$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-$$i_M < j \Rightarrow n_{r_j} = 0$$
-And Thus,
-$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$
-
-Call this locally optimal solution $e$. Then we define its `probability value' to be
-$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
-and its `Sampling Cost' to be
-$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Since there are at most $d^2$ possible values of $i_m$ and $i_M$, we can try out all of them and obtain all locally optimal solutions. 
-\end{comment}
-
 Then next step is to combine the solutions using dynamic programming.
@@
 This can be solved using dynamic programming, in $O(D\mathcal{S}3^d)$ time. 
+}
 
@@
 \subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
+\papertext{
+We now briefly describe the design of the {\em SampleHandler}, which given
+a certain memory capacity $M$, and a minimum sample size $minSS$,
+creates, maintains, retrieves, and removes samples, all in
+response to user interactions on the table. A detailed description, with some potential optimizations, can be found in the technical report.
+
+At all points, the SampleHandler maintains a set of samples in memory.
+Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ and (c) a set $T_s$ of tuples from the table.
+The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
+The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
+The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
+
+Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$.
+The $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$.
+The SampleHandler also computes $N_s$ when a sample is created.
+Then we run BRS on sample $s$ to obtain the list of rules to display.
+The SampleHandler uses one of the mechanisms below to obtain such a sample $s$:
+
+\squishlist
+\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e.
+$f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
+
+\item \textbf{Combine:} If \textbf{Find} doesn't work i.e., if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
+If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
+
+\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
+Since accessing the hard disk and making a pass through the entire table is a bottleneck, it also creates samples for rules other than $r$ and augments existing samples in the same pass.
+It first uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine how many tuples to sample for each rule, and uses reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to create a uniformly random sample of a given size for each rule.
+\squishend
+}
+
+\techreporttext{
 We now describe the design of the {\em SampleHandler}, which given
@@
 
-% Example from situation in a previous table. And good and bad samples to keep there.
 At all points, the SampleHandler maintains a set of samples in memory.
@@
 The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
-% TODO: Give example of what samples we'd keep.
 
@@
 The SampleHandler also computes $N_s$ when a sample is created.
-Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
+Then we run BRS on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display.
 The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table.
@@
 $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.
-Algorithm~\ref{algo:best-rule-set} can then be run on $s$.
+BRS can then be run on $s$.
 
@@
 
-Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by Algorithm~\ref{algo:best-rule-set}.
-Since all the tuples in the `new' sample are already present in existing samples, it can give Algorithm~\ref{algo:best-rule-set} a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by Algorithm~\ref{algo:best-rule-set}.
-In contrast, if we had created a new sample from hard disk, we would maintain the sample even after Algorithm~\ref{algo:best-rule-set} terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
+Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by BRS.
+Since all the tuples in the `new' sample are already present in existing samples, it can give BRS a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by BRS.
+In contrast, if we had created a new sample from hard disk, we would maintain the sample even after BRS terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
 %TODO: Not sure how clear/necessary this is.
@@
 \smallskip
-\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the Algorithm~\ref{algo:best-rule-set}, and display a rule-list to the user.
+\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the BRS, and display a rule-list to the user.
 If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
@@
 In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
+}

### 2015-02-10 TableSummarization/src/dataextraction/Marketing.java

@@ -224,7 +224,7 @@ public class Marketing {
 		TableInfo fullTable = Marketing.parseData();
 		//out.println(System.currentTimeMillis() - timer);
 		List<Integer> columns = new ArrayList<Integer>();
-		final Integer firstNumColumns = 9;
+		final Integer firstNumColumns = 7;//9;
 		for (int i = 1; i < firstNumColumns; i++) {
 			columns.add(i);
 		}

### 2015-02-12 TableSummarization/Paper/TableSummarization.tex
length reduction and small changes
changes to related, to intro
small changes
Changes to paper
Changed image font size, and some other changes

@@
-\documentclass{vldb}
+\documentclass{sig-alternate}
 \usepackage{graphicx}
@@
 \usepackage[normalem]{ulem}
+
+\usepackage[font={small,it}]{caption}
 
@@
 
-\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
-\newcommand{\mrj}[1]{\textcolor{green}{#1}}
-\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}
+\newcommand{\agp}[1]{\textcolor{green}{Aditya: #1}}
+\newcommand{\mrj}[1]{\textcolor{red}{#1}}
+\newcommand{\mrjdel}[1]{\textcolor{red}{\sout{#1}}}
 
@@
 \newcommand{\calD}{\mathcal{D}\xspace}
+
+\newenvironment{denselist}{
+    \begin{list}{\small{$\bullet$}}%
+    {\setlength{\itemsep}{0ex} \setlength{\topsep}{0ex}
+    \setlength{\parsep}{0pt} \setlength{\itemindent}{0pt}
+    \setlength{\leftmargin}{1.5em}
+    \setlength{\partopsep}{0pt}}}%
+    {\end{list}}
 
@@
 algorithm for finding the approximately optimal list of rules to display when the user uses a smart drill-down, and a dynamic sampling scheme for efficiently interacting with large tables. Finally, we
-perform experiments on real datasets to demonstrate the usefulness of smart drill-down and study the performance of our algorithms.
+perform experiments on real datasets on our experimental prototype to demonstrate the usefulness of smart drill-down and study the performance of our algorithms.
 \end{abstract}
@@
 
+\balance
 {\small

### 2015-02-12 TableSummarization/Paper/conc.tex
Changes to paper
Changed image font size, and some other changes

@@
 
-Our experimental results show that smart drill down
+Our experimental results on our experimental prototype show that smart drill down
 is fast enough to be interactive under various
@@
 when sampling is used, and when the maximum weight ($m_w$)
-approximation is used.
+approximation is used. Moreover, we have a tunable parameter $minSS$ that the user can tweak to tradeoff performance of smart drill down for the accuracy of the rules.

### 2015-02-12 TableSummarization/Paper/exp.tex
Changed image font size, and some other changes

@@
 \caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
+\end{figure}
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=90mm]{graphs/screenshots6col/age_drilldown.jpg}
+\caption{A regular drill down on Age\label{fig:drilldownexample}}
 \end{figure}
@@
 We first perform a qualitative study of smart drill down.
-We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
+We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
 \subsubsection{User interface testing}
@@
 This opens up Figure~\ref{fig:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
-
 Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed Figure~\ref{fig:uiexamplerule}.
@@
 As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Figure~\ref{fig:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
+
+A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules.
+We use this to perform a drill down on the `Age' column using our experimental prototype.
+The result is shown in Figure~\ref{fig:drilldownexample}.
+We can contrast it with Figure~\ref{fig:uiexample1}; the latter gives information about multiple columns at once and only displays high count values.
+Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.
 

### 2015-02-12 TableSummarization/Paper/sampling.tex
Changed image font size, and some other changes

@@
 BRS makes multiple passes over the table in order to determine the best set of rules to display.
-This can be really slow when the table is too large to fit in main memory.
+This can be slow when the table is too large to fit in main memory.
 We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance.
@@
 When the user drills down on a rule $r$, the sub-rules of $r$ obtained by running BRS are added as children of node $r$.
-Our system maintains multiple samples in memory, with upto one sample per rule in $U$.
+Our system maintains multiple samples in memory, with one sample per rule in $U$.
 Specifically, for each rule $r \in U$, we choose an integer $n_r$, and create a sample $s_r$ consisting of $n_r$ uniformly randomly chosen tuples covered by $r$ from the table.
@@
 \item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
-Since accessing the hard disk and making a pass through the entire table is a bottleneck, it also creates samples for rules other than $r$ and augments existing samples in the same pass.
-It first uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine how many tuples to sample for each rule, and uses reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to create a uniformly random sample of a given size for each rule.
+Since accessing the hard disk is expensive, it also creates samples for rules other than $r$ and augments existing samples in the same pass.
+It first deletes all existing samples, then uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine how many tuples to sample for each rule, and uses reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to create a uniformly random sample of a given size for each rule.
 \squishend

### 2015-02-17 TableSummarization/Paper/algorithms.tex
Changes to paper

@@
 
-Note that the reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently.
+The reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently.
 This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.

### 2015-02-17 TableSummarization/Paper/exp.tex
Changes to paper

@@
 \caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
+\vspace{-15pt}
 \end{figure}
@@
 
-\subsection{Qualitative Study}
+\subsection{Qualitative Study}\label{sec:exp_qualitative}
 We first perform a qualitative study of smart drill down.
@@
 The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples.
-We observe that the accuracy and running time of our algorithm depends largely on $m_w$ and $minSS$, rather than the underlying database size.
+The accuracy of our algorithm depends on $m_w$ and $minSS$, rather than the underlying database size.
+The worst case running time for large datasets is close to the time taken for making one pass on the dataset.
+When we expand a rule using an existing sample in memory, the running time is small and only depends on $minSS$ rather than on the dataset size.
 

### 2015-02-17 TableSummarization/Paper/formal.tex
Changes to paper

@@
 If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).}
-
-\papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
+Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).} \papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :
 Here we set weight equal to the number of non-starred values in the rule $r$ i.e.

### 2015-02-17 TableSummarization/Paper/intro.tex
Changes to paper

@@
 \squishend
+We have a prototype of the smart drill down operator that is completely implemented and usable.
+Starting from this point on, whenever we provide result snippets, these will be screenshots from our prototype.
+In Section~\ref{sec:exp_qualitative}, we will compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results; as a matter of fact, smart drill-down can be viewed as a generalization of traditional drill-down (when the weighting function is set to equal $1$ if the drill-down column has a non-$\star$ value and $0$ otherwise), so our implemented prototype also supports traditional drill-down.
 

### 2015-02-17 TableSummarization/Paper/related.tex
Changes to paper

@@
 This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values.
+A few vision papers~\cite{data-deluge, cetintemel2013query} suggest frameworks for building interactive data exploration systems.
+Some of these ideas, like maintaining user profiles, could be integrated into a smart drill down system.
+Reference~\cite{DBLP:conf/edbt/CandanCQS09} proposes an extension to OLAP drill-down that takes visualization real estate into account, by clustering attribute values.
+But it focuses on expanding a single column at a time, and relies on a given value hierarchy for clustering.
 

### 2015-02-17 TableSummarization/Paper/sampling.tex
Changes to paper

@@
 In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples.
-We show that the problem is NP-Hard, and mention an approximate solution using an additional assumption.
+We show that the problem is NP-Hard, and describe an approximate solution.
 In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process.
@@
 If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down.
+Note that this set of tuples forms a uniformly random sample of tuples covered by $r$.
 If not, then we need to access the hard disk to obtain more tuples covered by $r$.
@@
 Let $M_{0}, M_{1}, ...
-M_{D}$ be the sets for which we found locally optimal solutions.
+M_{D}$ be the sets for which we have found locally optimal solutions.
 We make $D$ iterations in all, where in the $i^{th}$ iteration, we try different locally optimal solutions for $M_i$ and use it to update the maximum attainable value for each cost.

### 2015-02-17 TableSummarization/Paper/algorithms.tex
small changes throughout

@@
 Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
-$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
+$\text{Score}(R^{\prime}) \geq \text{Score}(R)$.
 \end{lemma}
 \papertext{The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}}\techreporttext{The proof of this lemma, as well as other proofs, can be found in the appendix}.
-
-\noindent Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight.
-Or equivalently, we can define Score for a \emph{set} of rules as follows:
+Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:
 
@@
 Let $R$ be a set of rules. Then the Score of $R$ is
-$$\text{Score}(R) = \text{Score}(R^{\prime})$$
+$\text{Score}(R) = \text{Score}(R^{\prime})$
 where $R^{\prime}$ is the list of rules obtained by ordering the rules in the set $R$ in decreasing order by weight.
@@
 \end{lemma}
+\vspace{-10pt}
 

### 2015-02-17 TableSummarization/Paper/exp.tex
small changes throughout

@@
 \section{Experiments}\label{sec:experiments}
-We have implemented a fully-functional interactive system instrumented with the smart drill down operator, with a web interface.
-We now describe our experiments on this system with real datasets.
+We have implemented a fully-functional interactive tool instrumented with the smart drill down operator, having a web interface.
+We now describe our experiments on this tool with real datasets.
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\vspace{-10pt}
+\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
+\vspace{-15pt}
+\end{figure}
 
@@
 Continuous values, such as income, have been bucketized in the dataset, and each column has up to $10$ distinct values.
-
-\begin{figure}
-\vspace{-5pt}
-\centering
-\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
-\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
-\vspace{-15pt}
-\end{figure}
-
-\begin{figure}
-\vspace{-5pt}
-\centering
-\includegraphics[width=90mm]{graphs/screenshots6col/age_drilldown.jpg}
-\caption{A regular drill down on Age\label{fig:drilldownexample}}
-\end{figure}
-
-\begin{figure}
-\vspace{-5pt}
-\centering
-\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
-\vspace{-10pt}
-\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
-\vspace{-15pt}
-\end{figure}
-
-\begin{figure}
-\vspace{-5pt}
-\centering
-\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
-\vspace{-10pt}
-\caption{A rule expansion \label{fig:uiexamplerule}}
-\vspace{-15pt}
-\end{figure}
 
@@
 
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
+\vspace{-10pt}
+\caption{A regular drill down on Age\label{fig:drilldownexample}}
+\vspace{-5pt}
+\end{figure}
+
 \subsection{Qualitative Study}\label{sec:exp_qualitative}
@@
 Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill down, whereas smart drill down displays that information to the user with a single click.
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\vspace{-10pt}
+\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
+\vspace{-20pt}
+\end{figure}
 
@@
 %TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.
+
+\begin{figure}
+\vspace{-5pt}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\vspace{-10pt}
+\caption{A rule expansion \label{fig:uiexamplerule}}
+\vspace{-15pt}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=90mm]{graphs/screenshots6col/bits_scoring.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.jpg}
+\vspace{-10pt}
 \caption{Bits scoring\label{fig:weigtingbitwise}}
@@
 \centering
-\includegraphics[width=90mm]{graphs/screenshots6col/min2_scoring.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.jpg}
 \vspace{-10pt}

### 2015-02-17 TableSummarization/Paper/intro.tex
small changes throughout

--- 2015-02-10 TableSummarization/Paper/intro.tex
@@
 \squishend
+We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator.
+From this point on, when we provide result snippets, these will be screenshots from our prototype tool.
+Our prototype tool also supports traditional drill-down: smart drill-down
+can be viewed as a generalization of traditional drill-down (with the weighting function set appropriately).
+In Section~\ref{sec:exp_qualitative}, we compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results.
 

### 2015-02-17 TableSummarization/Paper/related.tex
small changes throughout

@@
 
-Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
-There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
+Some related work~\cite{DBLP:journals/debu/GolabKS11, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
+There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf/kdd/XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
 In both these cases, the emphasis is on completely covering or

### 2015-04-07 TableSummarization/Paper/KDDrebuttal.tex
made rebuttal file

\documentclass{article}
\date{\today}

\begin{document}

\title{Smart Drill-Down (KDD Rebuttal)}
\author{Manas Joglekar, Hector Garcia-Molina, Aditya Parameswaran}
\maketitle
To the editors,

We would like to thank the reviewers for their insightful comments. The primary comment made by every reviewer was that we had missed several papers related to our work. We are outsiders to the KDD community, and were not very familiar with previous work in the pattern mining area. We went through several of the papers pointed out by the reviewers, and will mention the differences between these papers and our work below. We will also do a more thorough literature review and significantly augment the related work section for the camera ready submission.

There are four primary ways in which existing work on pattern mining differs from our work:
\begin{enumerate}
\item {\bf Interactiveness}: user guided. fast (making approximation necessary).

\item {\bf Relational Data}: gives meaning to interestingness measures based on columns, such as `Bits'.

\item {\bf Tunability}: The primary attribute of our work is that we allow the user to specify the kind of rules/patterns he is interested in by providing any monotone weighting function. While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an approximate optimality guarantee. This is in contrast to several of the related papers, where the authors use a heuristic algorithm such as beam search, or a restricted top down search, with no theoretical optimality guarantees. 
% mention paper names/authors above

\item {\bf Theoretical Guarantees}:
\end{enumerate}

% address other reviewer comments, especially 4th reviewers.

\end{document}

### 2015-04-09 TableSummarization/Paper/KDDrebuttal.tex
changes

@@
 
-There are four primary ways in which existing work on pattern mining differs from our work:
+There are three primary ways in which existing work on pattern mining differs from our work:
 \begin{enumerate}
-\item {\bf Interactiveness}: user guided. fast (making approximation necessary).
+\item {\bf Interactiveness}: Smart drill-down allows the user to direct the data exploration according to his interests.
+This means that in each step, the number of rules displayed to the user should be small enough to be readable in a short amount of time, and the next set of rules to be displayed should depend on what part of the table the user expresses interest in.
+In addition, the response time of the system needs to be small ($< 10$ seconds) in order to prevent the user's flow of thought from being broken.
+In contrast, many of the mentioned related works (e.g.
+`The Chosen few', `Constraint based Pattern Mining') display {\em all} output patterns in one go, which would take a lot of processing time and potentially display unneeded information to the user.
+Some, such as `Krimp', `Summarizing Itemset Patterns: A Profile-Based Approach' display a smaller number of patterns, but are still not interactive.
 
-\item {\bf Relational Data}: gives meaning to interestingness measures based on columns, such as `Bits'.
+\item {\bf Relational Data}: Smart drill-down is used for summarizing data from a relational table, as opposed to a set of transactions.
+Different columns in a table can be of differing importance to a user, and smart drill-down allows a user to focus on expanding certain columns, or to ignore certain columns.
+In a transaction based database, it is much harder to express extra interest in certain `types' of items.
 
-\item {\bf Tunability}: The primary attribute of our work is that we allow the user to specify the kind of rules/patterns he is interested in by providing any monotone weighting function.
-While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an approximate optimality guarantee.
-This is in contrast to several of the related papers, where the authors use a heuristic algorithm such as beam search, or a restricted top down search, with no theoretical optimality guarantees.
-% mention paper names/authors above
-
-\item {\bf Theoretical Guarantees}:
+\item {\bf Tunability}: An important attribute of our work is that we allow the user to specify the kind of rules he is interested in by specifying any monotone weighting function.
+While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an {\em approximate optimality guarantee}.
+Thus we attempt to strike a fine balance between tunability and optimizability.
+This is in contrast to several of the related papers with a fully general scoring function (`Diverse subgroup set discovery', ), forcing the authors use an algorithm that is either heuristic with no optimality guarantees, or very time consuming.
+Some other papers (`MIME', `One click mining') are not tunable enough, and only provide a fixed set of interestingess parameters.
 \end{enumerate}
 
-% address other reviewer comments, especially 4th reviewers.
+In addition, the paper on Intelligent roll-ups by Sunita Sarawagi is about roll-ups as opposed to drill down.
+So it assumes that the user is already at a detailed level of the table, has found an anomaly, and wants to `zoom out' to the highest level where the anomaly still exists.
+`Formalizing Complex Prior Information to Quantify Subjective Interestingness of Frequent Pattern Sets' uses a max entropy based approach, instead of maximizing total coverage of the table.
+
+Based on Reviewer 1's suggestion, we will add a more detailed comparison between our contributions and existing work in Section 1 itself.
+Reviewers 1 and 2 also pointed out that some of the concepts and ideas introduced are simple.
+This is because we had moved a lot of the technical material, especially the sampling related work, to the technical report.
+
+Reviewer 4 correctly points out that its the processing requirements of the priori algorithm than make the sampling procedure necessary, for the 2.5 million tuple dataset.
+The rule finding algorithm makes several passes over the dataset/sample, and would be very slow if it had to make many passes over a 2.5 million tuples.
+In this case, the sampling helps with speed, but can be done away with otherwise.
+On the other hand, users might want to explore even bigger datasets, containing billions of tuples, where the dataset doesn't fit in main memory at all, in which case the sampling scheme becomes even more essential.
 

### 2015-04-09 TableSummarization/Paper/KDDrebuttal.tex
changes

@@
 We are outsiders to the KDD community, and were not very familiar with previous work in the pattern mining area.
-We went through several of the papers pointed out by the reviewers, and will mention the differences between these papers and our work below.
-We will also do a more thorough literature review and significantly augment the related work section for the camera ready submission.
+Indeed, our paper considers a type of pattern mining, and should have cited the papers pointed out by the reviewers.
+However, there are significant differences with existing pattern mining work, which we believe make our paper interesting.
+We describe these differences below.
+We plan to significantly augment our related work section, describing the differences with the referenced papers and other pattern mining work.
 
@@
 Some, such as `Krimp', `Summarizing Itemset Patterns: A Profile-Based Approach' display a smaller number of patterns, but are still not interactive.
+Note that our need for interactivity causes us to implement a smart memory manager that can use limited memory effectively while preparing for future drill down requests.
 
@@
 While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an {\em approximate optimality guarantee}.
-Thus we attempt to strike a fine balance between tunability and optimizability.
+Thus we strike a balance between tunability and optimizability.
 This is in contrast to several of the related papers with a fully general scoring function (`Diverse subgroup set discovery', ), forcing the authors use an algorithm that is either heuristic with no optimality guarantees, or very time consuming.

### 2015-04-10 TableSummarization/Paper/KDDrebuttal.tex
changes

@@
 There are three primary ways in which existing work on pattern mining differs from our work:
-\begin{enumerate}
-\item {\bf Interactiveness}: Smart drill-down allows the user to direct the data exploration according to his interests.
+
+Interactiveness: Smart drill-down allows the user to direct the data exploration according to his interests.
 This means that in each step, the number of rules displayed to the user should be small enough to be readable in a short amount of time, and the next set of rules to be displayed should depend on what part of the table the user expresses interest in.
-In addition, the response time of the system needs to be small ($< 10$ seconds) in order to prevent the user's flow of thought from being broken.
+In addition, the response time of the system needs to be small (upto 10 seconds) in order to prevent the user's flow of thought from being broken.
 In contrast, many of the mentioned related works (e.g.
-`The Chosen few', `Constraint based Pattern Mining') display {\em all} output patterns in one go, which would take a lot of processing time and potentially display unneeded information to the user.
+`The Chosen few', `Constraint based Pattern Mining') display all output patterns in one go, which would take a lot of processing time and potentially display unneeded information to the user.
 Some, such as `Krimp', `Summarizing Itemset Patterns: A Profile-Based Approach' display a smaller number of patterns, but are still not interactive.
@@
 
-\item {\bf Relational Data}: Smart drill-down is used for summarizing data from a relational table, as opposed to a set of transactions.
+Relational Data: Smart drill-down is used for summarizing data from a relational table, as opposed to a set of transactions.
 Different columns in a table can be of differing importance to a user, and smart drill-down allows a user to focus on expanding certain columns, or to ignore certain columns.
@@
 
-\item {\bf Tunability}: An important attribute of our work is that we allow the user to specify the kind of rules he is interested in by specifying any monotone weighting function.
-While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an {\em approximate optimality guarantee}.
+Tunability: An important attribute of our work is that we allow the user to specify the kind of rules he is interested in by specifying any monotone weighting function.
+While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an `approximate optimality guarantee'.
 Thus we strike a balance between tunability and optimizability.
 This is in contrast to several of the related papers with a fully general scoring function (`Diverse subgroup set discovery', ), forcing the authors use an algorithm that is either heuristic with no optimality guarantees, or very time consuming.
-Some other papers (`MIME', `One click mining') are not tunable enough, and only provide a fixed set of interestingess parameters.
-\end{enumerate}
+Some other papers (`MIME', `One click mining', `Finding Robust Itemsets Under Subsampling') are not tunable enough, and only provide a fixed set of interestingess parameters.
 
@@
 `Formalizing Complex Prior Information to Quantify Subjective Interestingness of Frequent Pattern Sets' uses a max entropy based approach, instead of maximizing total coverage of the table.
+Similarly, `A framework for mining interesting pattern sets' also uses a max entropy based approach and focuses on presenting `surprising' information to the user, based on the user's prior probability distribution.
 
-Based on Reviewer 1's suggestion, we will add a more detailed comparison between our contributions and existing work in Section 1 itself.
-Reviewers 1 and 2 also pointed out that some of the concepts and ideas introduced are simple.
+Based on Reviewer 2's suggestion, we will add a more detailed comparison between our contributions and existing work in Section 1 itself.
+Reviewer 2 also pointed out that some of the concepts and ideas introduced are simple.
 This is because we had moved a lot of the technical material, especially the sampling related work, to the technical report.
 
-Reviewer 4 correctly points out that its the processing requirements of the priori algorithm than make the sampling procedure necessary, for the 2.5 million tuple dataset.
+Reviewer 4 mentions that our experiments don't show scalability, which is essential for KDD.
+The performance of our rule finding algorithm depends only on the sample size it runs on, so once our system creates a sample (in time linear in table size), it should have the same processing time for different sized datasets.
+
+Reviewer 5 suggests we use a multivariate discretizer to preserve interactions, for continuous valued data. We thank the reviewer for this suggestion, and will incorporate this into our system.
+
+Reviewer 6 correctly points out that it's the processing requirements of the apriori algorithm that make the sampling procedure necessary, for the 2.5 million tuple dataset.
 The rule finding algorithm makes several passes over the dataset/sample, and would be very slow if it had to make many passes over a 2.5 million tuples.
@@
 
+% address state of the art comment
+% R6: number of attributes???
+
 \end{document}

### 2015-04-10 TableSummarization/Paper/KDDrebuttal.tex
changes

@@
 \maketitle
-To the editors,
 
-We would like to thank the reviewers for their insightful comments.
+We would like to thank the reviewers for their insightful and incredibly useful comments.
 The primary comment made by every reviewer was that we had missed several papers related to our work.
-We are outsiders to the KDD community, and were not very familiar with previous work in the pattern mining area.
+We are outsiders to the KDD community, and were not very familiar with previous work in pattern mining.
 Indeed, our paper considers a type of pattern mining, and should have cited the papers pointed out by the reviewers.
@@
 
-Interactiveness: Smart drill-down allows the user to direct the data exploration according to his interests.
+1) Interactiveness: Smart drill-down allows the user to direct the data exploration according to his interests.
 This means that in each step, the number of rules displayed to the user should be small enough to be readable in a short amount of time, and the next set of rules to be displayed should depend on what part of the table the user expresses interest in.
-In addition, the response time of the system needs to be small (upto 10 seconds) in order to prevent the user's flow of thought from being broken.
-In contrast, many of the mentioned related works (e.g.
+In addition, the response time of the system needs to be small (<10 seconds) in order to prevent the user's flow of thought from being broken.
+In contrast, many of the mentioned papers (e.g.
 `The Chosen few', `Constraint based Pattern Mining') display all output patterns in one go, which would take a lot of processing time and potentially display unneeded information to the user.
-Some, such as `Krimp', `Summarizing Itemset Patterns: A Profile-Based Approach' display a smaller number of patterns, but are still not interactive.
-Note that our need for interactivity causes us to implement a smart memory manager that can use limited memory effectively while preparing for future drill down requests.
+Some, such as `Krimp', `Summarizing Itemset Patterns' display a smaller number of patterns, but are still not interactive.
+Note that our need for interactivity causes us to implement a smart memory manager that can use limited memory effectively while preparing for future requests.
 
-Relational Data: Smart drill-down is used for summarizing data from a relational table, as opposed to a set of transactions.
-Different columns in a table can be of differing importance to a user, and smart drill-down allows a user to focus on expanding certain columns, or to ignore certain columns.
+2) Relational Data: Smart drill-down is used for summarizing data from a relational table, as opposed to a set of transactions.
+Different columns in a table can be of differing importance to a user, and smart drill-down allows a user to focus on expanding certain columns, while ignoring others.
 In a transaction based database, it is much harder to express extra interest in certain `types' of items.
 
-Tunability: An important attribute of our work is that we allow the user to specify the kind of rules he is interested in by specifying any monotone weighting function.
+3) Tunability: An important attribute of our work is that we allow the user to specify the kind of rules he/she is interested in by specifying any monotone weighting function.
 While this allows for a very wide variety of interestingness measures, the scoring function is restricted enough that it is submodular on the rule-set, allowing us to obtain an `approximate optimality guarantee'.
 Thus we strike a balance between tunability and optimizability.
-This is in contrast to several of the related papers with a fully general scoring function (`Diverse subgroup set discovery', ), forcing the authors use an algorithm that is either heuristic with no optimality guarantees, or very time consuming.
+This is in contrast to related papers with a fully general scoring function (`Diverse subgroup set discovery'), forcing the use of an algorithm that is either heuristic with no optimality guarantees, or very time consuming.
 Some other papers (`MIME', `One click mining', `Finding Robust Itemsets Under Subsampling') are not tunable enough, and only provide a fixed set of interestingess parameters.
 
-In addition, the paper on Intelligent roll-ups by Sunita Sarawagi is about roll-ups as opposed to drill down.
+Further, the paper on Intelligent roll-ups by Sunita Sarawagi is about roll-ups as opposed to drill down.
 So it assumes that the user is already at a detailed level of the table, has found an anomaly, and wants to `zoom out' to the highest level where the anomaly still exists.
-`Formalizing Complex Prior Information to Quantify Subjective Interestingness of Frequent Pattern Sets' uses a max entropy based approach, instead of maximizing total coverage of the table.
-Similarly, `A framework for mining interesting pattern sets' also uses a max entropy based approach and focuses on presenting `surprising' information to the user, based on the user's prior probability distribution.
+% `Formalizing Complex Prior Information to Quantify Subjective Interestingness of Frequent Pattern Sets' uses a max entropy based approach, instead of maximizing total coverage of the table.
+Similarly, `A framework for mining interesting pattern sets' also uses max entropy and focuses on presenting `surprising' information based on the user's prior probability distribution.
 
 Based on Reviewer 2's suggestion, we will add a more detailed comparison between our contributions and existing work in Section 1 itself.
-Reviewer 2 also pointed out that some of the concepts and ideas introduced are simple.
-This is because we had moved a lot of the technical material, especially the sampling related work, to the technical report.
+Reviewer 2 also pointed out that we spent a little too much space on somewhat simple explanations;
+this is again because we are outsiders to KDD and felt the need to explain things in detail.
+We'd be happy to move some of the more technical aspects and proofs from the technical report back into the paper once we shorten these explanations.
 
-Reviewer 4 mentions that our experiments don't show scalability, which is essential for KDD.
-The performance of our rule finding algorithm depends only on the sample size it runs on, so once our system creates a sample (in time linear in table size), it should have the same processing time for different sized datasets.
+Reviewer 4 mentions that our experiments do not sufficiently demonstrate scalability.
+In fact, the performance of our rule finding algorithm depends only on the sample size it runs on, so once our system creates a sample (in time linear in table size), it should have the same processing time for different sized datasets.
 
-Reviewer 5 suggests we use a multivariate discretizer to preserve interactions, for continuous valued data. We thank the reviewer for this suggestion, and will incorporate this into our system.
+Reviewer 5 suggests we use a multivariate discretizer to preserve interactions for continuous valued data. We thank the reviewer for this suggestion, and will incorporate this into our system.
 
-Reviewer 6 correctly points out that it's the processing requirements of the apriori algorithm that make the sampling procedure necessary, for the 2.5 million tuple dataset.
-The rule finding algorithm makes several passes over the dataset/sample, and would be very slow if it had to make many passes over a 2.5 million tuples.
-In this case, the sampling helps with speed, but can be done away with otherwise.
-On the other hand, users might want to explore even bigger datasets, containing billions of tuples, where the dataset doesn't fit in main memory at all, in which case the sampling scheme becomes even more essential.
+Reviewer 6 correctly points out that it is the processing requirements of the apriori algorithm that make the sampling procedure necessary for the 2.5M tuple dataset.
+The rule finding algorithm makes several passes over the dataset, and would be very slow if it had to make many passes over so many tuples.
+Here, sampling helps with speed, but can be done away with otherwise.
+On the other hand, users might want to explore even bigger datasets, containing billions of tuples, where the dataset doesn't fit in main memory at all; here, sampling is not just a nice-to-have, it is a must-have.
 

### 2015-04-11 TableSummarization/Paper/KDDrebuttal.tex
changes

@@
 This means that in each step, the number of rules displayed to the user should be small enough to be readable in a short amount of time, and the next set of rules to be displayed should depend on what part of the table the user expresses interest in.
-In addition, the response time of the system needs to be small (<10 seconds) in order to prevent the user's flow of thought from being broken.
+In addition, the response time of the system needs to be small (<10 seconds) to prevent the user's flow of thought from being broken.
 In contrast, many of the mentioned papers (e.g.
@@
 This is in contrast to related papers with a fully general scoring function (`Diverse subgroup set discovery'), forcing the use of an algorithm that is either heuristic with no optimality guarantees, or very time consuming.
-Some other papers (`MIME', `One click mining', `Finding Robust Itemsets Under Subsampling') are not tunable enough, and only provide a fixed set of interestingess parameters.
+Some other papers (`MIME', `One click mining', `Finding Robust Itemsets Under Subsampling', `A framework for mining interesting pattern sets') are not tunable enough, and only provide a fixed set of interestingess parameters.
 
-Further, the paper on Intelligent roll-ups by Sunita Sarawagi is about roll-ups as opposed to drill down.
-So it assumes that the user is already at a detailed level of the table, has found an anomaly, and wants to `zoom out' to the highest level where the anomaly still exists.
-% `Formalizing Complex Prior Information to Quantify Subjective Interestingness of Frequent Pattern Sets' uses a max entropy based approach, instead of maximizing total coverage of the table.
-Similarly, `A framework for mining interesting pattern sets' also uses max entropy and focuses on presenting `surprising' information based on the user's prior probability distribution.
+Further, the paper on `Intelligent roll-ups' is about roll-ups as opposed to drill down.
+It assumes that the user is already at a detailed level of the table, has found an anomaly, and wants to `zoom out' to the highest level where the anomaly still exists.
 
 Based on Reviewer 2's suggestion, we will add a more detailed comparison between our contributions and existing work in Section 1 itself.
-Reviewer 2 also pointed out that we spent a little too much space on somewhat simple explanations;
+Reviewer 2 also pointed out that we spent too much space on somewhat simple explanations;
 this is again because we are outsiders to KDD and felt the need to explain things in detail.
-We'd be happy to move some of the more technical aspects and proofs from the technical report back into the paper once we shorten these explanations.
+We'd be happy to move some of the more technical aspects and proofs from the technical report back to the paper once we shorten these explanations.
 
@@
 
-% address state of the art comment
-% R6: number of attributes???
-
 \end{document}

### 2015-06-23 TableSummarization/Paper/intro.tex
Added new related work

@@
 Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
+and smart drill down to provide an additional option of anomaly detection. 
+\mrjdel{We discuss related work in detail in  Section~\ref{sec:related}.}
+
+\mrj{Our work is also related to pattern mining.
+Several of the pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, and are not interactive.
+Moreover, to the best of our knowledge, all other pattern mining work is either not flexible enough~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117}, restricting the amount of tuning the user can perform, or so general~\cite{Leeuwen:2012:DSS:2347179.2347243} as to preclude efficient optimization.
+Our work also merges `interesting pattern mining' into the OLAP framework.
+We discuss related work in detail in Section~\ref{sec:related}.}
 

### 2015-06-23 TableSummarization/Paper/related.tex
Added new related work

@@
 Some related work~\cite{DBLP:journals/debu/GolabKS11, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
-There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf/kdd/XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
+There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf:kdd:XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
 In both these cases, the emphasis is on completely covering or
@@
 
+\mrj{There has also been plenty of work on pattern mining.
+Several works~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} are non-interactive, and attempt to find a one shot summary of the table.
+This usually consumes a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest.
+In contrast, our work is interactive, and includes a smart memory manager that can use limited memory effectively while preparing for future requests.}
+
+\mrj{Our Smart Drill-Down operator is tunable because of the flexible weighting function, but the monotonicity of the weighting function and the use of $MCount$, still make it possible for us to get an approximate optimality guarantee for the rules we display.
+In contrast, much of the existing pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117} is not not tunable enough, providing only a fixed set of interestingess parameters.
+On the other hand, reference~\cite{Leeuwen:2012:DSS:2347179.2347243} allows a fully general scoring function, necessitating the use of heuristics with no optimality guarantees, and very time consuming algorithms.
+A lot of pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Yan:2005:SIP:1081870.1081907,Vreeken:2011:KMI:1969593.1969615} also focuses on itemsets rather than Relational Data, which doesn't allow the user to express interest in certain `columns' over others.}
 % All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.

### 2015-07-09 TableSummarization/Paper/related.tex
changes to related work (and copy pasted intro changes

@@
 \section{Related Work}\label{sec:related}
+
+% intro copy-paste for editing
+%\mrj{Our work is also related to pattern mining.
+Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, but do not propose interactive mechanisms.
+Moreover, to the best of our knowledge, other pattern mining work is either not flexible enough~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117}, restricting the amount of tuning the user can perform, or so general~\cite{Leeuwen:2012:DSS:2347179.2347243} as to preclude efficient optimization.
+Our work also merges `interesting pattern mining' into the OLAP framework.
+We discuss related work in detail in Section~\ref{sec:related}.}
 
@@
 \mrj{There has also been plenty of work on pattern mining.
-Several works~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} are non-interactive, and attempt to find a one shot summary of the table.
-This usually consumes a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest.
+Several works~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} propose non-interactive schemes that attempt to find a one shot summary of the table.
+These schemes usually consume a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest.
 In contrast, our work is interactive, and includes a smart memory manager that can use limited memory effectively while preparing for future requests.}
@@
 On the other hand, reference~\cite{Leeuwen:2012:DSS:2347179.2347243} allows a fully general scoring function, necessitating the use of heuristics with no optimality guarantees, and very time consuming algorithms.
-A lot of pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Yan:2005:SIP:1081870.1081907,Vreeken:2011:KMI:1969593.1969615} also focuses on itemsets rather than Relational Data, which doesn't allow the user to express interest in certain `columns' over others.}
+A lot of pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Yan:2005:SIP:1081870.1081907,Vreeken:2011:KMI:1969593.1969615} also focuses on itemsets rather than Relational Data, which does not allow the user to express interest in certain `columns' over others.}
 % All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.

### 2015-07-09 TableSummarization/Paper/intro.tex
changes to related work and intro

--- 2015-02-17 TableSummarization/Paper/intro.tex
@@
 Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.
+and smart drill down to provide an additional option of anomaly detection. 
+\mrjdel{We discuss related work in detail in  Section~\ref{sec:related}.}
+
+\mrj{Our work is also related to pattern mining.
+Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, but do not propose interactive mechanisms.
+Moreover, to the best of our knowledge, other pattern mining work is either not flexible enough~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117}, restricting the amount of tuning the user can perform, or so general~\cite{Leeuwen:2012:DSS:2347179.2347243} as to preclude efficient optimization.
+Our work also merges `interesting pattern mining' into the OLAP framework.
+We discuss related work in detail in Section~\ref{sec:related}.}
 

### 2015-07-11 TableSummarization/Paper/algorithms.tex
added stuff from tech report

@@
 
-\techreporttext{
 \begin{algorithm}
@@
 \end{algorithm}
-}
 
@@
 
-\stitle{Parametrized Algorithm:} \techreporttext{Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.} \papertext{We describe our algorithm in detail below.
-The pseudo-code for our algorithm is in the technical report~\cite{tr}.} We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et).
+\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
+We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et).
 BRS takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
@@
 We now describe the function to find the best marginal rule.
-\techreporttext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.} \papertext{The pseudo-code for the function is in the technical report.} The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
-The function makes several iterations \techreporttext{(Step $3$)}, counting marginal values for size $j$ rules in the $j^{th}$ iteration.
+The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.
+The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
+The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration.
 We maintain three sets of rules.
@@
 We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$.
-Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$\techreporttext{, as shown in Step 3.3.2}.
+Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2.
 Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule.
-Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$\techreporttext{, as shown in Step 3.5}.
+Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5.
 If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

### 2015-07-11 TableSummarization/Paper/related.tex
changes to related work and intro
added stuff from tech report

--- 2015-02-17 TableSummarization/Paper/related.tex
@@
 
-% All the existing works solve a fixed optimization problem, whereas  we focus on finding an optimal summary for a flexible user-specified weighting function.
+There has also been plenty of work on pattern mining.
+Several works~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} propose non-interactive schemes that attempt to find a one shot summary of the table.
+These schemes usually consume a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest.
+In contrast, our work is interactive, and includes a smart memory manager that can use limited memory effectively while preparing for future requests.
+
+Our Smart Drill-Down operator is tunable because of the flexible weighting function, but the monotonicity of the weighting function and the use of $MCount$, still make it possible for us to get an approximate optimality guarantee for the rules we display.
+In contrast, much of the existing pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117} is not not tunable enough, providing only a fixed set of interestingess parameters.
+On the other hand, reference~\cite{Leeuwen:2012:DSS:2347179.2347243} allows a fully general scoring function, necessitating the use of heuristics with no optimality guarantees, and very time consuming algorithms.
+A lot of pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Yan:2005:SIP:1081870.1081907,Vreeken:2011:KMI:1969593.1969615} also focuses on itemsets rather than Relational Data, which does not allow the user to express interest in certain `columns' over others.
 

### 2015-07-11 TableSummarization/Paper/sampling.tex
small changes throughout
added stuff from tech report

@@
 
-\papertext{
-BRS makes multiple passes over the table in order to determine the best set of rules to display.
-This can be slow when the table is too large to fit in main memory.
-We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance.
-If we had obtained a sample $s$ by selecting each table tuple with probability $p$, and run BRS on $s$, then we multiply the count of each rule found by BRS, by $\frac{1}{p}$ to estimate its count over the full table.
-
-In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples.
-We show that the problem is NP-Hard, and describe an approximate solution.
-In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process.
-Detailed descriptions of the sampling problem and the SampleHandler are in the technical report.
-The technical report also contains some additional optimizations, and ways to set the minimum sample size.
-
-\subsubsection{Deciding what to sample}\label{sec:sampling_algorithms}
-We are given a memory capacity $M$, and a minimum sample size $minSS$, both specified by the user.
-$minSS$ is the minimum number of tuples on which we are allowed to run BRS, without accessing the hard disk.
-A higher value of $minSS$ increases both accuracy and computation cost of our system.
-
-At any point, we have a tree $U$ of rules displayed to the user.
-Initially, the tree consists of a single node corresponding to the empty rule.
-When the user drills down on a rule $r$, the sub-rules of $r$ obtained by running BRS are added as children of node $r$.
-Our system maintains multiple samples in memory, with one sample per rule in $U$.
-Specifically, for each rule $r \in U$, we choose an integer $n_r$, and create a sample $s_r$ consisting of $n_r$ uniformly randomly chosen tuples covered by $r$ from the table.
-Because of the memory constraint, we must have $\sum_{r\in U} n_r \leq M$.
-
-When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$.
-If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down.
-Note that this set of tuples forms a uniformly random sample of tuples covered by $r$.
-If not, then we need to access the hard disk to obtain more tuples covered by $r$.
-
-Leaves of tree $U$ correspond to rules that may be drilled down on next.
-We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next.
-This can be a uniform distribution, or a machine learned distribution using past user data.
-We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing the hard disk.
-
-If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$.
-We denote the ratio of selectivities $x$ by $S(r',r)$.
-$S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$.
-If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$.
-We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$.
-We now formally define our problem:
-\begin{problem}\label{prob:sample-sizes}
-Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
-choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
-where the $I$'s are indicator variables, subject to :
-$\sum_{r \in U} n_r \leq M$
-\end{problem}
-Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
-%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
-We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
-\begin{lemma}\label{lemma:sampling-hardness}
-Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
-\end{lemma}
-\begin{proof}(Sketch)
-Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
-We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
-We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.
-
-We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$.
-For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$.
-Tree $U$ has $m$ special nodes $r_1, r_2, ...
-r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$.
-All other leaves have expansion probability $0$.
-The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$.
-In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes.
-Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$.
-In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$.
-Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
-So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$.
-Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
-Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem.
-Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
-\end{proof}
-A more detailed proof sketch is given in the technical report.
-}
-
-\techreporttext{
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
@@
 how quickly we can return results to the user.
-\papertext{We describe how to set $minSS$ in the technical report.}
 
@@
 \end{proof}
-}
 
@@
 
+\begin{comment}
 \papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
@@
 }
+\end{comment}
 
-\techreporttext{Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
+Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
 Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
@@
 This can be solved using dynamic programming, in $O(D\mathcal{S}3^d)$ time.
-}
 
@@
 \subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
+\begin{comment}
 \papertext{
@@
 }
+\end{comment}
 
-\techreporttext{
+%\techreporttext{
 We now describe the design of the {\em SampleHandler}, which given
@@
 
+\techreporttext{
 \smallskip

### 2015-07-16 TableSummarization/Paper/intro.tex
added stuff from tech report
changes

@@
 In this paper, we present a new interaction operator 
-that is an extension to a traditional 
+that is an extension to the traditional 
 drill down operator, aimed at providing {\em complementary}
@@
 (e.g., the Walmart rule going from Table~\ref{table:introexample} to
-Table~\ref{table:introexample2}.
+Table~\ref{table:introexample2}).
 In Section~\ref{sec:interface} we describe another option
@@
 and smart drill down to provide an additional option of anomaly detection.
-\mrjdel{We discuss related work in detail in  Section~\ref{sec:related}.}
 
-\mrj{Our work is also related to pattern mining.
+Our work is also related to pattern mining.
 Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, but do not propose interactive mechanisms.
@@
 Our work also merges `interesting pattern mining' into the OLAP framework.
-We discuss related work in detail in Section~\ref{sec:related}.}
+We discuss related work in detail in Section~\ref{sec:related}.
 

### 2015-08-18 TableSummarization/Paper/formal.tex
new data files

@@
 Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule.
-Consider the examples in Table \ref{table:sizescoringexample}.
+Consider the examples in Table~\ref{table:sizescoringexample}.
 The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$.
@@
 the {\em size} of the rule.
-For example, in Table~\ref{table:introexample3}, the rule (Target, bicycles, $\star$) has weight $2$.
+For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.
 To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules.

### 2015-08-18 TableSummarization/src/dataextraction/Donations.java

@@ -1,0 +1,86 @@
+package dataextraction;
+
+import static java.lang.System.out;
+
+import java.io.BufferedReader;
+import java.io.FileReader;
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+
+public class Donations {
+	
+	final static String ALLWEBKFILELOCATION = "TestDatasets/donations_data/all_webk_2016.csv";
+	final static String CANDIDATESFILELOCATION = "TestDatasets/donations_data/candidates_2016.csv";
+	final static String SUPERPACFILELOCATION = "TestDatasets/donations_data/superpac_contribs_2016.csv";
+	final static String DATAFILELOCATION = SUPERPACFILELOCATION;
+	
+	public static void addNames (TableInfo table) {
+	}
+	
+	public static TableInfo parseData() throws IOException {
+		BufferedReader br = new BufferedReader(new FileReader(DATAFILELOCATION));
+		List<List<String>> dictionary = new ArrayList<List<String>>();
+		List<Map<String, Integer>> reverseDictionary = new ArrayList<Map<String, Integer>>();
+		List<List<Integer>> contents = new ArrayList<List<Integer>>();
+		String line = br.readLine();
+		out.println(line);
+		{
+			String[] vals = line.split(",");
+			for (int i = 0; i < vals.length; i++) {
+				dictionary.add(new ArrayList<String>());
+				reverseDictionary.add(new HashMap<String, Integer>());
+			}
+			out.println(vals.length);
+		}
+	
+		line = br.readLine(); // ignore first line since its column names.
+		do {
+			boolean inQuote = false;
+			for (int i = 0; i < line.length(); i++) {
+				if (line.charAt(i) == '\"') {
+					inQuote = inQuote ^ true;
+					continue;
+				}
+				if (inQuote && line.charAt(i) == ',') {
+					line= line.substring(0,i) + ' ' + line.substring(i + 1);
+				}
+			}
+			String[] vals = line.split(",");
+			List<Integer> tuple = new ArrayList<Integer>(vals.length);
+			for (int i = 0; i < vals.length; i++) {
+				final String value = vals[i];
+				/*if (value.equals("03")) {
+					out.println(line);
+					System.exit(1);
+				}*/
+				Map<String, Integer> columnDictionary = reverseDictionary.get(i);
+				if (columnDictionary.containsKey(value)) {
[…]

### 2015-08-19 TableSummarization/src/dataextraction/Donations.java
small change

@@ -12,10 +12,12 @@ import java.util.Map;
 
 public class Donations {
 	
+	final static String CONTRIBSBIGFILELOCATION = "TestDatasets/contribs_big/contributions.fec.2014.csv";
+	final static String CONTRIBSDENVERFILELOCATION = "TestDatasets/contribs_denver/contributions.csv";
 	final static String ALLWEBKFILELOCATION = "TestDatasets/donations_data/all_webk_2016.csv";
 	final static String CANDIDATESFILELOCATION = "TestDatasets/donations_data/candidates_2016.csv";
 	final static String SUPERPACFILELOCATION = "TestDatasets/donations_data/superpac_contribs_2016.csv";
-	final static String DATAFILELOCATION = SUPERPACFILELOCATION;
+	final static String DATAFILELOCATION = CONTRIBSBIGFILELOCATION;
 	
 	public static void addNames (TableInfo table) {
 	}
@@ -38,6 +40,10 @@ public class Donations {
 	
 		line = br.readLine(); // ignore first line since its column names.
 		do {
+			//out.println(line);
+			if (Math.random() > 0.1) {
+				continue;
+			}
 			boolean inQuote = false;
 			for (int i = 0; i < line.length(); i++) {
 				if (line.charAt(i) == '\"') {
@@ -49,6 +55,9 @@ public class Donations {
 				}
 			}
 			String[] vals = line.split(",");
+			if (vals.length > dictionary.size()) {
+				continue;
+			}
 			List<Integer> tuple = new ArrayList<Integer>(vals.length);
 			for (int i = 0; i < vals.length; i++) {
 				final String value = vals[i];

### 2015-08-19 TableSummarization/src/dataextraction/Donations.java
changes

@@ -3,8 +3,14 @@ package dataextraction;
 import static java.lang.System.out;
 
 import java.io.BufferedReader;
+import java.io.FileInputStream;
+import java.io.FileOutputStream;
 import java.io.FileReader;
+import java.io.FileWriter;
 import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.io.ObjectOutputStream;
+import java.io.PrintWriter;
 import java.util.ArrayList;
 import java.util.HashMap;
 import java.util.List;
@@ -12,36 +18,118 @@ import java.util.Map;
 
 public class Donations {
 	
+	final static String SERIALIZEDTABLEINFO = "TestDatasets/contribs_big/contributions.reduced.fec.2014.ser";
+	
+	final static String CONTRIBSSMALLFILELOCATION = "TestDatasets/contribs_big/contributions.reduced.fec.2014.csv";
 	final static String CONTRIBSBIGFILELOCATION = "TestDatasets/contribs_big/contributions.fec.2014.csv";
 	final static String CONTRIBSDENVERFILELOCATION = "TestDatasets/contribs_denver/contributions.csv";
 	final static String ALLWEBKFILELOCATION = "TestDatasets/donations_data/all_webk_2016.csv";
 	final static String CANDIDATESFILELOCATION = "TestDatasets/donations_data/candidates_2016.csv";
 	final static String SUPERPACFILELOCATION = "TestDatasets/donations_data/superpac_contribs_2016.csv";
-	final static String DATAFILELOCATION = CONTRIBSBIGFILELOCATION;
 	
+	// For CONTRIBSSMALLFILELOCATION only.
 	public static void addNames (TableInfo table) {
+List<Map<String, String>> names = new ArrayList<Map<String, String>>();
+		
+		Map<String, String> col1 = new HashMap<String, String>();
+		col1.put("column", "Import Ref. Id");
+		for (String val : table.dictionary.get(0)) {
+			col1.put(val, val);
+		}
+		names.add(col1);
+		
+		Map<String, String> col2 = new HashMap<String, String>();
+		col2.put("column", "Is Amendment");
+		for (String val : table.dictionary.get(0)) {
+			col2.put(val, val);
+		}
+		names.add(col2);
+		
+		Map<String, String> col3 = new HashMap<String, String>();
+		col3.put("column", "Contributor Type");
+		for (String val : table.dictionary.get(0)) {
+			col3.put(val, val);
+		}
+		names.add(col3);
+		
[…]

### 2015-08-19 TableSummarization/src/dataextraction/TableInfo.java

@@ -1,5 +1,6 @@
 package dataextraction;
 
+import java.io.Serializable;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.HashMap;
@@ -8,7 +9,7 @@ import java.util.List;
 import java.util.Map;
 import java.util.Set;
 
-public class TableInfo {
+public class TableInfo implements Serializable{
 	public List<List<String>> dictionary;
 	public List<Map<String, Integer>> reverseDictionary;
 	public List<List<Integer>> contents;

### 2015-09-24 TableSummarization/Paper/algorithms.tex
changes

@@
-%!TEX root = TableSummarization.tex
+%!TEX root = smartDrillDownICDE.tex
 
@@
 
-\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}.
+\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labeled Algorithm~\ref{algo:best-rule-set}.
 We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et).
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
+We later describe a strategy for setting $m_w$.
 
@@
 
+\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
+Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
+To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.
+
+Parameter $k$ is the number of new rules to display upon each smart drill-down.
+Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
+On the other hand, small values of $k$ may display too little information about the table.
+Fortunately, the BRS algorithm is incremental in nature.
+That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$.
+Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found.
+This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search.
+Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.
+
+$W$ is the weight function that determines which rules are interesting.
+This is a function specified by the user as a black box.
+Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
+In addition, the user can express interest or disinterest in certain columns by telling the system of favor or ignore those columns, via a simple graphical user interface.
+The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
+
+The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time.
+Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
+We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling.
+We create a small random sample of tuples from the table, and run the BRS algorithm on it.
+Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output.
+To account for sampling error, we can set $m_w$ to $2x$.
+
 \input{sampling}

### 2015-09-24 TableSummarization/Paper/exp.tex
added stuff from tech report
changes

@@
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
 \vspace{-10pt}
-\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
-\vspace{-15pt}
 \end{figure}
@@
 Continuous values, such as income, have been bucketized in the dataset, and each column has up to $10$ distinct values.
-
-\begin{comment}
-\begin{table*}
-\centering 
-\scriptsize
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\vspace{-10pt}
-\caption{Summary after clicking on the empty rule \label{table:uiexample1}} 
-\vspace{-10pt}
-\end{table*} 
-\begin{table*} 
-\centering 
-\scriptsize
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & High school & $\star$ & $\star$ & $1149$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & Grades 9-11 & $\star$ & $\star$ & $605$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & College graduate & $\star$ & $\star$ & $771$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Female & $\star$ & $\star$ & 1-3 years college & $\star$ & $\star$ & $1712$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\hline 
-\end{tabular} 
-\vspace{-10pt}
-\caption{Star expansion on `Education' Column \label{table:uiexamplestar}} 
-\vspace{-10pt}
-\end{table*} 
-\begin{table*} 
-\centering 
-\scriptsize
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4918$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $4075$ & $1$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & $\star$ & $1897$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & $\star$ & $1368$ & $2$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $3$ \\
-\cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\triangleright$ Male & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2242$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $2$ \\
-\hline 
-\end{tabular} 
-\vspace{-10pt}
-\caption{A rule expansion \label{table:uiexamplerule}} 
-\vspace{-10pt}
-\end{table*} 
-\end{comment}
 
@@
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
+\caption{A regular drill down on Age\label{fig:drilldownexample}}
 \vspace{-10pt}
-\caption{A regular drill down on Age\label{fig:drilldownexample}}
-\vspace{-5pt}
 \end{figure}
@@
 We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
-\subsubsection{User interface testing}
+\subsubsection{Testing the User Interface}
 We now present the rule-based summaries displayed as a result of a few different user actions.
@@
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
 \vspace{-10pt}
-\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
-\vspace{-20pt}
 \end{figure}
@@
 
-\vspace{-3pt}
 \subsubsection{Weighting functions}
@@
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\caption{A rule expansion \label{fig:uiexamplerule}}
 \vspace{-10pt}
-\caption{A rule expansion \label{fig:uiexamplerule}}
-\vspace{-15pt}
 \end{figure}
@@
 % TODO: Add the experiment where you compare to vanilla a priori.
-\vspace{-3pt}
 \subsection{Quantitative Study}\label{sec:quantitative_study}
@@
 \begin{figure}
-\vspace{-5pt}
 \centering
   \includegraphics[scale=0.46]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
 \vspace{-10pt}
-  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
-\vspace{-15pt}
 \end{figure}
@@
 
-The computation time increases approximately linearly with parameter $minSS$.
-This is expected, because BRS makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample.
-The running time at very small values of $minSS$ is simply the time taken to read the table and create the sample.
-This also increases linearly with table size, and is only about $2.5$ seconds for a table with $2.5$ million tuples.
-The running time when $minSS$ equals table size, is the time we would require without sampling.
-Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
-The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
-
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.jpg}
+\caption{Bits scoring\label{fig:weigtingbitwise}}
 \vspace{-10pt}
-\caption{Bits scoring\label{fig:weigtingbitwise}}
 \end{figure}
@@
 \begin{figure}
-\vspace{-5pt}
 \centering
 \includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.jpg}
+\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
 \vspace{-10pt}
-\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
-\vspace{-15pt}
 \end{figure}
 
-\begin{comment}
-\begin{table*} 
-\scriptsize
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $5182$ & $3$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & $\star$ & $\star$ & $\star$ & Professional / Managerial & $\star$ & $2820$ & $4$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Never married & $\star$ & $\star$ & Student & > 10 years & $742$ & $10$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ $\star$ & Married & $\star$ & $\star$ & Professional / Managerial & > 10 years & $825$ & $10$ \\
-\hline 
-\end{tabular} 
-\vspace{-10pt}
-\caption{Bits scoring\label{table:weigtingbitwise}} 
-\vspace{-10pt}
-\end{table*}
-
-\begin{table*} 
-\scriptsize
-\centering 
-\begin{tabular}{| p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | l | l |} 
-\hline Gender & Marital Status & Age & Education & Occupation & Time in Bay Area & Count & Weight \\ \hline 
-\cline{1-8} $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $\star$ & $8993$ & $0$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & $\star$ & $\star$ & $\star$ & $\star$ & > 10 years & $2940$ & $1$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Never married & $\star$ & $\star$ & $\star$ & > 10 years & $980$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Female & Married & $\star$ & $\star$ & $\star$ & > 10 years & $1230$ & $2$ \\
-\cline{1-1} \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5} \cline{6-6} \cline{7-8} $\triangleright$ Male & Married & $\star$ & $\star$ & $\star$ & > 10 years & $823$ & $2$ \\
-\hline 
-\end{tabular}
-\vspace{-10pt}
-\caption{Size minus one weighting \label{table:weightingsizeminusone}} 
-\vspace{-10pt}
-\end{table*}  
-\end{comment}
+Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
+The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
 
@@
 \includegraphics[scale=0.46]{graphs/minSS_error_rule.pdf}
-\vspace{-15pt}
 \caption{(a) Running time for different values of parameter $minSS$ (b) Error in Count for different values of parameter $minSS$ (c) Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_speed}\label{fig:minSS_error_percent} \label{fig:minSS_error_rule}}
@@
 \end{figure*}
+
+\subsubsection{Scaling properties of our algorithms}
+The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$.
+That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants.
+In the worst-case wherer we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk.
+That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk.
+The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.
+
+When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
+When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk).
+When we have a few million tuples, we only require a few seconds.
+But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time.
+To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).

### 2015-09-24 TableSummarization/src/dataextraction/Marketing.java

@@ -3,8 +3,12 @@ package dataextraction;
 import static java.lang.System.out;
 
 import java.io.BufferedReader;
+import java.io.FileInputStream;
+import java.io.FileOutputStream;
 import java.io.FileReader;
 import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.io.ObjectOutputStream;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.HashMap;
@@ -26,7 +30,11 @@ import solvers.RuleTree.RuleNode;
 import utils.Experiments;
 
 public class Marketing {
-	final static String DATAFILELOCATION = "TestDatasets/Marketing/marketing.data.txt";
+	public final static String DATAFILELOCATION = "TestDatasets/Marketing/marketing.data.txt";
+	
+	public static String name () {
+		return "Manas";
+	}
 	
 	public static void addNames (TableInfo table) {
 		List<Map<String, String>> names = new ArrayList<Map<String, String>>();
@@ -220,8 +228,21 @@ public class Marketing {
 		return table;
 	}
 	
-	public static void main(String[] args) throws IOException {
-		TableInfo fullTable = Marketing.parseData();
+	public static void main(String[] args) throws IOException, ClassNotFoundException {
+		boolean createTable = true;
+		TableInfo fullTable = null;
+		String tableFile = "TestDatasets/Marketing/marketing.data.ser";
+		if (createTable) {
+			fullTable = Marketing.parseData();
+			ObjectOutputStream oo = new ObjectOutputStream(new FileOutputStream(tableFile));
+			oo.writeObject(fullTable);
+			oo.close();
+		} else {
+			ObjectInputStream oi = new ObjectInputStream(new FileInputStream(tableFile));
+			fullTable = (TableInfo) oi.readObject();
+			oi.close();
+		}
+		if(1!=2) return;
 		//out.println(System.currentTimeMillis() - timer);
 		List<Integer> columns = new ArrayList<Integer>();
 		final Integer firstNumColumns = 7;//9;

### 2015-09-24 TableSummarization/src/dataextraction/TableInfo.java

@@ -26,7 +26,7 @@ public class TableInfo implements Serializable{
 		}
 	}
 	
-	TableInfo (int numColumns) {
+	public TableInfo (int numColumns) {
 		dictionary = new ArrayList<List<String>>(numColumns);
 		reverseDictionary = new ArrayList<Map<String, Integer>>(numColumns);
 		for (int i = 0; i < numColumns; i++) {
@@ -37,13 +37,13 @@ public class TableInfo implements Serializable{
 		bucketizeNumericalColumns(new ArrayList<Integer>(), new ArrayList<Integer>(), new ArrayList<String>());
 	}
 	
-	TableInfo () {
+	public TableInfo () {
 		dictionary = new ArrayList<List<String>>();
 		reverseDictionary = new ArrayList<Map<String, Integer>>();
 		contents = new ArrayList<List<Integer>>();
 	}
 	
-	TableInfo (List<List<String>> dictionary, List<Map<String, Integer>> reverseDictionary, List<List<Integer>> contents) {
+	public TableInfo (List<List<String>> dictionary, List<Map<String, Integer>> reverseDictionary, List<List<Integer>> contents) {
 		this.dictionary = dictionary;
 		this.reverseDictionary = reverseDictionary;
 		this.contents = contents;

### 2015-10-11 TableSummarization/Paper/exp.tex
changes

@@
 
-Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
-The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
+\agp{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
+The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}
 
@@
 
+\agp{
 \subsubsection{Scaling properties of our algorithms}
@@
 To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
+}

### 2015-10-11 TableSummarization/Paper/sampling.tex
changes

@@
 We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
+\agp{$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.}
 The parameter $minSS$ is the minimum number of sample tuples needed
@@
 This parameter determines how accurate our count estimates will be, and also
-how quickly we can return results to the user.
+how quickly we can return results to the user. \agp{The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.}
 

### 2015-10-12 TableSummarization/Paper/algorithms.tex
changes
tiny change

@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-We later describe a strategy for setting $m_w$.
+\mrj{We later describe a strategy for setting $m_w$.}
 
@@
 
-\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
+\stitle{Setting parameters $W$, $k$, $m_w$:} \mrj{Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
 Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
-To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.
+To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.}
 
-Parameter $k$ is the number of new rules to display upon each smart drill-down.
+\mrj{Parameter $k$ is the number of new rules to display upon each smart drill-down.
 Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
@@
 This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search.
-Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.
+Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.}
 
-$W$ is the weight function that determines which rules are interesting.
+\mrj{$W$ is the weight function that determines which rules are interesting.
 This is a function specified by the user as a black box.
@@
 In addition, the user can express interest or disinterest in certain columns by telling the system of favor or ignore those columns, via a simple graphical user interface.
-The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
+The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.}
 
-The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time.
+\mrj{The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time.
 Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
@@
 To account for sampling error, we can set $m_w$ to $2x$.
+}
 

### 2015-10-12 TableSummarization/Paper/exp.tex
tiny change

@@
 
-\agp{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
+\mrj{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
 The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}
@@
 
-\agp{
 \subsubsection{Scaling properties of our algorithms}
+\mrj{
 The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$.
@@
 That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk.
-The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.
+The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.}
 
-When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
+\mrj{When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
 When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk).

### 2015-10-12 TableSummarization/Paper/sampling.tex
tiny change

--- 2015-07-11 TableSummarization/Paper/sampling.tex
@@
 We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
+\mrj{$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.}
 The parameter $minSS$ is the minimum number of sample tuples needed
@@
 This parameter determines how accurate our count estimates will be, and also
-how quickly we can return results to the user.
+how quickly we can return results to the user. \mrj{The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.}
 

### 2015-10-16 TableSummarization/Paper/algorithms.tex
shortened significantly

@@
 the smart drill down operator.
-We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand).
+We assume that all columns are categorical (so numerical columns have been bucketized beforehand).
 \papertext{We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}}\techreporttext{We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}}.
@@
 \stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
-as listed below, has desirable approximation guarantees:
-\vspace{-5pt}
-\begin{framed}
-\vspace{-5pt}
-\begin{denselist}
-\item Set $S = \phi$
-\item For $i$ from $1$ to $k$
-\begin{denselist}
-\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
-\item $S = S \cup \left\lbrace r \right\rbrace$
-\end{denselist}
-\end{denselist}
-\vspace{-5pt}
-\end{framed}
-\vspace{-5pt}
-Since Score is a submodular function of the set $S$, 
+shown in Algorithm~\ref{algo:best-rule-set}, has desirable approximation guarantees. Since Score is a submodular function of the set $S$, 
 this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-\mrj{We later describe a strategy for setting $m_w$.}
+\papertext{In the technical report, we describe strategies for setting $m_w$ as well as other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
 
@@
 
+\techreporttext{
 \stitle{Setting parameters $W$, $k$, $m_w$:} \mrj{Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
@@
 Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
-In addition, the user can express interest or disinterest in certain columns by telling the system of favor or ignore those columns, via a simple graphical user interface.
+In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
 The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.}
 
-\mrj{The $m_w$ parameter lets the user trade off between the accuracy of the optimal rule-list, and running time.
+\mrj{The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
 Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
@@
 Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output.
-To account for sampling error, we can set $m_w$ to $2x$.
+To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
+}
 }

### 2015-10-16 TableSummarization/Paper/exp.tex
shortened significantly

@@
 That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants.
-In the worst-case wherer we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk.
+In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk.
 That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk.
@@
 When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk).
-When we have a few million tuples, we only require a few seconds.
+When we have a few million tuples, our total runtime is only a few seconds.
 But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time.

### 2015-10-16 TableSummarization/Paper/formal.tex
shortened significantly

@@
 We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ depending on how expressive the rule is (i.e., how much information
-it conveys). Let us consider some canonical forms for the function $W(r)$ and their interpretation; later, we describe the full
+it conveys). We mention some canonical forms for function $W(r)$ \papertext{(their detailed interpretations are described in the technical report)}; later, we specify the full
 family of weighting functions our techniques can handle:
@@
 the {\em size} of the rule.
-For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.
-To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules.
-We start with a blank value in every column of every table tuple.
-Then for each displayed rule $r$, we fill each of its values in the corresponding column of $MCount(r)$ tuples.
-We do this for each rule, filling a total of $\sum_{r} MCount(r)W(r)$ values.
-Thus maximising score leads to choosing the set of rules that will pre-fill as much of the table as possible.}
+For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.}
 
@@
 \stitle{Bits Weighting Function:}
+\techreporttext{
 $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
@@
 This scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table.
-If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.
+If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.}
+\papertext{$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
+This function weighs each column based on its inherent complexity, instead of equally like the Size function.}
 
@@
 
-Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$.
-This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
+Instead of clicking on a rule, the user can click on a $\star$, say in column $c$ of rule $r$.
+This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
 \techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.} This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

### 2015-10-16 TableSummarization/Paper/sampling.tex
shortened significantly

@@
 
+\papertext{
+BRS makes multiple passes over the table in order to determine the best set of rules to display.
+This can be slow when the table is too large to fit in main memory.
+We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance.
+If we had obtained a sample $s$ by selecting each table tuple with probability $p$, and run BRS on $s$, then we multiply the count of each rule found by BRS, by $\frac{1}{p}$ to estimate its count over the full table.
+
+In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples.
+We show that the problem is NP-Hard, and describe an approximate solution.
+In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process.
+Detailed descriptions of the sampling problem and the SampleHandler are in the technical report.
+The technical report also contains some additional optimizations, and ways to set the minimum sample size.
+
+\subsubsection{Deciding what to sample}\label{sec:sampling_algorithms}
+We are given a memory capacity $M$, and a minimum sample size $minSS$, both specified by the user.
+$minSS$ is the minimum number of tuples on which we are allowed to run BRS, without accessing the hard disk.
+A higher value of $minSS$ increases both accuracy and computation cost of our system.
+
+At any point, we have a tree $U$ of rules displayed to the user.
+Initially, the tree consists of a single node corresponding to the empty rule.
+When the user drills down on a rule $r$, the sub-rules of $r$ obtained by running BRS are added as children of node $r$.
+Our system maintains multiple samples in memory, with one sample per rule in $U$.
+Specifically, for each rule $r \in U$, we choose an integer $n_r$, and create a sample $s_r$ consisting of $n_r$ uniformly randomly chosen tuples covered by $r$ from the table.
+Because of the memory constraint, we must have $\sum_{r\in U} n_r \leq M$.
+
+When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$.
+If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down.
+Note that this set of tuples forms a uniformly random sample of tuples covered by $r$.
+If not, then we need to access the hard disk to obtain more tuples covered by $r$.
+
+Leaves of tree $U$ correspond to rules that may be drilled down on next.
+We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next.
+This can be a uniform distribution, or a machine learned distribution using past user data.
+We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing the hard disk.
+
+If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$.
+We denote the ratio of selectivities $x$ by $S(r',r)$.
+$S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$.
+If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$.
+We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$.
+We now formally define our problem:
+\begin{problem}\label{prob:sample-sizes}
+Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
+choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
+where the $I$'s are indicator variables, subject to :
+$\sum_{r \in U} n_r \leq M$
+\end{problem}
+Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
+%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
+We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
+\begin{lemma}\label{lemma:sampling-hardness}
+Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
+\end{lemma}
+\begin{proof}(Sketch)
+Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
+We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
+We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.
+
+We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$.
+For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$.
+Tree $U$ has $m$ special nodes $r_1, r_2, ...
+r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$.
+All other leaves have expansion probability $0$.
+The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$.
+In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes.
+Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$.
+In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$.
+Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
+So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$.
+Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
+Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem.
+Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
+\end{proof}
+\vspace{-10pt}
+\noindent A more detailed proof sketch is given in the technical report.}
+
+\techreporttext{
 Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
@@
 \end{proof}
+}
 
@@
 
-\begin{comment}
 \papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
@@
 }
-\end{comment}
 
+\techreporttext{
 Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
@@
 \subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
-\begin{comment}
 \papertext{
@@
 }
-\end{comment}
 
-%\techreporttext{
+\techreporttext{
 We now describe the design of the {\em SampleHandler}, which given
@@
 \squishend
+}
 

### 2015-10-17 TableSummarization/Paper/algorithms.tex
almost at length

@@
 \For {$i$ from $1$ to $k$}{
-$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}
+$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$
 
@@
 We now describe the function to find the best marginal rule.
-The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}.
+\papertext{The pesudo-code for the function is in the technical report}\techreporttext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}}.
 The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.

### 2015-10-19 TableSummarization/Paper/algorithms.tex
small changes throughout -- paper now under length

@@
 The expensive step in the above procedure is the step where the Score is computed for every
-single rule. Given the number of rules can be large, this can be an especially time-consuming process.
+single rule. Given the number of rules can be large, this can be especially time-consuming.
 
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-\papertext{In the technical report, we describe strategies for setting $m_w$ as well as other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
+\papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $m_w$ as well as other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
 

### 2015-10-19 TableSummarization/Paper/exp.tex
small changes throughout -- paper now under length

@@
 We now present the rule-based summaries displayed as a result of a few different user actions.
-
 To begin with, the user sees an empty rule with the total number of tuples as the count.
@@
 \begin{figure}
+%\vspace{-10pt}
 \centering
@@
 The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size).
-In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm.
-We use both the Marketing and Census datasets.
+We now study the effects of these parameters on the computation time and accuracy of our algorithm.
+We use the Marketing and Census datasets.
 The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples.
@@
 We consider one value of $minSS$ and one weight function $W$ at a time.
-For those values of $minSS$ and $W$, we drill down on the empty rule and measure the time taken for this operation.
+For those values of $minSS$ and $W$, we drill down on the empty rule and measure the time taken.
 We also measure the percent error in the estimated counts of the displayed rules.
@@
 \caption{Bits scoring\label{fig:weigtingbitwise}}
-\vspace{-10pt}
+\vspace{-15pt}
 \end{figure}
@@
 \begin{figure}
+\vspace{-10pt}
 \centering
@@
 In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk.
-That is, $a \times |T|$ is the time taken to make a single pass over the table on hard disk.
+That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk.
 The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.}

### 2015-10-19 TableSummarization/Paper/formal.tex
small changes throughout -- paper now under length

@@
 \stitle{Scoring:} We now define some additional properties of rules; these properties
-help us ``score'' individual rules as part of a rule-list.
+help us score individual rules in a rule-list.
 
@@
 The reason why we separate the scoring into these two portions is
-that they allow us to separate the inherent ``goodness'' of a rule from
+that they allow us to separate the inherent goodness of a rule from
 how much it captures the data in $\calD$.
@@
 $$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
-Overall, our goal is to choose the rule-list maximizing 
+Our goal is to choose the rule-list maximizing 
 total score.
@@
 We discuss how to modify our algorithms to use $Sum$ instead of $Count$
+\papertext{in our technical report~\cite{tr}}.
 \techreporttext{in the `Extensions' section}
-\papertext{in the technical report~\cite{tr}}.
 

### 2015-10-19 TableSummarization/Paper/intro.tex
changes
small changes throughout -- paper now under length

@@
 % useful complement.
-The best way to explain smart drill down is through a simple example.
+We now explain smart drill-down via a simple example.
 
@@
 Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
-he obtains Table~\ref{table:introexample}.
+she obtains Table~\ref{table:introexample}.
 The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
@@
 
+Say that after seeing the results of Table~\ref{table:introexample},
+the analyst wishes to dig deeper into the Walmart tuples
+represented by the last rule.
+For instance, the analyst may want to know
+which states Walmart has more sales in, or which products they sell
+the most. In this case, the analyst clicks on the Walmart rule,
+obtaining the expanded summary in Table~\ref{table:introexample2}.
+The three new rules in this table provide additional information
+about the $1000$ Walmart tuples.
+In particular, one of the new rules shows that
+Walmart sells a lot of cookies; the others show it sells a lot of products in
+the regions CA-1 and WA-5.
+
 \begin{table}
@@
 \end{tabular}
+\caption{Initial summary}\label{table:introexample0}
 \vspace{-10pt}
-\caption{Initial summary}\label{table:introexample0}
-\vspace{-5pt}
 \end{table}
@@
 \end{tabular}
-\vspace{-10pt}
 \caption{Result after first smart drill down}\label{table:introexample}
-\vspace{-10pt}
+\vspace{-20pt}
 \end{table}
@@
 \begin{table}
+\vspace{-10pt}
 \scriptsize
 \centering
-\vspace{-5pt}
 \begin{tabular}{| l | l | l | l | l |}
@@
 \end{tabular}
-\vspace{-10pt}
 \caption{Result after second smart drill down} \label{table:introexample2}
@@
 \end{table}
-
-Say that after seeing the results of Table~\ref{table:introexample},
-the analyst wishes to dig deeper into the Walmart tuples
-represented by the last rule.
-For instance, the analyst may want to know
-which states Walmart has more sales in, or which products they sell
-the most. In this case, the analyst clicks on the Walmart rule,
-obtaining the expanded summary in Table~\ref{table:introexample2}.
-The three new rules in this table provide additional information
-about the $1000$ Walmart tuples.
-In particular, one of the new rules shows that
-Walmart sells a lot of cookies; the others show it sells a lot of products in
-the regions CA-1 and WA-5.
 
@@
 
-Compared to traditional drill down, our smart drill down has two important advantages:
+\noindent {\bf Relationship to Other Work.} Compared to traditional drill down, our smart drill down has two important advantages:
 \squishlist
@@
 
-To reiterate, our chief contribution in this paper is the {\em smart drill down} interaction operator,
+\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} interaction operator,
 an extension of traditional drill down, aimed at allowing analysts to zoom into the
@@
 \squishend
-We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator.
+We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB this year~\cite{DBLP:journals/pvldb/JoglekarGP15}.
 From this point on, when we provide result snippets, these will be screenshots from our prototype tool.
@@
 
-\stitle{Overview of paper:} 
+Our tool and techniques are also part of a larger effort for
+building {\sc DataSpread}~\cite{dataspread}, a data analytics system with a spreadsheet-based front-end, and a database-based back-end, combining the benefits of spreadsheets and databases.
+
+\noindent {\bf Overview of paper:} 
 \squishlist

### 2015-10-19 TableSummarization/Paper/related.tex
small changes throughout -- paper now under length

@@
 
-There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}.
+There is work on constructing `explanation tables', sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}.
 This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values.
 A few vision papers~\cite{data-deluge, cetintemel2013query} suggest frameworks for building interactive data exploration systems.
-Some of these ideas, like maintaining user profiles, could be integrated into a smart drill down system.
+Some of these ideas, like maintaining user profiles, could be integrated into smart drill down.
 Reference~\cite{DBLP:conf/edbt/CandanCQS09} proposes an extension to OLAP drill-down that takes visualization real estate into account, by clustering attribute values.
@@
 There has also been plenty of work on pattern mining.
-Several works~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} propose non-interactive schemes that attempt to find a one shot summary of the table.
+Several papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} propose non-interactive schemes that attempt to find a one shot summary of the table.
 These schemes usually consume a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest.

### 2015-10-19 TableSummarization/Paper/sampling.tex
small changes throughout -- paper now under length

@@
 This can be a uniform distribution, or a machine learned distribution using past user data.
-We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing the hard disk.
+We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing disk.
 
@@
 $$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
-where the $I$'s are indicator variables, subject to :
+where the $I$'s are indicator variables, with :
 $\sum_{r \in U} n_r \leq M$
@@
 \end{lemma}
-\begin{proof}(Sketch)
+\begin{proof}(Sketch; details in~\cite{DBLP:journals/pvldb/JoglekarGP15})
 Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
@@
 \end{proof}
-\vspace{-10pt}
-\noindent A more detailed proof sketch is given in the technical report.}
+%\vspace{-5pt}
+%\noindent A more detailed proof sketch is given in the technical report.
+}
 
@@
 $$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
-where the $I$'s are indicator variables, subject to :
+where the $I$'s are indicator variables, subject to
 $\sum_{r \in U} n_r \leq M$
@@
 creates, maintains, retrieves, and removes samples, all in
-response to user interactions on the table. A detailed description, with some potential optimizations, can be found in the technical report.
+response to user interactions on the table. A detailed description, with some potential optimizations, can be found in~\cite{DBLP:journals/pvldb/JoglekarGP15}.
 
 At all points, the SampleHandler maintains a set of samples in memory.
-Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ and (c) a set $T_s$ of tuples from the table.
+Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ (c) a set $T_s$ of tuples from the table.
 The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$.
 The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table.
-The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.
+The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any time.
 

### 2015-12-25 TableSummarization/Paper/algorithms.tex
addressed most reviewer comments, brought back to length

--- 2015-10-17 TableSummarization/Paper/algorithms.tex
@@
 \end{lemma}
-\papertext{The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}}\techreporttext{The proof of this lemma, as well as other proofs, can be found in the appendix}.
+\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}}\techreporttext{
+\begin{proof}
+The score of rule list $R$ is given by $$\text{Score}(R) = \sum_{r \in R} W(r) \times MCount(r,R)$$
+For each tuple $t$, let $TOP(t,R)$ denote the first rule in rule-list $R$ that covers $t$.
+Then the MCount of rule $r$ in $R$ is simply $\sum_{t \in T : TOP(t,R) = r} 1$.
+Thus Score can be rewritten as:
+$$\text{Score}(R) = \sum_{t \in T} W(TOP(t,R))$$
+where we set $W(TOP(t)) = 0$ when $t$ is not covered by any rule in $R$.
+Now say we have rule lists $R$, $R'$ which have the same rules, but where $R'$ has rules in decreasing order by weight.
+For any tuple $t$ covered by $R$, let $r'$ be the highest weight rule in $R$ that covers $t$.
+Let $r$ be the first rule in $R$ that covers $t$.
+Then we have $TOP(t,R) = r$, $TOP(t, R') = r'$ and $W(r') \geq W(r)$.
+This gives us $W(TOP(t,R')) \geq W(TOP(t,R))$.
+Adding these inequalities for all $t$ covered by $R$ gives us $\text{Score}(R') \geq \text{Score}(R)$ as required.
+\end{proof}
+}. 
 Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:
@@
 
-\stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
+\stitle{High-Level Procedure:} Based on the submodularity property, the greedy procedure,
 shown in Algorithm~\ref{algo:best-rule-set}, has desirable approximation guarantees. Since Score is a submodular function of the set $S$, 
-this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.
+this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum (actually, it is $1 - (\frac{k-1}{k})^k$ for $k$ rules, which is much better for small $k$).
 
 The expensive step in the above procedure is the step where the Score is computed for every
-single rule. Given the number of rules can be large, this can be an especially time-consuming process.
+single rule. Given the number of rules can be as large as the table itself, this can be especially time-consuming.
 
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-\papertext{In the technical report, we describe strategies for setting $m_w$ as well as other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
+\papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
 
@@
 \techreporttext{
-\stitle{Setting parameters $W$, $k$, $m_w$:} \mrj{Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
+\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
 Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
-To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.}
+To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.
 
-\mrj{Parameter $k$ is the number of new rules to display upon each smart drill-down.
+Parameter $k$ is the number of new rules to display upon each smart drill-down.
 Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
@@
 This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search.
-Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.}
+Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.
 
-\mrj{$W$ is the weight function that determines which rules are interesting.
+$W$ is the weight function that determines which rules are interesting.
 This is a function specified by the user as a black box.
@@
 In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
-The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.}
+The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
 
-\mrj{The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
+The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
 Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
@@
 }
-}
 

### 2015-12-25 TableSummarization/Paper/exp.tex
addressed most reviewer comments, brought back to length

@@
 We now describe our experiments on this tool with real datasets.
-
-\begin{figure}
-\centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
-\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
-\vspace{-10pt}
-\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
-\caption{A regular drill down on Age\label{fig:drilldownexample}}
-\vspace{-10pt}
-\end{figure}
-
 \subsection{Qualitative Study}\label{sec:exp_qualitative}
@@
 Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill down, whereas smart drill down displays that information to the user with a single click.
+
+\begin{figure}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
+\vspace{-10pt}
+\end{figure}
 
@@
 
-A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules.
+A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\papertext{ (see technical report for details)}\techreporttext{.
+Specifically, if we want to perform a regular drill down on a column $C$ that has $n$ distinct values, then we set the number of rules to be displayed ($k$) to $n$.
+The weighting function $W$ is set such that $W(r) = 1$ if rule $r$ has $C$ instantiated, and $0$ otherwise.
+This ensures that all displayed rules have $C$ instantiated (with no other column $C'$ instantiated unless there is a functional dependency from $C$ to $C'$), and that each displayed rule has a distinct value of $C$.
+This effectively gives us a regular drill-down on $C$}.
 We use this to perform a drill down on the `Age' column using our experimental prototype.
@@
 \caption{A rule expansion \label{fig:uiexamplerule}}
+\vspace{-10pt}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
+\caption{A regular drill down on Age\label{fig:drilldownexample}}
 \vspace{-10pt}
@@
 
-\mrj{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
-The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}
+Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings.
+The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
 
@@
 \subsubsection{Scaling properties of our algorithms}
-\mrj{
 The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$.
@@
 That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk.
-The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.}
+The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.
 
-\mrj{When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
+When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
 When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk).
@@
 To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
-}

### 2015-12-25 TableSummarization/Paper/formal.tex
addressed most reviewer comments, brought back to length

@@
 As we will see, the weighting function does not depend on the specific
-tuples in $\calD$, but could, as we will see later, 
+tuples in $\calD$, but could 
 depend on the number of $\star$s in $r$,
@@
 $$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
-Our goal is to choose the rule-list maximizing 
+Our goal is to choose the rule-list of a given length that maximizes 
 total score.
@@
 \papertext{in our technical report~\cite{tr}}.
-\techreporttext{in the `Extensions' section}
+\techreporttext{in the `Extensions' section.}
 
@@
 \end{tabular}
-\vspace{-10pt}
 \caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
-\vspace{-15pt}
 \end{table}
@@
 Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
-By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-As an example of this operation, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
-Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.
-The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
+By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
+Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.} The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
 
@@
 This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.} This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.
-Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e.
+\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
+This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.} Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e.
 collapses it.
@@
 \end{tabular}
-\vspace{-10pt}
 \caption{Result of clicking on a $\star$ \label{table:introexample3}}
-\vspace{-10pt}
 \end{table}

### 2015-12-25 TableSummarization/Paper/intro.tex
addressed most reviewer comments, brought back to length

@@
 
-In addition to presenting this new operator called smart drill down, we also present novel sampling techniques to compute the results for this
+In addition to presenting the new smart drill down operator, we present novel sampling techniques to compute the results for this
 operator {\em in an interactive fashion} on increasingly larger databases. 
@@
 describes tuples that are not described by the first rule.
+\end{example}
 
@@
 by smart drill down to select the set of $k$ rules to display.
-
-\end{example}
 
@@
 focuses on rules with high counts, covering as
-much of the table as possible.
+much of the table as possible. Thus our work can be thought of as complementary to anomaly detection.
 Furthermore, our summarization is couched in
@@
 directs the drill down and can tailor the optimization criteria.
-Nevertheless, one can envision extending traditional
-and smart drill down to provide an additional option of anomaly detection.
 
@@
 \squishend
-We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB this year~\cite{DBLP:journals/pvldb/JoglekarGP15}.
+We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB 2015~\cite{DBLP:journals/pvldb/JoglekarGP15}.
 From this point on, when we provide result snippets, these will be screenshots from our prototype tool.

### 2015-12-25 TableSummarization/Paper/sampling.tex
addressed most reviewer comments, brought back to length

@@
 We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
-\mrj{$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.}
+$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.
 The parameter $minSS$ is the minimum number of sample tuples needed
@@
 This parameter determines how accurate our count estimates will be, and also
-how quickly we can return results to the user. \mrj{The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.}
+how quickly we can return results to the user. The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.
 

### 2016-06-25 TableSummarization/Paper/extensions.tex
added journal version with changed format

@@
 
-\begin{comment}
-The distribution of values for the $A$ may not be similar in different parts of the table.
-For instance, we may create buckets have approximately equal numbers of tuples in the table.
-But then suppose a user tries to expand a rule $r$.
-The part of the table covered by $r$ might have a very different distribution of age values.
-For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment.
-To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.
-\end{comment}
-
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}

### 2016-06-25 TableSummarization/journalVersion/smartDrillDown.tex
added journal version with changed format

%\documentclass[twoside,10pt]{IEEEtran}
\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{graphicx}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\let\proof\relax
\let\endproof\relax
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}
\usepackage[export]{adjustbox}
\usepackage[font={small,it}]{caption}

\newtheorem{theorem}{Theorem}
%\newtheorem*{lemma*}{Lemma}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{green}{Aditya: #1}}
\newcommand{\mrj}[1]{\textcolor{red}{#1}}
\newcommand{\mrjdel}[1]{\textcolor{red}{\sout{#1}}}
%\newcommand{\mrj}[1]{#1}
%\newcommand{\mrjdel}[1]{}

\newcommand{\squishlist}{
   \begin{list}{$\bullet$}
    { \setlength{\itemsep}{0pt}
      \setlength{\parsep}{2pt}
      \setlength{\topsep}{2pt}
      \setlength{\partopsep}{0pt}
    }
}
\newcommand{\stitle}[1]{\vspace{0.5em}\noindent\textbf{#1}}
\newcommand{\squishend}{\end{list}}
\newcommand{\eat}[1]{}
\newcommand{\papertext}[1]{}
\newcommand{\techreporttext}[1]{#1}

\newcommand{\calD}{\mathcal{D}\xspace}

\newenvironment{denselist}{
    \begin{list}{\small{$\bullet$}}%
    {\setlength{\itemsep}{0ex} \setlength{\topsep}{0ex}
    \setlength{\parsep}{0pt} \setlength{\itemindent}{0pt}
    \setlength{\leftmargin}{1.5em}
    \setlength{\partopsep}{0pt}}}%
    {\end{list}}
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Interactive Data Exploration with Smart Drill-Down}

\author{Manas~Joglekar,
        Hector~Garcia-Molina,~\IEEEmembership{Member,~IEEE,}
        Aditya~Parameswaran
%\thanks{H. Garcia-Molina, M. Joglekar and V. Verroios are with Stanford University}%
%\thanks{A. Marcus is with Unlimited Labs}%
%\thanks{A. Parameswaran is with the University of Illinois (UIUC)}
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem M. Joglekar is with Google.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: manasrj@google.com
\IEEEcompsocthanksitem H. Garcia-Molina is with Stanford University.\protect\\
E-mail: hector@cs.stanford.edu
\IEEEcompsocthanksitem A. Parameswaran is with the University of Illinois (UIUC).\protect\\
E-mail: adityagp@illinois.edu
}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
}% <-this % stops a space

% \author{Hector~Garcia-Molina,~\IEEEmembership{Member,~IEEE,}
%         Manas~Joglekar,
%         Adam~Marcus,
%         Aditya~Parameswaran,
%         Vasilis~Verroios\footnote{H. Garcia-Molina, M. Joglekar and V. Verroios are with Stanford University;
% A. Marcus is with Unlimited Labs,
% A. Parameswaran is with the University of Illinois (UIUC).}}

\IEEEtitleabstractindextext{%

\begin{abstract}
We present {\em smart drill-down},
an operator for interactively exploring a relational table
to discover and summarize ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a thousand tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill-down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively apply smart drill-down on an existing rule to
explore that part of the table.
We demonstrate that the underlying optimization problems 
are {\sc NP-Hard}, and describe an
algorithm for finding the approximately optimal list of rules to display when the user uses a smart drill-down, and a dynamic sampling scheme for efficiently interacting with large tables. Finally, we
perform experiments on real datasets on our experimental prototype to demonstrate the usefulness of smart drill-down and study the performance of our algorithms.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
ADD-OTHER-IEEE-KEYWORDS
\end{IEEEkeywords}
}

\maketitle

\section{Introduction}
Analysts often use OLAP (Online Analytical Processing) operations
such as drill down (and roll up)~\cite{export:69578} to explore
relational databases. 
These operations are very useful for analytics and data exploration and have stood the test of time;
all commercial OLAP systems 
in existence support these operations. (Recent reports estimate the size of the OLAP market to be \$10+ Billion~\cite{gartner}.)

However, there are cases where drill down is ineffective; 
for example, when the number of distinct values
in a column is large, vanilla drill down 
could easily overwhelm analysts by presenting them with too many 
results (i.e., aggregates). 
Further, drill down only allows us to instantiate values
 one column at a time, instead of allowing simultaneous drill downs
on multiple columns---this simultaneous drill down on multiple columns 
could once again suffer from the problem
of having too many results, stemming from many distinct combinations of column values.

In this paper, we present a new interaction operator 
that is an extension to the traditional 
drill down operator, aimed at providing {\em complementary}
functionality to drill down in cases where drill down is
ineffective. We call our operator {\em smart drill down}.
At a high level, smart drill down lets analysts zoom into
the more ``interesting'' parts of a table or a database,
with fewer operations, and without having to examine as much
data as traditional drill down.
Note that our goal is {\em not} to replace traditional 
drill down functionality, which we believe is fundamental;
instead, our goal is to provide auxiliary functionality 
which analysts are free to use whenever they find 
traditional drill downs ineffective.

In addition to presenting the new smart drill down operator, we present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

% drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
% relational tables.
% In this paper we study an extension of drill down
% that lets users zoom into the more ``interesting'' parts of a table,
% with fewer operations and without having to examine as much data
% as traditional drill down.
% We call our approach {\em smart drill down};
% it does not replace traditional drill down but we will argue is a very
% useful complement.
We now explain smart drill-down via a simple example.

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by a Sum aggregate over some measure column,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the operator displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
she obtains Table~\ref{table:introexample}.
The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
that display noteworthy or interesting drill downs.
The number $3$ is a user specified parameter, which we call $k$.

For example, the tuple (Target, bicycles, $\star$, $200$, $2$)
says that there are $200$ tuples (out of the 6000) with
Target as the first column value and bicycle as the second.
This fact tells the analyst that Target is selling a lot of bicycles.
The next tuple tells the analyst that comforters are selling well in
the MA-3 region, across multiple stores. The last tuple
states that Walmart is doing well in general over multiple products and regions.
We call each tuple in Table~\ref{table:introexample} a {\em rule}
to distinguish it from the tuples in the original table that is being explored.
Each rule summarizes the set of tuples that are described by it.
Again, instead of Count, the operator can display a Sum aggregate, such as
the total Sales.

Say that after seeing the results of Table~\ref{table:introexample},
the analyst wishes to dig deeper into the Walmart tuples
represented by the last rule.
For instance, the analyst may want to know
which states Walmart has more sales in, or which products they sell
the most. In this case, the analyst clicks on the Walmart rule,
obtaining the expanded summary in Table~\ref{table:introexample2}.
The three new rules in this table provide additional information
about the $1000$ Walmart tuples.
In particular, one of the new rules shows that
Walmart sells a lot of cookies; the others show it sells a lot of products in
the regions CA-1 and WA-5.

\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Initial summary}\label{table:introexample0}
\vspace{-10pt}
\end{table}

\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \hline
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Result after first smart drill down}\label{table:introexample}
\vspace{-20pt}
\end{table}

\begin{table}
\vspace{-10pt}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & CA-1 & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & WA-5 & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Result after second smart drill down} \label{table:introexample2}
\vspace{-10pt}
\end{table}

When the analyst clicks on a rule $r$, smart drill down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if it contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.
\end{example}

In this paper we describe how to combine or blend these three factors
in order to obtain a single desirability score for a set of rules.
Our score function can actually be tuned by the analyst
(by specifying how weights are computed),
providing significant flexibility in what is considered a good set of rules.
We also present an efficient optimization procedure to maximize score, invoked
by smart drill down to select the set of $k$ rules to display.

\noindent {\bf Relationship to Other Work.} Compared to traditional drill down, our smart drill down has two important advantages:
\squishlist
\item
Smart drill down limits the information displayed
to the most interesting $k$ facts (rules).
With traditional drill down, a column is expanded and {\em all}
attribute values are displayed in arbitrary order.
In our example, if we drill down on say the store attribute,
we would see all stores listed, which may be a very large number.
\item
Smart drill down explores several attributes to open up together,
and automatically selects combinations that are interesting.
For example, in Table~\ref{table:introexample},
the rule (Target, bicycles, $\star$, $200$, $2$)
is obtained after a single drill down;
with a traditional approach, the analyst would first have to drill down on
Store, examine the results, drill down on Product,
look through all the displayed rules and then find the interesting rule
(Target, bicycles, $\star$, $200$, $2$).
\squishend

Incidentally, note that in the example we only described one type of smart drill down,
where the analyst selects a {\em rule} to drill down on
(e.g., the Walmart rule going from Table~\ref{table:introexample} to
Table~\ref{table:introexample2}).
In Section~\ref{sec:interface} we describe another option
where the analyst clicks on a $\star$ in a column to obtain
rules that have non-$\star$ values in that column.

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible. Thus our work can be thought of as complementary to anomaly detection.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.

Our work is also related to pattern mining. Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, but do not propose interactive mechanisms. Moreover, to the best of our knowledge, other pattern mining work is either not flexible enough~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117}, restricting the amount of tuning the user can perform, or so general~\cite{Leeuwen:2012:DSS:2347179.2347243} as to preclude efficient optimization. Our work also merges `interesting pattern mining' into the OLAP framework. We discuss related work in detail in Section~\ref{sec:related}.

\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
\squishend
We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB 2015~\cite{DBLP:journals/pvldb/JoglekarGP15}. From this point on, when we provide result snippets, these will be screenshots from our prototype tool. Our prototype tool also supports traditional drill-down: smart drill-down 
can be viewed as a generalization of traditional drill-down (with the weighting function set appropriately). In Section~\ref{sec:exp_qualitative}, we compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results.

Our tool and techniques are also part of a larger effort for
building {\sc DataSpread}~\cite{dataspread}, a data analytics system with a spreadsheet-based front-end, and a database-based back-end, combining the benefits of spreadsheets and databases.

\noindent {\bf Overview of paper:} 
\squishlist

\item In Section~\ref{sec:formal}, we formally define smart drill down. After that, we describe different schemes for weighting rules, and our interactive user interface.

\item In Section~\ref{sec:algorithms}, we present our algorithms for
finding optimal sets of rules, as well as our dynamic sampling schemes
for dealing with large tables.

\item Based on our implemented smart drill down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on real datasets,
and show additional examples of smart drill down in action.

\item We describe related work in Section~\ref{sec:related}, and conclude in Section~\ref{sec:conclusion}.
\squishend

\section{Formal Description}\label{sec:formal}
We describe our formal problem
in Section~\ref{sec:preliminaries},
describe different scoring functions
in Section~\ref{sec:weighting},
and describe our operator interfaces
in Section~\ref{sec:interface}. 
\subsection{Preliminaries and Definitions}
\label{sec:preliminaries}

\stitle{Tables and Rules:} As in a traditional OLAP setting, we assume we are given a 
star or snowflake schema;
for simplicity, we represent this schema using a single denormalized relational table,
which we call $\calD$. 
For the purpose of the rest of the discussion, we will operate on this
table $\calD$.
We let $T$ denote the set of tuples in $\calD$, and $C$ denote 
the set of columns in $\calD$.

Our objective (formally defined later) is to 
enable smart drill downs on this table or on portions of it:
the result of our drill downs are lists of {\em rules}. 
A {\em rule} is a tuple with a value for each column of the table. 
In addition, a rule has other attributes, such as count and weight 
(which we define later) associated with it. 
The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. The {\em trivial rule} is one that has a $\star$ value in all columns. The {\em Size} of a rule is defined as the number of non-starred values in that rule.

\stitle{Coverage:} A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$. At a high level, we are interested
in identifying rules that cover many tuples. We next define the concept of subsumption that allow us to 
relate the coverage of different rules to each other.

We say that rule $r_1$ is a {\em sub-rule} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) is a sub-rule of ($a$, $b$). If $r_1$ is a sub-rule of $r_2$, then we also say that $r_2$ is a {\em super-rule} of $r_1$. If $r_1$ is a sub-rule of $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system in response to a smart drill down operation. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a new rule-list below $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

\stitle{Scoring:} We now define some additional properties of rules; these properties
help us score individual rules in a rule-list.

There are two portions that constitute our scores for a rule as part of a rule list. 
The first portion dictates how much the rule $r$ ``covers'' the tuples in $\calD$;
the second portion dictates how ``good'' the rule $r$ is (independent of how many
tuples it covers). 
The reason why we separate the scoring into these two portions is
that they allow us to separate the inherent goodness of a rule from
how much it captures the data in $\calD$.

We now describe the first portion:
we define {\em Count}($r$) as the total number of tuples $t \in T$ that are covered by $r$. 
Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') as the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user
as part of the smart drill down result, to increase the coverage of the rule-list.

Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
on monotonic weighting functions because we prefer 
rules that are more ``specific''
rather than those that are more ``general'' 
(thereby conveying less information).
We further describe our weighting functions in Section~\ref{sec:weighting}.

Thus, the total score for our list of rules is given by 
$$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
Our goal is to choose the rule-list of a given length that maximizes 
total score.

We use $MCount$ rather than Count in the above equation to ensure that we do not redundantly cover
the same tuples multiple times using multiple rules, and thereby increase coverage of the table. 
If we had defined total score as $\sum_{r \in R} \text{Count}(r)W(r)$, then our optimal rule-list could 
contain rules that repeatedly refer to the most `summarizable' part of the table. 
For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then 
for some weighting functions $W$, 
the summary may potentially consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.

Our smart drill downs still display the Count of each rule rather than the $MCount$. This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret. In any case, it would be a simple extension to display MCount in another column.

\stitle{Formal Problem:} We now formally define our problem:
\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r,R)$$
for one of the following smart drill down operations:
\begin{denselist}
\item $[$Rule drill down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star drill down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{denselist}
\end{problem}
\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ 
\papertext{in our technical report~\cite{tr}}.
\techreporttext{in the `Extensions' section.}

\subsection{Weighting Rules}
\label{sec:weighting}
We now describe our weighting function $W$ that is used to score individual rules. 
At a high level, we want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ depending on how expressive the rule is (i.e., how much information
it conveys). We mention some canonical forms for function $W(r)$ \papertext{(their detailed interpretations are described in the technical report)}; later, we specify the full
family of weighting functions our techniques can handle:

\stitle{Size Weighting Function:} 
\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table~\ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).} \papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$ i.e. the {\em size} of the rule. For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.}

\techreporttext{To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus, this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those tuples. Thus, $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus, this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.}

\stitle{Bits Weighting Function:}
\techreporttext{
$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$. This scoring function weighs each column depending on its inherent complexity, instead of equally like the Size function. The reason behind is: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Except that this time, we count the number of `bits' of the table that are pre-filled by the rules. This is because for column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information. This scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.}
\papertext{$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$. This function weighs each column based on its inherent complexity, instead of equally like the Size function.}

\smallskip

\stitle{Other Weighting Functions:}
Even though we have given two example weighting functions here, our algorithms allow the user to leverage any weighting function $W$, subject to two conditions:
\squishlist
\item Non-negativity: For all rules $r$, $W(r) \geq 0$.
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\squishend
A weight function can be used in several ways, including expressing a higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or expressing indifference towards a column (by adding zero weight for having non-$\star$ value in that column).

\techreporttext{
\begin{table}
\centering
\scriptsize
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\end{table}
}

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in Table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.} The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on a $\star$, say in column $c$ of rule $r$. This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. \techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.} Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. \techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}. This operation is equivalent to a traditional
roll up, but for smart drill downs instead of traditional drill downs.}

\techreporttext{
\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}
}

\section{Smart drill down Algorithms} \label{sec:algorithms}
We now describe online algorithms for implementing
the smart drill down operator. We assume that all columns are categorical (so numerical columns have been bucketized beforehand). \papertext{We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}}\techreporttext{We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}}.

\subsection{Problem Reduction and Important Property} \label{sec:reduction}
When the user drills down on a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$. We can reduce Problem~\ref{prob:optimal-subrule-list} to the following simpler problem by removing the user-interaction based constraints:

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a monotonic weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r,R)$$
\end{problem}

\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\begin{enumerate}
\item $[$Rule drill down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star drill down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\end{enumerate}

\begin{algorithm}
\scriptsize
\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
\KwOut{$S$ (Solution set of rules)}
$S = \phi$

\For {$i$ from $1$ to $k$}{
$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$

$S = S \cup \left\lbrace R_m \right\rbrace$
}
\Return $S$
\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}\label{algo:best-rule-set}}
\end{algorithm}

As a first step towards solving Problem~\ref{prob:optimal-rule-list}, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight. Note that the weight of a rule is independent of its $MCount$. The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score. Thus, if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$\text{Score}(R^{\prime}) \geq \text{Score}(R)$.
\end{lemma}
\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}}\techreporttext{
\begin{proof}
The score of rule list $R$ is given by $$\text{Score}(R) = \sum_{r \in R} W(r) \times MCount(r,R)$$
For each tuple $t$, let $TOP(t,R)$ denote the first rule in rule-list $R$ that covers $t$. Then the MCount of rule $r$ in $R$ is simply $\sum_{t \in T : TOP(t,R) = r} 1$. Thus Score can be rewritten as:
$$\text{Score}(R) = \sum_{t \in T} W(TOP(t,R))$$
where we set $W(TOP(t)) = 0$ when $t$ is not covered by any rule in $R$. Now say we have rule lists $R$, $R'$ which have the same rules, but where $R'$ has rules in decreasing order by weight. For any tuple $t$ covered by $R$, let $r'$ be the highest weight rule in $R$ that covers $t$. Let $r$ be the first rule in $R$ that covers $t$. Then we have $TOP(t,R) = r$, $TOP(t, R') = r'$ and $W(r') \geq W(r)$. This gives us $W(TOP(t,R')) \geq W(TOP(t,R))$. Adding these inequalities for all $t$ covered by $R$ gives us $\text{Score}(R') \geq \text{Score}(R)$ as required.
\end{proof}
}. 
Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:

\begin{definition}\label{def:set-score}
Let $R$ be a set of rules. Then the Score of $R$ is
$\text{Score}(R) = \text{Score}(R^{\prime})$
where $R^{\prime}$ is the list of rules obtained by ordering the rules in the set $R$ in decreasing order by weight.
\end{definition}

This gives us a reduced version of Problem~\ref{prob:optimal-rule-list}: 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a monotonic weight function $W$, and a number $k$, find the set (not list) $R$ of $k$ rules which maximizes Score($R$) as defined in Definition~\ref{def:set-score}.
\end{problem}

\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now first show that Problem~\ref{prob:optimal-rule-set}, and consequently Problem~\ref{prob:optimal-subrule-list} and Problem~\ref{prob:optimal-rule-list} are {\sc NP-Hard}, and then present an approximation algorithm for solving Problem~\ref{prob:optimal-rule-set}.

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximizes $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus, the goal of the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We can reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
\begin{lemma}
Problem~\ref{prob:optimal-rule-set} is {\sc NP-Hard}.
\end{lemma}

\subsection{Algorithm Overview}\label{sec:alg-overview}
Given that the problem is {\sc NP-Hard}, we now present our algorithms 
for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
The problem consists of finding a set of rules, given size $k$, that maximizes 
Score.

The next few sections fully develop the
details of our solution:

\squishlist
\item We show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm. At a high level, this greedy algorithm is simple to state. The algorithm runs for $k$ steps;
we start with an empty rule set $R$, and then at each step, we add the next best rule that maximizes Score

\item In order to find the rule $r$ to add in each step, we need to measure the impact on Score for each $r$. This is done in several passes over the table, using ideas from the a-priori algorithm~\cite{apriori} for frequent item-set mining. 
\squishend

\techreporttext{
\begin{algorithm}
\scriptsize
\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
$H = 0$ \tcc*{Threshold for deciding if to count for a rule.}
$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
\For{$j$ from $1$ to number of columns in $T$}{
\If {$j = 1$}{
$C_n = \text{ all rules of size } 1$
}
\Else{ 
$C_n = \text{all size-}i$ super-rules of rules from $C_o$
}
\ForEach {$R \in C_n$}{
$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
}
\If {$(M < H)$}{ 
$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to be in the solution}
}
}
\If {$C_n = \phi$}{
break;
}
\ForEach {$R \in C_n$}{
Count$(R) = 0$ \tcc*{Initialize}
MarginalValue$(R) = 0$ \tcc*{Initialize}
}
\ForEach {$t \in T$}{
Let $R_S$ be the highest weight rule in $S$ that covers $t$

\ForEach {$R \in C_n$ that covers $t$}{
Count($R$) $++$

MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
}
}
$C = C \cup C_n$

$C_o = C_n$

$C_n = \phi$

$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
}
\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
\caption{Find best marginal rule\label{algo:best-marginal-rule}}
\end{algorithm}
}

In some cases, the dataset may still be too large for us to return a good rule set in
a reasonable time; in such cases, we may want to run our algorithm on a sample of the table
rather than the entire table. In Section~\ref{sec:sampling}, we describe a scheme 
for maintaining multiple samples in memory and using them to improve response 
time for different drill down operations performed by the user. Our sampling scheme dynamically adapts to the current interaction scenario that the user is in; drawing from ideas in approximation algorithms and optimization theory.

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
\stitle{Submodularity:} We will now show that the Score function over sets of rules has a property called {\em submodularity}, giving us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$: $f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, it is well known that the solution to the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.

\begin{lemma}\label{lemma:submodular}
For a given table $T$, the Score function over sets $S$ of rules, defined by the following is submodular:
$$\text{Score}(S) = \sum_{r \in S} MCount(r,S)W(r)$$
\end{lemma}
\vspace{-10pt}

\stitle{High-Level Procedure:} Based on the submodularity property, the greedy procedure,
shown in Algorithm~\ref{algo:best-rule-set}, has desirable approximation guarantees. Since Score is a submodular function of the set $S$, 
this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum (actually, it is $1 - (\frac{k-1}{k})^k$ for $k$ rules, which is much better for small $k$).

The expensive step in the above procedure is the step where the Score is computed for every
single rule. Given the number of rules can be as large as the table itself, this can be especially time-consuming.

Instead of using the procedure described above directly, we instead develop a ``parameterized'' version 
that will admit further approximation (depending on the parameter) in order to reduce computation further. We describe
this algorithm next.

\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labeled Algorithm~\ref{algo:best-rule-set}. We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et). BRS takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}

% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary. Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.

BRS initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step. To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.

\stitle{Finding the Best Marginal Rule:} In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

In order to avoid counting too many rules, we leverage a technique inspired by the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}. Recall that the a-priori algorithm is used to find all frequent itemsets that have a support greater than a threshold. Unlike the a-priori algorithm, our goal is to find the single best marginal rule. Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset.

We compute the best marginal rule over multiple passes on the dataset, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts and marginal values for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the counts and marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts and marginal values of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

We now describe the function to find the best marginal rule. \papertext{The pesudo-code for the function is in the technical report}\techreporttext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}}. The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far. The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted in all previous iterations. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous iteration. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

For the second pass onwards, we are more selective about which rules to consider for marginal value evaluation. We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$. Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2. Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule. Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5. If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

The reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently. This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.

\techreporttext{
\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

Parameter $k$ is the number of new rules to display upon each smart drill-down. Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information. On the other hand, small values of $k$ may display too little information about the table. Fortunately, the BRS algorithm is incremental in nature. That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$. Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found. This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search. Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.

$W$ is the weight function that determines which rules are interesting. This is a function specified by the user as a black box. Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu. In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface. The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.

The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
}

\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}

\papertext{
BRS makes multiple passes over the table in order to determine the best set of rules to display. This can be slow when the table is too large to fit in main memory. We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance. If we had obtained a sample $s$ by selecting each table tuple with probability $p$, and run BRS on $s$, then we multiply the count of each rule found by BRS, by $\frac{1}{p}$ to estimate its count over the full table.

In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples. We show that the problem is NP-Hard, and describe an approximate solution. In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process. Detailed descriptions of the sampling problem and the SampleHandler are in the technical report. The technical report also contains some additional optimizations, and ways to set the minimum sample size.

\subsubsection{Deciding what to sample}\label{sec:sampling_algorithms}
We are given a memory capacity $M$, and a minimum sample size $minSS$, both specified by the user. $minSS$ is the minimum number of tuples on which we are allowed to run BRS, without accessing the hard disk. A higher value of $minSS$ increases both accuracy and computation cost of our system.

At any point, we have a tree $U$ of rules displayed to the user. Initially, the tree consists of a single node corresponding to the empty rule. When the user drills down on a rule $r$, the sub-rules of $r$ obtained by running BRS are added as children of node $r$. 
Our system maintains multiple samples in memory, with one sample per rule in $U$. Specifically, for each rule $r \in U$, we choose an integer $n_r$, and create a sample $s_r$ consisting of $n_r$ uniformly randomly chosen tuples covered by $r$ from the table. Because of the memory constraint, we must have $\sum_{r\in U} n_r \leq M$.

When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$. If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down. Note that this set of tuples forms a uniformly random sample of tuples covered by $r$. If not, then we need to access the hard disk to obtain more tuples covered by $r$.

Leaves of tree $U$ correspond to rules that may be drilled down on next. We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next. This can be a uniform distribution, or a machine learned distribution using past user data. We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing disk.

If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$. We denote the ratio of selectivities $x$ by $S(r',r)$. $S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$. If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$. We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$. We now formally define our problem: 
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, with :
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch; details in~\cite{DBLP:journals/pvldb/JoglekarGP15})
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$. All other leaves have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes. Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$. In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$. Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem. Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}
%\vspace{-5pt}
%\noindent A more detailed proof sketch is given in the technical report.
}

\techreporttext{
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be expensive if the table is large and the table does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

But if we are willing to accept approximate counts rather than exact counts, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. Thankfully since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.

We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms}. 
Then, in Section~\ref{sec:sample-handler}, we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Finally, we mention some additional optimizations we can make, and
describe how we can set the minimum sample size required from the SampleHandler.

\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}

We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user). $M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.
The parameter $minSS$ is the minimum number of sample tuples needed
such that we can use the sample in memory 
instead of having to resort to the entire table stored
on hard disk. 
This parameter determines how accurate our count estimates will be, and also
how quickly we can return results to the user. The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.

We now consider the following problem: say we currently have no
samples in memory (we describe the scenario where there are already 
some samples in Section~\ref{sec:sample-handler}), and say the user is currently viewing some rules; how do we
materialize the ``best possible'' samples in memory that fit within the capacity $M$,
such that we can respond to as many user interactions as possible using the stored samples,
without having to resort to retrieving the entire table.
That is, we want to maximize the probability that the next user interaction
can be answered using the existing samples, without reading from the hard disk. 
We leverage techniques from approximation algorithms and optimization theory, formalized
below.

\stitle{Tree of Rules.}
At any stage, we have a tree $U$ of rules displayed to the user, 
with each node of the tree corresponding to a displayed rule. 
In the rest of this section, we will refer to nodes in $U$ and rules interchangeably.
The tree is formed as follows: The root of the tree corresponds to the trivial rule. 
And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. 
Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$, and so on.

Even though the rules displayed to the user can be considered a tree, 
multiple nodes of the tree may correspond to the same rule. 
For example, $(a, b)$ may be a child of both $(\star, b)$ and $(a, \star)$.
Thus, even though the structure displayed to the user is a tree, 
the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering. 
However, for the purposes of this section, we focus on the tree representation.

Internal nodes of the tree $U$ are ones that have been expanded (drilled down on), 
while the leaves are nodes that have not been expanded. 
Let $L$ be the set of leaves. 
Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.

\stitle{Probability Distribution.} 
We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded. In the absence of additional data, we can assume a uniform probability distribution. That is, we can assume that every leaf is equally likely to be expanded next. If we have data on past user behaviour available, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `distance from last expanded node' to get a better probability estimate of each node being expanded next.

% When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to make a pass through the entire table, as well as potentially shrink some existing samples to free up memory. Since making a pass over the the table from the hard disk is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass. Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also potentially creates other new samples, or resizes existing samples. For each displayed rule $r^{\prime}$, it may create a new sample for $r^{\prime}$ in the same pass through the table. Based on the current leaves and their expansion probabilities, it finds an optimal set of sizes of samples to create for each displayed rule. More specifically, for every displayed rule $r^{\prime}$, it determines an integer $n_{r^{\prime}}$, and then creates a fresh sample $s_{r^{\prime}}$ with $f_{s_{r^{\prime}}} = r^{\prime}$ and $|T_{s_{r^{\prime}}}| = n_{r^{\prime}}$ while making its pass through the table.

\stitle{Sampling Strategy.} The sampling strategy we adopt involves storing, 
for every displayed rule $r'$, a sample of $r'$ of size $n_{r'}$, i.e., 
containing $n_{r'}$ randomly chosen tuples $\in r'$.
Note that we may choose to not store a sample for some rules, in which case $n_{r'}$
will be zero for those rules. 
We need to pick the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory. 
As it turns out, picking a sample for some rules can help not just that rule, but
also all of its sub-rules, all the while preserving uniform randomness of samples. We formalize this notion next.

\stitle{Selectivity.} 
Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule. 
For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples. We denote this quantity as $S(r_1, r_2)$. 
We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. 
If the same rule occurs in multiple nodes $r_1$, $r_2$ of the tree, then $S(r_1, r_2)$ is naturally $1$. 
Essentially, $S(r_1, r_2)$ denotes how much $r_1$'s sample helps $r_2$. 
If $r_1$ is a strict super-rule of $r_2$, then $S$ will be $0$ since using a sample of $r_1$
for $r_2$ will lead to bias.

Then, if we have an $n_r$ sized uniformly random sample of tuples covered by $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is denoted as $ess(r')$ (for `effective sample size'), as below: 
\begin{equation} \label{def:ess}
ess(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r
\end{equation}
Basically, $ess(r')$ captures how much of an unbiased sample of $r'$ can be retrieved using all the samples for the rules in $U$.
If $ess(r^{\prime}) \geq minSS$, then if the user expands $r^{\prime}$, we do not need to make another 
pass through the table.

We wish to set the sample size $n_r$ of each rule so as to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, subject to
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch)
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two children $r_{i,1}, r_{i,2}$. These $2m$ children are all leaves, and all leaves other than these have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child. Therefore, each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else. Thus, $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$. In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$. As a result, in any optimal solution, $ess(r_{i,1})$ must be equal to $minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$. For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$). The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$). Moreover, the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$. Hence, solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}
}

\stitle{Approximate DP Solution.} Even though the problem as stated is {\sc NP-Hard}, with an additional simplifying assumption, we can make the problem approximately solvable using a Dynamic Programming algorithm. The assumption is: for each $r \in L$, we assume that its $ess$ can get sample tuples only from samples obtained for itself and its immediate parent. That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$. This is similar to what we had for tree $U$ in our proof of Lemma~\ref{lemma:sampling-hardness}. $ess(r)$ now becomes $ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$ where $r$ is the parent of $r^{\prime}$.

\papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
\squishlist
\item \textbf{Locally Optimal Solutions :} For each rule $r_0 \in U \setminus L$, let $M_{r_0}$ be the set consisting of $r_0$ and its leaf children. We consider the number of tuples assigned to members of $M_{r_0}$. Each such assignment of numbers has a `cost' equal to total number of tuples assigned, and `value' equal to sum of probabilities of $M_{r_0}$ members with $ess > minSS$. We consider all `locally optimal' assignments, where probability value cannot be increased without increasing the cost.
\item \textbf{Combining Solutions :} Then, we use dynamic programming to pick one optimal solution per $r_0$ so as to get maximum probability value with the given cost constraint. We discretize costs, say to have granularity $100$. Let $M_{0}, M_{1}, ... M_{D}$ be the sets for which we have found locally optimal solutions. We make $D$ iterations in all, where in the $i^{th}$ iteration, we try different locally optimal solutions for $M_i$ and use it to update the maximum attainable value for each cost. 
\squishend
}

\techreporttext{
Now consider a rule in $r_0 \in U \setminus L$ along with all its children. Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children. Then, by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the $ess$ value of rules in $M_{r_0}$. This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$. Thus, for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$. Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$. Then, we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s. We describe both these steps in detail below:

Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$. For any child $r_i$, $n_{r_i}$ only contributes to its own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ... r_d$. Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
\squishlist
\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$. Both these decreases would decrease sampling cost without affecting probability score. 
\squishend
Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess = minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution. Consider any such locally optimal solution $e$. For $e$ let children $r_{i_1}, r_{i_2}, ... r_{i_m}$ be in the first category, $r_{i_{m+1}}, .. r_{i_M}$ be in the second category, and $r_{i_{M+1}}, .. r_{i_d}$ in the third.  Then the `probability value' of solution $e$ is given by : 
$P(e) = \sum_{j = 1}^{i_M}p_{j}$,
and its `Sampling Cost' is
$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
Thus, there are at most $3^d$ locally optimal solutions; $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.

Then next step is to combine the solutions using dynamic programming. Let the $M$ sets be called $M_0, M_1, ... M_D$. Let our possible sample sizes range from $0$ to $\mathcal{S}$. The number of sample sizes can be pretty large ($\mathcal{S}$), but we can make it smaller by discretizing the sample sizes, say to have granularity $100$. Then we create a $D \times \mathcal{S}$ array $A$. The value $A\left[i\right]\left[j\right]$ contains the best probability score we can get from $M_0, M_1, ... M_i$ with total sample size at most $j$. We can populate $A\left[0\right]\left[j\right] \forall j$ using the locally optimal solutions for $M_0$. Let $E_{i+1}$ denote the set of locally optimal solutions for $M_{i+1}$. Then we have,
$$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
This can be solved using dynamic programming, in $O(D\mathcal{S}3^d)$ time. 
}

\techreporttext{
\subsubsection{Alternative Convex-Optimization based solution}
We noted earlier that Problem~\ref{prob:sample-sizes} is NP-Hard, but can be approximately solved with an additional simplifying assumption regarding the $S(r_1, r_2)$ values. Instead of making this simplification, we can make the problem convex (and hence tractable) with two different simplifications. The first simplification is, we modify our objective function to use hinge-loss instead of a step function. That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$. The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers. After obtaining our optimal sample sizes, we can round them up to get integer sample sizes. This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree. $|U|$ is usually negligible compared to the memory capacity $M$, or $minSS$.

In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function). Thus, our new optimization problem becomes
\begin{problem}\label{prob:sample-sizes-hinge-loss}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
subject to :
$$\sum_{r \in U} n_r \leq M$$
\end{problem}
The constraint is linear in the $n_r$ variables, and hence convex. Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex. The constant function $-1$ is convex as well. Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum. Thus, we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.

The main weakness of this approach is that the hinge-loss objective rewards values of $ess < minSS$, which may lead us to all leaves having large $ess$ values that are nonetheless less than $minSS$, and thus gives lower quality count estimates than required by the user.

\subsubsection{Additional optimizations}
There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
Suppose we have a sample $s$, and say its filter rule $f_s$ has value $v$ in column $c$. Then we know that each tuple $t$ in $T_s$ must also have value $v$ in column $c$, since it is covered by $f_s$. So we do not need to explicitly store the column $c$ value of any tuple in $T_s$. We only need to store the tuple values of columns that have a $\star$ value in $f_s$.
In addition, we may have a tuple occur in multiple samples. Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.

\subsubsection{Setting $minSS$}
Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e. $x|T|$ tuples. Say we have a uniform random sample $s$ of $T$. The samples has size $|T_s|$, and let $X_{r,s}$ be the random variable denoting the number of tuples of $T_s$ covered by $r$. Then $E\left[ X_{r,s} \right] = x|T_s|$, and $\text{Dev}(X_{r,s}) \approx \sqrt{|T_s|x(1-x)}$. In order to get a good estimate of $x$ (and hence of Count$(r) = x|T|$), we want $E\left[X_{r,s}\right] >> \text{Dev}(X_{r,s})$. That is, $x|T_s| >> \sqrt{|T_s|x(1-x)} \Leftrightarrow \frac{x|T_s|}{1-x} >> 1$.

We want to set the parameter minSS such that we get good count estimates for rules when using a sample of size $|T_s| = minSS$. If a rule displayed in our summary has covers $x$ fraction of the tuples, we want minSS to be at least $\rho\frac{1-x}{x}$, So the value of minSS must be at least $\rho\frac{1-x}{x}$ where $\rho$ is a constant chosen by us based on how accurate we want the count estimate to be. Moreover, since we want good Count estimates for all rules displayed in the summary, we want $minSS >> \rho\frac{1-x}{x}$ where $x$ is the minimum fraction of tuples covered by any of the rules displayed in our summary.

Thus, a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$. This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule). Let $c$ be the column with the fewest distinct values. Say it has $|c|$ values. Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$. For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times. So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.

The highest scoring rule can have weight at most $|C|$ (the total number of columns). Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$. Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS. For example, if $|T| = 10000$, $|c| = 5$, $|C| = 10$, then we want $minSS >> 5 \times 10$.
}

\subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
\papertext{
We now briefly describe the design of the {\em SampleHandler}, which given
a certain memory capacity $M$, and a minimum sample size $minSS$,
creates, maintains, retrieves, and removes samples, all in
response to user interactions on the table. A detailed description, with some potential optimizations, can be found in~\cite{DBLP:journals/pvldb/JoglekarGP15}.

At all points, the SampleHandler maintains a set of samples in memory. Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ (c) a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any time.

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. The $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$. The SampleHandler also computes $N_s$ when a sample is created. Then we run BRS on sample $s$ to obtain the list of rules to display. The SampleHandler uses one of the mechanisms below to obtain such a sample $s$:

\squishlist
\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e. $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$.

\item \textbf{Combine:} If \textbf{Find} doesn't work i.e., if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$. If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.

\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.  Since accessing the hard disk is expensive, it also creates samples for rules other than $r$ and augments existing samples in the same pass. It first deletes all existing samples, then uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine how many tuples to sample for each rule, and uses reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to create a uniformly random sample of a given size for each rule.
\squishend
}

\techreporttext{
We now describe the design of the {\em SampleHandler}, which given
a certain memory capacity $M$, and a minimum sample size $minSS$,
creates, maintains, retrieves, and removes samples, all in
response to user interactions on the table.
It uses the algorithms from Section~\ref{sec:sampling_algorithms} 
to decide which samples to create, as we will see below.

At all points, the SampleHandler maintains a set of samples in memory. For instance, it may keep a sample of tuples used to expand the first (trivial) rule, and another sample used to expand the rule last clicked on by the user. 
Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ and (c) a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. Thus, the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$. The SampleHandler also computes $N_s$ when a sample is created. Then we run BRS on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table. In addition, since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.

When the SamplerHandler gets called with argument $r$, it needs to find or create a sample with $r$ as the filter rule. At the beginning when it gets called with the empty rule as an argument, there are no samples in memory and it must make a pass through the data to generate a sample. Creating a new sample by making a pass through the table is called \textbf{Create} (further described below). At later stages, when there are potentially multiple samples available, there are multiple mechanisms it could use to return a sample for rule $r$:
\squishlist
\item \textbf{Find:} If the SampleHandler finds an existing sample $s$ in memory, which has $r$ as its filter rule (i.e. $f_s = r$) and at least $minSS$ tuples ($|T_s| \geq minSS$, then it simply returns sample $s$. BRS can then be run on $s$.

\item \textbf{Combine:} If \textbf{Find} doesn't work i.e., if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$. If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution. That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.

Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by BRS. Since all the tuples in the `new' sample are already present in existing samples, it can give BRS a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by BRS. In contrast, if we had created a new sample from hard disk, we would maintain the sample even after BRS terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
%TODO: Not sure how clear/necessary this is.

\item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table. Making a pass can be expensive for big tables, so we only use \textbf{Create} when \textbf{Find} and \textbf{Combine} cannot be used. We can use reservoir sampling~\cite{maibdr1983,Vitter:1985:RSR:3147.3165} to get a uniformly random sample of given size in a single pass through the table.

When the SamplerHandler uses \textbf{Create} for a rule $r$, it needs to access the hard disk to make a pass through the entire table. Since accessing the hard disk and making a pass through the entire table is usually a bottleneck, it can also do things like creating samples for rules other than $r$, and augmenting existing samples, in the same pass. Hence, we assume that in a \textbf{Create} phase, the SampleHandler not only creates one new sample for $r$, but also uses the algorithm from Section~\ref{sec:sampling_algorithms} to determine the new optimal allocation of memory $n_r$ for each displayed rule $r$. Then in a single pass, it creates a sample of size $n_r$ for each displayed $r$. 
\squishend
}

\techreporttext{
\smallskip
\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the BRS, and display a rule-list to the user. If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table. But after expanding $r$, there is a high chance that the user goes further and drills down on one of the sub-rules $r^{\prime}$ of $r$. We may not be able to use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ with the existing samples. So while the user is busy reading the current rule-list obtained from drilling down on $r$, we can start running the algorithm from Section~\ref{sec:sampling_algorithms} in the background, and then making a pass through the table to create a new samples. That way, when the user expands the next rule $r^{\prime}$, there will be a high chance of a sample being pre-fetched for $r^{\prime}$, increasing the chance that we can use \textbf{Find} or \textbf{Combine} on $r^{\prime}$ and reducing our response time.
In addition, while we are making the pass in the background, we can find the exact counts for currently displayed rules (which only have estimated counts shown), and update them when our pass is complete.
}

\section{Experiments}\label{sec:experiments}
We have implemented a fully-functional interactive tool instrumented with the smart drill down operator, having a web interface.
We now describe our experiments on this tool with real datasets.

\stitle{Datasets.} The first dataset, denoted `Marketing', contains demographic information about potential customers~\cite{dataset1}. A total of $N=9409$ questionnaires containing $502$ questions were filled out by shopping mall customers in the San Francisco Bay area. This dataset is the summarized result of this survey. Each tuple in the table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and each column has up to $10$ distinct values.

The columns (in order) are as follows:
annual household income, gender, marital status, age, education, occupation, time lived in the Bay Area, dual incomes?, persons in household, persons in household under $18$, householder status, type of home, ethnic classification, language most spoken in home.

The second dataset, denoted `Census', is a US 1990 Census dataset from the UCI Machine Learning repository~\cite{uciml},
consisting of about $2.5$ million tuples, with each tuple corresponding to a person. It has $68$ columns, including ancestry, age, and citizenship. Numerical columns, such as age, have been bucketized beforehand in the dataset. We use this dataset in Section~\ref{sec:quantitative_study} in order to study the accuracy and performance of sampling on a large dataset.

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page. We use the current implementation of our the smart drill down operator, and insert cropped screenshots of its output in this paper. 
We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}). Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.

\subsection{Qualitative Study}\label{sec:exp_qualitative}
We first perform a qualitative study of smart drill down. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
\subsubsection{Testing the User Interface}
We now present the rule-based summaries displayed as a result of a few different user actions.
To begin with, the user sees an empty rule with the total number of tuples as the count. Suppose the user expands the rule. Then the user will see Figure~\ref{fig:uiexample1}. The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples. The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Area for $> 10$ years. Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill down, whereas smart drill down displays that information to the user with a single click.

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
\vspace{-10pt}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
\vspace{-10pt}
\end{figure}

Now suppose the user decides to further explore the table, by looking at education related information of females in the dataset. Say the user clicks on the $\star$ in the `Education' column of the second rule. This opens up Figure~\ref{fig:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females. Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed Figure~\ref{fig:uiexamplerule}.

\subsubsection{Weighting functions}
Our system can display optimal rule lists for any monotonic weighting function. By default, we assign a rule weight equal to its size. In this section, we consider other weighting functions.

We first try the weighting function given by:
$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct possible values in column $c$. This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values. The rule summary for this weighting is in Figure~\ref{fig:weigtingbitwise} (contrast with Figure~\ref{fig:uiexample1}). The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column. Thus, this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Figure~\ref{fig:uiexample1}.

The other weighting function we try is given by:
$$W(r) = \text{Min}(0, \text{Size}(r) - 1)$$
This gives us Figure~\ref{fig:weightingsizeminusone}. This weighting gives a $0$ weight to rules with a single non-$\star$ value, and thus forces the algorithm to finds good rules having at least $2$ non-$\star$ values. As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Figure~\ref{fig:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.

A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\papertext{ (see technical report for details)}\techreporttext{. Specifically, if we want to perform a regular drill down on a column $C$ that has $n$ distinct values, then we set the number of rules to be displayed ($k$) to $n$. The weighting function $W$ is set such that $W(r) = 1$ if rule $r$ has $C$ instantiated, and $0$ otherwise. This ensures that all displayed rules have $C$ instantiated (with no other column $C'$ instantiated unless there is a functional dependency from $C$ to $C'$), and that each displayed rule has a distinct value of $C$. This effectively gives us a regular drill-down on $C$}. We use this to perform a drill down on the `Age' column using our experimental prototype. The result is shown in Figure~\ref{fig:drilldownexample}. We can contrast it with Figure~\ref{fig:uiexample1}; the latter gives information about multiple columns at once and only displays high count values. Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.

%TODO: Give application examples for different weighting functions
%TODO: Say somewhere, that we haev drop-down list of common weighting functions in the user interface, with column selects and unselects.

\begin{figure}
%\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
\caption{A rule expansion \label{fig:uiexamplerule}}
\vspace{-10pt}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
\caption{A regular drill down on Age\label{fig:drilldownexample}}
\vspace{-10pt}
\end{figure}

% TODO: Add the experiment where you compare to vanilla a priori.
\subsection{Quantitative Study}\label{sec:quantitative_study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). We now study the effects of these parameters on the computation time and accuracy of our algorithm. We use the Marketing and Census datasets. The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples. The accuracy of our algorithm depends on $m_w$ and $minSS$, rather than the underlying database size. The worst case running time for large datasets is close to the time taken for making one pass on the dataset. When we expand a rule using an existing sample in memory, the running time is small and only depends on $minSS$ rather than on the dataset size.

\subsubsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as its weight is $\leq m_w$, but runs faster for smaller values of $m_w$. We now study the effect of varying $m_w$ on the speed of our algorithm running on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

We fix a weighting function $W$, and a value of $m_w$. For that value of the $W$ and $m_w$ parameters, we find the time taken for expanding the empty rule. We repeat this procedure $10$ times and take the average value of the running times across the $10$ iterations. This time is plotted against $m_w$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:mw_speed}. The figure shows that running time seems to be approximately linear in $m_w$.

For the Census dataset, the running time is dominated by time spent in making a pass through the $2.5$ million tuples to create the first sample. The response time for the next user click should be quite small, as the sample created for the first expansion can usually be re-used for the next rule expansion.

The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule. Thus, for size scoring on the Marketing dataset, according to Figure~\ref{fig:uiexample1}, we require $m_w \geq 3$. For the second weighting function, according to Figure~\ref{fig:weigtingbitwise}, the minimum required value of $m_w$ is $10$. At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively. Of course, the minimum value of $m_w$ we can use is not known to us beforehand. But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.

\begin{figure}
\centering
  \includegraphics[scale=0.46]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
\vspace{-10pt}
\end{figure}

\subsubsection{Effects of $minSS$}
We now study the effects of the sampling parameter $minSS$. This parameter tells the SampleHandler the minimum sample size on which we run BRS. Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.

We consider one value of $minSS$ and one weight function $W$ at a time. For those values of $minSS$ and $W$, we drill down on the empty rule and measure the time taken. We also measure the percent error in the estimated counts of the displayed rules. That is, for each displayed rule $r$, if the displayed (estimated) count if $c_1$ and the actual count (computed separately on the entire table) is $c_2$, then the percent error for rule $r$ is $\frac{100 \times |c_1-c_2|}{c_2}$. We consider the average of percent errors over all displayed rules. For each value of $minSS$ and $W$, we drill down on the empty rule and find the computation time and percent error $50$ times, and take the average value for time and error over those $50$ iterations. This average time is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_speed}(a). The average percent error is plotted against $minSS$, for $W(r) = \text{Size}(r)$ and $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ in Figure~\ref{fig:minSS_error_percent}(b).

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.jpg}
\caption{Bits scoring\label{fig:weigtingbitwise}}
\vspace{-15pt}
\end{figure}

\begin{figure}
\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.jpg}
\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
\vspace{-10pt}
\end{figure}

Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.

In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}(c). This number for the Marketing dataset is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. For the Census dataset, it is around $1$ for $minSS \leq 1000$ and falls to about $0.3$ for larger values of $minSS$. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

\begin{figure*}[!ht]
\centering
\includegraphics[scale=0.46]{graphs/minSS_speed.pdf}
\includegraphics[scale=0.46]{graphs/minSS_error_percent.pdf}
\includegraphics[scale=0.46]{graphs/minSS_error_rule.pdf}
\caption{(a) Running time for different values of parameter $minSS$ (b) Error in Count for different values of parameter $minSS$ (c) Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_speed}\label{fig:minSS_error_percent} \label{fig:minSS_error_rule}}
\vspace{-10pt}
\end{figure*}

\subsubsection{Scaling properties of our algorithms}
The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$. That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants. In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk. That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk. The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.

When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a). When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk). When we have a few million tuples, our total runtime is only a few seconds. But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time. To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).

\techreporttext{
\section{Extensions}\label{sec:extensions}
\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
Our algorithm assumes that all attributes are categorical in nature. Attributes that have a large domain tend to have a smaller tuple count per value, and hence don't appear in rule summaries. Thus our algorithm does not summarise information about numerical attributes.

However, we can modify the algorithm to deal with numerical attributes. Suppose we have a numerical attribute $A$. A simple approach is to create buckets for values of $A$. We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, each corresponding to a bucket. We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket. Then we can use our algorithm, treating the bucket number as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).

\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.
}

\section{Related Work}\label{sec:related}

There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}. This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution. The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones. Some other data exploration work~\cite{sellam:meet} focuses on finding attribute values that divide the database in equal sized parts, while we focus on values that occur as frequently as possible.

There is work on constructing `explanation tables', sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}. This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values. A few vision papers~\cite{data-deluge, cetintemel2013query} suggest frameworks for building interactive data exploration systems. Some of these ideas, like maintaining user profiles, could be integrated into smart drill down. Reference~\cite{DBLP:conf/edbt/CandanCQS09} proposes an extension to OLAP drill-down that takes visualization real estate into account, by clustering attribute values. But it focuses on expanding a single column at a time, and relies on a given value hierarchy for clustering.

Some related work~\cite{DBLP:journals/debu/GolabKS11, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies. There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf:kdd:XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables. In both these cases, the emphasis is on completely covering or 
summarizing the table, suffering from the same problems as traditional drill down in that the user may be presented with
too many results. The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case do not scale well to a large number of attributes (in their case, $\geq 4$).

Several existing papers also deal with the problem of frequent itemset mining~\cite{apriori, 1411744, Han:2000:MFP:342009.335372}. Vanilla frequent itemset mining is not directly applicable to our problem because the flexible user-specified objective function emphasizes coverage of the table rather than simply frequent itemsets.  However, we do leverage ideas from the a-priori algorithm~\cite{apriori} as applicable. Several extensions have been proposed to the a-priori algorithm, including those for dealing with numerical attributes~\cite{Srikant:1996:MQA:233269.233311, Miller:1997:ARO:253260.253361}. We can potentially use these ideas to improve handing of numerical attributes in our work. Unlike our paper, there has been no work on dynamically maintaining samples for interaction in the frequent itemset literature, since frequent itemset mining is a one-shot problem.

There has also been plenty of work on pattern mining. Several papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} propose non-interactive schemes that attempt to find a one shot summary of the table. These schemes usually consume a large amount of time processing the whole table, rather than allowing the user to slowly steer into portions of interest. In contrast, our work is interactive, and includes a smart memory manager that can use limited memory effectively while preparing for future requests.

Our Smart Drill-Down operator is tunable because of the flexible weighting function, but the monotonicity of the weighting function and the use of $MCount$, still make it possible for us to get an approximate optimality guarantee for the rules we display. In contrast, much of the existing pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117} is not not tunable enough, providing only a fixed set of interestingess parameters. On the other hand, reference~\cite{Leeuwen:2012:DSS:2347179.2347243} allows a fully general scoring function, necessitating the use of heuristics with no optimality guarantees, and very time consuming algorithms. 
A lot of pattern mining work~\cite{Goethals:2011:MFI:2020408.2020529,Yan:2005:SIP:1081870.1081907,Vreeken:2011:KMI:1969593.1969615} also focuses on itemsets rather than Relational Data, which does not allow the user to express interest in certain `columns' over others.

We use sampling to find approximate estimates of rule counts. Various other database systems~\cite{Acharya:1999:AAQ:304182.304581, Agarwal:2013:BQB:2465351.2465355} use samples to find approximate results to SQL aggregation queries. These systems create samples in advance and only update them when the database changes. In contrast, we keep updating our samples on the fly, as the user interacts with our system. There is work on using weighted sampling~\cite{6691587} to create samples favouring data that is of interest to a user, based on the user's history. In contrast, we create samples at run time in response to the user's commands.

\section{Conclusion}\label{sec:conclusion}
We have presented a new data exploration operator called smart drill down.
Like traditional drill down, it allows an analyst
to quickly discover interesting value patterns (rules) that occur frequently
(or that represent high values of some metric attribute)
across diverse parts of a table.

We presented an algorithm for optimally selecting rules
to display, as well as a scheme for performing such selections
based on data samples. Working with samples makes smart drill down
relatively insensitive to the size of the table.

Our experimental results on our experimental prototype show that smart drill down
is fast enough to be interactive under various
realistic scenarios. We also showed that the accuracy is high
when sampling is used, and when the maximum weight ($m_w$)
approximation is used. Moreover, we have a tunable parameter $minSS$ that the user can tweak to tradeoff performance of smart drill down for the accuracy of the rules.

\bibliographystyle{abbrv}
\bibliography{smartDrillDown}

\vspace{17pt}

\begin{comment}
\begin{wrapfigure}{l}{0.15\textwidth}
%   \begin{center}
\includegraphics[height=3.5cm, width=2.7cm]{./figures/hector.pdf}
%   \end{center}
\end{wrapfigure}

Hector Garcia-Molina is the Leonard Bosack and Sandra Lerner Professor in the Departments of Computer Science and Electrical Engineering at Stanford University, Stanford, California. He was the chairman of the Computer Science Department from January 2001 to December 2004. From 1997 to 2001 he was a member the President's Information Technology Advisory Committee (PITAC). From August 1994 to December 1997 he was the Director of the Computer Systems Laboratory at Stanford. From 1979 to 1991 he was on the faculty of the Computer Science Department at Princeton University, Princeton, New Jersey. His research interests include distributed computing systems, digital libraries and database systems. He received a BS in electrical engineering from the Instituto Tecnologico de Monterrey, Mexico, in 1974. From Stanford University, Stanford, California, he received in 1975 a MS in electrical engineering and a PhD in computer science in 1979. He holds an honorary PhD from ETH Zurich (2007). Garcia-Molina is a Fellow of the Association for Computing Machinery and of the American Academy of Arts and Sciences; is a member of the National Academy of Engineering; received the 1999 ACM SIGMOD Innovations Award; is a Venture Advisor for Onset Ventures, is a member of the Board of Directors of Oracle, and is a member of the State Farm Technical Advisory Council.
\medskip

\begin{wrapfigure}{l}{0.15\textwidth}
%   \begin{center}
\includegraphics[height=3.5cm, width=2.7cm]{./figures/Manas.pdf}
%   \end{center}
\end{wrapfigure}

Manas Joglekar is a fifth year PhD student in the Computer Science Department at Stanford University. Manas' advisor is Prof. Hector Garcia-Molina, and his research interests include Crowdsourcing, Data Quality Evaluation and Database Theory. Prior to that, he received a B. Tech in Computer Science and Engineering from the Indian Institute of Technology Bombay, in 2011.
\medskip

\begin{wrapfigure}{l}{0.15\textwidth}
%   \begin{center}
\includegraphics[height=3.5cm, width=2.7cm]{./figures/aditya.pdf}
%   \end{center}
\end{wrapfigure}

Aditya Parameswaran is an Assistant Professor in Computer Science at the University of Illinois (UIUC). He spent the 2013-14 year visiting MIT CSAIL and Microsoft Research New England, after completing his Ph.D. from Stanford University, advised by Prof. Hector Garcia-Molina. He is broadly interested in data analytics, with research results in human computation, visual analytics, information extraction and integration, and recommender systems. 
Aditya is a recipient of the Arthur Samuel award for the best dissertation in CS at Stanford (2014), the SIGMOD Jim Gray dissertation award (2014), the SIGKDD dissertation award runner up (2014), a Google Faculty Research Award (2015), the Key Scientific Challenges Award from Yahoo! Research (2010), three best-of-conference citations (VLDB 2010, KDD 2012 and ICDE 2014), the Terry Groswith graduate fellowship at Stanford (2007), and the Gold Medal in Computer Science at IIT Bombay (2007). His research group is supported with funding from by the NIH, the NSF, and Google.
\medskip
\end{comment}

\end{document}

### 2016-07-01 TableSummarization/journalVersion/smartDrillDown.tex
removing widows

@@
 where $X$ is a Store in the answer table, and $C$
-is the number of tuples for $X$ (or the aggregate sales for $X$).
+is the number of tuples for $X$ (or aggregate sales for $X$).
 
@@
 
-Say that after seeing the results of Table~\ref{table:introexample},
+Suppose after seeing the results of Table~\ref{table:introexample},
 the analyst wishes to dig deeper into the Walmart tuples
 represented by the last rule.
-For instance, the analyst may want to know
+The analyst may want to know
 which states Walmart has more sales in, or which products they sell
@@
 expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
-(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
 There are three factors that make a rule set interesting.
-One is if it contains rules with high Count (or total sales) fields,
+One is if it contains rules with high Count,
 since the larger the count, the more tuples are summarized.
@@
 For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
-seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
+is more interesting than (Walmart, cookies, $*$, $200$, $2$)
 since the former tells us the high sales are concentrated in a single region.
-A third desirability factor is diversity:
-For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
-in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
+The third factor is diversity: For example, if our set already has the rule (Walmart, $\star$, $\star$, $1000$, $1$), 
+we would rather have add rule (Target, bicycles, $\star$, $200$, $2$)
 than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
@@
 
-Incidentally, note that in the example we only described one type of smart drill down,
+Note that in the example we only described one type of smart drill down,
 where the analyst selects a {\em rule} to drill down on
@@
 These papers mostly focus on
-giving the most ``surprising'' information to the user, i.e., information
+giving ``surprising'' information to the user, i.e., information
 that would minimize the Kullback-Liebler(KL) divergence between the
@@
 Our work is also related to pattern mining.
-Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, but do not propose interactive mechanisms.
+Several pattern mining papers~\cite{Vreeken:2011:KMI:1969593.1969615,Bringmann:2007:ICDM,Yan:2005:SIP:1081870.1081907} focus on providing one shot summaries of data, and do not propose interactive mechanisms.
 Moreover, to the best of our knowledge, other pattern mining work is either not flexible enough~\cite{Goethals:2011:MFI:2020408.2020529,Tatti:2014:FRI:2676651.2656261,DeBie:2010:FMI:1816112.1816117}, restricting the amount of tuning the user can perform, or so general~\cite{Leeuwen:2012:DSS:2347179.2347243} as to preclude efficient optimization.
@@
 
-\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} interaction operator,
+\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} operator,
 an extension of traditional drill down, aimed at allowing analysts to zoom into the
@@
 
-\item We describe related work in Section~\ref{sec:related}, and conclude in Section~\ref{sec:conclusion}.
+\item Section~\ref{sec:extensions} covers extensions of our work. We describe related work in Section~\ref{sec:related}, and conclude in Section~\ref{sec:conclusion}.
 \squishend
@@
 In addition, a rule has other attributes, such as count and weight 
-(which we define later) associated with it. 
+(defined later) associated with it. 
 The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column.
@@
 Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
-As we will see, the weighting function does not depend on the specific
+The weighting function does not depend on the specific
 tuples in $\calD$, but could 
@@
 \noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user.
-We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead.
+We can also use a {\em Sum} of values over a given `measure column' instead.
 We discuss how to modify our algorithms to use $Sum$ instead of $Count$
-\papertext{in our technical report~\cite{tr}}.
-\techreporttext{in the `Extensions' section.}
+\papertext{in our technical report~\cite{tr}.}
+\techreporttext{in Section~\ref{sec:extensions}.}
 
@@
 \stitle{Size Weighting Function:} 
+\begin{comment}
 \techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
@@
 If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$.
-Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).} \papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :
+Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).}
+\end{comment}
+%\papertext{
+$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
 Here we set weight equal to the number of non-starred values in the rule $r$ i.e.
 the {\em size} of the rule.
-For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.}
+For example, in Table~\ref{table:introexample}, the rule (Target, bicycles, $\star$) has weight $2$.
+%}
 
@@
 \stitle{Bits Weighting Function:}
+\begin{comment}
 \techreporttext{
@@
 If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.}
-\papertext{$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
-This function weighs each column based on its inherent complexity, instead of equally like the Size function.}
+\end{comment}
+%\papertext{
+$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
+This function weighs each column based on its inherent complexity, instead of equally like the Size function.
+%}
 
@@
 Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$.
-By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
-Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.} The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
+By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
+%\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
+Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.}
+The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).
 
@@
 This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
-This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.} Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e.
-collapses it.
+%\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
+This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.}
+Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. 
 \techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}.
-This operation is equivalent to a traditional
-roll up, but for smart drill downs instead of traditional drill downs.}
+This operation is equivalent to a traditional roll up, but for smart drill downs instead of traditional drill downs.}
+
+\begin{comment}
 
@@
 }
+\end{comment}
 
@@
 \end{lemma}
-\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}}\techreporttext{
+\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}.}\techreporttext{
 \begin{proof}
@@
 where we set $W(TOP(t)) = 0$ when $t$ is not covered by any rule in $R$.
-Now say we have rule lists $R$, $R'$ which have the same rules, but where $R'$ has rules in decreasing order by weight.
+Now say two rule lists $R$, $R'$ have the same rules, but $R'$ has rules in decreasing order by weight.
 For any tuple $t$ covered by $R$, let $r'$ be the highest weight rule in $R$ that covers $t$.
 Let $r$ be the first rule in $R$ that covers $t$.
-Then we have $TOP(t,R) = r$, $TOP(t, R') = r'$ and $W(r') \geq W(r)$.
-This gives us $W(TOP(t,R')) \geq W(TOP(t,R))$.
+Then $TOP(t,R) = r$, $TOP(t, R') = r'$ and $W(r') \geq W(r)$, so $W(TOP(t,R')) \geq W(TOP(t,R))$.
 Adding these inequalities for all $t$ covered by $R$ gives us $\text{Score}(R') \geq \text{Score}(R)$ as required.
 \end{proof}
-}. 
+}
 Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:
@@
 \subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
-We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
-thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
+We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} (MCP) to a special case of Problem~\ref{prob:optimal-rule-set};
+thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. MCP is given below: 
 \begin{problem}\label{prob:maximum-coverage}
@@
 \end{problem}
-Thus, the goal of the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
-We can reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
+Thus, the goal of MCP is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible.
+We can reduce an instance of MCP (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
 \begin{lemma}
@@
 \subsection{Algorithm Overview}\label{sec:alg-overview}
-Given that the problem is {\sc NP-Hard}, we now present our algorithms 
-for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
+Given that Problem~\ref{prob:optimal-rule-set} is {\sc NP-Hard}, we now present our algorithms 
+for approximating the solution to it. 
 The problem consists of finding a set of rules, given size $k$, that maximizes 
@@
 The first simplification is, we modify our objective function to use hinge-loss instead of a step function.
-That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}(1, \frac{ess(r^{\prime})}{minSS} $$
+That is, our new objective function to maximise is $$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{min}\left(1, \frac{ess(r^{\prime})}{minSS}\right)$$
 Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$,  though we still prefer bigger sample sizes upto $minSS$.
@@
 choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
-$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
+$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}\left(-1, -\frac{ess(r^{\prime})}{minSS}\right) $$
 subject to :

### 2016-07-02 TableSummarization/journalVersion/smartDrillDown.tex
widow hunting

@@
 
-The expensive step in the above procedure is the step where the Score is computed for every
-single rule. Given the number of rules can be as large as the table itself, this can be especially time-consuming.
+The expensive step in the above procedure is where the Score is computed for every
+single rule. Given the number of rules can be as large as the table itself, this is very time-consuming.
 
@@
 We call our algorithm {\em BRS} (for \textbf{B}est \textbf{R}ule \textbf{S}et).
-BRS takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.
+BRS takes four parameters as input: the table $T$, the number of required rules $k$, a parameter $m_w$ (described in the next paragraph), and the weight function $W$.
 
 The parameter $m_w$ stands for \textit{Max Weight}.
-The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$.
+This parameter tells the algorithm to assume that all rules that appear in the optimal solution are going to have weight $\leq m_w$.
 Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$.
@@
 This is because as the size (and weight) of a rule increases, its Count falls rapidly.
-The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$).
+The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as Size).
 Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$.
@@
 The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
-The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration.
-We maintain three sets of rules.
-$C$ is the set of all rules whose marginal values have been counted in all previous iterations.
-$C_n$ is the set of rules whose marginal values are going to be counted in the current pass.
-And $C_o$ is the set of rules whose marginal values were counted in the previous iteration.
+The function makes several passes over the table \techreporttext{(Step $3$)}, counting marginal values for size $j$ rules in the $j^{th}$ pass.
+We maintain three sets of rules: $C$, the set of rules whose marginal values have been counted in all previous passes, $C_n$, the set of rules whose marginal values will be counted in the current pass, and $C_o$, the set of rules whose marginal values were counted in the previous pass.
 For the first pass, we set $C_n$ to be all rules of size $1$.
@@
 Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
-On the other hand, small values of $k$ may display too little information about the table.
+Very small values of $k$ may display too little information about the table.
 Fortunately, the BRS algorithm is incremental in nature.
@@
 But if we are willing to accept approximate counts rather than exact counts, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-Thankfully since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
+Since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us.
+If we had obtained the sample by sampling each tuple with probability $p$, then we must multiply the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
 
@@
 We now consider the following problem: say we currently have no
-samples in memory (we describe the scenario where there are already 
-some samples in Section~\ref{sec:sample-handler}), and say the user is currently viewing some rules; how do we
+samples in memory (we describe the scenario where we already have samples in Section~\ref{sec:sample-handler}), 
+and say the user is currently viewing some rules; how do we
 materialize the ``best possible'' samples in memory that fit within the capacity $M$,
 such that we can respond to as many user interactions as possible using the stored samples,
-without having to resort to retrieving the entire table.
+without having to retrieve the entire table.
 That is, we want to maximize the probability that the next user interaction
@@
 Let $L$ be the set of leaves. 
-Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
+Each leaf is something the user can potentially expand in the next step, so we would like to have pre-fetched samples for rules corresponding to leaf nodes.
 
@@
 \stitle{Selectivity.} 
-Let the `selectivity' of a rule be the fraction of tuples in $T$ that are covered by the rule. 
+Let the `selectivity' of a rule $r$ be the fraction of tuples in $T$ that are covered by $r$. 
 For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
@@
 Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$.
-We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$.
+We are also given a weight limit $W$.
+Our objective is to choose a set of objects with maximum value and total weight $< W$.
 We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.
 
-We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem.
+We scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem.
 For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$.
@@
 The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$.
-In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
+The $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
 Therefore, each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
@@
 Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
-As a result, in any optimal solution, $ess(r_{i,1})$ must be equal to $minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+So in any optimal solution, $ess(r_{i,1})$ must equal $minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
-The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
-Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$).
+The latter option consumes $w_i \times minSS$ extra memory and gives $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ extra probability value.
+So having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (it consumes additional memory $\propto w_i$ and gives additional value $\propto v_i$).
 Moreover, the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
-Hence, solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem.
+Hence, solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ solves the instance of the knapsack problem.
 \end{proof}
@@
 
-\stitle{Approximate DP Solution.} Even though the problem as stated is {\sc NP-Hard}, with an additional simplifying assumption, we can make the problem approximately solvable using a Dynamic Programming algorithm.
+\stitle{Approximate DP Solution.} The problem as stated is {\sc NP-Hard}, but with an additional simplifying assumption we can make the problem approximately solvable using  Dynamic Programming.
 The assumption is: for each $r \in L$, we assume that its $ess$ can get sample tuples only from samples obtained for itself and its immediate parent.
@@
 $$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
-This can be solved using dynamic programming, in $O(D\mathcal{S}3^d)$ time. 
+Dynamic programming solves this in $O(D\mathcal{S}3^d)$ time. 
 }
@@
 Here we assume that it is acceptable to run our algorithm on samples smaller than $minSS$, though we still prefer bigger sample sizes upto $minSS$.
-The other simplification we make is assuming that the sample sizes $n_r$ are real numbers instead of integers.
-After obtaining our optimal sample sizes, we can round them up to get integer sample sizes.
-This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree.
-$|U|$ is usually negligible compared to the memory capacity $M$, or $minSS$.
+The other simplification we make is assuming that sample sizes are real numbers instead of integers.
+After determining optimal sample sizes, we can round them up to get integer sample sizes.
+This will increase the memory usage by at most $|U|$, the number of nodes in displayed tree, which is negligible compared to the memory capacity $M$, or $minSS$.
 
@@
 
-Thus, a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$.
-This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule).
+Thus, a reasonable value of minSS can be found by bounding $\frac{1-x}{x}$.
+This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function.
 Let $c$ be the column with the fewest distinct values.
@@
 The highest scoring rule can have weight at most $|C|$ (the total number of columns).
-Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, the Count of the highest scoring rule must be at least $\frac{|T|}{|C||c|}$.
+Since the score of the highest scoring rule is at least $\frac{|T|}{|c|}$, its Count must be at least $\frac{|T|}{|C||c|}$.
 Thus if minSS is significantly larger than $|C||c|$, then the Count of the first few highest scoring rules should be well-approximated in a sample of size more than minSS.
@@
 a certain memory capacity $M$, and a minimum sample size $minSS$,
-creates, maintains, retrieves, and removes samples, all in
+creates, maintains, retrieves, and removes samples, in
 response to user interactions on the table.
-It uses the algorithms from Section~\ref{sec:sampling_algorithms} 
+It uses algorithms from Section~\ref{sec:sampling_algorithms} 
 to decide which samples to create, as we will see below.
@@
 \item \textbf{Combine:} If \textbf{Find} doesn't work i.e., if the SampleHandler cannot find an existing sample with filter $r$ and $\geq minSS$ tuples, then it looks at all existing samples $s^{\prime}$ such that $f_{s^{\prime}}$ is a sub-rule of $r$.
-If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for rule $r$.
-We can show that tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
+If the set of all tuples that are covered by $r$, from all such $T_{s^{\prime}}$'s combined, exceeds $minSS$ in size, then we can simply treat that set as our sample for $r$.
+Tuples that are covered by $r$, from the combination of $T_{s^{\prime}}$s, follow a uniform distribution.
 That is, each table tuple $t$ that is covered by $r$ is equally likely to appear in a $T_{s^{\prime}}$.
@@
 \techreporttext{
-\smallskip
-\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (on the rule itself or on a $\star$ in the rule), we need to get a sample, run the BRS, and display a rule-list to the user.
+\noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (or on a $\star$ in $r$), we need to get a sample, run the BRS, and display a rule-list to the user.
 If we use \textbf{Find} or \textbf{Combine}, then we can display the rule-list much faster because we don't have to read the entire table.
@@
 
-We first try the weighting function given by:
-$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct possible values in column $c$.
-This function gives higher weight to rules that have non-$\star$ values in columns that have many possible values.
+We first try the `Bits' weighting function, given by:
+$$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$$ where $|c|$ refers to the number of distinct values in column $c$.
+This function gives higher weight to rules that have non-$\star$ values in columns that have many distinct values.
 The rule summary for this weighting is in Figure~\ref{fig:weigtingbitwise} (contrast with Figure~\ref{fig:uiexample1}).
-The weighting scheme gives low weight for non-$\star$ values in binary columns, like the gender column.
+Bits weighting gives low weight for non-$\star$ values in binary columns, like the gender column.
 Thus, this summary instead gives us information about the Marital Status/Time in Bay Area/Occupation columns instead of the Gender column like in Figure~\ref{fig:uiexample1}.
@@
 \subsubsection{Effects of $minSS$}
-We now study the effects of the sampling parameter $minSS$.
-This parameter tells the SampleHandler the minimum sample size on which we run BRS.
-Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.
+We now study the effects of sampling parameter $minSS$.
+This parameter determines the minimum sample size on which we run BRS.
+Higher values of $minSS$ cause our system to use bigger samples, increasing the accuracy of count estimates for displayed rules, but also correspondingly increasing computation time.
 
@@
 The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$.
-That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants.
+That is, the runtime can be written as $a \times |T| + b \times minSS$ for some constants $a$ and $b$.
 In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk.
-That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk.
-The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.
-
-When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
-When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk).
+That is, $a \times |T|$ is the time taken to make a single scan over the table on disk.
+$b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.
+
+When $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a).
+When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by $a \times |T|$ (this is when we need to create a fresh sample from hard disk).
 When we have a few million tuples, our total runtime is only a few seconds.
-But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time.
-To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
+But if the dataset contained billions of tuples, the process of reading the table to create a sample could itself take a very long time.
+To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also summarizes the billion tuple table).
 
@@
 \subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
-Our algorithm assumes that all attributes are categorical in nature.
-Attributes that have a large domain tend to have a smaller tuple count per value, and hence don't appear in rule summaries.
+Our algorithm assumes that all attributes are categorical.
+Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
 Thus our algorithm does not summarise information about numerical attributes.
@@
 Suppose we have a numerical attribute $A$.
-A simple approach is to create buckets for values of $A$.
-We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, each corresponding to a bucket.
-We can create buckets having an equal range size, or decide their range such that there is an approximately equal number of tuples in each bucket.
-Then we can use our algorithm, treating the bucket number as a categorical attribute.
+We can create buckets for values of $A$.
+We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, with one bucket per interval.
+We can create buckets having equal range sizes, or decide their range such that there is an approximately equal number of tuples in each bucket.
+Then we run BRS, treating the bucket id as a categorical attribute.
 This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).
@@
 There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
-This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution.
+This and other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution.
 The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
@@
 Some related work~\cite{DBLP:journals/debu/GolabKS11, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies.
-There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf:kdd:XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
+There is some work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, DBLP:conf:kdd:XiangJFD08, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables.
 In both these cases, the emphasis is on completely covering or
 summarizing the table, suffering from the same problems as traditional drill down in that the user may be presented with
-too many results.
-The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case do not scale well to a large number of attributes (in their case, $\geq 4$).
+too many results. The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case do not scale well to a $\geq 4$ attributes.
 

### 2016-07-24 TableSummarization/journalVersion/smartDrillDown.tex
changes

@@
 
-\title{Interactive Data Exploration with Smart Drill-Down}
+\textbf{Note to Reviewers}: We have made the following additions to the extended version of our paper.
+
+\begin{enumerate}
+\item We have added pseudo-code for the `Find Best Marginal Rule' subroutine of the BRS algorithm.
+\item We have added methods for setting good values of our tunable parameters $W$, $k$, and $m_w$ in Section~\ref{sec:setting-paramters}.
+\item Section~\ref{sec:sampling} has been significantly expanded, with details on how to utilize samples stored in memory and when to form new samples.
+We have also added a more detailed description and proof for our approximate dynamic-programming based algorithm for determining the best samples to form.
+\item We have added an alternative convex-optimization based solution for determining the best samples.
+\item Additionally, Section~\ref{sec:sampling} contains additional optimizations such as pre-fetching, and guidelines for setting the sampling parameter $minSS$. 
+\item Section~\ref{sec:extensions} contains two simple extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
+\end{enumerate}
+
+\newpage
+
+\title{Interactive Data Exploration with Smart Drill-Down (Extended Version)}
 
@@
         Hector~Garcia-Molina,~\IEEEmembership{Member,~IEEE,}
-        Aditya~Parameswaran
+        Aditya~Parameswaran,~\IEEEmembership{Member,~IEEE,}
 %\thanks{H. Garcia-Molina, M. Joglekar and V. Verroios are with Stanford University}%
@@
 \item In Section~\ref{sec:algorithms}, we present our algorithms for
-finding optimal sets of rules, as well as our dynamic sampling schemes
-for dealing with large tables.
+finding optimal sets of rules.
+
+\item In Section~\ref{sec:sampling}, we present our dynamic sampling schemes
+for dealing with large tables
 
@@
 
-\begin{algorithm}
-\scriptsize
-\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$S$ (Solution set of rules)}
-$S = \phi$
-
-\For {$i$ from $1$ to $k$}{
-$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$
-
-$S = S \cup \left\lbrace R_m \right\rbrace$
-}
-\Return $S$
-\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}\label{algo:best-rule-set}}
-\end{algorithm}
-
 As a first step towards solving Problem~\ref{prob:optimal-rule-list}, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
@@
 
-\techreporttext{
-\begin{algorithm}
-\scriptsize
-\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
-\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
-$H = 0$ \tcc*{Threshold for deciding if to count for a rule.}
-$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
-\For{$j$ from $1$ to number of columns in $T$}{
-\If {$j = 1$}{
-$C_n = \text{ all rules of size } 1$
-}
-\Else{ 
-$C_n = \text{all size-}i$ super-rules of rules from $C_o$
-}
-\ForEach {$R \in C_n$}{
-$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
-\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
-$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
-}
-\If {$(M < H)$}{ 
-$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to be in the solution}
-}
-}
-\If {$C_n = \phi$}{
-break;
-}
-\ForEach {$R \in C_n$}{
-Count$(R) = 0$ \tcc*{Initialize}
-MarginalValue$(R) = 0$ \tcc*{Initialize}
-}
-\ForEach {$t \in T$}{
-Let $R_S$ be the highest weight rule in $S$ that covers $t$
-
-\ForEach {$R \in C_n$ that covers $t$}{
-Count($R$) $++$
-
-MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
-}
-}
-$C = C \cup C_n$
-
-$C_o = C_n$
-
-$C_n = \phi$
-
-$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
-}
-\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
-\caption{Find best marginal rule\label{algo:best-marginal-rule}}
-\end{algorithm}
-}
-
 In some cases, the dataset may still be too large for us to return a good rule set in
@@
 this algorithm next.
+
+\begin{algorithm}
+\scriptsize
+\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$S$ (Solution set of rules)}
+$S = \phi$
+
+\For {$i$ from $1$ to $k$}{
+$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$
+
+$S = S \cup \left\lbrace R_m \right\rbrace$
+}
+\Return $S$
+\caption{BRS\label{algo:best-rule-set}}
+\end{algorithm}
 
@@
 
-\stitle{Finding the Best Marginal Rule:} In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
+\subsection{Finding the Best Marginal Rule}\label{sec:finding-best-marginal}
+In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one.
 A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data.
@@
 Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset.
+
+\techreporttext{
+\begin{algorithm}
+\scriptsize
+\KwIn{$S$ (Current solution set), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
+\KwOut{$R_m$ (Rule which adds the highest marginal value among rules with weight $\leq m_w$)}
+$H = 0$ \tcc*{Threshold for deciding if to count for a rule.}
+$C = C_o = C_n = \phi$ \tcc*{Set of all, old and new candidate rules respectively. }
+\For{$j$ from $1$ to number of columns in $T$}{
+\If {$j = 1$}{
+$C_n = \text{ all rules of size } 1$
+}
+\Else{ 
+$C_n = \text{all size-}i$ super-rules of rules from $C_o$
+}
+\ForEach {$R \in C_n$}{
+$M =\infty$ \tcc*{Upper bound on marginal value count of $R$}
+\ForEach {$R$-sub-rule $R^{\prime} \in C$}{  
+$M = \text{min}(M, \text{MarginalVal}(R^{\prime}) + \text{Count}(R^{\prime})(m_w - W(R^{\prime}))$
+}
+\If {$(M < H)$}{ 
+$C_n = C_n \setminus \left\lbrace R \right\rbrace$ \tcc{Delete $R$ if its max count is too small for $R$ to be in the solution}
+}
+}
+\If {$C_n = \phi$}{
+break;
+}
+\ForEach {$R \in C_n$}{
+Count$(R) = 0$ \tcc*{Initialize}
+MarginalValue$(R) = 0$ \tcc*{Initialize}
+}
+\ForEach {$t \in T$}{
+Let $R_S$ be the highest weight rule in $S$ that covers $t$
+
+\ForEach {$R \in C_n$ that covers $t$}{
+Count($R$) $++$
+
+MarginalValue($R$) $+=$ $W(R) - \text{min}(W(R), W(R_S))$
+}
+}
+$C = C \cup C_n$
+
+$C_o = C_n$
+
+$C_n = \phi$
+
+$H = \textrm{max}_{R \in C}(\text{MarginalValue}(R))$
+}
+\Return $\textrm{argmax}_{r \in C} \text{MarginalValue}(r)$
+\caption{Find best marginal rule\label{algo:best-marginal-rule}}
+\end{algorithm}
+}
 
@@
 \techreporttext{
-\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
+\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
+Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
 Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
@@
 
-\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
+\section{Dynamic Sampling for Large Tables}\label{sec:sampling}
 
@@
 
-\subsubsection{Deciding what to sample}\label{sec:sampling_algorithms}
+\subsection{Deciding what to sample}\label{sec:sampling_algorithms}
 We are given a memory capacity $M$, and a minimum sample size $minSS$, both specified by the user.
@@
 \techreporttext{
-Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules.
+The BRS algorithm need to to make multiple passes over the entire table in order to find counts of rules.
 These passes can be expensive if the table is large and the table does not fit in main memory.
@@
 
-We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms}.
+We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms} through Section~\ref{sec:alternative-convex}.
 Then, in Section~\ref{sec:sample-handler}, we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
@@
 
-\subsubsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
+\subsection{Algorithms for deciding what to sample}\label{sec:sampling_algorithms}
 
@@
 \stitle{Sampling Strategy.} The sampling strategy we adopt involves storing, 
-for every displayed rule $r'$, a sample of $r'$ of size $n_{r'}$, i.e., 
+for every displayed rule $r'$ in $U$, a sample of $r'$ of size $n_{r'}$, i.e., 
 containing $n_{r'}$ randomly chosen tuples $\in r'$.
@@
 will be zero for those rules.
+Also note that we may store samples for $r' \in U \setminus L$, because even though the user cannot expand an internal node of $U$, a sample of an internal $r'$ can help us create samples for each descendant of $r'$ in $L$.
+
 We need to pick the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory. 
@@
 \techreporttext{
-\subsubsection{Alternative Convex-Optimization based solution}
+\subsection{Alternative Convex-Optimization based solution}\label{sec:alternative-convex}
 We noted earlier that Problem~\ref{prob:sample-sizes} is NP-Hard, but can be approximately solved with an additional simplifying assumption regarding the $S(r_1, r_2)$ values.
@@
 
-\subsubsection{Additional optimizations}
+\stitle{Additional optimizations:}
 There are some additional minor optimizations we can make to reduce the memory cost per sample, allowing us to store more and bigger samples. 
@@
 
-\subsubsection{Setting $minSS$}
+\stitle{Setting $minSS$:}
 Suppose a rule $r$ covers $x$ fraction of the tuples of $T$ i.e.
@@
 
-\subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
+\subsection{Design of the SampleHandler}\label{sec:sample-handler}
 \papertext{
@@
 \centering
-  \includegraphics[scale=0.46]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
+  \includegraphics[scale=0.5]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
   \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
@@
 \subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
-Our algorithm assumes that all attributes are categorical.
+Our framework assumes that all attributes are categorical.
 Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
-Thus our algorithm does not summarise information about numerical attributes.
+Thus our algorithm does not summarize information about numerical attributes.
 

### 2016-07-25 TableSummarization/journalVersion/smartDrillDown.tex
added couple of keywords, and some changes

@@
 \item We have added pseudo-code for the `Find Best Marginal Rule' subroutine of the BRS algorithm.
-\item We have added methods for setting good values of our tunable parameters $W$, $k$, and $m_w$ in Section~\ref{sec:setting-paramters}.
+\item The new Section~\ref{sec:setting-paramters} contains methods for setting good values of our tunable parameters $W$, $k$, and $m_w$ .
 \item Section~\ref{sec:sampling} has been significantly expanded, with details on how to utilize samples stored in memory and when to form new samples.
 We have also added a more detailed description and proof for our approximate dynamic-programming based algorithm for determining the best samples to form.
-\item We have added an alternative convex-optimization based solution for determining the best samples.
+\item We have added an alternative convex-optimization based solution for determining the best samples in Section~\ref{sec:sampling}.
 \item Additionally, Section~\ref{sec:sampling} contains additional optimizations such as pre-fetching, and guidelines for setting the sampling parameter $minSS$. 
-\item Section~\ref{sec:extensions} contains two simple extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
+\item We added Section~\ref{sec:extensions}, which contains two simple extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
 \end{enumerate}
@@
 \begin{IEEEkeywords}
-ADD-OTHER-IEEE-KEYWORDS
+Data Exploration, Data Summarization
 \end{IEEEkeywords}
@@
 
+\begin{comment}
 \stitle{Additional optimizations:}
@@
 Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
+\end{comment}
 

### 2016-07-30 TableSummarization/journalVersion/smartDrillDown.tex
r
change
back to size

@@
 
-\textbf{Note to Reviewers}: We have made the following additions to the extended version of our paper.
+\textbf{Note to Reviewers}: This paper is an extended version of the paper "Interactive Data Exploration with Smart Drill Down" that appeared in the 2016 IEEE International Conference on Data Engineering.
+This extended version contains about $30\%$ new material compared to the original version.
+In particular, the additions include:
 
@@
 \item Additionally, Section~\ref{sec:sampling} contains additional optimizations such as pre-fetching, and guidelines for setting the sampling parameter $minSS$. 
-\item We added Section~\ref{sec:extensions}, which contains two simple extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
+\item We added Section~\ref{sec:extensions}, which contains two extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
 \end{enumerate}
@@
 \IEEEcompsocitemizethanks{
-\IEEEcompsocthanksitem M. Joglekar is with Google.\protect\\
+\IEEEcompsocthanksitem M. Joglekar is with Stanford University.\protect\\
 % note need leading \protect in front of \\ to get a newline within \thanks as
 % \\ is fragile and will error, could use \hfil\break instead.
-E-mail: manasrj@google.com
+E-mail: joglekarmanas@gmail.com
 \IEEEcompsocthanksitem H. Garcia-Molina is with Stanford University.\protect\\
@@
 
+\begin{comment}
+\techreporttext{
+\begin{table}
+\centering
+\scriptsize
+\begin{tabular}{ | l | c | }
+ \hline Rule-MCount list & Score \\ \hline
+  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
+  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
+  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
+\end{tabular}
+\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
+\end{table}
+}
+\end{comment}
+
 \stitle{Size Weighting Function:} 
@@
 
-\techreporttext{
-\begin{table}
-\centering
-\scriptsize
-\begin{tabular}{ | l | c | }
- \hline Rule-MCount list & Score \\ \hline
-  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
-  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
-  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
-\end{tabular}
-\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
-\end{table}
-}
-
 \subsection{Smart drill down Operations}
@@
 We now formally define our problem:
+
 \begin{problem}\label{prob:sample-sizes}
@@
 \techreporttext{
-The BRS algorithm need to to make multiple passes over the entire table in order to find counts of rules.
-These passes can be expensive if the table is large and the table does not fit in main memory.
-If we want exact counts for rules, we have no choice but to read the entire table.
-
-But if we are willing to accept approximate counts rather than exact counts, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count.
-Since our goal is to find a representative coverage of the table, if we miss out on a few rare tuples, it does not hurt us.
-If we had obtained the sample by sampling each tuple with probability $p$, then we must multiply the sample count of each rule by $\frac{1}{p}$ to get an estimate of its count over the full table.
-
-Thus, we use sampling to trade-off a small amount of accuracy for a faster response time.
-We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory.
+BRS makes multiple passes over the table in order to determine the best set of rules to display.
+This can be slow when the table is too large to fit in main memory.
+We can reduce the response time of smart drill down by running BRS on a sample of the table instead, trading off accuracy of our rules for performance.
+If we had obtained a sample $s$ by selecting each table tuple with probability $p$, and run BRS on $s$, then we multiply the count of each rule found by BRS, by $\frac{1}{p}$ to estimate its count over the full table.
 
@@
 
-We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user).
-$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.
-The parameter $minSS$ is the minimum number of sample tuples needed
+We are given a user-specified memory capacity $M$, and a minimum sample size $minSS$.
+$M$ can be default to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.
+$minSS$ is the minimum number of sample tuples needed
 such that we can use the sample in memory 
@@
 on hard disk. 
-This parameter determines how accurate our count estimates will be, and also
-how quickly we can return results to the user. The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.
-
-We now consider the following problem: say we currently have no
+$minSS$ determines our count estimates' accuracy as well as runtime. The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.
+
+Now consider the following problem: say we currently have no
 samples in memory (we describe the scenario where we already have samples in Section~\ref{sec:sample-handler}), 
-and say the user is currently viewing some rules; how do we
-materialize the ``best possible'' samples in memory that fit within the capacity $M$,
-such that we can respond to as many user interactions as possible using the stored samples,
+and the user is currently viewing some rules; how do we
+materialize the ``best possible'' samples in memory that fit within the capacity $M$, such that we can respond to as many user interactions as possible using the stored samples,
 without having to retrieve the entire table.
 That is, we want to maximize the probability that the next user interaction
-can be answered using the existing samples, without reading from the hard disk. 
-We leverage techniques from approximation algorithms and optimization theory, formalized
-below.
+can be answered using existing samples, without reading the hard disk.
 
@@
 At any stage, we have a tree $U$ of rules displayed to the user, 
-with each node of the tree corresponding to a displayed rule. 
-In the rest of this section, we will refer to nodes in $U$ and rules interchangeably.
-The tree is formed as follows: The root of the tree corresponds to the trivial rule. 
-And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. 
-Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$, and so on.
-
-Even though the rules displayed to the user can be considered a tree, 
-multiple nodes of the tree may correspond to the same rule. 
+with each node of $U$ corresponding to a displayed rule. 
+We will refer to nodes in $U$ and rules interchangeably.
+The tree is formed as follows: The root of $U$ corresponds to the trivial rule. 
+And when the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed, we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$, and so on.
+
+Even though the rules displayed to the user form a tree $U$, 
+multiple nodes of $U$ may correspond to the same rule. 
 For example, $(a, b)$ may be a child of both $(\star, b)$ and $(a, \star)$.
-Thus, even though the structure displayed to the user is a tree, 
-the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering. 
-However, for the purposes of this section, we focus on the tree representation.
-
-Internal nodes of the tree $U$ are ones that have been expanded (drilled down on), 
+Thus the user is shown a tree structure, but
+the set of displayed rules forms a partially ordered set (poset) using the sub-rule ordering. In this section, we focus on the tree representation.
+
+Internal nodes of $U$ are ones that have been expanded (drilled down on), 
 while the leaves are nodes that have not been expanded. 
@@
 \stitle{Probability Distribution.} 
-We assume that we have a probability distribution over leaves, which assigns a probability that each leaf will be the next one to be expanded.
+We assume that we have a probability distribution over $L$, which assigns a probability that $l \in L$ will be the next leaf to be expanded.
 In the absence of additional data, we can assume a uniform probability distribution.
-That is, we can assume that every leaf is equally likely to be expanded next.
-If we have data on past user behaviour available, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `distance from last expanded node' to get a better probability estimate of each node being expanded next.
+If we have data on past user behaviour, we can use Machine Learning on node features such as `node depth in tree', `weight of node rule' and `distance from last expanded node' to get a better probability estimate of each node being expanded next.
 
@@
 
-\stitle{Sampling Strategy.} The sampling strategy we adopt involves storing, 
-for every displayed rule $r'$ in $U$, a sample of $r'$ of size $n_{r'}$, i.e., 
+\stitle{Sampling Strategy.} Our sampling strategy involves storing, 
+for every displayed rule $r' \in U$, a sample of $r'$ of size $n_{r'}$, i.e., 
 containing $n_{r'}$ randomly chosen tuples $\in r'$.
-Note that we may choose to not store a sample for some rules, in which case $n_{r'}$
-will be zero for those rules.
-Also note that we may store samples for $r' \in U \setminus L$, because even though the user cannot expand an internal node of $U$, a sample of an internal $r'$ can help us create samples for each descendant of $r'$ in $L$.
-
-We need to pick the $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory. 
-As it turns out, picking a sample for some rules can help not just that rule, but
-also all of its sub-rules, all the while preserving uniform randomness of samples. We formalize this notion next.
+Note that we may choose to not store a sample for some rules $r'$; then $n_{r'}$ will be set to zero.
+We may store samples for $r' \in U \setminus L$, because even though the user cannot expand an internal node $r'$, a sample of $r'$ can help us create samples for each descendant of $r'$ in $L$.
+
+We need to choose our $n_{r^{\prime}}$ values so as to maximize the probability that the next user drill down can be satisfied using samples available in memory.
+Picking a sample for some rules can help not just that rule, but
+also its sub-rules, while preserving uniform randomness of samples. We formalize this notion next.
 
@@
 Let the `selectivity' of a rule $r$ be the fraction of tuples in $T$ that are covered by $r$. 
-For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples.
-We denote this quantity as $S(r_1, r_2)$.
+For each pair $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of their selectivities using existing samples. We denote this quantity as $S(r_1, r_2)$. 
 We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. 
@@
 The $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child.
-Therefore, each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else.
+So each leaf $r_{i,j}$ gets tuples from its own sample or it's parents' sample, and from nowhere else.
 Thus, $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$.
-In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$.
+In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close), and $S(r_i, r_{i,2}) = 1 - w_i$.
 Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
-Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$.
-So in any optimal solution, $ess(r_{i,1})$ must equal $minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
+Thus, each individual $p_{r_{i,1}}$ is higher than all $p_{r_{i,2}}$s combined, and $M$ is high enough to cover all $r_{i,1}$.
+So in an optimal solution, $ess(r_{i,1}) = minSS$ for all $i$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$.
 For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$).
 The latter option consumes $w_i \times minSS$ extra memory and gives $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ extra probability value.
-So having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (it consumes additional memory $\propto w_i$ and gives additional value $\propto v_i$).
+So having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem.
 Moreover, the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$.
-Hence, solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ solves the instance of the knapsack problem.
+Hence, solving Problem~\ref{prob:sample-sizes} and picking $i$'s for which $ess(r_{i,2}) = minSS$ solves our knapsack problem instance.
 \end{proof}
@@
 
-\stitle{Approximate DP Solution.} The problem as stated is {\sc NP-Hard}, but with an additional simplifying assumption we can make the problem approximately solvable using  Dynamic Programming.
-The assumption is: for each $r \in L$, we assume that its $ess$ can get sample tuples only from samples obtained for itself and its immediate parent.
+\stitle{Approximate DP Solution.} The problem as stated is {\sc NP-Hard}, but with a simplifying assumption we can make the problem approximately solvable using  Dynamic Programming.
+The assumption is: for each $r \in L$, we assume that its $ess$ can get tuples only from samples obtained for itself and its parent.
 That is, we set $S(r_1, r_2)$ to be zero if $r_1 \neq r_2$ and $r_2$ is not a child of $r_1$.
 This is similar to what we had for tree $U$ in our proof of Lemma~\ref{lemma:sampling-hardness}.
-$ess(r)$ now becomes $ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$ where $r$ is the parent of $r^{\prime}$.
+So now $ess(r^{\prime}) = n_{r^{\prime}} + n_rS(r, r^{\prime})$ where $r$ is the parent of $r^{\prime}$.
 
@@
 \techreporttext{
-Now consider a rule in $r_0 \in U \setminus L$ along with all its children.
+Consider a rule $r_0 \in U \setminus L$.
 Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children.
-Then, by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the $ess$ value of rules in $M_{r_0}$.
-This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
-Thus, for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
-Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
-Then, we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s.
-We describe both these steps in detail below:
+By our assumption, the number of tuples $n_{r^{\prime}}$ for any rule$r^{\prime} \in M_{r_0}$ only affects the $ess$ value of rules in $M_{r_0}$.
+This allows us to split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$.
+For each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$.
+Locally optimal means that we cannot get a higher `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$.
+Then we use dynamic programming to combine locally optimal solutions of different $M_{r_0}$s.
+We describe these steps in detail below:
 
 Let $r_0 \in U \setminus L$.
-Let $d$ be the number of leaf children of $r_0$.
-Let the children be $r_1, r_2, ...
-r_d$.
-For any child $r_i$, $n_{r_i}$ only contributes to its own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
+Let $r_1, r_2, ...
+r_d$ be the leaf children of $r_0$.
+For any child $r_i$, $n_{r_i}$ only contributes to its own $ess$, while $n_{r_0}$ contributes to the $ess$ of all children $r_1, ...
 r_d$.
@@
 \squishend
-Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess = minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
+Thus, there are three kinds of children $r_i$: Those with (i) $ess \geq minSS$ but $n_{r_i} = 0$, (ii) $ess < minSS$ and $n_{r_i} = 0$, (iii) $ess = minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$.
 There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution.
@@
 $$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
-Thus, there are at most $3^d$ locally optimal solutions; $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.
+Thus, there are $\leq 3^d$ locally optimal solutions; $d$ is usually small ($\leq k$), even when $U$ is big.
+So we enumerate all locally optimal solutions and find their sampling cost and probability scores.
 
@@
 
-\begin{comment}
 \stitle{Additional optimizations:}
@@
 Instead of storing the entire tuple repeatedly, we could create a dictionary of common tuples, and only store a pointer to the tuple's dictionary entry in $T_s$.
-\end{comment}
 
@@
 \techreporttext{
+\vspace{-5pt}
 \section{Extensions}\label{sec:extensions}
@@
 
+\vspace{-5pt}
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}
@@
 
+\vspace{-10pt}
 \section{Related Work}\label{sec:related}
 
-There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
-This and other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution.
-The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.
+There has been work on finding cubes for OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}.
+This and other work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution.
+The work does not guarantee good coverage of the table, since it rates infrequent sets of values as highly as frequent ones.
 Some other data exploration work~\cite{sellam:meet} focuses on finding attribute values that divide the database in equal sized parts, while we focus on values that occur as frequently as possible.

### 2016-08-03 TableSummarization/journalVersion/smartDrillDown.tex
Added bios

@@
 \usepackage{times}
+\usepackage{wrapfig}
 \usepackage{subfigure}
@@
 \begin{document}
-
-\textbf{Note to Reviewers}: This paper is an extended version of the paper "Interactive Data Exploration with Smart Drill Down" that appeared in the 2016 IEEE International Conference on Data Engineering.
-This extended version contains about $30\%$ new material compared to the original version.
-In particular, the additions include:
-
-\begin{enumerate}
-\item We have added pseudo-code for the `Find Best Marginal Rule' subroutine of the BRS algorithm.
-\item The new Section~\ref{sec:setting-paramters} contains methods for setting good values of our tunable parameters $W$, $k$, and $m_w$ .
-\item Section~\ref{sec:sampling} has been significantly expanded, with details on how to utilize samples stored in memory and when to form new samples.
-We have also added a more detailed description and proof for our approximate dynamic-programming based algorithm for determining the best samples to form.
-\item We have added an alternative convex-optimization based solution for determining the best samples in Section~\ref{sec:sampling}.
-\item Additionally, Section~\ref{sec:sampling} contains additional optimizations such as pre-fetching, and guidelines for setting the sampling parameter $minSS$. 
-\item We added Section~\ref{sec:extensions}, which contains two extensions, namely dealing with numerical attributes, and using aggregates other than Count. 
-\end{enumerate}
-
-\newpage
 
@@
 
-\begin{comment}
 \begin{wrapfigure}{l}{0.15\textwidth}
@@
 Aditya is a recipient of the Arthur Samuel award for the best dissertation in CS at Stanford (2014), the SIGMOD Jim Gray dissertation award (2014), the SIGKDD dissertation award runner up (2014), a Google Faculty Research Award (2015), the Key Scientific Challenges Award from Yahoo!
-Research (2010), three best-of-conference citations (VLDB 2010, KDD 2012 and ICDE 2014), the Terry Groswith graduate fellowship at Stanford (2007), and the Gold Medal in Computer Science at IIT Bombay (2007).
+Research (2010), four best-of-conference citations (VLDB 2010, KDD 2012, ICDE 2014, ICDE 2016), the Terry Groswith graduate fellowship at Stanford (2007), and the Gold Medal in Computer Science at IIT Bombay (2007).
 His research group is supported with funding from by the NIH, the NSF, and Google.
 \medskip
-\end{comment}
 

### 2016-08-06 TableSummarization/journalVersion/smartDrillDown.tex
back to length

@@
 
-\papertext{
+\techreporttext{
 BRS makes multiple passes over the table in order to determine the best set of rules to display.
@@
 
-\techreporttext{
+\papertext{
 BRS makes multiple passes over the table in order to determine the best set of rules to display.
@@
 
-Note that the \textbf{Combine} procedure doesn't really require additional memory apart from the temporary memory used by BRS.
-Since all the tuples in the `new' sample are already present in existing samples, it can give BRS a set of temporary pointers to the tuples, and the memory for the pointers can be freed as soon as the sample has been processed by BRS.
-In contrast, if we had created a new sample from hard disk, we would maintain the sample even after BRS terminated, and would hence need to use memory from the SampleHandler's capacity $M$.
-%TODO: Not sure how clear/necessary this is.
-
 \item \textbf{Create:} If \textbf{Combine} doesn't work either, then the SampleHandler needs to create a new sample $s$ with $f_s = r$ by making a pass through the table.
@@
 Tech in Computer Science and Engineering from the Indian Institute of Technology Bombay, in 2011.
+Manas is a recipient of International Mathematics Olympiad Silver (2007) and Bronze (2006) Medals, and two best-of-conference citations (ICDT 2016, ICDE 2016).
 \medskip
@@
 \begin{wrapfigure}{l}{0.15\textwidth}
-%   \begin{center}
 \includegraphics[height=3.5cm, width=2.7cm]{./figures/aditya.pdf}
-%   \end{center}
 \end{wrapfigure}
-
 Aditya Parameswaran is an Assistant Professor in Computer Science at the University of Illinois (UIUC).
@@
 His research group is supported with funding from by the NIH, the NSF, and Google.
+
 \medskip

### 2016-12-02 TableSummarization/journalVersion/smartDrillDown.tex
Added content based on reviewer comments

@@
 \end{lemma}
+%%%%%%%%%%%%%%%%%%%%
+\begin{proof}
+Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise.
+And consider a weighting function with $W(r) = 1$ if there is at least one $1$ in $r$, and $0$ otherwise.
+Let $k \leq m$.
+
+Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$.
+Moreover, a rule-list that has two rules with a $1$ in the same column is dominated by one that that $1$s in different columns.
+Then the score of a rule list that has rules with $1$s in column set $C \subset \{1, 2,\ldots m\}$ with $|C| = k$ has score equal to the size of union $\bigcup_{c \in C} S_c$.
+Thus maximizing score is equivalent to maximizing the size of the union of $k$ sets, which is the Maximum Coverage Problem.
+Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.
+\end{proof}
+%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{lemma}
-\vspace{-10pt}
+\papertext{\vspace{-10pt}}
+\techreporttext{
+\begin{proof}
+Let $R(S)$ be the rule list obtained by sorting $S$ in descending weight order.
+For each tuple $t$, let $TOP(t,S)$ denote the first rule in rule-list $R(S)$ that covers $t$.
+By definition of $R(S)$, $r$ must also have the highest weight out of all rules in $S$ that cover $t$.
+Then $MCount(r,S)$ is simply $\sum_{t \in T : TOP(t,S) = r} 1$.
+Thus Score can be rewritten as:
+$$\text{Score}(S) = \sum_{t \in T} W(TOP(t,S))$$
+Now let $S \subsetneq S'$ and $s \notin S$.
+Then to prove submodularity, we simply need to prove that $$\text{Score}(S \cup \{s\}) - \text{Score}(S) \geq \text{Score}(S' \cup \{s\}) - \text{Score}(S')$$.
+We have
+$\text{Score}(S \cup \{s\}) - \text{Score}(S) = \sum_{t \in T} W(TOP(t,S \cup \{s\})) - W(TOP(t,S))$ 
+and an analogous equation for $S'$. Consider two cases: 
+\begin{enumerate}
+\item $W(TOP(t,S' \cup \{s\})) - W(TOP(t,S')) > 0$: This means $s$ covers $t$ and has higher weight than any rule in $S'$ that covers $t$.
+That is, $TOP(t,S' \cup \{s\}) = s$.
+Since $S \subseteq S'$, we must have $W(TOP(t,S')) \geq W(TOP(t,S))$, and also $TOP(t,S \cup \{s\}) = s$.
+Thus we have $W(TOP(t,S' \cup \{s\})) - W(TOP(t,S')) \leq W(TOP(t,S \cup \{s\})) - W(TOP(t,S))$.
+\item $W(TOP(t,S' \cup \{s\})) - W(TOP(t,S')) = 0$: In this case, clearly 
+$W(TOP(t,S' \cup \{s\})) - W(TOP(t,S')) \leq W(TOP(t,S \cup \{s\})) - W(TOP(t,S))$ since the latter is non-negative. 
+\end{enumerate}
+Either way, we have $W(TOP(t,S' \cup \{s\})) - W(TOP(t,S')) \leq W(TOP(t,S \cup \{s\})) - W(TOP(t,S))$, and summing over all $t \in T$ gives us 
+$$\text{Score}(S \cup \{s\}) - \text{Score}(S) \geq \text{Score}(S' \cup \{s\}) - \text{Score}(S')$$
+as required.
+\end{proof}
+}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Runtime analysis:}
+Our algorithm runs calls the function for finding best marginal rule $k$ times.
+Thus the total computational cost is about $k$ times the cost of computing the best marginal rule.
+Let $c_n$ be the number of counts in $C_n$ at the end of the $n^{th}$ iteration (that is, after eliminating rules that have no chance of having the highest marginal value).
+Also, let $c_0 = 1$.
+
+The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
+(ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$.
+The number of super-rules we can consider for (i) can be no larger than $c_{n-1}|T|$, since each rule in $C_{n-1}$ can give us at most $|T|$ new candidate super-rules for $C_n$.
+The cost of (ii) is $O(|C_n|)$.
+Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively.
+Thus our total cost is $O(|T|\sum_i c_i)$.
+
+In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$.
+However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.
+
+In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$.
+This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
+Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$.
+If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e.
+not scale with table size), then our cost is $O(|T|)$.
+
+\textbf{Approximation ratio:}
+If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
+On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.
+
+We now bound the sub-optimality of our chosen rule.
+Suppose the best rule to pick is $r^*$, and we pick rule $r$.
+Since we estimated $r^*$'s weight to be at most $m_w$, we would have estimated $r^*$ to have marginal value $\text{Est}(r^*) = \sum_{t \in r^*}(m_w - w(t))$ where $t \in r^*$ iterates over all tuples covered by $r^*$, and $w(t)$ is the weight of the current highest weight rule covering $t$.
+Since we picked $r$ over $r^*$, it means $\text{MarginalValue}(r) \geq \text{Est}(r^*)$.
+Suppose $m$ is the highest weight of a rule selected in the previous steps of BRS (so $w(t) \leq m$ for all $t$).
+Then the actual marginal value of $r^*$ is $\leq \frac{m_w - m}{m_w^* - m}\text{Est}(r^*)$.
+Thus our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick.
+If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \techreporttext{
@@
 }
+
+%%%%%%%%%%%%%%%%%%
+While the setting of $W$ (and hence $m_w$ is mostly specific to the user, we will consider a simple model of how frequently values occur in a table and determine a weighting function to favor rules of a particular size based on that model.
+
+Let us consider a parametric weight function which generalizes the `Bits Weighting Function' from Section~\ref{sec:weighting}. 
+$W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
+The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.
+
+We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
+Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time.
+If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$.
+Thus the score of a rule set containing $r$ alone can be approximated by:
+$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
+
+Now consider the rule that has the highest weight times count according to the above weighting function.
+We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$.
+According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise.
+For each $c$, we have:
+$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$
+
+Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
+Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e.
+highest frequency values.
+In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e.
+values that are frequent relative to the number of distinct values in the column.
+
+If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$.
+This implies that for all $c$, we must have $w_c \propto \ln f_c$.
+Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$).
+This provides an intuitive justification for we might want to use the Bits weighting function.
+
+We can also estimate the size of the optimal rule under the above weighting function.
+Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
+Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.
+
+The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.
+
+We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function.
+But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations.
+As a result, our estimate of the maximum size is a significant under-estimate.
+In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$.
+As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.
+
+%This evaluates to $1.44$, while the observed value in our experiments is $10$. 
+%%%%%%%%%%%%%%%%%%
 
@@
 
-Then next step is to combine the solutions using dynamic programming.
+The% next step is to combine the solutions using dynamic programming.
 Let the $M$ sets be called $M_0, M_1, ...
@@
 \caption{Bits scoring\label{fig:weigtingbitwise}}
-\vspace{-15pt}
 \end{figure}

### 2016-12-04 TableSummarization/journalVersion/smartDrillDown.tex
combined setting parameters with Extensions

@@
 \end{lemma}
-%%%%%%%%%%%%%%%%%%%%
 \begin{proof}
-Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise.
+\mrj{Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise.
 And consider a weighting function with $W(r) = 1$ if there is at least one $1$ in $r$, and $0$ otherwise.
-Let $k \leq m$.
-
-Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$.
+Let $k \leq m$.}
+
+\mrj{Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$.
 Moreover, a rule-list that has two rules with a $1$ in the same column is dominated by one that that $1$s in different columns.
@@
 Thus maximizing score is equivalent to maximizing the size of the union of $k$ sets, which is the Maximum Coverage Problem.
-Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.
+Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.}
 \end{proof}
-%%%%%%%%%%%%%%%%%%%%
 
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-\papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}
-
-% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary.
-Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.
+\papertext{In the Extensions Section of the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in the extensions section, we describe strategies for setting $m_w$ as well as other parameters.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%
 \textbf{Runtime analysis:}
-Our algorithm runs calls the function for finding best marginal rule $k$ times.
+\mrj{Our algorithm runs calls the function for finding best marginal rule $k$ times.
 Thus the total computational cost is about $k$ times the cost of computing the best marginal rule.
 Let $c_n$ be the number of counts in $C_n$ at the end of the $n^{th}$ iteration (that is, after eliminating rules that have no chance of having the highest marginal value).
-Also, let $c_0 = 1$.
-
-The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
+Also, let $c_0 = 1$.}
+
+\mrj{The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
 (ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$.
@@
 Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively.
-Thus our total cost is $O(|T|\sum_i c_i)$.
-
-In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$.
-However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.
-
-In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$.
+Thus our total cost is $O(|T|\sum_i c_i)$.}
+
+\mrj{In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$.
+However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.}
+
+\mrj{In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$.
 This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
@@
 If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e.
-not scale with table size), then our cost is $O(|T|)$.
+not scale with table size), then our cost is $O(|T|)$.}
 
 \textbf{Approximation ratio:}
-If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
-On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.
-
-We now bound the sub-optimality of our chosen rule.
+\mrj{If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
+On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.}
+
+\mrj{We now bound the sub-optimality of our chosen rule.
 Suppose the best rule to pick is $r^*$, and we pick rule $r$.
@@
 Thus our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick.
-If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\techreporttext{
-\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
-Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
-Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
-To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.
-
-Parameter $k$ is the number of new rules to display upon each smart drill-down.
-Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
-Very small values of $k$ may display too little information about the table.
-Fortunately, the BRS algorithm is incremental in nature.
-That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$.
-Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found.
-This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search.
-Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.
-
-$W$ is the weight function that determines which rules are interesting.
-This is a function specified by the user as a black box.
-Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
-In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
-The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
-
-The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
-Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
-We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling.
-We create a small random sample of tuples from the table, and run the BRS algorithm on it.
-Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output.
-To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
-}
-
-%%%%%%%%%%%%%%%%%%
-While the setting of $W$ (and hence $m_w$ is mostly specific to the user, we will consider a simple model of how frequently values occur in a table and determine a weighting function to favor rules of a particular size based on that model.
-
-Let us consider a parametric weight function which generalizes the `Bits Weighting Function' from Section~\ref{sec:weighting}. 
-$W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
-The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.
-
-We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
-Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time.
-If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$.
-Thus the score of a rule set containing $r$ alone can be approximated by:
-$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
-
-Now consider the rule that has the highest weight times count according to the above weighting function.
-We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$.
-According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise.
-For each $c$, we have:
-$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$
-
-Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
-Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e.
-highest frequency values.
-In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e.
-values that are frequent relative to the number of distinct values in the column.
-
-If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$.
-This implies that for all $c$, we must have $w_c \propto \ln f_c$.
-Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$).
-This provides an intuitive justification for we might want to use the Bits weighting function.
-
-We can also estimate the size of the optimal rule under the above weighting function.
-Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
-Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.
-
-The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.
-
-We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function.
-But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations.
-As a result, our estimate of the maximum size is a significant under-estimate.
-In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$.
-As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.
-
-%This evaluates to $1.44$, while the observed value in our experiments is $10$. 
-%%%%%%%%%%%%%%%%%%
+If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.}
 
@@
 \section{Extensions}\label{sec:extensions}
+\techreporttext{
+\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
+Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
+Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user.
+To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.
+
+Parameter $k$ is the number of new rules to display upon each smart drill-down.
+Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information.
+Very small values of $k$ may display too little information about the table.
+Fortunately, the BRS algorithm is incremental in nature.
+That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$.
+Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found.
+This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search.
+Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.
+
+$W$ is the weight function that determines which rules are interesting.
+This is a function specified by the user as a black box.
+Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
+In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
+The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
+
+The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
+Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time.
+We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling.
+We create a small random sample of tuples from the table, and run the BRS algorithm on it.
+Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output.
+To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
+
+\mrj{We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
+Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
+The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.}
+
+\mrj{We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
+Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time.
+If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$.
+Thus the score of a rule set containing $r$ alone can be approximated by:
+$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
+}
+
+\mrj{Now consider the rule that has the highest weight times count according to the above weighting function.
+We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$.
+According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise.
+For each $c$, we have:
+$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$}
+
+\mrj{Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
+Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e.
+highest frequency values.
+In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e.
+values that are frequent relative to the number of distinct values in the column.}
+
+\mrj{If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$.
+This implies that for all $c$, we must have $w_c \propto \ln f_c$.
+Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$).
+This provides an intuitive justification for we might want to use the Bits weighting function.}
+
+\mrj{We can also estimate the size of the optimal rule under the above weighting function.
+Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
+Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.}
+
+\mrj{The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.}
+
+\mrj{We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function.
+But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations.
+As a result, our estimate of the maximum size is a significant under-estimate.
+In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$.
+As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.}
+}
+%This evaluates to $1.44$, while the observed value in our experiments is $10$.
+
 \subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
-Our framework assumes that all attributes are categorical.
+\mrjdel{Our framework assumes that all attributes are categorical.
 Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
-Thus our algorithm does not summarize information about numerical attributes.
-
-However, we can modify the algorithm to deal with numerical attributes.
+Thus our algorithm does not summarize information about numerical attributes.}
+
+\mrjdel{However, we can modify the algorithm to deal with numerical attributes.
 Suppose we have a numerical attribute $A$.
@@
 Then we run BRS, treating the bucket id as a categorical attribute.
-This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).
+This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}
+
+\mrj{Our framework assumes that all attributes are categorical.
+Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
+Thus our algorithm does not summarize information about numerical attributes.
+But it can easily be modified to do so by bucketizing a numerical attribute and treating the bucket id as a categorical attribute.
+This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}
 
@@
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}
-Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
+\mrjdel{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
 However, if we have a numerical column (i.e.
@@
 The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$.
-Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.
+Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
+
+\mrj{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
+However, if we have a numerical column (i.e.
+a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead.
+We can define $Sum$ and $MSum$ for a rule analogously to $Count$ and $MCount$, and $Score$ can be modified to use $MSum$ instead of $MCount$.
+We can then modify Algorithm~\ref{algo:best-rule-set} to maximize the new score by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
 }

### 2016-12-04 TableSummarization/journalVersion/smartDrillDown.tex
missing red color

@@
 \begin{proof}
+\mrj{
 Let $R(S)$ be the rule list obtained by sorting $S$ in descending weight order.
@@
 $$\text{Score}(S \cup \{s\}) - \text{Score}(S) \geq \text{Score}(S' \cup \{s\}) - \text{Score}(S')$$
-as required.
+as required.}
 \end{proof}

### 2016-12-07 TableSummarization/journalVersion/smartDrillDown.tex
typos, moved some stuff to techreport

--- 2016-12-04 TableSummarization/journalVersion/smartDrillDown.tex
@@
 \newcommand{\eat}[1]{}
-\newcommand{\papertext}[1]{}
-\newcommand{\techreporttext}[1]{#1}
+\newcommand{\papertext}[1]{#1}
+\newcommand{\techreporttext}[1]{}
 
@@
 We discuss how to modify our algorithms to use $Sum$ instead of $Count$
-\papertext{in our technical report~\cite{tr}.}
-\techreporttext{in Section~\ref{sec:extensions}.}
+\eat{in our technical report~\cite{tr}.}
+\papertext{in Section~\ref{sec:extensions}.}
 
@@
 We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ depending on how expressive the rule is (i.e., how much information
-it conveys). We mention some canonical forms for function $W(r)$ \papertext{(their detailed interpretations are described in the technical report)}; later, we specify the full
+it conveys). We mention some canonical forms for function $W(r)$ \eat{(their detailed interpretations are described in the technical report)}; later, we specify the full
 family of weighting functions our techniques can handle:
@@
 \begin{comment}
-\techreporttext{
+\papertext{
 \begin{table}
@@
 \begin{comment}
-\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
+\papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
 Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule.
@@
 \end{comment}
-%\papertext{
+%\eat{
 $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
@@
 
-\techreporttext{To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
+\papertext{To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules.
 Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule.
@@
 \begin{comment}
-\techreporttext{
+\papertext{
 $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
@@
 \end{comment}
-%\papertext{
+%\eat{
 $W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$.
@@
 By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.
-%\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
+%\papertext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}.
 Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.}
@@
 This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$.
-%\techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
+%\papertext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.
 This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.}
 Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. 
-\techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}.
+\papertext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}.
 This operation is equivalent to a traditional roll up, but for smart drill downs instead of traditional drill downs.}
@@
 
-\techreporttext{
+\papertext{
 \begin{table}
@@
 We assume that all columns are categorical (so numerical columns have been bucketized beforehand).
-\papertext{We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}}\techreporttext{We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}}.
+\eat{We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}}\papertext{We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}}.
 
@@
 \end{lemma}
-\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}.}\techreporttext{
+\eat{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}.}\papertext{
 \begin{proof}
@@
 \end{lemma}
+Proofs can be found in the technical report~\cite{tr}.
+\techreporttext{
 \begin{proof}
-\mrj{Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise.
+Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise.
 And consider a weighting function with $W(r) = 1$ if there is at least one $1$ in $r$, and $0$ otherwise.
-Let $k \leq m$.}
-
-\mrj{Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$.
+Let $k \leq m$.
+
+Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$.
 Moreover, a rule-list that has two rules with a $1$ in the same column is dominated by one that that $1$s in different columns.
@@
 Thus maximizing score is equivalent to maximizing the size of the union of $k$ sets, which is the Maximum Coverage Problem.
-Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.}
+Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.
 \end{proof}
+}
 
@@
 \end{lemma}
-\papertext{\vspace{-10pt}}
 \techreporttext{
 \begin{proof}
+
 Let $R(S)$ be the rule list obtained by sorting $S$ in descending weight order.
@@
 Then to prove submodularity, we simply need to prove that $$\text{Score}(S \cup \{s\}) - \text{Score}(S) \geq \text{Score}(S' \cup \{s\}) - \text{Score}(S')$$.
-We have
+This gives us
 $\text{Score}(S \cup \{s\}) - \text{Score}(S) = \sum_{t \in T} W(TOP(t,S \cup \{s\})) - W(TOP(t,S))$ 
@@
 Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small.
-\papertext{In the Extensions Section of the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in the extensions section, we describe strategies for setting $m_w$ as well as other parameters.}
+\eat{In the Extensions Section of the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \papertext{Later in the extensions section, we describe strategies for setting $m_w$ as well as other parameters.}
 
@@
 
-\techreporttext{
+\papertext{
 \begin{algorithm}
@@
 We now describe the function to find the best marginal rule.
-\papertext{The pesudo-code for the function is in the technical report}\techreporttext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}}.
+\eat{The pesudo-code for the function is in the technical report}\papertext{The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}}.
 The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far.
-The function makes several passes over the table \techreporttext{(Step $3$)}, counting marginal values for size $j$ rules in the $j^{th}$ pass.
+The function makes several passes over the table \papertext{(Step $3$)}, counting marginal values for size $j$ rules in the $j^{th}$ pass.
 We maintain three sets of rules: $C$, the set of rules whose marginal values have been counted in all previous passes, $C_n$, the set of rules whose marginal values will be counted in the current pass, and $C_o$, the set of rules whose marginal values were counted in the previous pass.
@@
 
+\noindent
 \textbf{Runtime analysis:}
-\mrj{Our algorithm runs calls the function for finding best marginal rule $k$ times.
+\mrj{Our algorithm calls the function for finding best marginal rule $k$ times.
 Thus the total computational cost is about $k$ times the cost of computing the best marginal rule.
@@
 Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively.
-Thus our total cost is $O(|T|\sum_i c_i)$.}
+Thus, our total cost is $O(|T|\sum_i c_i)$.}
 
@@
 If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e.
-not scale with table size), then our cost is $O(|T|)$.}
-
+they do not scale with table size), then our cost is $O(|T|)$.}
+
+\noindent
 \textbf{Approximation ratio:}
-\mrj{If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
+\mrj{If $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
 On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.}
@@
 Suppose $m$ is the highest weight of a rule selected in the previous steps of BRS (so $w(t) \leq m$ for all $t$).
-Then the actual marginal value of $r^*$ is $\leq \frac{m_w - m}{m_w^* - m}\text{Est}(r^*)$.
-Thus our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick.
-If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.}
+Then, the actual marginal value of $r^*$ is $\leq \frac{m_w - m}{m_w^* - m}\text{Est}(r^*)$.
+Thus, our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick.
+If we always choose  our $m_w$ such that we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be within a $w(1-1/e)$ factor of the optimal score.}
 
@@
 In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process.
-Detailed descriptions of the sampling problem and the SampleHandler are in the technical report.
-The technical report also contains some additional optimizations, and ways to set the minimum sample size.
 
@@
 
-\papertext{
+\eat{
 BRS makes multiple passes over the table in order to determine the best set of rules to display.
@@
 
-\papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
+\eat{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
 \squishlist
@@
 
-\techreporttext{
+\papertext{
 Consider a rule $r_0 \in U \setminus L$.
@@
 
-\techreporttext{
+\papertext{
 \subsection{Alternative Convex-Optimization based solution}\label{sec:alternative-convex}
@@
 \subsection{Design of the SampleHandler}\label{sec:sample-handler}
-\papertext{
+\eat{
 We now briefly describe the design of the {\em SampleHandler}, which given
@@
 
-\techreporttext{
+\papertext{
 We now describe the design of the {\em SampleHandler}, which given
@@
 
-\techreporttext{
+\papertext{
 \noindent \textbf{Pre-fetching:} When the user clicks on rule $r$ (or on a $\star$ in $r$), we need to get a sample, run the BRS, and display a rule-list to the user.
@@
 
-A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\papertext{ (see technical report for details)}\techreporttext{.
+A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\eat{ (see technical report for details)}\papertext{.
 Specifically, if we want to perform a regular drill down on a column $C$ that has $n$ distinct values, then we set the number of rules to be displayed ($k$) to $n$.
@@
 
-\techreporttext{
+\papertext{
 \vspace{-5pt}
 \section{Extensions}\label{sec:extensions}
-\techreporttext{
+\papertext{
 \subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
@@
 \mrj{We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
-Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time.
+Suppose that for each column $c$, the most frequent value occurs a $f_c$ fraction of the time.
 If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$.
@@
 We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$.
-According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise.
+According to the KKT conditions~\cite{Boyd:2004:CO:993483}, at the optimum $Weight \times Count$, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise.
 For each $c$, we have:
@@
 
-\mrj{Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
-Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e.
-highest frequency values.
-In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e.
-values that are frequent relative to the number of distinct values in the column.}
+\mrj{Thus, if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
+Thus, for Size weighting, the highest score rule should select columns with high values of $\ln f_c$, i.e., the highest frequency values.
+In contrast, Bits weighting will select columns with highest values of $\log_{|c|}(f_c)$, i.e., values that are frequent relative to the number of distinct values in the column.}
 
@@
 Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$).
-This provides an intuitive justification for we might want to use the Bits weighting function.}
+This provides an intuitive justification for why we might want to use the Bits weighting function.}
 
@@
 Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
-Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.}
+Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.}
 
@@
 \subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
-\mrjdel{Our framework assumes that all attributes are categorical.
-Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
-Thus our algorithm does not summarize information about numerical attributes.}
-
-\mrjdel{However, we can modify the algorithm to deal with numerical attributes.
-Suppose we have a numerical attribute $A$.
-We can create buckets for values of $A$.
-We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, with one bucket per interval.
-We can create buckets having equal range sizes, or decide their range such that there is an approximately equal number of tuples in each bucket.
-Then we run BRS, treating the bucket id as a categorical attribute.
-This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}
-
 \mrj{Our framework assumes that all attributes are categorical.
@@
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}
-\mrjdel{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
-However, if we have a numerical column (i.e.
-a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead.
-Suppose we are given a measure column $c_m$.
-Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule.
-{\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$.
-The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$.
-Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
-
 \mrj{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.

### 2016-12-11 TableSummarization/journalVersion/smartDrillDown.tex
hector comments

@@
 \textbf{Runtime analysis:}
-\mrj{Our algorithm calls the function for finding best marginal rule $k$ times.
+Our algorithm calls the function for finding best marginal rule $k$ times.
 Thus the total computational cost is about $k$ times the cost of computing the best marginal rule.
 Let $c_n$ be the number of counts in $C_n$ at the end of the $n^{th}$ iteration (that is, after eliminating rules that have no chance of having the highest marginal value).
-Also, let $c_0 = 1$.}
-
-\mrj{The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
+Also, let $c_0 = 1$.
+
+The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
 (ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$.
@@
 Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively.
-Thus, our total cost is $O(|T|\sum_i c_i)$.}
-
-\mrj{In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$.
-However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.}
-
-\mrj{In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$.
+Thus, our total cost is $O(|T|\sum_i c_i)$.
+
+In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$.
+However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.
+
+If fact, say that values in columns occur independently, and let $x$ be the frequency of the most common value (note $x<1$).
+Then the number of candidate rules is bounded by $c_i \le c_1 x^i$, where $c_1$ is the number of distinct values in the table.
 This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
-Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$.
-If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e.
-they do not scale with table size), then our cost is $O(|T|)$.}
+Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$. If $c_1$, $k$ and $x$ are taken to be constants (i.e. they do not scale with table size), then our cost is $O(|T|)$.
 
@@
 \textbf{Approximation ratio:}
-\mrj{If $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
-On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.}
-
-\mrj{We now bound the sub-optimality of our chosen rule.
+If $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$.
+On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.
+
+We now bound the sub-optimality of our chosen rule, for the case where $m_w < m_w^*$.
 Suppose the best rule to pick is $r^*$, and we pick rule $r$.
-Since we estimated $r^*$'s weight to be at most $m_w$, we would have estimated $r^*$ to have marginal value $\text{Est}(r^*) = \sum_{t \in r^*}(m_w - w(t))$ where $t \in r^*$ iterates over all tuples covered by $r^*$, and $w(t)$ is the weight of the current highest weight rule covering $t$.
+Since we estimated $r^*$'s weight to be at most $m_w$, we would have estimated $r^*$ to have marginal value $\text{Est}(r^*) = \sum_{t \in r^*}(m_w - W(t))$ where $t \in r^*$ iterates over all tuples covered by $r^*$, and $w(t)$ is the weight of the current highest weight rule covering $t$.
 Since we picked $r$ over $r^*$, it means $\text{MarginalValue}(r) \geq \text{Est}(r^*)$.
@@
 Thus, our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick.
-If we always choose  our $m_w$ such that we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be within a $w(1-1/e)$ factor of the optimal score.}
+If we always choose  our $m_w$ such that we have $\frac{m_w - m}{m_w^* - m} \geq a$, for some constant $a < 1$, then our final score is guaranteed to be within a $a(1-1/e)$ factor of the optimal score.
 
@@
 
-We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms} through Section~\ref{sec:alternative-convex}.
+We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms}\techreporttext{ through Section~\ref{sec:alternative-convex}}.
 Then, in Section~\ref{sec:sample-handler}, we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
@@
 
-\papertext{
+\techreporttext{
 \subsection{Alternative Convex-Optimization based solution}\label{sec:alternative-convex}
@@
 
-\mrj{We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
+We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
 Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
-The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.}
-
-\mrj{We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
+The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.
+
+We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule.
 Suppose that for each column $c$, the most frequent value occurs a $f_c$ fraction of the time.
@@
 $$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
-}
-
-\mrj{Now consider the rule that has the highest weight times count according to the above weighting function.
+
+Now consider the rule that has the highest weight times count according to the above weighting function.
 We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$.
@@
 For each $c$, we have:
-$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$}
-
-\mrj{Thus, if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
+$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$
+
+Thus, if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
 Thus, for Size weighting, the highest score rule should select columns with high values of $\ln f_c$, i.e., the highest frequency values.
-In contrast, Bits weighting will select columns with highest values of $\log_{|c|}(f_c)$, i.e., values that are frequent relative to the number of distinct values in the column.}
-
-\mrj{If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$.
+In contrast, Bits weighting will select columns with highest values of $\log_{|c|}(f_c)$, i.e., values that are frequent relative to the number of distinct values in the column.
+
+If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$.
 This implies that for all $c$, we must have $w_c \propto \ln f_c$.
 Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$).
-This provides an intuitive justification for why we might want to use the Bits weighting function.}
-
-\mrj{We can also estimate the size of the optimal rule under the above weighting function.
+This provides an intuitive justification for why we might want to use the Bits weighting function.
+
+We can also estimate the size of the optimal rule under the above weighting function.
 Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
-Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.}
-
-\mrj{The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.}
-
-\mrj{We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function.
+Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.
+
+The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.
+
+We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function.
 But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations.
@@
 In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$.
-As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.}
+As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.
 }
@@
 \subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
-\mrj{Our framework assumes that all attributes are categorical.
+Our framework assumes that all attributes are categorical.
 Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
@@
 But it can easily be modified to do so by bucketizing a numerical attribute and treating the bucket id as a categorical attribute.
-This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}
+This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).
 
@@
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}
-\mrj{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
+Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
 However, if we have a numerical column (i.e.
@@
 We can then modify Algorithm~\ref{algo:best-rule-set} to maximize the new score by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
-}
 

### 2016-12-18 TableSummarization/journalVersion/cover-letter.tex
Changed techreport contents, added letter

\documentclass[11pt,a4paper]{letter}

\usepackage{microtype} % Improves typography
\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage{lmodern}

\textwidth 6.75in
\textheight 11.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0in

\begin{document}
\begin{letter}
{{\bf Subject:} {\it TKDE submission, titled ``Interactive Data Exploration with Smart Drill-Down (Extended Version)''}}
\opening{To the Editors, and the Reviewers:}

We would like to thank the reviewers for their insightful comments for our TKDE submission, titled ``Interactive Data Exploration with Smart Drill-Down (Extended Version)''. We have made changes to our paper based on the suggestions made by the reviewers.

The primary concern raised by the reviewers and the editor was
that the additions in Section 3.6 and Section 6 were relatively straightforward. To address this, we have added an analysis of a general family of weighting functions, that includes our Size and Bits weighting functions as special cases. Our analysis provides theoretical justification for these two weighting functions. We also estimate the right values for parameters $k$ and $m_w$ for this family. This has been added to the parameter settings section (formerly 3.6), which has now been combined with Section 6.

Other major changes in our revision include: 
\begin{enumerate}
\item Reviewer 3 pointed out that we were missing proofs for some lemmas. This apologize for this oversight; proofs for the sub-modularity lemma and the NP-hardness results have now been added to our arxiv technical report. Unfortunately, we could not make space for these proofs in the main manuscript.
\item Reviewer 3 also mentioned that we did not have a formal runtime and approximation ratio analysis for the BRS algorithm. We have now added this analysis at the end of Section 3.5.
\item In order to make space for the new additions to Section 6 and the runtime and approximation analyses, we have moved the `Setting $minSS$' part of Section 4.2, and the Qualitative experiments with weighting functions (Section 5.1.2) to the technical report.
\item Our extensions in Section 6 were straightforward, as pointed out by the reviewers. So we have shortened the previous contents of Section 6, and combined them with the (augmented) Section 3.6.
\end{enumerate}

In addition, Reviewer 3 commented on how weight and count seem to favor different kinds of rules. This was intentional on our part. There is a trade-off between a rule's coverage and how informative it is of the portion it covers, as pointed out by the reviewer. Very small and very large rules are both bad (due to being too vague/based on only a tiny part of the table respectively). So we attempt to have the Score function strike a balance between these two aspects. 
%Reviewer 3 had also mentioned comparing Smart Drill-Down with OLAP drill-down. Our paper contains an example of a regular drill-down on the Age column of our dataset, but we could .

\closing{Sincerely, \\
The Authors}
\end{letter}
\end{document}

### 2016-12-18 TableSummarization/journalVersion/smartDrillDown.tex
Changed techreport contents, added letter

@@
 
-We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms}\techreporttext{ through Section~\ref{sec:alternative-convex}}.
+We describe our technique to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk in Section~\ref{sec:sampling_algorithms} through Section~\ref{sec:alternative-convex}.
 Then, in Section~\ref{sec:sample-handler}, we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints.
@@
 
-\techreporttext{
 \subsection{Alternative Convex-Optimization based solution}\label{sec:alternative-convex}
@@
 
+\techreporttext{
+
 \stitle{Setting $minSS$:}
@@
 We first perform a qualitative study of smart drill down.
-We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
+We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset)\techreporttext{, and then try out different weight functions to study their effects}.
+\papertext{In the technical report, we try out different weight functions to study their effects.}
 \subsubsection{Testing the User Interface}
@@
 
+\techreporttext{
 \subsubsection{Weighting functions}
@@
 Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.
-
+}
 %TODO: Give application examples for different weighting functions
@@
 
+\techreporttext{
 \begin{figure}
@@
 \end{figure}
+}
 
@@
 
-We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
+\noindent
+\textbf{Generalizing our weight functions:} We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
 Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
@@
 
+Hector Garcia-Molina is the Leonard Bosack and Sandra Lerner Professor in the Departments of Computer Science and Electrical Engineering at Stanford University, Stanford, California. 
 \begin{wrapfigure}{l}{0.15\textwidth}
@@
 \end{wrapfigure}
-
-Hector Garcia-Molina is the Leonard Bosack and Sandra Lerner Professor in the Departments of Computer Science and Electrical Engineering at Stanford University, Stanford, California.
 He was the chairman of the Computer Science Department from January 2001 to December 2004.

### 2016-12-18 TableSummarization/journalVersion/smartDrillDown.tex
serious typos

@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/education_expansion.jpg}
 \caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
@@
 This opens up Figure~\ref{fig:uiexamplestar} that shows the number of females with different levels of education, for the $4$ most frequent levels of education among females.
-Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed Figure~\ref{fig:uiexamplerule}.
+\techreporttext{Instead of expanding the `Education' column, if the user had simply expanded the third rule, it would have displayed Figure~\ref{fig:uiexamplerule}.
+}
 
@@
 As a result, we can see that our system only displays rules having $2$ or $3$ non-$\star$ values, unlike Figure~\ref{fig:uiexample1} which has two rules displaying the total number of males and females, that have size $1$.
-
-A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\eat{ (see technical report for details)}\papertext{.
+}
+
+A regular drill down is a special case of smart drill-down with the right weighting function and number of rules\papertext{ (see technical report for details)}\techreporttext{.
 Specifically, if we want to perform a regular drill down on a column $C$ that has $n$ distinct values, then we set the number of rules to be displayed ($k$) to $n$.
@@
 Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.
-}
 %TODO: Give application examples for different weighting functions
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/rule_expansion.jpg}
 \caption{A rule expansion \label{fig:uiexamplerule}}
@@
 \subsection{Using Sum instead of Count}\label{sec:extensions-sum}
-Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary.
-However, if we have a numerical column (i.e.
-a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead.
-We can define $Sum$ and $MSum$ for a rule analogously to $Count$ and $MCount$, and $Score$ can be modified to use $MSum$ instead of $MCount$.
+We defined the total score of a rule-list using the marginal counts of rules in the list, and display the counts of rules in our table summary.
+However, if we have a numerical (i.e.
+a `measure') column in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column.
+We can define $Sum$ and $MSum$ analogously to $Count$ and $MCount$, and $Score$ can be modified to use $MSum$ instead of $MCount$.
 We can then modify Algorithm~\ref{algo:best-rule-set} to maximize the new score by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
@@
 
+\vspace{-10pt}
 \section{Conclusion}\label{sec:conclusion}
@@
 He holds an honorary PhD from ETH Zurich (2007).
-Garcia-Molina is a Fellow of the Association for Computing Machinery and of the American Academy of Arts and Sciences; is a member of the National Academy of Engineering; received the 1999 ACM SIGMOD Innovations Award; is a Venture Advisor for Onset Ventures, is a member of the Board of Directors of Oracle, and is a member of the State Farm Technical Advisory Council.
+He is a Fellow of the Association for Computing Machinery and of the American Academy of Arts and Sciences; is a member of the National Academy of Engineering; received the 1999 ACM SIGMOD Innovations Award; is a Venture Advisor for Onset Ventures, is a member of the Board of Directors of Oracle, and is a member of the State Farm Technical Advisory Council.
 \medskip

### 2017-03-10 TableSummarization/journalVersion/smartDrillDown.tex
Addressed comments, changed title, change image formats

@@
 
-\title{Interactive Data Exploration with Smart Drill-Down (Extended Version)}
+\title{Interactive Data Exploration with Smart Drill-Down}
 
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.png}
 \caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/education_expansion.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/education_expansion.png}
 \caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/rule_expansion.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/rule_expansion.png}
 \caption{A rule expansion \label{fig:uiexamplerule}}
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.png}
 \caption{A regular drill down on Age\label{fig:drilldownexample}}
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.png}
 \caption{Bits scoring\label{fig:weigtingbitwise}}
@@
 \centering
-\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.jpg}
+\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.png}
 \caption{Size minus one weighting \label{fig:weightingsizeminusone}}
@@
 \papertext{
-\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
+\subsection{Setting parameters $k$, $m_w$, $W$}\label{sec:setting-paramters}
 Our system allows the user to tune the smart drill-down by adjusting a number of parameters.
@@
 
-$W$ is the weight function that determines which rules are interesting.
-This is a function specified by the user as a black box.
-Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
-In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
-The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
-
 The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time.
@@
 
+$W$ is the weight function that determines which rules are interesting.
+This is a function specified by the user as a black box.
+Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
+In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
+The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
+
 \noindent
 \textbf{Generalizing our weight functions:} We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them.
-Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
-The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.
+Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^{\mathsf{p}}$, where the $w_c$'s and $\mathsf{p}$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise.
+The Size and Bits weighting functions can be seen as special cases of the above function with $\mathsf{p}=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.
 
@@
 Thus the score of a rule set containing $r$ alone can be approximated by:
-$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
+$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^{\mathsf{p}}|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
 
@@
 For each $c$, we have:
-$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$
-
-Thus, if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
+$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{\mathsf{p}w_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$
+
+Thus, if $x = \frac{\mathsf{p}}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$.
 Thus, for Size weighting, the highest score rule should select columns with high values of $\ln f_c$, i.e., the highest frequency values.
@@
 We can also estimate the size of the optimal rule under the above weighting function.
-Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$.
-Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.
-
-The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.
+Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-\mathsf{p}/(\sum_{c \in C} \ln f_c)$.
+Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $\mathsf{p} = -s\sum_{c \in C} \ln f_c$.
+
+The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{\mathsf{p}w_c})^{\mathsf{p}}$.
 

### 2017-03-10 TableSummarization/journalVersion/smartDrillDown.tex
Aditya comments

@@
 giving ``surprising'' information to the user, i.e., information
-that would minimize the Kullback-Liebler(KL) divergence between the
+that would minimize the Kullback-Liebler (KL) divergence between the
 resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
@@
 
-The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration.
-(ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$.
+The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration; (ii) Determining which of the rules in $C_n$ to eliminate; (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$.
 The number of super-rules we can consider for (i) can be no larger than $c_{n-1}|T|$, since each rule in $C_{n-1}$ can give us at most $|T|$ new candidate super-rules for $C_n$.
@@
 This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
-Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$. If $c_1$, $k$ and $x$ are taken to be constants (i.e. they do not scale with table size), then our cost is $O(|T|)$.
+Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$. If $c_1$, $k$, and $x$ are taken to be constants (i.e., they do not scale with table size), then our cost is $O(|T|)$.
 
@@
 Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$.
-So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$.
+So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we have to decide which $i$'s also have $ess(r_{i,2}) = minSS$.
 Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value.
@@
 
-The% next step is to combine the solutions using dynamic programming.
+Next, we combine the solutions using dynamic programming.
 Let the $M$ sets be called $M_0, M_1, ...
-M_D$.
-Let our possible sample sizes range from $0$ to $\mathcal{S}$.
-The number of sample sizes can be pretty large ($\mathcal{S}$), but we can make it smaller by discretizing the sample sizes, say to have granularity $100$.
+M_D$, while possible sample sizes range from $0$ to $\mathcal{S}$.
+The number of sample sizes can be pretty large, but we can make it smaller by discretizing, say to have granularity $100$.
 Then we create a $D \times \mathcal{S}$ array $A$.
@@
 $$A\left[i+1\right] \left[j \right] = \textrm{max} (A\left[i\right]\left[j\right], \textrm{max}_{e \in E_{i+1}}(A\left[i\right]\left[j-S(e)\right] + P(e)))$$
-Dynamic programming solves this in $O(D\mathcal{S}3^d)$ time. 
+Dynamic programming solves this recurrence in $O(D\mathcal{S}3^d)$ time. 
 }
@@
 The columns (in order) are as follows:
-annual household income, gender, marital status, age, education, occupation, time lived in the Bay Area, dual incomes?, persons in household, persons in household under $18$, householder status, type of home, ethnic classification, language most spoken in home.
+annual household income, gender, marital status, age, education, occupation, time lived in the Bay Area, dual incomes status, persons in household, persons in household under $18$, householder status, type of home, ethnic classification, and language most spoken in home.
 
@@
 Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page.
-We use the current implementation of our the smart drill down operator, and insert cropped screenshots of its output in this paper.
+We use the current implementation of our smart drill down operator, and insert cropped screenshots of its output in this paper.
 We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}).
@@
 Thus, for size scoring on the Marketing dataset, according to Figure~\ref{fig:uiexample1}, we require $m_w \geq 3$.
-For the second weighting function, according to Figure~\ref{fig:weigtingbitwise}, the minimum required value of $m_w$ is $10$.
+For the second weighting function, \techreporttext{according to Figure~\ref{fig:weigtingbitwise}, }the minimum required value of $m_w$ is $10$.
 At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively.
@@
 $W$ is the weight function that determines which rules are interesting.
-This is a function specified by the user as a black box.
 Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu.
-In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface.
-The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.
 
@@
 Our framework assumes that all attributes are categorical.
-Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries.
-Thus our algorithm does not summarize information about numerical attributes.
+It does not summarize numerical values.
 But it can easily be modified to do so by bucketizing a numerical attribute and treating the bucket id as a categorical attribute.
